### 2015-10-22 writing/gist/gist.tex
Working paper.

\documentclass[11pt,letterpaper]{article}
\usepackage{acl2015}
\usepackage{times}
\usepackage{latexsym}
% \setlength\titlebox{5cm}    % Expanding the titlebox

%%% Custom additions %%%
\usepackage{url}
\usepackage[leqno, fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{qtree}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{color}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{ifthen}
\usepackage{framed}

\newcount\colveccount
\newcommand*\colvec[1]{
        \global\colveccount#1
        \begin{bmatrix}
        \colvecnext
}
\def\colvecnext#1{
        #1
        \global\advance\colveccount-1
        \ifnum\colveccount>0
                \\
                \expandafter\colvecnext
        \else
                \end{bmatrix}
        \fi
}

\newcommand{\nateq}{\equiv}
\newcommand{\natind}{\mathbin{\#}}
\newcommand{\natneg}{\mathbin{^{\wedge}}}
\newcommand{\natfor}{\sqsubset}
\newcommand{\natrev}{\sqsupset}
\newcommand{\natalt}{\mathbin{|}}
\newcommand{\natcov}{\mathbin{\smallsmile}}

\newcommand{\plneg}{\mathop{\textit{not}}}
\newcommand{\pland}{\mathbin{\textit{and}}}
\newcommand{\plor}{\mathbin{\textit{or}}}

\newcommand{\shift}{\textsc{shift}}
\newcommand{\reduce}{\textsc{reduce}}

% Strikeout
\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}

\newcommand{\True}{\texttt{T}}
\newcommand{\False}{\texttt{F}}
\usepackage{stmaryrd}
\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}

\newcommand{\mynote}[1]{{\color{blue}#1}}
\newcommand{\tbchecked}[1]{{\color{red}#1}}

\usepackage{gb4e}
\noautomath

\def\ii#1{\textit{#1}}
\newcommand{\word}[1]{\emph{#1}}
\newcommand{\fulllabel}[2]{\b{#1}\newline\textsc{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Code to simulate natbib's citealt, which prints citations with
%%%%% no parentheses:

\makeatletter
\def\citealt{\def\citename##1{{\frenchspacing##1} }\@internalcitec}
\def\@citexc[#1]#2{\if@filesw\immediate\write\@auxout{\string\citation{#2}}\fi
  \def\@citea{}\@citealt{\@for\@citeb:=#2\do
    {\@citea\def\@citea{;\penalty\@m\ }\@ifundefined
       {b@\@citeb}{{\bf ?}\@warning
       {Citation `\@citeb' on page \thepage \space undefined}}%
{\csname b@\@citeb\endcsname}}}{#1}}
\def\@internalcitec{\@ifnextchar [{\@tempswatrue\@citexc}{\@tempswafalse\@citexc[]}}
\def\@citealt#1#2{{#1\if@tempswa, #2\fi}}
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% %%%

\title{NSHRDLU?}

\author{
Samuel R.\ Bowman$^{\ast\dag}$ \\
\texttt{sbowman@stanford.edu} \\
\And
Jon Gauthier$^{\dag\ddag}$ \\
\texttt{angeli@stanford.edu} \\
\AND
Christopher Potts$^{\ast}$\\
\texttt{cgpotts@stanford.edu}
\And
Christopher D.\ Manning$^{\ast\dag\S}$\\
\texttt{manning@stanford.edu}\\
\AND\\[-3ex]
{$^{\ast}$Stanford Linguistics\quad
$^{\dag}$Stanford NLP Group}\\
{$^{\ddag}$Stanford Symbolic Systems\quad
$^{\S}$Stanford Computer Science}
}

\date{}

\makeatletter
\newcommand{\@BIBLABEL}{\@emptybiblabel}
\newcommand{\@emptybiblabel}[1]{}
\definecolor{black}{rgb}{0,0,0}
\makeatother
\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black]{hyperref}

\def\t#1{#1}
\def\b#1{\t{\textbf{#1}}}
\def\colspaceS{2.25mm}
\def\colspaceM{4.0mm}
\def\colspaceL{4.25mm}

\newcommand\todo[1]{\textcolor{blue}{\textbf{TODO:} #1}}

\begin{document}
\maketitle

This project aims to build neural network models that can jointly learn to parse sentences and to use those parses to guide semantic composition.

Table~\ref{models-table} shows the sequence of model designs that we plan to build.

\begin{table*}[t]
  \centering\small
  \begin{tabular}{ccccc} 
    \toprule
    Name & Stack Representation & Input Representation & Ops Classifier & Op Predictions Used in Training \\ 
    \midrule
Model 0 & Discrete & Op. sequence & N & -- \\
Model 1 & Discrete & \bf Discrete Buffer & \bf Y: Directly supervised & \bf N \\
Model 2 & Discrete & Discrete Buffer & Y: Directly supervised & \bf Y \\
Model 3 & \bf Soft & \bf Soft Buffer & Y: Directly supervised & Y \\
Model 4 & Soft & Soft Buffer & \bf Y: Indirectly supervised & Y \\
    \bottomrule
  \end{tabular}
  \protect\caption{\protect\label{models-table}Model variants, ordered by increasing reliance on learning. Bolding indicates the differences between each model and its parent model.} 
\end{table*}

\section{Models}

\subsection{Model 0}

Model 0 in its simplest form computes the same function as a plain TreeRNN, but uses a stack and a recurrent input reader instead of a tree as its graph structure. \todo{Model 0 figure.} We expect it to be substantially faster than conventional TreeRNN implementations, since it readily supports batched matrix--vector multiplications, allowing it to fast CPU libraries and to function well on GPUs.

\subsection{Model 1}

\input{model1_fig.tex}

Model 1 adapts Model 0 to use a stack and a buffer, making it more closely resemble a shift--reduce parser, and laying the groundwork for a model which can parse novel sentences at test time. The structre of the model is shown in Figure~\ref{m1-views}.

The model runs for a fixed number of transition steps: $2N - 3$. In its starting configuration, it contains a stack that is prepopulated with the first two words of the sentence (since \shift \shift is the only legal operation sequence for the first two timesteps of a true shift-reduce parser), as well as a buffer (a queue) prepopulated with all of the remaining words in the sentence. Both the stack and buffer represent words using their embeddings.

At each timestep at test time, the model combines views of the stack and buffer (the top element of the buffer and the top two elements of the stack, highlighted in yellow) as the input to a tracking LSTM (red). This LSTM's output is fed into a sigmoid operation classifier (blue) which chooses between the \shift and \reduce operations. If \shift is chosen, one word embedding is popped from the buffer and pushed onto the stack. If \reduce is chosen, the buffer is left as is, and the top two elements of the stack are popped and composed using a learned composition function (green), with the result placed back on top of the stack.

\paragraph{Supervision} The model is trained using two objective functions simultaneously. The semantic objective function is computed by feeding the value from the top of the stack at the final timestep---the full sentence encoding---into a downstream neural network model for some semantic task, like a sentiment classifier or an entailment classifier. The gradients from that classifier propagate to every part of the model except the operation classifier (blue). The syntactic objective function takes the form of direct supervision on the operation classifier (blue) which encourages that classifier to produce the same sequence of operations that an existing parser would produce for that sentence. The gradients from the syntactic objective function propagate to every part of the model but the downstream semantic model.

At training time, following the strategy used in LSTM text decoders, the decisions made by the operation classifier (blue) is discarded, and the model instead uses the correct operation as specified in the (already parsed) training corpus. At test time, this signal is not available, and the model uses its own predicted operations.

\subsection{Model 2}

Model 2 makes a small change to Model 1 that is likely to substantially change the dymanics of learning: It uses the operation sequence predicted by the operation classifier (blue) at training time as well as at test time. It may be possible to accelerate Model 2 training by initializing it with parameters learned by Model 1.

\subsection{Model 3}

Model 3 modifies Model 2 by introducing the soft stack/soft queue from \cite{grefenstette2015learning} in place of the hard, conventional stack and buffer. The soft stack makes it possible to for the model to predict smooth distributions over operations of the form (0.93 \shift, 0.07 \reduce), instead of making hard decisions. These soft decisions allow for gradient information to flow from the stack and the buffer back into the operation classifier (blue). This is crucial to our ultimate goal, as it makes it possible for semantic considerations to influence the model's parsing decisions.

\subsection{Model 4}

Model 4 modifies Model 3 by removing the direct supervision signal from the operation classifier (blue), instead forcing the operation classifier to learn solely from the gradient provided by the downstream supervision task. It may be possible to accelerate or otherwise improve Model 4 training by initializing it with parameters learned by Model 3.

\section{Other possible model features}

\subsection{Encoding the contents of the stack and buffer}

The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to make even minimally informed decisions about whether to shift or reduce. It could benefit further from additional information about broader sentential context. This can be provided by running new LSTMs along the elements of each of the stack and the buffer (following \citealt{dyer-EtAl:2015:ACL-IJCNLP}) and feeding the result into the tracking LSTM.

\subsection{Contextually-informed composition}

The composition function in the basic model (green) combines only the top elements of the stack, without using any further information. It may be possible to encourage the composition function to learn to do some amount of context-sensitive interpretation/disambiguation by adding a connection from the tracking LSTM (red) directly into the composition function.

\subsection{Constituent labels}

It would be possible to train any of the parse-supervised models (1--3) to learn explicit part of speech operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\}.

\subsubsection*{Acknowledgments}

\bibliographystyle{acl}
\bibliography{MLSemantics}

\end{document}

### 2015-10-22 writing/gist/model1_fig.tex
Working paper.

\begin{figure*}[t]
\begin{subfigure}[t]{\textwidth}
\centering
\scalebox{0.6}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{11pt}
    \def\sy{13*\dy}
    \def\oxb{8*\dx}
    \def\by{1*\dy}
    \def\ox{0*\oxb}

    \tikzstyle{word}=[text width=35mm,align=center,text height=2mm]    
    \tikzstyle{tracker}=[fill=red!40,text width=15mm,align=center,text height=2mm]
    \tikzstyle{softmax}=[fill=blue!40,text width=15mm,align=center,text height=2mm]
    \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
    \tikzstyle{result}=[line width=1pt,draw=black,text width=15mm,align=center,text height=2mm]    
    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=10mm]
    \tikzstyle{bbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=6.5mm]
    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=2mm]
    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=5.5mm]

    \node[focus1] (0bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (0b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (0b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (0b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (0bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (0sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (0s1) at (\ox+0*\dx,\sy-1*\dy) {cat};
    \node[word]  (0s2) at (\ox+0*\dx,\sy+0*\dy) {the};
    \node[word]  (0s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (0sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[comp] (0c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};

    \node[tracker] (0t) at  (\ox+0*\dx,5*\dy) {tracking};
    \node[softmax] (0sm) at  (\ox+3*\dx,7*\dy) {$\sigma$};
    \node[result] (0so) at  (\ox+3*\dx,9*\dy) {\reduce};

    \def\ox{1*\oxb}

    \node[focus1] (1bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (1b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (1b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (1b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (1bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (1sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (1s1) at (\ox+0*\dx,\sy-1*\dy) {the cat};
    \node[word]  (1s2) at (\ox+0*\dx,\sy+0*\dy) {};
    \node[word]  (1s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (1sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[comp] (1c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};

    \node[tracker] (1t) at  (\ox+0*\dx,5*\dy) {tracking};
    \node[softmax] (1sm) at  (\ox+3*\dx,7*\dy) {$\sigma$};
    \node[result] (1so) at  (\ox+3*\dx,9*\dy) {\shift};

    \def\ox{2*\oxb}

    \node[focus1] (2bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (2b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (2b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (2b1) at (\ox+0*\dx,\by+1*\dy) {down};
    \node[bbox] (2bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (2sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (2s1) at (\ox+0*\dx,\sy-1*\dy) {sat};
    \node[word]  (2s2) at (\ox+0*\dx,\sy+0*\dy) {the cat};
    \node[word]  (2s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (2sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[tracker] (2t) at  (\ox+0*\dx,5*\dy) {tracking};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

    \draw [fwd] (0sb) -- (0t);
    \draw [fwd] (0bb) -- (0t);
    \draw [fwd] (0t) -- (0sm);
    \draw [fwd] (0sm) -- (0so);
    \draw [fwd] (0sb) -- (0c);

    \draw [fwd] (0t) -- (1t);
    \draw [fwd] (0sb) -- (1sb);
    \draw [fwd] (0bb) -- (1bb);
    \draw [fwd] (0so) -- (1sb);
    \draw [fwd] (0so) -- (1bb);
    \draw [fwd] (0bb) to[out=20,in=-110] (1sb);
    \draw [fwd] (0c) -- (1sb);

    \draw [fwd] (1sb) -- (1t);
    \draw [fwd] (1bb) -- (1t);
    \draw [fwd] (1t) -- (1sm);
    \draw [fwd] (1sm) -- (1so);
    \draw [fwd] (1sb) -- (1c);

    \draw [fwd] (1t) -- (2t);
    \draw [fwd] (1sb) -- (2sb);
    \draw [fwd] (1bb) -- (2bb);
    \draw [fwd] (1so) -- (2sb);
    \draw [fwd] (1so) -- (2bb);
    \draw [fwd] (1bb) to[out=20,in=-110] (2sb);
    \draw [fwd] (1c) -- (2sb);

    \draw [fwd] (2sb) -- (2t);
    \draw [fwd] (2bb) -- (2t);

  \end{tikzpicture}}

 \caption{The Model 1/2 network unrolled for two transitions on the input \word{the cat sat down}.}\label{fig:model:top}

\end{subfigure}\\\\\\
\begin{subfigure}[t]{\textwidth}
\centering
\scalebox{0.6}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{11pt}
    \def\sy{7*\dy}
    \def\oxb{5.5*\dx}
    \def\by{1*\dy}
    \def\ox{0*\oxb}

    \tikzstyle{word}=[text width=35mm,align=center,text height=2mm]    
    \tikzstyle{tracker}=[fill=red!40,text width=15mm,align=center,text height=2mm]
    \tikzstyle{softmax}=[fill=orange!40,text width=33mm,align=center,text height=2mm]
    \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
    \tikzstyle{result}=[line width=1pt,draw=black,text width=15mm,align=center,text height=2mm]    
    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=10mm]
    \tikzstyle{bbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=6.5mm]
    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=33mm,align=center,text height=2mm]
    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=33mm,align=center,text height=5.5mm]

    \def\ox{0*\oxb}

    \node[focus1] (1bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (1b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (1b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (1b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (1bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (1sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (1s1) at (\ox+0*\dx,\sy-1*\dy) {cat};
    \node[word]  (1s2) at (\ox+0*\dx,\sy+0*\dy) {the};
    \node[word]  (1s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (1sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (1so) at  (\ox+0.5*\dx,4*\dy) {\reduce};

    \def\ox{1*\oxb}

    \node[focus1] (2bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (2b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (2b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (2b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (2bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (2sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (2s1) at (\ox+0*\dx,\sy-1*\dy) {the cat};
    \node[word]  (2s2) at (\ox+0*\dx,\sy+0*\dy) {};
    \node[word]  (2s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (2sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (2so) at  (\ox+0.5*\dx,4*\dy) {\shift};

    \def\ox{2*\oxb}

    \node[focus1] (3bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (3b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (3b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (3b1) at (\ox+0*\dx,\by+1*\dy) {down};
    \node[bbox] (3bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (3sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (3s1) at (\ox+0*\dx,\sy-1*\dy) {sat};
    \node[word]  (3s2) at (\ox+0*\dx,\sy+0*\dy) {the cat};
    \node[word]  (3s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (3sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (3so) at  (\ox+0.5*\dx,4*\dy) {\shift};

    \def\ox{3*\oxb}

    \node[focus1] (4bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (4b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (4b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (4b1) at (\ox+0*\dx,\by+1*\dy) {};
    \node[bbox] (4bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (4sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (4s1) at (\ox+0*\dx,\sy-1*\dy) {down};
    \node[word]  (4s2) at (\ox+0*\dx,\sy+0*\dy) {sat};
    \node[word]  (4s3) at (\ox+0*\dx,\sy+1*\dy) {the cat};
    \node[sbox] (4sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (4so) at  (\ox+0.5*\dx,4*\dy) {\reduce};

    \def\ox{4*\oxb}

    \node[focus1] (5bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (5b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (5b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (5b1) at (\ox+0*\dx,\by+1*\dy) {};
    \node[bbox] (5bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (5sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (5s1) at (\ox+0*\dx,\sy-1*\dy) {sat down};
    \node[word]  (5s2) at (\ox+0*\dx,\sy+0*\dy) {the cat};
    \node[word]  (5s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (5sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (5so) at  (\ox+0.5*\dx,4*\dy) {\reduce};

    \def\ox{5*\oxb}

    \node[focus2] (6sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (6s1) at (\ox+0*\dx,\sy-1*\dy) {(the cat) (sat down)};
    \node[word]  (6s2) at (\ox+0*\dx,\sy+0*\dy) {};
    \node[word]  (6s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (6sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[softmax] (6sm) at  (\ox+0.5*\dx,2*\dy) {output to semantic task model};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

   \draw [fwd] (1sb) -- (1so);
   \draw [fwd] (1bb) -- (1so);

    \draw [fwd] (1sb) -- (2sb);
    \draw [fwd] (1bb) -- (2bb);
    \draw [fwd] (1so) -- (2sb);
    \draw [fwd] (1so) -- (2bb);
    \draw [fwd] (1bb) -- (2sb);

   \draw [fwd] (2sb) -- (2so);
   \draw [fwd] (2bb) -- (2so);

    \draw [fwd] (2sb) -- (3sb);
    \draw [fwd] (2bb) -- (3bb);
    \draw [fwd] (2so) -- (3sb);
    \draw [fwd] (2so) -- (3bb);
    \draw [fwd] (2bb) -- (3sb);

   \draw [fwd] (3sb) -- (3so);
   \draw [fwd] (3bb) -- (3so);

    \draw [fwd] (3sb) -- (4sb);
    \draw [fwd] (3bb) -- (4bb);
    \draw [fwd] (3so) -- (4sb);
    \draw [fwd] (3so) -- (4bb);
    \draw [fwd] (3bb) -- (4sb);

   \draw [fwd] (4sb) -- (4so);
   \draw [fwd] (4bb) -- (4so);

    \draw [fwd] (4sb) -- (5sb);
    \draw [fwd] (4bb) -- (5bb);
    \draw [fwd] (4so) -- (5sb);
    \draw [fwd] (4so) -- (5bb);
    \draw [fwd] (4bb) -- (5sb);

   \draw [fwd] (5sb) -- (5so);
   \draw [fwd] (5bb) -- (5so);

    \draw [fwd] (5sb) -- (6sb);
    \draw [fwd] (5so) -- (6sb);

   \draw [fwd] (6sb) -- (6sm);

  \end{tikzpicture}}

 \caption{The fully unrolled Model 1/2 network for \word{the cat sat down} with some layers omitted for clarity.}\label{fig:model:top}  
\end{subfigure}
\caption{\label{m1-views}Two views of Models 1 and 2 (which use equivalent model graphs). In both views, the lower boxes represent the input buffer, and the upper boxes represent the stack. Yellow highlighting indicates which portions of these data structures are visible to the tracking LSTM and to the composition function.}
\end{figure*}

### 2015-10-22 writing/gist/model1_fig.tex
Writing: Buffer size change.

@@
     \tikzstyle{result}=[line width=1pt,draw=black,text width=15mm,align=center,text height=2mm]    
-    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=10mm]
+    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=13.3mm]
     \tikzstyle{bbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=6.5mm]
@@
     \node[word]  (0s3) at (\ox+0*\dx,\sy+1*\dy) {};
-    \node[sbox] (0sb) at  (\ox+0*\dx,\sy+0*\dy) {};
+    \node[sbox] (0sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
 
@@
     \node[word]  (1s3) at (\ox+0*\dx,\sy+1*\dy) {};
-    \node[sbox] (1sb) at  (\ox+0*\dx,\sy+0*\dy) {};
+    \node[sbox] (1sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
 
@@
     \node[word]  (2s3) at (\ox+0*\dx,\sy+1*\dy) {};
-    \node[sbox] (2sb) at  (\ox+0*\dx,\sy+0*\dy) {};
+    \node[sbox] (2sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
 
@@
     \tikzstyle{result}=[line width=1pt,draw=black,text width=15mm,align=center,text height=2mm]    
-    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=10mm]
+    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=13.3mm]
     \tikzstyle{bbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=6.5mm]
@@
     \node[word]  (3s3) at (\ox+0*\dx,\sy+1*\dy) {};
-    \node[sbox] (3sb) at  (\ox+0*\dx,\sy+0*\dy) {};
+    \node[sbox] (3sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
 
@@
     \node[word]  (4s3) at (\ox+0*\dx,\sy+1*\dy) {the cat};
-    \node[sbox] (4sb) at  (\ox+0*\dx,\sy+0*\dy) {};
+    \node[sbox] (4sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
 
@@
     \node[word]  (5s3) at (\ox+0*\dx,\sy+1*\dy) {};
-    \node[sbox] (5sb) at  (\ox+0*\dx,\sy+0*\dy) {};
+    \node[sbox] (5sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
 
@@
     \node[word]  (6s3) at (\ox+0*\dx,\sy+1*\dy) {};
-    \node[sbox] (6sb) at  (\ox+0*\dx,\sy+0*\dy) {};
+    \node[sbox] (6sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
 

### 2015-10-22 rembed/models/classifier.py
Fix broken logging. Start setting up cluster use.

@@ -72,6 +72,8 @@ def train():
         logging.error("Bad data type.")
         return
 
+    logging.info("Flag values: " + str(FLAGS.FlagValuesDict()))
+
     # Load the data
     training_data_iter, vocabulary = data_manager.load_data(
         FLAGS.training_data_path, seq_length=FLAGS.seq_length, batch_size=FLAGS.batch_size)
@@ -132,22 +134,22 @@ def train():
                              (step, acc_accum / eval_batches, eval_set[0]))
 
 if __name__ == '__main__':
-    # Experiment naming
+    # Experiment naming.
     gflags.DEFINE_string("experiment_name", "experiment", "")
 
-    # Data types
+    # Data types.
     gflags.DEFINE_string("data_type", "bl", "Values: bl, sst")
 
-    # Data settings
+    # Data settings.
     gflags.DEFINE_string("training_data_path", None, "")
     gflags.DEFINE_string("eval_data_path", None, "")
     gflags.DEFINE_integer("seq_length", 29, "")
 
-    # Model architecture settings
+    # Model architecture settings.
     gflags.DEFINE_integer("embedding_dim", 5, "")
     gflags.DEFINE_integer("num_composition_layers", 1, "")
 
-    # Optimization settings
+    # Optimization settings.
     gflags.DEFINE_integer("training_steps", 50000, "")
     gflags.DEFINE_integer("batch_size", 32, "")
     gflags.DEFINE_float("learning_rate", 0.01, "")
@@ -156,18 +158,26 @@ if __name__ == '__main__':
     gflags.DEFINE_float("l2_lambda", 1e-6, "")
     gflags.DEFINE_float("init_range", 0.1, "")
 
-    # Display settings
+    # Display settings.
     gflags.DEFINE_integer("statistics_interval_steps", 50, "")
     gflags.DEFINE_integer("eval_interval_steps", 50, "")
 
-    # Parse command line flags
+    # Parse command line flags.
     FLAGS(sys.argv)
 
-    logging.basicConfig(level=logging.DEBUG,
-                        format='%(asctime)s %(levelname)-8s %(message)s',
-                        datefmt='%m-%d-%y %H:%M',
-                        handlers=[logging.FileHandler(
-                            FLAGS.experiment_name + ".log"),
-                        logging.StreamHandler()])
+    # Set up logging.
+    logFormatter = logging.Formatter(
[…]

### 2015-10-22 run.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+### Generic job script for all experiments.
+
+#PBS -l nodes=1:ppn=6 	### Request at least 6 cores
+#PBS -l walltime=99:00:00	### Die after four days
+#PBS -l mem=6000MB
+#PBS -q nlp
+
+# Usage example:
+# export REMBED_FLAGS="--learning_rate 0.2"; qsub -v MATLABCMD quant/run.sh
+
+# Change to the submission directory.
+cd $PBS_O_WORKDIR 
+
+# Log what we're running and where.
+echo `hostname` - $PBS_JOBID - $REMBED_FLAGS >> ~/rembed_machine_assignments.txt
+
+python rembed/models/classifier.py $REMBED_FLAGS

### 2015-10-23 writing/gist/gist.tex
Writing: Buffer size change.
Update to gist.

@@
 
-\title{NSHRDLU?}
+\title{NSHRDLU:\thanks{Name is provisional, and stolen from Geoff Hinton.}\\Work in progress on the joint learning of parsing and semantic encoding}
 
@@
 \AND
+Christopher D.\ Manning$^{\ast\dag\S}$\\
+\texttt{manning@stanford.edu}\\
+\And
 Christopher Potts$^{\ast}$\\
 \texttt{cgpotts@stanford.edu}
-\And
-Christopher D.\ Manning$^{\ast\dag\S}$\\
-\texttt{manning@stanford.edu}\\
 \AND\\[-3ex]
@@
 
-This project aims to build neural network models that can jointly learn to parse sentences and to use those parses to guide semantic composition.
+\section{Introduction}
 
-Table~\ref{models-table} shows the sequence of model designs that we plan to build.
+This project aims to use ideas from greedy transition-based parsing to build neural network models that can jointly learn to parse sentences and to use those parses to guide semantic composition.
+
+Table~\ref{models-table} shows the sequence of model designs that we plan to build, and Figure~\ref{m1-views} depicts some representative models.
+
+We see three reasons to pursue this approach:
+\begin{itemize}
+\item Simply adapting a greedy transition-based approach to sentence encoding makes it possible to exploit semantic compositionality in the same manner as in a TreeRNN, but using a static graph structure that can take advantage of existing neural network libraries like Theano for both automatic differentiation and for highly optimized matrix computations on both CPUs and GPUs.
+Model 0 pursues this property directly, and all subsequent models share it.
+\item When a parsing system and an interpretation system are trained jointly and forced to share representations, it is likely that the performance of both models will benefit: the semantic composition function will have better access to syntactic type information that could provide additional evidence on the functional behavior of rare words, and the syntactic parser will have access to information about the semantic interpretability of constituents, making semantically-conditioned parsing decisions like PP attachment easier to learn.
+Models 1--3 have this property.
+\item When a system that uses constituent structure information as a latent variable is initialized from scratch and trained solely to perform some semantic task, it will likely learn some coherent syntax for the genre of language on which it is trained.
+If this syntax is reasonably stable across similar training sets and across random initializations of training, it can reasonably be said to be latent in the text itself, offering a new kind of evidence about natural language syntax.
+Model 4 aims to collect this kind of evidence.
+
+\end{itemize}
+
+All five models function as sentence encoding models: their inputs are sentences (input as sequences of words, with the help of a learned embedding matrix), and their outputs are single sentence encoding vectors which can be used as the inputs to downstream models for tasks like sentiment analysis, translation, or inference.
+Provided that these downstream models are differentiable---as is the case for neural network models and simpler regression models---their gradient signals can be used to train the sentence encoding models.
 
@@
 
-Model 0 in its simplest form computes the same function as a plain TreeRNN, but uses a stack and a recurrent input reader instead of a tree as its graph structure.
-\todo{Model 0 figure.} We expect it to be substantially faster than conventional TreeRNN implementations, since it readily supports batched matrix--vector multiplications, allowing it to fast CPU libraries and to function well on GPUs.
+Model 0, depicted in Figure~\ref{fig:model:0}, is the simplest instantiation of our design, using only a conventional stack and a learned composition function to incrementally build a sentence representation over a series of timesteps.
+For a sentence of $N$ words its input is a sequence of $2N-3$ inputs.
+These inputs can take two types.
+At some timesteps, the input will be a word embedding vector.
+This triggers the \shift~operation, in which the vector is pushed onto the top of a stack.
+At other timesteps, the input will be the special \reduce~token, which triggers the reduction operation.
+In that operation, the top two word vectors are popped from the stack, fed into a learned composition function that maps them to a single vector (in the simplest case, this is a single neural network layer), and then pushed back onto the stack.
+
+If we add no additional features, this model computes the same function as a plain TreeRNN.
+However, we expect it to be substantially faster than conventional TreeRNN implementations.
+Unlike a TreeRNN, the Model 0 computaiton graph is essentially static across examples, so examples of varying structures and lengths can be batched together and run on the same graph in a single step.
+This simply requires ensuring that the graph is run for enough timesteps to finish all of the sentences.
+This involves some wasted computation, since the composition function will be run $2N-3$ times (with the output of composition at non-\reduce~steps discarded), rather than $N-1$ times in a TreeRNN.
+However, this loss can be dramatically offset by the gains of batching, which stem from the ability to exploit highly optimized CPU and GPU libraries for batched matrix mulitiplication.
 
@@
 
-Model 1 adapts Model 0 to use a stack and a buffer, making it more closely resemble a shift--reduce parser, and laying the groundwork for a model which can parse novel sentences at test time.
-The structre of the model is shown in Figure~\ref{m1-views}.
+Model 1, depicted in Figures~\ref{fig:model:1d} and \ref{fig:model:1b}, adapts Model 0 to use a stack and a buffer, making it more closely resemble a shift--reduce parser, and laying the groundwork for a model which can parse novel sentences at test time.
 
 The model runs for a fixed number of transition steps: $2N - 3$.
-In its starting configuration, it contains a stack that is prepopulated with the first two words of the sentence (since \shift \shift is the only legal operation sequence for the first two timesteps of a true shift-reduce parser), as well as a buffer (a queue) prepopulated with all of the remaining words in the sentence.
+In its starting configuration, it contains a stack that is prepopulated with the first two words of the sentence (since \shift~\shift~is the only legal operation sequence for the first two timesteps of a true shift-reduce parser), as well as a buffer (a queue) prepopulated with all of the remaining words in the sentence.
 Both the stack and buffer represent words using their embeddings.
@@
 At each timestep at test time, the model combines views of the stack and buffer (the top element of the buffer and the top two elements of the stack, highlighted in yellow) as the input to a tracking LSTM (red).
-This LSTM's output is fed into a sigmoid operation classifier (blue) which chooses between the \shift and \reduce operations.
-If \shift is chosen, one word embedding is popped from the buffer and pushed onto the stack.
-If \reduce is chosen, the buffer is left as is, and the top two elements of the stack are popped and composed using a learned composition function (green), with the result placed back on top of the stack.
+This LSTM's output is fed into a sigmoid operation classifier (blue) which chooses between the \shift~and \reduce~operations.
+If \shift~is chosen, one word embedding is popped from the buffer and pushed onto the stack.
+If \reduce~is chosen, the buffer is left as is, and the top two elements of the stack are popped and composed using a learned composition function (green), with the result placed back on top of the stack.
 
@@
 It may be possible to accelerate Model 2 training by initializing it with parameters learned by Model 1.
+
+For discussion of related issues in sequence-to-sequence models, see \citealt{bengio2015scheduled}.
 
@@
 
-The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to make even minimally informed decisions about whether to shift or reduce.
+The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to m ake even minimally informed decisions about whether to shift or reduce.
 It could benefit further from additional information about broader sentential context.
@@
 
+For model zero, no tracking LSTM is needed for the ordinary operation of the model, but it would be possible to add one for this purpose, taking as inputs the top two values of the stack at each time point and emitting as output a context vector that can be used to condition the composition function.
+
 \subsection{Constituent labels}
@@
 It would be possible to train any of the parse-supervised models (1--3) to learn explicit part of speech operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\}.
+
+\section{Implementation notes}
+
+The size of the stack should be $N$ for sentences of $N$ words, in case the first reduce merges the final two words. The size of the buffer should be $N - 2$.
+
+\subsection{Data preparation}
+
+For Models 0--3, all training data must be parsed in advance into an unlabeled binary constituency tree.
+In addition, Model 0 requires that  parses be available at test time as well.
+For both SST and SNLI we use the parses included with the corpus distributions whenever parses are needed.
+
+For model 0, training data can be prepared by linearizing the provided parse, then deleting left brackets and replacing right brackets with \reduce~instructions.
+That is demonstrated here with the example sentence \ii{the cat sat down}:
+
+\begin{quote}\small
+( ( the cat ) ( sat down ) )$\Rightarrow$\\
+the cat \reduce~sat down \reduce~\reduce
+\end{quote}
+
+The input for models 1--4 is simply the word sequence from the parse, with the first two words moved into the stack.
+The syntactic supervision labels for models 1--3 are simply a binarized version of the Model 0 inputs, with the first two tokens (which are necessarily \shift~\shift) omitted:
+
+\begin{quote}\small
+( ( the cat ) ( sat down ) )$\Rightarrow$ \\
+stack: $\langle$the, cat$\rangle$\\
+buffer: $\langle$sat, down$\rangle$\\
+ops: \reduce~\shift~\shift~\reduce~\reduce
+\end{quote}
+
+\section{Experiments}
+
+\subsection{Step time}
+
+Comparing model step time to the plain RNN of \citealt{li2015tree}. We use the small \citealt{pang2004sentimental} sentiment corpus that they use, and train with identical hyperparameters: ...
+
+Evaluation metrics: Time per minibatch on a jag machine, with and without GPU access.
+
+\subsection{Sentiment}
+
+Learning to jointly parse and to predict logical relations between sentences over SST  
+\cite{socher2013recursive}.
+
+Evaluation metrics: accuracy, F1 (for all models but 0).
+
+\subsection{Natural language inference}
+
+Learning to jointly parse and to predict logical relations between sentences over SNLI  \cite{snli:emnlp2015}.
+
+Evaluation metrics: accuracy, F1 (for all models but 0).
+
+\section{Discussion}
+
+\subsection{Inferred Model 4 parses?}
 

### 2015-10-23 writing/gist/model1_fig.tex
Update to gist.

@@
 
+    \tikzstyle{label}=[text width=35mm,align=center,text height=2mm]    
+    \tikzstyle{word}=[text width=35mm,align=center,text height=2mm]    
+    \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
+    \tikzstyle{input}=[line width=1pt,text width=20mm,align=center,text height=2mm]    
+    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=13.3mm]
+    \tikzstyle{bbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=6.5mm]
+    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=2mm]
+    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=5.5mm]
+
+    \node[label]  (sl) at (\ox-0.35*\oxb+0*\dx,\sy+0.5*\dy) {stack};
+
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=0$};
+
+    \node[focus2] (0sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
+    \node[word]  (0s1) at (\ox+0*\dx,\sy-1*\dy) {cat};
+    \node[word]  (0s2) at (\ox+0*\dx,\sy+0*\dy) {the};
+    \node[word]  (0s3) at (\ox+0*\dx,\sy+1*\dy) {};
+    \node[sbox] (0sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
+
+    \node[comp] (0c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};
+
+    \node[input] (0so) at  (\ox+0.75*\oxb,9*\dy) {\reduce};
+
+    \def\ox{1*\oxb}
+
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=1$};
+
+    \node[focus2] (1sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
+    \node[word]  (1s1) at (\ox+0*\dx,\sy-1*\dy) {the cat};
+    \node[word]  (1s2) at (\ox+0*\dx,\sy+0*\dy) {};
+    \node[word]  (1s3) at (\ox+0*\dx,\sy+1*\dy) {};
+    \node[sbox] (1sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
+
+    \node[comp] (1c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};
+
+    \node[input] (1so) at  (\ox+0.75*\oxb,9*\dy) {\shift:~sat};
+
+    \def\ox{2*\oxb}
+
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=2$};
+
+    \node[focus2] (2sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
+    \node[word]  (2s1) at (\ox+0*\dx,\sy-1*\dy) {sat};
+    \node[word]  (2s2) at (\ox+0*\dx,\sy+0*\dy) {the cat};
+    \node[word]  (2s3) at (\ox+0*\dx,\sy+1*\dy) {};
+    \node[sbox] (2sb) at  (\ox+0*\dx,\sy+0.5*\dy) {};
+
+    \pgfsetarrowsend{latex}
+    \tikzstyle{fwd} = [draw=black, line width=1pt]
+    \tikzstyle{gated} = [draw=black!50, line width=0.8pt]
+
+    \draw [fwd] (0sb) -- (0c);
+
+    \draw [fwd] (0sb) -- (1sb);
+    \draw [fwd] (0so) -- (1sb);
+    \draw [fwd] (0c) -- (1sb);
+
+    \draw [fwd] (1sb) -- (1c);
+
+    \draw [fwd] (1sb) -- (2sb);
+    \draw [fwd] (1so) -- (2sb);
+    \draw [gated] (1c) -- (2sb);
+
+  \end{tikzpicture}}
+
+ \caption{\label{fig:model:0}The Model 0 network unrolled for two transitions on the input \word{the cat sat down}.
+The operations at the bottom of the figure`  are given as inputs to the models, as is the start state of the stack at $t=0$.}
+
+\end{subfigure}\\\\\\
+\begin{subfigure}[t]{\textwidth}
+\centering
+\scalebox{0.6}{
+ \begin{tikzpicture}
+    \def\dx{21pt}
+    \def\dy{11pt}
+    \def\sy{13*\dy}
+    \def\oxb{8*\dx}
+    \def\by{1*\dy}
+    \def\ox{0*\oxb}
+
+    \tikzstyle{label}=[text width=35mm,align=center,text height=2mm]    
     \tikzstyle{word}=[text width=35mm,align=center,text height=2mm]    
@@
 
+    \node[label]  (sl) at (\ox-0.35*\oxb+0*\dx,\by+0.5*\dy) {buffer};
+
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=0$};
+
     \node[focus1] (0bb) at  (\ox+0*\dx,2*\dy) {};
@@
     \node[bbox] (0bb) at  (\ox+0*\dx,\by+0.5*\dy) {};
+
+    \node[label]  (sl) at (\ox-0.35*\oxb+0*\dx,\sy+0.5*\dy) {stack};
 
@@
     \def\ox{1*\oxb}
+
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=1$};
 
@@
 
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=2$};
+
     \node[focus1] (2bb) at  (\ox+0*\dx,2*\dy) {};
@@
     \tikzstyle{fwd} = [draw=black, line width=1pt]
+    \tikzstyle{gated} = [draw=black!50, line width=0.8pt]
 
@@
     \draw [fwd] (0so) -- (1bb);
-    \draw [fwd] (0bb) to[out=20,in=-110] (1sb);
+    \draw [gated] (0bb) to[out=20,in=-110] (1sb);
     \draw [fwd] (0c) -- (1sb);
@@
     \draw [fwd] (1bb) to[out=20,in=-110] (2sb);
-    \draw [fwd] (1c) -- (2sb);
+    \draw [gated] (1c) -- (2sb);
 
@@
 
- \caption{The Model 1/2 network unrolled for two transitions on the input \word{the cat sat down}.}\label{fig:model:top}
+ \caption{The Model 1/2 network unrolled for two transitions on the input \word{the cat sat down}.
+The start state of the stack and buffer at the initial step $t=0$ are the sole inputs to the model.}\label{fig:model:1d}
 
@@
 
+    \tikzstyle{label}=[text width=35mm,align=center,text height=2mm]    
     \tikzstyle{word}=[text width=35mm,align=center,text height=2mm]    
     \tikzstyle{tracker}=[fill=red!40,text width=15mm,align=center,text height=2mm]
-    \tikzstyle{softmax}=[fill=orange!40,text width=33mm,align=center,text height=2mm]
+    \tikzstyle{softmax}=[text width=40mm,align=center,text height=2mm]
     \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
@@
 
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=0$};
+
+    \node[label]  (sl) at (\ox-0.65*\oxb+0*\dx,\by+0.5*\dy) {buffer};
+
     \node[focus1] (1bb) at  (\ox+0*\dx,2*\dy) {};
@@
     \node[bbox] (1bb) at  (\ox+0*\dx,\by+0.5*\dy) {};
+
+    \node[label]  (sl) at (\ox-0.65*\oxb+0*\dx,\sy+0.5*\dy) {stack};
 
@@
 
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=2$};
+
     \node[focus1] (3bb) at  (\ox+0*\dx,2*\dy) {};
@@
     \def\ox{3*\oxb}
+
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=3$};
 
@@
 
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=4$};
+
     \node[focus1] (5bb) at  (\ox+0*\dx,2*\dy) {};
@@
 
+    \node[label]  (1l) at (\ox+0*\dx,\sy+3*\dy) {$t=5$};
+
     \node[focus2] (6sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
@@
 
-    \node[softmax] (6sm) at  (\ox+0.5*\dx,2*\dy) {output to semantic task model};
+    \node[softmax] (6sm) at  (\ox+0.5*\dx,2*\dy) {output to model for semantic task};
 
@@
 
- \caption{The fully unrolled Model 1/2 network for \word{the cat sat down} with some layers omitted for clarity.}\label{fig:model:top}  
+ \caption{The fully unrolled Model 1/2 network for \word{the cat sat down} with some layers omitted for clarity.}\label{fig:model:1b}  
 \end{subfigure}
@@
 In both views, the lower boxes represent the input buffer, and the upper boxes represent the stack.
-Yellow highlighting indicates which portions of these data structures are visible to the tracking LSTM and to the composition function.}
+Yellow highlighting indicates which portions of these data structures are visible to the tracking LSTM and to the composition function.
+Thin gray arrows indicate connections which are blocked by a gating function, and so contribute no information.}
 \end{figure*}

### 2015-10-24 writing/gist/gist.tex
Small additions and minor fixes to the gist writeup

@@
 However, we expect it to be substantially faster than conventional TreeRNN implementations.
-Unlike a TreeRNN, the Model 0 computaiton graph is essentially static across examples, so examples of varying structures and lengths can be batched together and run on the same graph in a single step.
+Unlike a TreeRNN, the Model 0 computation graph is essentially static across examples, so examples of varying structures and lengths can be batched together and run on the same graph in a single step.
 This simply requires ensuring that the graph is run for enough timesteps to finish all of the sentences.
 This involves some wasted computation, since the composition function will be run $2N-3$ times (with the output of composition at non-\reduce~steps discarded), rather than $N-1$ times in a TreeRNN.
-However, this loss can be dramatically offset by the gains of batching, which stem from the ability to exploit highly optimized CPU and GPU libraries for batched matrix mulitiplication.
+However, this loss can be dramatically offset by the gains of batching, which stem from the ability to exploit highly optimized CPU and GPU libraries for batched matrix multiiplication.
 
@@
 
-Model 2 makes a small change to Model 1 that is likely to substantially change the dymanics of learning: It uses the operation sequence predicted by the operation classifier (blue) at training time as well as at test time.
+Model 2 makes a small change to Model 1 that is likely to substantially change the dynamics of learning: It uses the operation sequence predicted by the operation classifier (blue) at training time as well as at test time.
 It may be possible to accelerate Model 2 training by initializing it with parameters learned by Model 1.
 
-For discussion of related issues in sequence-to-sequence models, see \citealt{bengio2015scheduled}.
+By exposing Model 2 to the results of its own decisions during training, we encourage it to become more robust to its own prediction errors.
+\citealt{bengio2015scheduled} applied a similar strategy\footnote{The authors experiment with several strategies which interpolate between oracle-driven training and oracle-free training (Models 1 and 2 in our presentation, respectively).
+It may be useful to adopt a similar interpolating approach.} to an image captioning model.
+They suggest that the resulting model can avoid propagating prediction errors through long sequences due to this training regime.
 
@@
 
+Model 3 still receives a direct supervision signal from some existing parser.
+In order to train the soft stack, we must represent the hard supervision signal from the parser by a soft prediction which matches the soft stack operation output.
+The supervision signal simply assigns 100\% weight to the ground-truth operation.
+
 \subsection{Model 4}
@@
 
+By removing an external parser signal, we allow Model 4 to fully exploit the soft stack representation.
+It is free to predict soft parse operations in the case of ambiguous parses.
+This distinguishes it from Model 3, which is encouraged to replicate the 100\%-certain ground truth parse predictions provided by the external parser.
+
 \section{Other possible model features}
@@
 
-The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to m ake even minimally informed decisions about whether to shift or reduce.
+The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to make even minimally informed decisions about whether to shift or reduce.
 It could benefit further from additional information about broader sentential context.
@@
 
-For model zero, no tracking LSTM is needed for the ordinary operation of the model, but it would be possible to add one for this purpose, taking as inputs the top two values of the stack at each time point and emitting as output a context vector that can be used to condition the composition function.
+For Model 0, no tracking LSTM is needed for the ordinary operation of the model, but it would be possible to add one for this purpose, taking as inputs the top two values of the stack at each time point and emitting as output a context vector that can be used to condition the composition function.
 
-\subsection{Constituent labels}
+\subsection{Typed \reduce~operations}
 
-It would be possible to train any of the parse-supervised models (1--3) to learn explicit part of speech operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\}.
+Shift-reduce parsers for natural language typically operate with a restricted set of typed \reduce~operations (also known as ``arc'' operations).
+These operations specify the precise relation between the elements being merged.
+It would be possible to train any of the parse-supervised models (1--3) to learn such typed arc operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\} (in the case of constituency parse supervision).
+The model can then learn a distinct composition function depending on the relation of the two elements being merged.
 

### 2015-10-24 writing/gist/model1_fig.tex
Small additions and minor fixes to the gist writeup
Gist update.

@@
     \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
+    \tikzstyle{compoff}=[fill=green!10!black!10,text width=20mm,align=center,text height=2mm]
     \tikzstyle{input}=[line width=1pt,text width=20mm,align=center,text height=2mm]    
@@
 
-    \node[comp] (1c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};
+    \node[compoff] (1c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};
 
@@
  \caption{\label{fig:model:0}The Model 0 network unrolled for two transitions on the input \word{the cat sat down}.
-The operations at the bottom of the figure`  are given as inputs to the models, as is the start state of the stack at $t=0$.}
+The operations at the bottom of the figure are given as inputs to the models, as is the start state of the stack at $t=0$.}
 
@@
     \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
+    \tikzstyle{compoff}=[fill=green!10!black!10,text width=20mm,align=center,text height=2mm]
     \tikzstyle{result}=[line width=1pt,draw=black,text width=15mm,align=center,text height=2mm]    

### 2015-10-24 writing/gist/model1_fig.tex
Ditch opacity. It's not well supported.

@@
     \tikzstyle{bbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=6.5mm]
-    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=2mm]
-    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=5.5mm]
+    \tikzstyle{focus1}=[fill=yellow!40,text width=25mm,align=center,text height=2mm]
+    \tikzstyle{focus2}=[fill=yellow!40,,text width=25mm,align=center,text height=5.5mm]
 
@@
     \tikzstyle{bbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=6.5mm]
-    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=2mm]
-    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=5.5mm]
+    \tikzstyle{focus1}=[fill=yellow!40,text width=25mm,align=center,text height=2mm]
+    \tikzstyle{focus2}=[fill=yellow!40,text width=25mm,align=center,text height=5.5mm]
 
@@
     \tikzstyle{bbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=6.5mm]
-    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=33mm,align=center,text height=2mm]
-    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=33mm,align=center,text height=5.5mm]
+    \tikzstyle{focus1}=[fill=yellow!40,text width=33mm,align=center,text height=2mm]
+    \tikzstyle{focus2}=[fill=yellow!40,text width=33mm,align=center,text height=5.5mm]
 

### 2015-11-17 pick_gpu.py
Lots of things!

@@ -1,0 +1,49 @@
+import subprocess
+import re
+import random
+
+USAGE_THRESHOLD = 90
+
+output = """Mon Nov 16 19:41:43 2015       
++------------------------------------------------------+                       
+| NVIDIA-SMI 352.39     Driver Version: 352.39         |                       
+|-------------------------------+----------------------+----------------------+
+| GPU  Name        Persistence-M| Bus-Id        Disp.A | Volatile Uncorr. ECC |
+| Fan  Temp  Perf  Pwr:Usage/Cap|         Memory-Usage | GPU-Util  Compute M. |
+|===============================+======================+======================|
+|   0  Tesla K40c          Off  | 0000:02:00.0     Off |                    0 |
+| 29%   59C    P0   136W / 235W |   4897MiB / 11519MiB |     83%      Default |
++-------------------------------+----------------------+----------------------+
+|   1  Tesla K40m          Off  | 0000:81:00.0     Off |                    0 |
+| N/A   61C    P0   158W / 235W |   6892MiB / 11519MiB |     99%      Default |
++-------------------------------+----------------------+----------------------+
+|   2  Tesla K40c          Off  | 0000:82:00.0     Off |                    0 |
+| 25%   48C    P0    94W / 235W |   9365MiB / 11519MiB |     31%      Default |
++-------------------------------+----------------------+----------------------+
+                                                                               
++-----------------------------------------------------------------------------+
+| Processes:                                                       GPU Memory |
+|  GPU       PID  Type  Process name                               Usage      |
+|=============================================================================|
+|    0     19867    C   ...b-r2014b/matlab/r2014b/bin/glnxa64/MATLAB  4872MiB |
+|    1     19774    C   ...b-r2014b/matlab/r2014b/bin/glnxa64/MATLAB  6867MiB |
+|    2     20285    C   ...b-r2014b/matlab/r2014b/bin/glnxa64/MATLAB  9340MiB |
++-----------------------------------------------------------------------------+"""
+
+p = subprocess.Popen("nvidia-smi", stdout=subprocess.PIPE, 
+                                   stderr=subprocess.PIPE)
+output, error = p.communicate()
+if error:
+    print "cpu"
+
+usage_re = re.compile(r"(?<=   )\d{1,3}(?=%)")
+matches = usage_re.findall(output)
+usage_amts = [int(usage_amt) for usage_amt in matches]
+print usage_amts
+
[…]

### 2015-11-17 rembed/models/classifier.py

@@ -421,7 +421,7 @@ if __name__ == '__main__':
     # gflags.DEFINE_integer("num_composition_layers", 1, "")
 
     # Optimization settings.
-    gflags.DEFINE_integer("training_steps", 100000, "")
+    gflags.DEFINE_integer("training_steps", 1000000, "")
     gflags.DEFINE_integer("batch_size", 32, "")
     gflags.DEFINE_float("learning_rate", 0.001, "Used in RMSProp.")
     # gflags.DEFINE_float("momentum", 0.9, "")

### 2015-11-17 pick_gpu.py
pick_gpu update

@@ -42,8 +42,12 @@ usage_amts = [int(usage_amt) for usage_amt in matches]
 print usage_amts
 
 # Get GPU
+gpus = [index for index in range(len(usage_amts))]
 open_gpus = [index for index in range(len(usage_amts)) if usage_amts[index] < USAGE_THRESHOLD]
+
 if open_gpus:
-    print random.choice(open_gpus)
+    print "gpu" + str(random.choice(open_gpus))
+elif gpus:
+    print "gpu" + str(random.choice(gpus))
 else:
     print "cpu"

### 2015-11-17 pick_gpu.py
Tidy up pick_gpu.py

@@ -1,39 +1,24 @@
 import subprocess
-import re
+import sys
 import random
+import re
+
+# Print the name of a device to use, either 'cpu' or 'gpu0', 'gpu1',...
+# GPUs with usage under the constant threshold will be chosen first,
+# but subject to that constraint, selection is random.
+#
+# Warning: This is hacky and brittle, and can break if nvidia-smi changes 
+# in the way it formats its output.
+#
+# Maintainer: sbowman@stanford.edu
 
 USAGE_THRESHOLD = 90
 
-output = """Mon Nov 16 19:41:43 2015       
-+------------------------------------------------------+                       
-| NVIDIA-SMI 352.39     Driver Version: 352.39         |                       
-|-------------------------------+----------------------+----------------------+
-| GPU  Name        Persistence-M| Bus-Id        Disp.A | Volatile Uncorr. ECC |
-| Fan  Temp  Perf  Pwr:Usage/Cap|         Memory-Usage | GPU-Util  Compute M. |
-|===============================+======================+======================|
-|   0  Tesla K40c          Off  | 0000:02:00.0     Off |                    0 |
-| 29%   59C    P0   136W / 235W |   4897MiB / 11519MiB |     83%      Default |
-+-------------------------------+----------------------+----------------------+
-|   1  Tesla K40m          Off  | 0000:81:00.0     Off |                    0 |
-| N/A   61C    P0   158W / 235W |   6892MiB / 11519MiB |     99%      Default |
-+-------------------------------+----------------------+----------------------+
-|   2  Tesla K40c          Off  | 0000:82:00.0     Off |                    0 |
-| 25%   48C    P0    94W / 235W |   9365MiB / 11519MiB |     31%      Default |
-+-------------------------------+----------------------+----------------------+
-                                                                               
-+-----------------------------------------------------------------------------+
-| Processes:                                                       GPU Memory |
-|  GPU       PID  Type  Process name                               Usage      |
-|=============================================================================|
-|    0     19867    C   ...b-r2014b/matlab/r2014b/bin/glnxa64/MATLAB  4872MiB |
-|    1     19774    C   ...b-r2014b/matlab/r2014b/bin/glnxa64/MATLAB  6867MiB |
-|    2     20285    C   ...b-r2014b/matlab/r2014b/bin/glnxa64/MATLAB  9340MiB |
[…]

### 2015-11-17 pick_gpu.py
More pick_gpu cleanup.

@@ -19,20 +19,19 @@ proc = subprocess.Popen("nvidia-smi", stdout=subprocess.PIPE,
 output, error = proc.communicate()
 if error:
     sys.stderr.write()
-    print "cpu"
+    sys.stdout.write("cpu")
 
 usage_re = re.compile(r"(?<=   )\d{1,3}(?=%)")
 matches = usage_re.findall(output)
 usage_amts = [int(usage_amt) for usage_amt in matches]
-print usage_amts
 
 # Get GPU
 gpus = [index for index in range(len(usage_amts))]
 open_gpus = [index for index in range(len(usage_amts)) if usage_amts[index] < USAGE_THRESHOLD]
 
 if open_gpus:
-    print "gpu" + str(random.choice(open_gpus))
+    sys.stdout.write("gpu" + str(random.choice(open_gpus)))
 elif gpus:
-    print "gpu" + str(random.choice(gpus))
+    sys.stdout.write("gpu" + str(random.choice(gpus)))
 else:
-    print "cpu"
+    sys.stdout.write("cpu")

### 2015-11-17 train_rembed_classifier.sh
Run script update

@@ -19,4 +19,4 @@ echo Flags: $REMBED_FLAGS
 # Log what we're running and where.
 echo `hostname` - $PBS_JOBID - $REMBED_FLAGS - at `git log --pretty=format:'%h' -n 1` >> ~/rembed_machine_assignments.txt
 
-THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=gpu2,floatX=float32 python -m rembed.models.classifier $REMBED_FLAGS
+THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=`python pick_gpu.py`,floatX=float32 python -m rembed.models.classifier $REMBED_FLAGS

### 2015-11-17 train_rembed_classifier.sh
More script tidying.

@@ -15,8 +15,9 @@
 cd $PBS_O_WORKDIR
 echo Lauching from working directory: $PBS_O_WORKDIR
 echo Flags: $REMBED_FLAGS
+DEVICE=`python pick_gpu.py`
 
 # Log what we're running and where.
-echo `hostname` - $PBS_JOBID - $REMBED_FLAGS - at `git log --pretty=format:'%h' -n 1` >> ~/rembed_machine_assignments.txt
+echo `hostname` - $PBS_JOBID - $REMBED_FLAGS - $DEVICE - at `git log --pretty=format:'%h' -n 1` >> ~/rembed_machine_assignments.txt
 
-THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=`python pick_gpu.py`,floatX=float32 python -m rembed.models.classifier $REMBED_FLAGS
+THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=$DEVICE,floatX=float32 python -m rembed.models.classifier $REMBED_FLAGS

### 2015-11-17 pick_gpu.py
Formatting tweak.

@@ -15,7 +15,7 @@ import re
 USAGE_THRESHOLD = 90
 
 proc = subprocess.Popen("nvidia-smi", stdout=subprocess.PIPE, 
-                                   stderr=subprocess.PIPE)
+                                      stderr=subprocess.PIPE)
 output, error = proc.communicate()
 if error:
     sys.stderr.write()

### 2016-01-18 writing/gist/gist.tex
Lots of things!
Incomplete update to gist.tex

@@
   \centering\small
-  \begin{tabular}{ccccc} 
+  \begin{tabular}{cccccc} 
     \toprule
-    Name & Stack Representation & Input Representation & Ops Classifier & Op Predictions Used in Training \\ 
+    Name & Stack Representation & Input Representation & Ops Classifier & \multicolumn{2}{l}{Op Predictions Used In} \\ 
+      &   &  &   & Training & Testing \\ 
     \midrule
-Model 0 & Discrete & Op. sequence & N & -- \\
-Model 1 & Discrete & \bf Discrete Buffer & \bf Y: Directly supervised & \bf N \\
-Model 2 & Discrete & Discrete Buffer & Y: Directly supervised & \bf Y \\
-Model 3 & \bf Soft & \bf Soft Buffer & Y: Directly supervised & Y \\
-Model 4 & Soft & Soft Buffer & \bf Y: Indirectly supervised & Y \\
+Model 0 & Discrete & Op. sequence & N & -- & -- \\
+Model 1 & Discrete & \bf Discrete Buffer & \bf Y: Directly supervised & \bf N & Y \\
+Model 2 & Discrete & Discrete Buffer & Y: Directly supervised & \bf Y & Y \\
+Model 2S & Discrete & Discrete Buffer & Y: Directly supervised & \bf Sched. sampling & Y \\
+Model 3 & \bf Soft & \bf Soft Buffer & Y: Directly supervised & Y & Y \\
+Model 4 & Soft & Soft Buffer & \bf Y: Indirectly supervised & Y & Y \\
     \bottomrule
@@
 \end{table*}
+
+Our first paper will be focused on Model 0 and Model 2S, with no discussion of soft stack models.
+Our primary goal with this paper will be to achieve state-of-the-art performance on SNLI with Model 0, and performance with Model 2S that surpasses that of a baseline LSTM encoder.
+We will pursue the state of the art both in the Siamese architecture regime with sentence encoders and in the \citealt{rocktaschel2015reasoning}-style word-by-word attention regime (reimplemented as constituent-by-constituent attention (See \S\ref{sec:c-by-c}).
 
@@
 
+\subsection{Constituent-by-constituent attention}\label{sec:c-by-c}
+
+We aim to build on the results of \citealt{rocktaschel2015reasoning} and \citealt{wang2015learning}, who find that neural soft attention models is an extremely effective technique for learning natural language inference.
+In particular, both papers use versions of word-by-word entailment, in which a latent alignment is learned from every word in the hypothesis to one or more words of the premise.
+We propose to borrow this basic idea, but to adapt it to a tree-structured setting, proposing \textit{constituent-by-constituent} attention.
+While these models do attention over a matrix $\mathbf{Y}$ of word-in-context representations from the premise encoder, we will perform attention instead over our own primary data structure, $\mathbf{Y}^{st}$, the matrix of vectors that have appeared at the top of the stack during premise encoding, which correspond one-to-one to the constituents in the tree structure representing the premise.
+Similarly, while the previous models perform one instance of soft attention conditioned on each word in the hypothesis, we perform one instance of soft attention conditioned on each stack top in the hypothesis encoder, representing the constituents of the hypothesis tree.
+
+In our model, attention is performed at each step $t$ of the premise encoder. At step $t$, the query vector that drives attention will be $S^t_0$, the top of the stack.
+
 \section{Implementation notes}
@@
 
+\subsection{Memory management and the thin stack}
+
+\todo{JG, fill out.}
+
 \section{Experiments}
@@
 
+(some of) the Tesla K40(s) used for this research was/were donated by the NVIDIA Corporation.
+
 \bibliographystyle{acl}

### 2016-01-19 make_snli_sweep.py
Sweep file cleanup.

@@ -32,8 +32,8 @@ FIXED_PARAMETERS = {
     "clipping_max_value":  "5.0",
     "batch_size":  "32",
     "init_range": "0.005",
-    "ckpt_path":    os.path.join("/scr", getpass.getuser())  # Launching user's home scr dir
-    "log_path":    os.path.join("/scr", getpass.getuser())  # Launching user's home scr dir
+    "ckpt_path":    os.path.join("/scr/", getpass.getuser(), "/"),  # Launching user's home scr dir
+    "log_path":    os.path.join("/scr/", getpass.getuser(),  "/")  # Launching user's home scr dir
 }
 
 # Tunable parameters.

### 2016-01-19 writing/gist/gist.tex
Gist update.

@@
 
-\newcommand\todo[1]{\textcolor{blue}{\textbf{TODO:} #1}}
+\newcommand\todo[1]{\textcolor{red}{\textbf{TODO:} #1}}
+\newcommand\note[1]{\textcolor{blue}{\textbf{NOTE:} #1}}
 
@@
 
+\subsection{Publication plans}
+
+\subsubsection{Winter 2016}
+
 Our first paper will be focused on Model 0 and Model 2S, with no discussion of soft stack models.
@@
 We will pursue the state of the art both in the Siamese architecture regime with sentence encoders and in the \citealt{rocktaschel2015reasoning}-style word-by-word attention regime (reimplemented as constituent-by-constituent attention (See \S\ref{sec:c-by-c}).
+
+\subsubsection{Spring 2015}
+
+Our second paper will be focused on the soft stack models, Models 3 and 4.
+It will investigate the role of differentiable stack data structures in sentence encoding and the possibility of learning consistent and useful syntactic parsers using only semantic supervision.
 
@@
 
-\subsection{Encoding the contents of the stack and buffer}
-
-The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to make even minimally informed decisions about whether to shift or reduce.
-It could benefit further from additional information about broader sentential context.
-This can be provided by running new LSTMs along the elements of each of the stack and the buffer (following \citealt{dyer-EtAl:2015:ACL-IJCNLP}) and feeding the result into the tracking LSTM.
-
 \subsection{Contextually-informed composition}
@@
 The composition function in the basic model (green) combines only the top elements of the stack, without using any further information.
-It may be possible to encourage the composition function to learn to do some amount of context-sensitive interpretation/disambiguation by adding a connection from the tracking LSTM (red) directly into the composition function.
+We can encourage the composition function to learn to do some amount of context-sensitive interpretation/disambiguation by adding a connection from the tracking LSTM (red) directly into the composition function.
 
-For Model 0, no tracking LSTM is needed for the ordinary operation of the model, but it would be possible to add one for this purpose, taking as inputs the top two values of the stack at each time point and emitting as output a context vector that can be used to condition the composition function.
+For Model 0, no tracking LSTM is needed for the ordinary operation of the model, but we can simply add one for this purpose, taking as inputs the top two values of the stack at each time point and emitting as output a context vector that can be used to condition the composition function.
 
-\subsection{Typed \reduce~operations}
-
-Shift-reduce parsers for natural language typically operate with a restricted set of typed \reduce~operations (also known as ``arc'' operations).
-These operations specify the precise relation between the elements being merged.
-It would be possible to train any of the parse-supervised models (1--3) to learn such typed arc operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\} (in the case of constituency parse supervision).
-The model can then learn a distinct composition function depending on the relation of the two elements being merged.
+So far, we have found this technique to yield non-trivial performance gain. A clearer evaluation on converged models is forthcoming.
 
@@
 In our model, attention is performed at each step $t$ of the premise encoder. At step $t$, the query vector that drives attention will be $S^t_0$, the top of the stack.
+
+\todo{[AR, RG] Write up the attention equations that you use.}
+
+\todo{[Anyone -- Sam can] Draw a diagram indicating the structure of constituent by constituent attention.}
+
+\subsection{Encoding the contents of the stack and buffer}
+
+\note{Not currently planned.}
+
+The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to make even minimally informed decisions about whether to shift or reduce.
+It could benefit further from additional information about broader sentential context.
+This can be provided by running new LSTMs along the elements of each of the stack and the buffer (following \citealt{dyer-EtAl:2015:ACL-IJCNLP}) and feeding the result into the tracking LSTM.
+
+\subsection{Typed \reduce~operations}
+
+\note{Not currently planned.}
+
+Shift-reduce parsers for natural language typically operate with a restricted set of typed \reduce~operations (also known as ``arc'' operations).
+These operations specify the precise relation between the elements being merged.
+It would be possible to train any of the parse-supervised models (1--3) to learn such typed arc operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\} (in the case of constituency parse supervision).
+The model can then learn a distinct composition function depending on the relation of the two elements being merged.
 
@@
 
-\todo{JG, fill out.}
+\todo{[JG] Explain.}
 
@@
 \subsection{Step time}
+
+\todo{Does anyone know of a better baseline TreeRNN implementation that we can use on Jagupard?
+We can use the CoreNLP SST model, but using a Java model as a baseline seems worrying unless we're guaranteed that it's competitively fast.}
 
@@
 \subsection{Sentiment}
+
+\note{We're running sentiment experiments to evaluate the behavior of converged models, since SNLI takes too long to train to convergence. However, we don't currently plan to publish SNLI results.}
 

### 2016-02-20 rembed/stack.py
thin-stack bprop: start a pretty refactor

@@ -545,31 +545,16 @@ class HardStack(object):
         stack_shift = T.cast(batch_range, theano.config.floatX)
         buffer_shift = T.cast(batch_range * self.seq_length, theano.config.floatX)
 
-        def step_b(# sequences
-                   t_f, transitions_t_f, stack_2_ptrs_t, buffer_cur_t,
-                   dE,
-                   # rest (incl. outputs_info, non_sequences)
-                   *accum_and_non_sequences):
-            # Separate the accum arguments from the non-sequence arguments.
-            n_extras = len(extra_bwd_init)
-            accum_deltas = accum_and_non_sequences[:-2 - n_extras - 1]
-            stack_bwd_t = accum_and_non_sequences[-2 - n_extras - 1]
-            extra_bwd = accum_and_non_sequences[-2 - n_extras:-2]
-            id_buffer, stack_final = accum_and_non_sequences[-2:]
+        def lookup(t_f, stack_fwd, stack_2_ptrs_t, buffer_cur_t,
+                  stack_bwd_t, extra_bwd):
+            """Retrieve all relevant bwd inputs/outputs at time `t`."""
 
-            # At first iteration, drop the external error signal into the main
-            # backward stack.
-            stack_bwd_next = ifelse(T.eq(t_f, self.seq_length - 1),
-                                    T.set_subtensor(stack_bwd_t[-self.batch_size:], error_signal),
-                                    stack_bwd_t)
-
-            err_prev = cuda_util.AdvancedSubtensor1Floats("B_errprev")(
-                stack_bwd_next, t_f * batch_size + stack_shift)
-
-            # Retrieve gradient of cost w.r.t. "extra" output
+            grad_cursor = t_f * batch_size + stack_shift
+            main_grad = cuda_util.AdvancedSubtensor1Floats("B_maingrad")(
+                stack_bwd_t, grad_cursor)
             extra_grads = tuple([
                 cuda_util.AdvancedSubtensor1Floats("B_extragrad_%i" % i)(
-                    extra_bwd_i, t_f * batch_size + stack_shift)
+                    extra_bwd_i, grad_cursor)
                 for i, extra_bwd_i in enumerate(extra_bwd)])
 
             # Find the timesteps of the two elements involved in the potential
@@ -577,34 +562,66 @@ class HardStack(object):
             t_c1 = (t_f - 1.0) * batch_size + stack_shift
             t_c2 = stack_2_ptrs_t
 
[…]

### 2016-02-22 rembed/models/classifier.py
Attention using only h states now

@@ -182,7 +182,7 @@ def build_sentence_pair_model(cls, vocab_size, seq_length, tokens, transitions,
         context_sensitive_use_relu=FLAGS.context_sensitive_use_relu,
         use_attention=FLAGS.use_attention)
 
-    premise_stack_tops = premise_model.stack_tops if FLAGS.use_attention else None
+    premise_stack_tops = premise_model.stack_tops if FLAGS.use_attention != "None" else None
 
     hypothesis_model = cls(
         FLAGS.model_dim, FLAGS.word_embedding_dim, vocab_size, seq_length,
@@ -214,9 +214,10 @@ def build_sentence_pair_model(cls, vocab_size, seq_length, tokens, transitions,
     mlp_input_dim = 2 * FLAGS.model_dim
 
     # Use the attention weighted representation
-    if FLAGS.use_attention:
-        mlp_input = hypothesis_model.final_weighed_representation.reshape((-1, FLAGS.model_dim))
-        mlp_input_dim = FLAGS.model_dim
+    if FLAGS.use_attention != "None":
+        h_dim = FLAGS.model_dim / 2
+        mlp_input = hypothesis_model.final_weighed_representation.reshape((-1, h_dim))
+        mlp_input_dim = h_dim
 
     if FLAGS.use_difference_feature:
         mlp_input = T.concatenate([mlp_input, premise_vector - hypothesis_vector], axis=1)

### 2016-02-22 rembed/stack.py

@@ -223,7 +223,7 @@ class HardStack(object):
 
     def _step(self, transitions_t, ss_mask_gen_matrix_t, stack_t, buffer_cur_t, 
             tracking_hidden, attention_hidden, stack_pushed, stack_merged, buffer, 
-            ground_truth_transitions_visible, premise_stack_tops, weighted_stack_tops):
+            ground_truth_transitions_visible, premise_stack_tops, projected_stack_tops):
         batch_size, _ = self.X.shape
 
         # Extract top buffer values.
@@ -294,9 +294,10 @@ class HardStack(object):
 
         # If attention is to be used and premise_stack_tops is not None i.e.
         # we're processing the hypothesis- Calculate the attention weighed representation
-        if self.use_attention and self.is_hypothesis:
-            attention_hidden = self._attention_unit(attention_hidden, stack_next[:, 0], premise_stack_tops, 
-                weighted_stack_tops, self.model_dim, self._vs, name="attention_unit")
+        if self.use_attention != "None" and self.is_hypothesis:
+            h_dim = self.model_dim / 2
+            attention_hidden = self._attention_unit(attention_hidden, stack_next[:, 0, :h_dim], premise_stack_tops, 
+                projected_stack_tops, h_dim, self._vs, name="attention_unit")
 
         # Move buffer cursor as necessary. Since mask == 1 when merge, we
         # should increment each buffer cursor by 1 - mask.
@@ -367,7 +368,8 @@ class HardStack(object):
 
         # Initialize the attention representation if needed
         if self.use_attention:
-            attention_init = T.zeros((batch_size, self.model_dim))
+            h_dim = self.model_dim / 2
+            attention_init = T.zeros((batch_size, h_dim))
         else:
             attention_init = DUMMY
 
@@ -393,9 +395,10 @@ class HardStack(object):
 
         non_sequences = [stack_pushed, stack_merged, buffer_t, self.ground_truth_transitions_visible]
 
-        if self.use_attention and self.is_hypothesis:
-            weighted_stack_tops = util.AttentionUnitInit(self.premise_stack_tops, self.model_dim, self._vs)
-            non_sequences = non_sequences + [self.premise_stack_tops, weighted_stack_tops]
+        if self.use_attention != "None" and self.is_hypothesis:
+            h_dim = self.model_dim / 2
[…]

### 2016-02-22 rembed/util.py

@@ -326,14 +326,14 @@ def TrackingUnit(state_prev, inp, inp_dim, hidden_dim, vs, name="track_unit", ma
     return state, logits
 
 
-def WangJiangAttentionUnit(attention_state_prev, lstm_input, premise_stack_tops, weighted_stack_tops, model_dim, 
+def WangJiangAttentionUnit(attention_state_prev, current_stack_top, premise_stack_tops, projected_stack_tops, model_dim, 
                     vs, name="attention_unit", initializer=None):
     """
     Args:
       attention_state_prev: The output of this unit at the previous time step.
-      lstm_input: The current stack top. (Strange name?)
+      current_stack_top: The current stack top.
       premise_stack_tops: The values to do attention over.
-      weighted_stack_tops: Projected (not weighted?) vectors to use to produce an attentive
+      projected_stack_tops: Projected vectors to use to produce an attentive
           weighting alpha_t.
       model_dim: The dimension of the vectors over which to do attention.
       vs: A variable store for the learned parameters.
@@ -349,11 +349,11 @@ def WangJiangAttentionUnit(attention_state_prev, lstm_input, premise_stack_tops,
     W_r = vs.add_param("%s_W_r" % name, (model_dim, model_dim), initializer=initializer)
     w = vs.add_param("%s_w" % name, (model_dim,), initializer=initializer)
 
-    W_h__h_t = T.dot(lstm_input, W_h)
+    W_h__h_t = T.dot(current_stack_top, W_h)
     W_r__r_t_prev = T.dot(attention_state_prev, W_r)
 
     # Shape: L x B x k
-    M_t = T.tanh(weighted_stack_tops + (W_h__h_t + W_r__r_t_prev))
+    M_t = T.tanh(projected_stack_tops + (W_h__h_t + W_r__r_t_prev))
 
     # Shape: B x L
     alpha_t = T.nnet.softmax(T.dot(M_t, w).T)
@@ -361,21 +361,21 @@ def WangJiangAttentionUnit(attention_state_prev, lstm_input, premise_stack_tops,
     # Shape B x k
     Y__alpha_t = T.sum(premise_stack_tops * alpha_t.T[:, :, np.newaxis], axis=0)
 
-    mlstm_input = T.concatenate([Y__alpha_t, lstm_input], axis=1)
+    mlstm_input = T.concatenate([Y__alpha_t, current_stack_top], axis=1)
 
[…]

### 2016-02-22 writing/gist/tree_attn_fig.tex
Tree attention figure to gist.

%!TEX root = gist.tex

\begin{figure*}[t]
\centering
\scalebox{0.8}{
 \begin{tikzpicture}
    \tikzstyle{word}=[fill=black!20,text height=2mm]    
    \tikzstyle{nonleaf}=[fill=black!20,text height=2mm]

    \begin{scope}[shift={(0in,0in)}, frontier/.style={distance from root=60pt}]

    \Tree [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ]

    \end{scope}

    \begin{scope}[shift={(2in,0in)}, frontier/.style={distance from root=60pt}]

    \Tree [.\node[nonleaf](2thekittenmeowed){the kitten meowed}; [.\node[nonleaf](2thekitten){the kitten}; \node[word](2the){the}; \node[word](2kitten){kitten}; ] \node[word](2meowed){meowed}; ]

    \end{scope}

    \usetikzlibrary{arrows,scopes}

    \tikzstyle{attn} = [line width=1pt,draw=black!40,opacity=0.25,line cap=round]
    \tikzstyle{heavy} = [line width=2pt]
    \tikzstyle{light} = [line width=0.5pt]
    \tikzstyle{focus} = [draw=blue!40,opacity=1.0]

    \draw [attn,heavy] (1the) to[in=155,out=25] (2the);
    \draw [attn,light] (1thecat) to[in=155,out=25] (2the);

    \draw [attn,heavy] (1cat) to[in=155,out=25] (2kitten);
    \draw [attn,light] (1thecat) to[in=155,out=25] (2kitten);
    \draw [attn,light] (1thecatsatdown) to[in=155,out=25] (2kitten);

    \draw [attn,light] (1satdown) to[in=155,out=25] (2thekittenmeowed);
    \draw [attn,light] (1thecat) to[in=155,out=25] (2thekittenmeowed);
    \draw [attn,heavy] (1thecatsatdown) to[in=155,out=25] (2thekittenmeowed);

    \draw [attn] (1satdown) to[in=155,out=25] (2meowed);
    \draw [attn] (1sat) to[in=155,out=25] (2meowed);
    \draw [attn,light] (1down) to[in=155,out=25] (2meowed);

    \draw [attn,focus] (1cat) to[in=155,out=25,] (2thekitten);
    \draw [attn,heavy,focus] (1thecat) to[in=155,out=25] (2thekitten);
    \draw [attn,light,focus] (1the) to[in=155,out=25] (2thekitten);
    \draw [attn,light,focus] (1thecatsatdown) to[in=155,out=25] (2thekitten);

  \end{tikzpicture}}

 \caption{Soft alignments between the nodes of two trees, with alignments to ``the kitten'' highlighted in blue.}\label{fig:tree_attn}
\end{figure*}

### 2016-02-22 rembed/stack.py
Fix silly masking typo in input gradient calculation

@@ -663,7 +663,7 @@ class ThinStack(object):
                 base = new_stacks.get(stack, stack)
 
                 mask_i = masks[m_delta.ndim - 1]
-                delta = mask_i + m_delta + (1. - mask_i) * p_delta
+                delta = mask_i * m_delta + (1. - mask_i) * p_delta
 
                 # Run subtensor update on associated structure using the
                 # current cursor.

### 2016-02-24 rembed/stack.py
Add flag compute_embedding_gradients to make_backprop_scan

@@ -472,7 +472,9 @@ class ThinStack(object):
 
         return wrt, f_p_delta, f_m_delta
 
-    def make_backprop_scan(self, error_signal, extra_cost_inputs=None):
+    def make_backprop_scan(self, error_signal,
+                           extra_cost_inputs=None,
+                           compute_embedding_gradients=True):
         """
         Args:
             error_signal: The external gradient d(cost)/d(stack top). A Theano
@@ -503,10 +505,14 @@ class ThinStack(object):
         self._zero_updates += wrt_vars
 
         # Also accumulate embedding gradients separately
-        dE = theano.shared(np.zeros(self.embeddings.get_value().shape,
-                                    dtype=np.float32),
-                           name="bwd/wrt/embeddings")
-        self._zero_updates.append(dE)
+        if compute_embedding_gradients:
+            dE = theano.shared(np.zeros(self.embeddings.get_value().shape,
+                                        dtype=np.float32),
+                               name="bwd/wrt/embeddings")
+            self._zero_updates.append(dE)
+        else:
+            # Make dE a dummy variable.
+            dE = T.zeros((1,))
 
         # Useful batch zero-constants.
         zero_stack = T.zeros((self.batch_size, self.model_dim))
@@ -603,10 +609,17 @@ class ThinStack(object):
                      mask.dimshuffle(0, "x", "x")]
 
             # Accumulate inp deltas, switching over push/merge decision.
-            stacks = (stack_bwd_next, stack_bwd_next, dE) + extra_bwd
+            stacks = (stack_bwd_next, stack_bwd_next,
+                      (compute_embedding_gradients and dE) or None)
+            cursors = (t_c1, t_c2,
+                       (compute_embedding_gradients and buffer_ids_t) or None)
+            # Handle potential aux bwd stacks.
+            stacks += extra_bwd
+            cursors += ((t_c1,)) * len(extra_bwd)
             new_stacks = {}
-            cursors = (t_c1, t_c2, buffer_ids_t) + ((t_c1,) * len(extra_bwd))
             for stack, cursor, m_delta, p_delta in zip(stacks, cursors, m_delta_inp, p_delta_inp):
+                if stack is None or cursor is None:
+                    continue
                 base = new_stacks.get(stack, stack)
 
                 mask_i = masks[m_delta.ndim - 1]
[…]

### 2016-02-24 rembed/tests/test_stack.py

@@ -177,7 +177,7 @@ class ThinStackBackpropTestCase(unittest.TestCase, BackpropTestMixin):
 
         return c2
 
-    def test_backprop(self):
+    def _test_backprop(self, embedding_grad=True):
         # Simulate a batch of two token sequences, each with the same
         # transition sequence
         X = np.array([[0, 1, 2, 3, 1], [2, 1, 3, 0, 1]], dtype=np.int32)
@@ -199,19 +199,31 @@ class ThinStackBackpropTestCase(unittest.TestCase, BackpropTestMixin):
         error_signal = T.grad(cost, top)
 
         # Build automatic backprop function.
-        self.stack.make_backprop_scan(error_signal, [self.y])
+        self.stack.make_backprop_scan(error_signal, [self.y],
+                                      compute_embedding_gradients=embedding_grad)
+        outputs = (cost, self.stack.gradients[W], self.stack.gradients[b])
+        if embedding_grad:
+            outputs = outputs + (self.stack.embedding_gradients,)
         f = theano.function(
-            [self.X, self.transitions, self.y],
-            (cost, self.stack.gradients[W], self.stack.gradients[b], self.stack.embedding_gradients),
+            [self.X, self.transitions, self.y], outputs,
             updates=self.stack.scan_updates + self.stack.bscan_updates)
 
         b_cost_sim, b_dW_sim, b_db_sim, b_embedding_gradients_sim = f_simulated(X, y)
-        b_cost, b_dW, b_db, b_embedding_gradients = f(X, transitions, y)
+        real_ret = f(X, transitions, y)
+        b_cost, b_dW, b_db = real_ret[:3]
 
         np.testing.assert_almost_equal(b_cost_sim, b_cost)
         np.testing.assert_almost_equal(b_dW_sim, b_dW)
         np.testing.assert_almost_equal(b_db_sim, b_db)
-        np.testing.assert_almost_equal(b_embedding_gradients_sim, b_embedding_gradients)
+
+        if embedding_grad:
+            np.testing.assert_almost_equal(b_embedding_gradients_sim, real_ret[3])
+
+    def test_backprop_with_embedding_gradients(self):
+        self._test_backprop(True)
+
+    def test_backprop_without_embedding_gradients(self):
+        self._test_backprop(False)
 
 
 class ThinStackTrackingBackpropTestCase(unittest.TestCase, BackpropTestMixin):

### 2016-02-24 rembed/models/classifier.py
Use new compute_embedding_gradients flag in classifier.py

@@ -609,10 +609,12 @@ def run(only_forward=False):
         # Train
         if data_manager.SENTENCE_PAIR_DATA:
             premise_error_signal = T.grad(total_cost, premise_stack_top)
-            premise_model.make_backprop_scan(premise_error_signal)
+            premise_model.make_backprop_scan(premise_error_signal,
+                                             compute_embedding_gradients=False)
 
             hypothesis_error_signal = T.grad(total_cost, hypothesis_stack_top)
-            hypothesis_model.make_backprop_scan(hypothesis_error_signal)
+            hypothesis_model.make_backprop_scan(hypothesis_error_signal,
+                                                compute_embedding_gradients=False)
 
             gradients = premise_model.gradients
             hypothesis_gradients = hypothesis_model.gradiennts
@@ -623,7 +625,8 @@ def run(only_forward=False):
             new_values += hypothesis_model.scan_updates.items() + hypothesis_model.bscan_updates.items()
         else:
             error_signal = T.grad(total_cost, stack_top)
-            model.make_backprop_scan(error_signal)
+            model.make_backprop_scan(error_signal,
+                                     compute_embedding_gradients=train_embeddings)
             if train_embeddings:
                 model.gradients[model.embeddings] = model.embedding_gradients
             gradients = model.gradients

### 2016-02-24 writing/hard_stack_paper/paper.tex
Initial outline of the paper.

--- 2016-01-18 writing/gist/gist.tex
@@
 \documentclass[11pt,letterpaper]{article}
-\usepackage{acl2015}
+\usepackage{../acl2015}
 \usepackage{times}
 \usepackage{latexsym}
-% \setlength\titlebox{5cm}    % Expanding the titlebox
+% \setlength\titlebox{7.5cm}    % Expanding the titlebox
 
@@
 \usepackage{tikz}
+\usepackage{tikz-qtree}
 \usepackage{ifthen}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%% %%%
-
-\title{NSHRDLU:\thanks{Name is provisional, and stolen from Geoff Hinton.}\\Work in progress on the joint learning of parsing and semantic encoding}
+\title{Tree-structured neural networks revisited}
 
@@
 Jon Gauthier$^{\dag\ddag}$ \\
-\texttt{angeli@stanford.edu} \\
+\texttt{jgauthie@stanford.edu} \\
+\And
+Abhinav Rastogi$^{\dag\S}$ \\
+\texttt{arastogi@stanford.edu} \\
 \AND
-Christopher D.\ Manning$^{\ast\dag\S}$\\
+Raghav Gupta$^{\dag\mathparagraph}$ \\
+\texttt{rgupta93@stanford.edu} \\
+\And
+Christopher D.\ Manning$^{\ast\dag\mathparagraph}$\\
 \texttt{manning@stanford.edu}\\
@@
 {$^{\ast}$Stanford Linguistics\quad
-$^{\dag}$Stanford NLP Group}\\
-{$^{\ddag}$Stanford Symbolic Systems\quad
-$^{\S}$Stanford Computer Science}
+$^{\dag}$Stanford NLP Group\quad
+$^{\ddag}$Stanford Symbolic Systems}\\
+{$^{\S}$Stanford Electrical Engineering\quad
+$^{\mathparagraph}$Stanford Computer Science}
 }
@@
 \makeatother
-\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black]{hyperref}
+\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black, draft]{hyperref}
 
@@
 \maketitle
+\todo{Anonymize.}
+
+\begin{abstract}
+\begin{itemize}
+\item TreeRNNs are great, don't work for big data.
+\item We present the transition-based TreeRNN implementation -- makes efficient training possible, show that it helps relative to seq models.
+\item Further, attention has proven powerful for extending sequence models. We introduce constit-by-constit attention, and find it to be effective here, too.
+\end{itemize}
+\end{abstract}
 
@@
 
-This project aims to use ideas from greedy transition-based parsing to build neural network models that can jointly learn to parse sentences and to use those parses to guide semantic composition.
-
-Table~\ref{models-table} shows the sequence of model designs that we plan to build, and Figure~\ref{m1-views} depicts some representative models.
-
-We see three reasons to pursue this approach:
 \begin{itemize}
-\item Simply adapting a greedy transition-based approach to sentence encoding makes it possible to exploit semantic compositionality in the same manner as in a TreeRNN, but using a static graph structure that can take advantage of existing neural network libraries like Theano for both automatic differentiation and for highly optimized matrix computations on both CPUs and GPUs.
-Model 0 pursues this property directly, and all subsequent models share it.
-\item When a parsing system and an interpretation system are trained jointly and forced to share representations, it is likely that the performance of both models will benefit: the semantic composition function will have better access to syntactic type information that could provide additional evidence on the functional behavior of rare words, and the syntactic parser will have access to information about the semantic interpretability of constituents, making semantically-conditioned parsing decisions like PP attachment easier to learn.
-Models 1--3 have this property.
-\item When a system that uses constituent structure information as a latent variable is initialized from scratch and trained solely to perform some semantic task, it will likely learn some coherent syntax for the genre of language on which it is trained.
-If this syntax is reasonably stable across similar training sets and across random initializations of training, it can reasonably be said to be latent in the text itself, offering a new kind of evidence about natural language syntax.
-Model 4 aims to collect this kind of evidence.
-
+\item Brief history of TreeRNNs
+\item Speed issues
+\item Attention issues
 \end{itemize}
 
-All five models function as sentence encoding models: their inputs are sentences (input as sequences of words, with the help of a learned embedding matrix), and their outputs are single sentence encoding vectors which can be used as the inputs to downstream models for tasks like sentiment analysis, translation, or inference.
-Provided that these downstream models are differentiable---as is the case for neural network models and simpler regression models---their gradient signals can be used to train the sentence encoding models.
+\section{The transition-based sentence model}
 
-\begin{table*}[t]
-  \centering\small
-  \begin{tabular}{cccccc} 
-    \toprule
-    Name & Stack Representation & Input Representation & Ops Classifier & \multicolumn{2}{l}{Op Predictions Used In} \\ 
-      &   &  &   & Training & Testing \\ 
-    \midrule
-Model 0 & Discrete & Op. sequence & N & -- & -- \\
-Model 1 & Discrete & \bf Discrete Buffer & \bf Y: Directly supervised & \bf N & Y \\
-Model 2 & Discrete & Discrete Buffer & Y: Directly supervised & \bf Y & Y \\
-Model 2S & Discrete & Discrete Buffer & Y: Directly supervised & \bf Sched. sampling & Y \\
-Model 3 & \bf Soft & \bf Soft Buffer & Y: Directly supervised & Y & Y \\
-Model 4 & Soft & Soft Buffer & \bf Y: Indirectly supervised & Y & Y \\
-    \bottomrule
-  \end{tabular}
-  \protect\caption{\protect\label{models-table}Model variants, ordered by increasing reliance on learning. Bolding indicates the differences between each model and its parent model.} 
-\end{table*}
+\subsection{Intuition: Transition-based parsing}
 
-Our first paper will be focused on Model 0 and Model 2S, with no discussion of soft stack models.
-Our primary goal with this paper will be to achieve state-of-the-art performance on SNLI with Model 0, and performance with Model 2S that surpasses that of a baseline LSTM encoder.
-We will pursue the state of the art both in the Siamese architecture regime with sentence encoders and in the \citealt{rocktaschel2015reasoning}-style word-by-word attention regime (reimplemented as constituent-by-constituent attention (See \S\ref{sec:c-by-c}).
+\subsection{The model}
 
-\section{Models}
+\todo{[SB] Reframe in terms of Model 0 alone.}
 
@@
 
-\input{model1_fig.tex}
+\input{../model1_fig.tex}
 
@@
 
-\subsection{Model 3}
+\subsection{Tracking left context with an auxiliary LSTM} Includes both the tracking LSTM and context-sensitive composition.
 
-Model 3 modifies Model 2 by introducing the soft stack/soft queue from \cite{grefenstette2015learning} in place of the hard, conventional stack and buffer.
-The soft stack makes it possible to for the model to predict smooth distributions over operations of the form (0.93 \shift, 0.07 \reduce), instead of making hard decisions.
-These soft decisions allow for gradient information to flow from the stack and the buffer back into the operation classifier (blue).
-This is crucial to our ultimate goal, as it makes it possible for semantic considerations to influence the model's parsing decisions.
+\subsection{Building a model}
 
-Model 3 still receives a direct supervision signal from some existing parser.
-In order to train the soft stack, we must represent the hard supervision signal from the parser by a soft prediction which matches the soft stack operation output.
-The supervision signal simply assigns 100\% weight to the ground-truth operation.
+\paragraph{The size of the stack}
+The size of the stack should be $N$ for sentences of $N$ words, in case the first reduce merges the final two words. The size of the buffer should be $N$.
 
-\subsection{Model 4}
-
-Model 4 modifies Model 3 by removing the direct supervision signal from the operation classifier (blue), instead forcing the operation classifier to learn solely from the gradient provided by the downstream supervision task.
-It may be possible to accelerate or otherwise improve Model 4 training by initializing it with parameters learned by Model 3.
-
-By removing an external parser signal, we allow Model 4 to fully exploit the soft stack representation.
-It is free to predict soft parse operations in the case of ambiguous parses.
-This distinguishes it from Model 3, which is encouraged to replicate the 100\%-certain ground truth parse predictions provided by the external parser.
-
-\section{Other possible model features}
-
-\subsection{Encoding the contents of the stack and buffer}
-
-The tracking LSTM (red) needs access to the top of the buffer and the top two elements of the stack in order to make even minimally informed decisions about whether to shift or reduce.
-It could benefit further from additional information about broader sentential context.
-This can be provided by running new LSTMs along the elements of each of the stack and the buffer (following \citealt{dyer-EtAl:2015:ACL-IJCNLP}) and feeding the result into the tracking LSTM.
-
-\subsection{Contextually-informed composition}
-
-The composition function in the basic model (green) combines only the top elements of the stack, without using any further information.
-It may be possible to encourage the composition function to learn to do some amount of context-sensitive interpretation/disambiguation by adding a connection from the tracking LSTM (red) directly into the composition function.
-
-For Model 0, no tracking LSTM is needed for the ordinary operation of the model, but it would be possible to add one for this purpose, taking as inputs the top two values of the stack at each time point and emitting as output a context vector that can be used to condition the composition function.
-
-\subsection{Typed \reduce~operations}
-
-Shift-reduce parsers for natural language typically operate with a restricted set of typed \reduce~operations (also known as ``arc'' operations).
-These operations specify the precise relation between the elements being merged.
-It would be possible to train any of the parse-supervised models (1--3) to learn such typed arc operations, expanding the op set dramatically to something like \{\shift, \reduce-NP, \reduce-S, \reduce-PP, ...\} (in the case of constituency parse supervision).
-The model can then learn a distinct composition function depending on the relation of the two elements being merged.
-
-\subsection{Constituent-by-constituent attention}\label{sec:c-by-c}
-
-We aim to build on the results of \citealt{rocktaschel2015reasoning} and \citealt{wang2015learning}, who find that neural soft attention models is an extremely effective technique for learning natural language inference.
-In particular, both papers use versions of word-by-word entailment, in which a latent alignment is learned from every word in the hypothesis to one or more words of the premise.
-We propose to borrow this basic idea, but to adapt it to a tree-structured setting, proposing \textit{constituent-by-constituent} attention.
-While these models do attention over a matrix $\mathbf{Y}$ of word-in-context representations from the premise encoder, we will perform attention instead over our own primary data structure, $\mathbf{Y}^{st}$, the matrix of vectors that have appeared at the top of the stack during premise encoding, which correspond one-to-one to the constituents in the tree structure representing the premise.
-Similarly, while the previous models perform one instance of soft attention conditioned on each word in the hypothesis, we perform one instance of soft attention conditioned on each stack top in the hypothesis encoder, representing the constituents of the hypothesis tree.
-
-In our model, attention is performed at each step $t$ of the premise encoder. At step $t$, the query vector that drives attention will be $S^t_0$, the top of the stack.
-
-\section{Implementation notes}
-
-The size of the stack should be $N$ for sentences of $N$ words, in case the first reduce merges the final two words. The size of the buffer should be $N - 2$.
+\paragraph{Handling overly long sentences}
 
@@
 
-\subsection{Memory management and the thin stack}
+\paragraph{Representing the stack in memory} Or, thin stack. 
+\todo{[JG] Explain.}
 
-\todo{JG, fill out.}
+\paragraph{Optimization and hyperparameters}
 
-\section{Experiments}
+\paragraph{Software infrastructure} We will make Theano code available.
 
 \subsection{Step time}
+
+\todo{Does anyone know of a better baseline TreeRNN implementation that we can use on Jagupard?
+We can use the CoreNLP SST model, but using a Java model as a baseline seems worrying unless we're guaranteed that it's competitively fast.}
 
@@
 
-\subsection{Sentiment}
+\section{Attention over trees}
 
-Learning to jointly parse and to predict logical relations between sentences over SST  
-\cite{socher2013recursive}.
+\input{../tree_attn_fig.tex}
 
-Evaluation metrics: accuracy, F1 (for all models but 0).
+\paragraph{Constituent-by-constituent attention}
+We aim to build on the results of \citealt{rocktaschel2015reasoning} and \citealt{wang2015learning}, who find that neural soft attention models is an extremely effective technique for learning natural language inference.
+In particular, both papers use versions of word-by-word entailment, in which a latent alignment is learned from every word in the hypothesis to one or more words of the premise.
+We propose to borrow this basic idea, but to adapt it to a tree-structured setting, proposing \textit{constituent-by-constituent} attention.
+While these models do attention over a matrix $\mathbf{Y}$ of word-in-context representations from the premise encoder, we will perform attention instead over our own primary data structure, $\mathbf{Y}^{st}$, the matrix of vectors that have appeared at the top of the stack during premise encoding, which correspond one-to-one to the constituents in the tree structure representing the premise.
+Similarly, while the previous models perform one instance of soft attention conditioned on each word in the hypothesis, we perform one instance of soft attention conditioned on each stack top in the hypothesis encoder, representing the constituents of the hypothesis tree.
 
-\subsection{Natural language inference}
+In our model, attention is performed at each step $t$ of the premise encoder. At step $t$, the query vector that drives attention will be $S^t_0$, the top of the stack.
 
-Learning to jointly parse and to predict logical relations between sentences over SNLI  \cite{snli:emnlp2015}.
+\todo{[AR, RG] Write up the attention equations that you use.}
 
-Evaluation metrics: accuracy, F1 (for all models but 0).
+\subsection{Accumulating the results of attention}
+
+\paragraph{Accumulating results sequentially} (After Rocktäschel, Wang and Jiang)
+
+\paragraph{Accumulating results using a second tree layer}
+
+\section{Experiments}
+
+\paragraph{The Stanford Natural Language Inference task} \cite{snli:emnlp2015}
+
+\paragraph{Results}
 
@@
 
-\subsection{Inferred Model 4 parses?}
+\paragraph{Patterns in the examples that the tree model got right that the LSTM didn't}
+
+\paragraph{Induced alignments}
+
+\section{Conclusions and future work}
+
+\begin{itemize}
+\item Future work: Incorporate the parser
+\item Future work: Latent parses
+\end{itemize}
 
@@
 
-\bibliographystyle{acl}
-\bibliography{MLSemantics}
+\bibliographystyle{../acl}
+\bibliography{../MLSemantics}
 

### 2016-02-24 writing/tree_attn_fig.tex
Initial outline of the paper.

--- 2016-02-22 writing/gist/tree_attn_fig.tex
@@
-%!TEX root = gist.tex
+%!TEX root = hard_stack_paper/paper.tex
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\scalebox{0.8}{
+\scalebox{0.9}{
  \begin{tikzpicture}
-    \tikzstyle{word}=[fill=black!20,text height=2mm]    
-    \tikzstyle{nonleaf}=[fill=black!20,text height=2mm]
+    \tikzstyle{word}=[fill=yellow!40,text height=2mm]    
+    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm]
 
@@
 
-    \tikzstyle{attn} = [line width=1pt,draw=black!40,opacity=0.25,line cap=round]
-    \tikzstyle{heavy} = [line width=2pt]
-    \tikzstyle{light} = [line width=0.5pt]
-    \tikzstyle{focus} = [draw=blue!40,opacity=1.0]
+    \tikzstyle{attn} = [line width=1pt,draw=black!80,opacity=0.25,line cap=round]
+    \tikzstyle{heavy} = [line width=3pt]
+    \tikzstyle{light} = [line width=0.33pt]
+    \tikzstyle{focus} = [draw=red,opacity=1.0]
 
@@
 
- \caption{Soft alignments between the nodes of two trees, with alignments to ``the kitten'' highlighted in blue.}\label{fig:tree_attn}
-\end{figure*}
+ \caption{Soft alignments between the nodes of two trees, with alignments to ``the kitten'' highlighted in red.}\label{fig:tree_attn}
+\end{figure}

### 2016-02-27 rembed/util.py
ensure_2d_arguments: broadcast rather than subtensor; this cleans up the backprop graph (removes 25% of time...)

@@ -864,7 +864,7 @@ def ensure_2d_arguments(f, squeeze_ret=True):
         for arg in args:
             if isinstance(arg, T.TensorVariable):
                 if arg.ndim == 1:
-                    arg = arg[np.newaxis, :]
+                    arg = arg.dimshuffle("x", 0)
                 elif arg.ndim > 2:
                     raise RuntimeError("ensure_2d_arguments wrapped a function"
                                        " which received an %i-d argument. "

### 2016-02-27 rembed/util.py
Make some nice graph names for symbolic grads_above inputs

@@ -749,8 +749,8 @@ def batch_subgraph_gradients(g_in, wrt, f_g_out, batch_size=None,
     g_out = [g_out] if not isinstance(g_out, (list, tuple)) else g_out
 
     # Prepare symbolic gradient variables for each of the graph outputs.
-    grads_above = [tensorx("%s/grad_%s" % (name, out_i.name), out_i.ndim)
-                   for out_i in g_out]
+    grads_above = [tensorx("%s/grad_%s" % (name, out_i.name or i), out_i.ndim)
+                   for i, out_i in enumerate(g_out)]
 
     # Compute gradients of subgraph beginning at `g_in` and ending at `g_out`,
     # where the cost gradient w.r.t. each `g_out` is given by the corresponding

### 2016-02-27 rembed/tests/test_stack.py
debugprint in SpeedTestCase

@@ -501,6 +501,7 @@ class ThinStackSpeedTestCase(unittest.TestCase, BackpropTestMixin):
             [self.X, self.transitions, self.y],
             [cost] + self.stack.gradients.values(),
             updates=self.stack.scan_updates + self.stack.bscan_updates)
+        theano.printing.debugprint(f.maker.fgraph.outputs[1])
 
         for t in range(10):
             self._run_batch(f)

### 2016-02-27 writing/hard_stack_paper/paper.tex
Even more affiliations.

@@
 \usepackage{framed}
+
+\newcommand\todo[1]{\textcolor{blue}{\textbf{TODO:} #1}}
+\newcommand\result[1]{\textcolor{red}{\textbf{RESULT NEEDED:} #1}}
+\newcommand\question[1]{\textcolor{orange}{\textbf{OPEN QUESTION:} #1}}
 
@@
 \And
-Abhinav Rastogi$^{\dag\S}$ \\
+Abhinav Rastogi$^{\S\mathparagraph}$ \\
 \texttt{arastogi@stanford.edu} \\
 \AND
-Raghav Gupta$^{\dag\mathparagraph}$ \\
+Raghav Gupta$^{\parallel}$ \\
 \texttt{rgupta93@stanford.edu} \\
 \And
-Christopher D.\ Manning$^{\ast\dag\mathparagraph}$\\
+Christopher D.\ Manning$^{\ast\dag\parallel}$\\
 \texttt{manning@stanford.edu}\\
@@
 {$^{\S}$Stanford Electrical Engineering\quad
-$^{\mathparagraph}$Stanford Computer Science}
+$^{\mathparagraph}$Stanford AI Lab\quad
+$^{\parallel}$Stanford Computer Science}
 }
@@
 
-\newcommand\todo[1]{\textcolor{red}{\textbf{TODO:} #1}}
-\newcommand\note[1]{\textcolor{blue}{\textbf{NOTE:} #1}}
-
 \begin{document}
 \maketitle
-\todo{Anonymize.}
-
 \begin{abstract}
@@
 \end{abstract}
+
+\todo{Anonymize.}
 

### 2016-02-27 rembed/recurrences.py
Refactor composition / layer functions to accept tuple inputs and avoid concat+split ops

@@ -98,33 +98,31 @@ class SharedRecurrenceMixin(object):
     def _tracking_lstm_predict(self, inputs, network):
         # TODO(SB): Offer more buffer content than just the top as input.
         c1, c2, buffer_top, tracking_hidden = inputs[:4]
-        inp = T.concatenate([c1, c2, buffer_top], axis=1)
-        return network(tracking_hidden, inp, self._spec.model_dim * 3,
+        inp = (c1, c2, buffer_top)
+        return network(tracking_hidden, inp, (self._spec.model_dim,) * 3,
                        self.tracking_lstm_hidden_dim, self._vs,
                        name="prediction_and_tracking")
 
     def _predict(self, inputs, network):
         # TODO(SB): Offer more buffer content than just the top as input.
-        c1, c2, buffer_top = inputs[:3]
-        inp = T.concatenate([c1, c2, buffer_top], axis=1)
-        return network(inp, self._spec.model_dim * 3,
+        inp = tuple(inputs[:3])
+        return network(inp, (self._spec.model_dim,) * 3,
                        util.NUM_TRANSITION_TYPES, self._vs,
                        name="prediction_and_tracking")
 
     def _merge(self, inputs, network):
-        c1, c2 = inputs[:2]
-        merge_items = T.concatenate([c1, c2], axis=1)
+        merge_items = tuple(inputs[:2])
         if self.use_tracking_lstm:
-            # NB: Unlike in the previous implementation, context-sensitive 
-            # composition (aka the tracking--composition connection) is not 
-            # optional here. It helps performance, so this shouldn't be a 
+            # NB: Unlike in the previous implementation, context-sensitive
+            # composition (aka the tracking--composition connection) is not
+            # optional here. It helps performance, so this shouldn't be a
             # big problem.
             tracking_h_t = inputs[3][:, :self.tracking_lstm_hidden_dim]
             return network(merge_items, tracking_h_t, self._spec.model_dim,
                            self._vs, name="compose",
                            external_state_dim=self.tracking_lstm_hidden_dim)
         else:
-            return network(merge_items, self._spec.model_dim * 2,
+            return network(merge_items, (self._spec.model_dim,) * 2,
                            self._spec.model_dim, self._vs, name="compose")
 
 

### 2016-02-27 rembed/tests/test_stack.py

@@ -24,10 +24,9 @@ class ThinStackTestCase(unittest.TestCase):
         spec = util.ModelSpec(embedding_dim, embedding_dim, self.batch_size,
                               vocab_size, seq_length)
 
-        def compose_network(inp, inp_dim, outp_dim, vs, name="compose"):
+        def compose_network((c1, c2), inp_dim, outp_dim, vs, name="compose"):
             # Just add the two embeddings!
-            W = T.concatenate([T.eye(outp_dim), T.eye(outp_dim)], axis=0)
-            return inp.dot(W)
+            return c1 + c2
 
         X = T.imatrix("X")
         transitions = T.imatrix("transitions")
@@ -146,11 +145,11 @@ class ThinStackBackpropTestCase(unittest.TestCase, BackpropTestMixin):
                               self.seq_length)
 
         self.vs = vs = VariableStore()
-        def compose_network(inp, *args, **kwargs):
+        def compose_network((c1, c2), *args, **kwargs):
             W = vs.add_param("W", (self.model_dim * 2, self.model_dim))
             b = vs.add_param("b", (self.model_dim,),
                              initializer=util.ZeroInitializer())
-            return T.dot(inp, W) + b
+            return T.dot(T.concatenate([c1, c2], axis=1), W) + b
 
         self.compose_network = compose_network
         recurrence = Model0(spec, vs, compose_network)
@@ -170,11 +169,8 @@ class ThinStackBackpropTestCase(unittest.TestCase, BackpropTestMixin):
         # seq_length * batch_size * emb_dim
         X_emb = self.stack.embeddings[self.X].dimshuffle(1, 0, 2)
 
-        z1 = T.concatenate([X_emb[1], X_emb[0]], axis=1)
-        c1 = self.compose_network(z1)
-
-        z2 = T.concatenate([X_emb[2], c1], axis=1)
-        c2 = self.compose_network(z2)
+        c1 = self.compose_network((X_emb[1], X_emb[0]))
+        c2 = self.compose_network((X_emb[2], c1))
 
         return c2
 
@@ -241,8 +237,8 @@ class ThinStackTrackingBackpropTestCase(unittest.TestCase, BackpropTestMixin):
 
         self.vs = VariableStore()
 
-        def compose_network(inp, hidden, *args, **kwargs):
-            conc = T.concatenate([hidden, inp], axis=1)
+        def compose_network((c1, c2), hidden, *args, **kwargs):
[…]

### 2016-02-27 rembed/util.py

@@ -206,9 +206,20 @@ def ReLULayer(inp, inp_dim, outp_dim, vs, name="relu_layer", use_bias=True, init
 
 
 def Linear(inp, inp_dim, outp_dim, vs, name="linear_layer", use_bias=True, initializer=None):
-    W = vs.add_param("%s_W" %
-                     name, (inp_dim, outp_dim), initializer=initializer)
-    outp = inp.dot(W)
+    if isinstance(inp, tuple):
+        assert isinstance(inp_dim, tuple)
+        Ws = [vs.add_param("%s_W%i" % (name, i), (dim_i, outp_dim),
+                           initializer=initializer)
+              for i, dim_i in enumerate(inp_dim)]
+
+        outp = T.dot(inp[0], Ws[0])
+        for inp_i, W_i in zip(inp[1:], Ws[1:]):
+            # TODO inplace add?
+            outp += T.dot(inp_i, W_i)
+    else:
+        W = vs.add_param("%s_W" %
+                         name, (inp_dim, outp_dim), initializer=initializer)
+        outp = inp.dot(W)
 
     if use_bias:
         b = vs.add_param("%s_b" % name, (outp_dim,),
@@ -248,8 +259,16 @@ def TreeLSTMLayer(lstm_prev, external_state, full_memory_dim, vs, name="tree_lst
     assert full_memory_dim % 2 == 0, "Input is concatenated (h, c); dim must be even."
     hidden_dim = full_memory_dim / 2
 
-    W = vs.add_param("%s/W" % name, (hidden_dim * 2 + external_state_dim, hidden_dim * 5),
-                     initializer=initializer)
+    assert isinstance(lstm_prev, tuple)
+    l_prev, r_prev = lstm_prev
+
+    W_l = vs.add_param("%s/W_l" % name, (hidden_dim, hidden_dim * 5),
+                       initializer=initializer)
+    W_r = vs.add_param("%s/W_r" % name, (hidden_dim, hidden_dim * 5),
+                       initializer=initializer)
+    if external_state_dim > 0:
+        W_ext = vs.add_param("%s/W_ext", (external_state_dim, hidden_dim * 5),
+                             initializer=initializer)
     b = vs.add_param("%s/b" % name, (hidden_dim * 5,),
                      initializer=TreeLSTMBiasInitializer())
 
[…]

### 2016-03-01 writing/hard_stack_paper/paper.tex
Very rough paper draft.

@@
 \author{
-Samuel R.\ Bowman$^{\ast\dag}$ \\
-\texttt{sbowman@stanford.edu} \\
+Samuel R.\ Bowman$^{1,2,5,}$\thanks{The first two authors contributed equally.} \\
+\texttt{\small sbowman@stanford.edu} \\
 \And
-Jon Gauthier$^{\dag\ddag}$ \\
-\texttt{jgauthie@stanford.edu} \\
+Jon Gauthier$^{2,3,5,*}$ \\
+\texttt{\small jgauthie@stanford.edu} \\
 \And
-Abhinav Rastogi$^{\S\mathparagraph}$ \\
-\texttt{arastogi@stanford.edu} \\
+Abhinav Rastogi$^{4,5}$ \\
+\texttt{\small arastogi@stanford.edu} \\
 \AND
-Raghav Gupta$^{\parallel}$ \\
-\texttt{rgupta93@stanford.edu} \\
+Raghav Gupta$^{6}$ \\
+\texttt{\small rgupta93@stanford.edu} \\
 \And
-Christopher D.\ Manning$^{\ast\dag\parallel}$\\
-\texttt{manning@stanford.edu}\\
+Christopher D.\ Manning$^{1,2,5,6}$\\
+\texttt{\small manning@stanford.edu}\\
 \And
-Christopher Potts$^{\ast}$\\
-\texttt{cgpotts@stanford.edu}
+Christopher Potts$^{1}$\\
+\texttt{\small cgpotts@stanford.edu}
 \AND\\[-3ex]
-{$^{\ast}$Stanford Linguistics\quad
-$^{\dag}$Stanford NLP Group\quad
-$^{\ddag}$Stanford Symbolic Systems}\\
-{$^{\S}$Stanford Electrical Engineering\quad
-$^{\mathparagraph}$Stanford AI Lab\quad
-$^{\parallel}$Stanford Computer Science}
+{$^{1}$Stanford Linguistics\quad
+$^{2}$Stanford NLP Group\quad
+$^{3}$Stanford Symbolic Systems}\\
+{$^{4}$Stanford Electrical Engineering\quad
+$^{5}$Stanford AI Lab\quad
+$^{6}$Stanford Computer Science}
 }
@@
 
+\begin{figure}[t]
+\begin{subfigure}[t]{0.9\columnwidth}
+\vspace{14em}
+\todo{[SB] Figure showing why trees can't be batched.}
+
+\caption{\label{fig:batching:bad}Trees are hard to batch.}
+\end{subfigure}
+
+\vspace{2em}
+
+\begin{subfigure}[t]{0.9\columnwidth}
+\vspace{14em}
+\todo{[SB] Figure showing why trees can't be batched.}
+
+\caption{\label{fig:batching:good}Sequences are easy to batch.}
+\end{subfigure}
+\end{figure}
+
 \begin{itemize}
+
 \item Brief history of TreeRNNs
+\begin{itemize}
+\item Lx theory suggests should be better than comparable sequence models via compositionality \cite{Partee84,Janssen97}.
+\item In practice, results mixed but positive.
+\citealt{tai2015improved} shows dramatic gains on sentiment analysis using tree-structured models over sequence models.
+\citealt{li2015tree} shows mixed results.
+Both of these evaluations are limited to small datasets with only hundreds or thousands of training examples.
+\item Larger evaluations have been impossible due to issues of computational efficiency. One aim of this paper is to mitigate these issues to make larger-scale experiments possible.
+\item Note, though, that while tree structure can be seen as a bias, simply adding more data within reason may not be enough to get a sequence model to learn without that bias \citealt{bowman2015trees}.
+\end{itemize}
+
 \item Speed issues
+\begin{itemize}
+\item Efficient batched computation is a crucial enabling technology in allowing NNs to be used on large datasets.
+Since NNs require large datasets to be able to model complex problems like language understanding, this is crucial in NLP.
+\item Tree structured neural networks can't be batched.
+As the schematic in Figure~\ref{fig:batching:bad} indicates, different sentences are assigned different trees, and there is no straightforward way of aligning these trees to synchronize computation in the conventional style of implementation.
+\item Our proposed model builds on ideas from transition-based parsing to build a model which implements the same computations as a tree-structured model, but which can be efficiently batched.
+See Figure~\ref{fig:batching:good}.
+
+\end{itemize}
+
 \item Attention issues
+\begin{itemize}
+\item The technique of soft attention has proven to be exceptionally effective as a way of training neural network models for tasks like translation \cite{bahdanau2014neural,luong-pham-manning:2015:EMNLP} and natural language inference \cite{rocktaschel2015reasoning}.
+\item This paper proposes two complementary extensions of the idea of attention to tree-structured models:
+\begin{itemize}
+\item Constituent-by-constituent attention, in which each node in the tree of one sentence is given a soft alignment to some node or nodes in a second sentence.
+\item The mTreeLSTM, which uses a third tree structured model to accumulate information about the alignment between two sentences into a single feature vector.
+This builds on the notion of a matching LSTM or mLSTM from \citealt{DBLP:journals/corr/WangJ15b}.
+\end{itemize}
+\end{itemize}
 \end{itemize}
@@
 
+Our model is based on a formalism adapted from transition-based parsing (\todo{What's a good classic citation for this?}).
+
+\todo{[SB, JG] Brief into to transition-based parsing.}
+
 \subsection{The model}
 
-\todo{[SB] Reframe in terms of Model 0 alone.}
+Our model, which is shown in Figure~\ref{fig:model:0}, takes the form of a transition-based parser, but it is designed to produce sentence representations given parse structures, rather than producing parse structures as its output, leading to the following differences:
+\begin{itemize}
+\item The model takes a sequence of transitions (\shift s or \reduce s) as input, rather than incorporating a decision function which chooses which transition to perform at each step.
+\item The intermediate representations in the parser, including the representations of the tree structures on the stack, are vectors of a fixed length.
+\item The \reduce~operation, which combines two trees or tree nodes on the stack into a single larger tree, is represented using a parametric neural network layer function (in particular, a TreeLSTM layer, after \citealt{tai2015improved}).
+\item The intermediate states of the stack and buffer during parsing are stored during training.
+This makes it possible to learn the parameters of the composition function and the input word representations that are fed into the buffer using backpropagation on an objective function that depends on the final sentence representation that emerges at the head of the stack.
+\item A novel tracking LSTM component is added to improve the model's ability to efficiently track the context in which each step is being performed.
+(\todo{Is this similar enough to anything from Dyer that we should add a citation here?})
+\end{itemize}
 
-\subsection{Model 0}
+\input{../model0_fig.tex}
 
-Model 0, depicted in Figure~\ref{fig:model:0}, is the simplest instantiation of our design, using only a conventional stack and a learned composition function to incrementally build a sentence representation over a series of timesteps.
-For a sentence of $N$ words its input is a sequence of $2N-3$ inputs.
-These inputs can take two types.
-At some timesteps, the input will be a word embedding vector.
-This triggers the \shift~operation, in which the vector is pushed onto the top of a stack.
-At other timesteps, the input will be the special \reduce~token, which triggers the reduction operation.
-In that operation, the top two word vectors are popped from the stack, fed into a learned composition function that maps them to a single vector (in the simplest case, this is a single neural network layer), and then pushed back onto the stack.
-
+\paragraph{Equivalence with existing models}
 If we add no additional features, this model computes the same function as a plain TreeRNN.
@@
 This simply requires ensuring that the graph is run for enough timesteps to finish all of the sentences.
-This involves some wasted computation, since the composition function will be run $2N-3$ times (with the output of composition at non-\reduce~steps discarded), rather than $N-1$ times in a TreeRNN.
+This involves some wasted computation, since the composition function will be run $2N-1$ times (with the output of composition at non-\reduce~steps discarded), rather than $N-1$ times in a TreeRNN.
 However, this loss can be dramatically offset by the gains of batching, which stem from the ability to exploit highly optimized CPU and GPU libraries for batched matrix multiiplication.
+\todo{Update.}
 
-\subsection{Model 1}
+\paragraph{The buffer and word representations}
 
-\input{../model1_fig.tex}
+We draw our word representations from the standard 300-dimensional vector package provided with GloVe \cite{pennington2014glove}.
+We do not update these vectors during training.
+Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors that can be used as inputs to a TreeLSTM composition function:
 
-Model 1, depicted in Figures~\ref{fig:model:1d} and \ref{fig:model:1b}, adapts Model 0 to use a stack and a buffer, making it more closely resemble a shift--reduce parser, and laying the groundwork for a model which can parse novel sentences at test time.
+\begin{equation}
+\colvec{2}
+    {\vec{h}}
+    {\vec{c}}
+= W_{\text{wd}} \vec{x}
+\end{equation}
 
-The model runs for a fixed number of transition steps: $2N - 3$.
-In its starting configuration, it contains a stack that is prepopulated with the first two words of the sentence (since \shift~\shift~is the only legal operation sequence for the first two timesteps of a true shift-reduce parser), as well as a buffer (a queue) prepopulated with all of the remaining words in the sentence.
-Both the stack and buffer represent words using their embeddings.
+\paragraph{The representations in the stack}
 
-At each timestep at test time, the model combines views of the stack and buffer (the top element of the buffer and the top two elements of the stack, highlighted in yellow) as the input to a tracking LSTM (red).
-This LSTM's output is fed into a sigmoid operation classifier (blue) which chooses between the \shift~and \reduce~operations.
-If \shift~is chosen, one word embedding is popped from the buffer and pushed onto the stack.
-If \reduce~is chosen, the buffer is left as is, and the top two elements of the stack are popped and composed using a learned composition function (green), with the result placed back on top of the stack.
+In our model, each row of the stack contains a pair of vectors $(h, c)$, which together represent some node in the parse tree of the sentence.
+Since the \shift~operation simply copies entries from the buffer onto the stack, this means that the word representations in the buffer also have two parts, $h$ and $c$.
 
-\paragraph{Supervision} The model is trained using two objective functions simultaneously.
-The semantic objective function is computed by feeding the value from the top of the stack at the final timestep---the full sentence encoding---into a downstream neural network model for some semantic task, like a sentiment classifier or an entailment classifier.
-The gradients from that classifier propagate to every part of the model except the operation classifier (blue).
-The syntactic objective function takes the form of direct supervision on the operation classifier (blue) which encourages that classifier to produce the same sequence of operations that an existing parser would produce for that sentence.
-The gradients from the syntactic objective function propagate to every part of the model but the downstream semantic model.
+\paragraph{The composition function}
 
-At training time, following the strategy used in LSTM text decoders, the decisions made by the operation classifier (blue) is discarded, and the model instead uses the correct operation as specified in the (already parsed) training corpus.
-At test time, this signal is not available, and the model uses its own predicted operations.
+From \citealt{tai2015improved}.
 
-\subsection{Model 2}
+The inputs to the composition function are the representations of the two words or phrases closest to the head of the, $(\vec{h}^1_s, \vec{c}^1_s)$ and $(\vec{h}^2_s, \vec{c}^2_s)$, which are treated as the two children of the tree node whose representation the composition function is constructing.
+Each vector is of the same dimension $D$.
 
-Model 2 makes a small change to Model 1 that is likely to substantially change the dynamics of learning: It uses the operation sequence predicted by the operation classifier (blue) at training time as well as at test time.
-It may be possible to accelerate Model 2 training by initializing it with parameters learned by Model 1.
+\begin{gather}
+\colvec{4}
+    {\vec{i}}
+    {\vec{f}_l}
+    {\vec{f}_r}
+    {\vec{o}}
+= \sigma(
+W_{\text{iffo}}
+\colvec{3}
+    {\vec{h}_s^1}
+    {\vec{h}_s^2}
+    {\vec{e}}
+)
+\\
+\vec{g}
+= \tanh(
+W_{g}
+\colvec{3}
+    {\vec{h}_s^1}
+    {\vec{h}_s^2}
+    {\vec{e}}
+)
+\\
+\vec{c} = \vec{f}_l * \vec{c}_s^2 + \vec{f}_r * \vec{c}_s^1 + \vec{i} * \vec{g}  
+\\
+\vec{h} = \vec{o} * \vec{c}
+\end{gather}
 
-By exposing Model 2 to the results of its own decisions during training, we encourage it to become more robust to its own prediction errors.
-\citealt{bengio2015scheduled} applied a similar strategy\footnote{The authors experiment with several strategies which interpolate between oracle-driven training and oracle-free training (Models 1 and 2 in our presentation, respectively).
-It may be useful to adopt a similar interpolating approach.} to an image captioning model.
-They suggest that the resulting model can avoid propagating prediction errors through long sequences due to this training regime.
+$\vec{e}$ is an optional input argument which is either the empty vector $[]$ or a vector from an external source like the tracking LSTM (see below).
+$*$ denotes the elementwise product.
+Each vector-valued variable listed here is of dimension $D$.
+The values $(h, c)$ represent the local subtree, and both will be placed on the stack during a \reduce~operation.
 
-\subsection{Tracking left context with an auxiliary LSTM} Includes both the tracking LSTM and context-sensitive composition.
+\paragraph{The classifier}
 
-\subsection{Building a model}
+This paper presents the transition-based sentence model within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task.
+To classify a sentence pair, a feature vector is first constructed.
+This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition.
+In particular, the $\vec{h}$ portions of these representations are used.
+The final feature vector consists of the concatenation of these two vectors, $\vec{x}_{\text{premise}}$ and $\vec{x}_{\text{hypothesis}}$, their difference, and their elementwise product:
+
+\begin{equation}
+\vec{x}_{classifier} = 
+\colvec{4}
+    {\vec{h}_{\text{premise}}}
+    {\vec{h}_{\text{hypothesis}}}
+    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
+    {\vec{h}_{\text{premise}} * \vec{h}_{\text{hypothesis}}}
+\end{equation}
+
+This feature vector is then passed to a series of two (\todo{Update}) ReLU neural network layers, then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+
+\subsection{Tracking left context with an auxiliary LSTM}
+
+The tracking LSTM captures left context. \todo{Why?}
+
+The tracking LSTM is a standard LSTM RNN.
+At each step it takes three vectors as inputs: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the two two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
+Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step.
+
+\subsection{Implementing the transition-based sentence model}
 
@@
 
-\paragraph{Handling overly long sentences}
-
-\subsection{Data preparation}
+\paragraph{Converting parses to transition sequences}
 
@@
 
+\paragraph{Handling overly long sentences}
+
+To efficiently implement batched computation, we must stipulate a fixed number of transitions (50) that the model can perform before producing an output.
+
+\todo{[SB] Why cropping is sensible if done well.}\\
+
 \paragraph{Representing the stack in memory} Or, thin stack. 
@@
 
+We used random search to tune \todo{[SB] List tuned hyperparameters.}
+
+We trained each model for 250k steps, using a minibatch size of 64 for each step.
+We tracked each model's performance on the development set during training and saving parameters when this performance reached a new peak.
+We used early stopping, evaluating on the test set using the parameters that performed best on the development set.
+
 \paragraph{Software infrastructure} We will make Theano code available.
 
-\subsection{Step time}
+\subsection{Related work: Transition-based parsing and neural networks.}
 
-\todo{Does anyone know of a better baseline TreeRNN implementation that we can use on Jagupard?
-We can use the CoreNLP SST model, but using a Java model as a baseline seems worrying unless we're guaranteed that it's competitively fast.}
-
-Comparing model step time to the plain RNN of \citealt{li2015tree}. We use the small \citealt{pang2004sentimental} sentiment corpus that they use, and train with identical hyperparameters: ...
-
-Evaluation metrics: Time per minibatch on a jag machine, with and without GPU access.
+\begin{itemize}
+\item \cite{chen2014}
+\item \cite{dyer-EtAl:2015:ACL-IJCNLP}, \cite{dyer2016rnn}
+\end{itemize}
 
@@
 
-\paragraph{The Stanford Natural Language Inference task} \cite{snli:emnlp2015}
+\paragraph{The Stanford Natural Language Inference task} 
+We evaluate our model on the Stanford Natural Language Inference task \cite{snli:emnlp2015}.
+
+550k/10k/10k split
+
+We skip unlabeled examples.
+
+\todo{[SB] Pull out an example from SNLI.}
+
+We evaluate five models on the SNLI task. Each uses 150d hidden states: 
+\begin{itemize}
+\item A baseline LSTM model (similar to that of \citealt{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM RNN sequence model.
+\item The unaugmented transition-based sentence model.
+\item The transition-based sentence model with the tracking LSTM.
+\item The transition-based sentence model with the tracking LSTM and sequence-based attention over premise states.
+\item The transition-based sentence model with the tracking LSTM and tree-based attention over premise states.
+\end{itemize}
+
+\begin{table*}[t]
+  \centering\small
+  \begin{tabular}{lcccc} 
+    \toprule
+Model                   & Trained params.    &   Train  &   Test \\
+\midrule
+\multicolumn{4}{c}{\bf Non-NN results}\\
+\midrule
+Lexicalized classifier \cite{snli:emnlp2015}
+                        & --                 &   99.7   &   78.2      \\
+\midrule
+\multicolumn{4}{c}{\bf Previous NN results without attention}\\
+\midrule
+100d LSTM encoders \cite{snli:emnlp2015}
+                        & 221k               &   84.8   &   77.6      \\
+1024d pretrained GRU encoders \cite{DBLP:journals/corr/VendrovKFU15}
+                        & 15m                &   98.8   &   81.4       \\
+\todo{[SB] Import baselines from page}  
+                        & ?                  &   ?      &   81?       \\
+\todo{[SB] Import baselines from page}  
+                        & ?                  &   ?      &   81?       \\
+\midrule
+\multicolumn{4}{c}{\bf Previous NN results with attention}\\
+\midrule
+\todo{[SB] Import baselines from page}  
+                        & ?                  &   ?      &   81?       \\
+\todo{[SB] Import baselines from page}  
+                        & ?                  &   ?      &   81?       \\
+\todo{[SB] Import baselines from page}  
+                        & ?                  &   ?      &   81?       \\
+\midrule
+\multicolumn{4}{c}{\bf New results}\\
+\midrule150d LSTM RNN           & ?                  &   ?      &   81?       \\
+150d TB TreeLSTM    
+                        & ?                  &   ?      &   82?       \\
+150d TB TreeLSTM w/ tracking    
+                        & ?                  &   ?      &   82?       \\
+150d TB TreeLSTM w/ tracking, sequence-based attn.        
+                        & ?                  &   ?      &   82?       \\
+150d TB TreeLSTM w/ tracking, tree-based attn.            
+                        & ?                  &   ?      &   \textbf{82?}\\
+    \bottomrule
+  \end{tabular}
+  \protect\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification.
+All reported figures are percent accuracy.
+\todo{Add seconds-per-step *if* we get a reasonably well-optimized model in time.}}
+\end{table*}
 
 \paragraph{Results}
+
+Table~\ref{tab:results} shows our results.
+
+\subsection{Step time}
+
+\todo{[JG,SB] Does anyone know of a better baseline TreeRNN implementation that we can use on Jagupard?
+We can use the CoreNLP SST model, but using a Java model as a baseline seems worrying unless we're guaranteed that it's competitively fast.}
+
+Comparing model step time to the plain RNN of \citealt{li2015tree}. We use the small \citealt{pang2004sentimental} sentiment corpus that they use, and train with identical hyperparameters: ...
+
+Evaluation metrics: Time per minibatch on a jag machine, with and without GPU access.
 
@@
 
+\begin{figure}
+\vspace{14em}
+\caption{\label{fig:induced-alignment}The alignment induced between two sentences by \todo{some model.}. Compare with a similar alignment presented by \protect\citealt{DBLP:journals/corr/WangJ15b}.}
+\end{figure}
+
 \section{Conclusions and future work}
@@
 \begin{itemize}
+\item It is possible to train a tree-structured model, without approximation, on large-scale language learning tasks.
+\item Tree structure helps in large-scale language learning tasks. Not massively, but measurably.
+\item The advantages of attention carry over to tree-structured models.
 \item Future work: Incorporate the parser

### 2016-03-01 rembed/cuda_util.py
Remove unnecessary debug logging from cuda_util

@@ -1155,11 +1155,9 @@ def local_gpua_row_switch(node):
     row switch.
     """
 
-    print "-------------", node, node.op.scalar_op.__class__
     if (node.op.__class__ == T.Elemwise
         and node.op.scalar_op.__class__ != theano.scalar.Switch):
         return False
-    print "\tstill here!"
 
     cond, ift, iff = node.inputs
     out, = node.outputs
@@ -1167,7 +1165,6 @@ def local_gpua_row_switch(node):
     # Only applies to Switch instances where a vector mask broadcasts over
     # matrices.
     bcast = cond.broadcastable
-    print "\t", bcast
     if not bcast or not (not bcast[0] and all(bcast[1:])
                          and ift.ndim in [2, 3]):
         return False

### 2016-03-01 rembed/cuda_util.py
Remove an extraneous print

@@ -1420,7 +1420,6 @@ def local_gpu_masked_careduce(node):
     # The graph looks okay. Check the dims.
     if node.op.reduce_mask != (1, 0, 0):
         return False
-    print "\t", node.op.pre_scalar_op
     if node.op.pre_scalar_op:
         return False
 

### 2016-03-01 writing/hard_stack_paper/paper.tex
Results table fleshing out.

@@
 \midrule
-\multicolumn{4}{c}{\bf Non-NN results}\\
+\multicolumn{4}{c}{Previous non-NN results}\\
 \midrule
@@
 \midrule
-\multicolumn{4}{c}{\bf Previous NN results without attention}\\
+\multicolumn{4}{c}{Previous NN results without attention}\\
 \midrule
@@
                         & 15m                &   98.8   &   81.4       \\
-\todo{[SB] Import baselines from page}  
-                        & ?                  &   ?      &   81?       \\
+300d Tree-based CNN encoders \cite{mou2015recognizing}
+                        & 3.5m               &   83.4   &   82.1       \\
 \todo{[SB] Import baselines from page}  
@@
 \midrule
-\multicolumn{4}{c}{\bf Previous NN results with attention}\\
+\multicolumn{4}{c}{Previous NN results with attention}\\
 \midrule
-\todo{[SB] Import baselines from page}  
-                        & ?                  &   ?      &   81?       \\
-\todo{[SB] Import baselines from page}  
-                        & ?                  &   ?      &   81?       \\
-\todo{[SB] Import baselines from page}  
-                        & ?                  &   ?      &   81?       \\
+100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
+                        & 252k               &   85.3   &   83.5       \\
+300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
+                        & 1.9m               &   92.0   &   86.1      \\
+300d LSTMN with deep attention fusion \cite{cheng2016long} (1.4m params) 92.3    89.1
+
+                        & 1.4m               &   92.3   &   89.1      \\
 \midrule
-\multicolumn{4}{c}{\bf New results}\\
+\multicolumn{4}{c}{New results}\\
 \midrule150d LSTM RNN           & ?                  &   ?      &   81?       \\

### 2016-03-01 rembed/models/classifier.py
Tentative SNLI gradient bugfix.

@@ -650,7 +650,10 @@ def run(only_forward=False):
             gradients = premise_model.gradients
             hypothesis_gradients = hypothesis_model.gradients
             for key in hypothesis_gradients:
-                gradients[key] += hypothesis_gradients[key]
+                if key in gradients:
+                    gradients[key] += hypothesis_gradients[key]
+                else:
+                    gradients[key] = hypothesis_gradients[key]
 
             new_values = util.merge_updates(
                 premise_model.scan_updates + premise_model.bscan_updates,

### 2016-03-01 rembed/models/classifier.py
Tentative SNLI gradient bugfix.

@@ -649,7 +649,10 @@ def run(only_forward=False):
             gradients = premise_model.gradients
             hypothesis_gradients = hypothesis_model.gradients
             for key in hypothesis_gradients:
-                gradients[key] += hypothesis_gradients[key]
+                if key in gradients:
+                    gradients[key] += hypothesis_gradients[key]
+                else:
+                    gradients[key] = hypothesis_gradients[key]
 
             new_values = util.merge_updates(
                 premise_model.scan_updates + premise_model.bscan_updates,

### 2016-03-01 train_rembed_fat_classifier.sh
One training script per model binary.

@@ -19,4 +19,4 @@ source /u/nlp/packages/anaconda/bin/activate conda-common
 export LD_LIBRARY_PATH=/usr/local/cuda/lib64:$LD_LIBRARY_PATH
 export PYTHONPATH=/scr/jgauthie/tmp/theano-nshrdlu:$PYTHONPATH
 
-THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=$DEVICE,floatX=float32 python -m rembed.models.classifier $REMBED_FLAGS
+THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=$DEVICE,floatX=float32 python -m rembed.models.fat_classifier $REMBED_FLAGS

### 2016-03-02 writing/hard_stack_paper/paper.tex
More comparison, comments on padding/cropping.

@@
 \item A novel tracking LSTM component is added to improve the model's ability to efficiently track the context in which each step is being performed.
-(\todo{Is this similar enough to anything from Dyer that we should add a citation here?})
 \end{itemize}
@@
 
-\paragraph{Handling overly long sentences}
+\paragraph{Handling variable sentence lengths}
 
 To efficiently implement batched computation, we must stipulate a fixed number of transitions (50) that the model can perform before producing an output.
+Sentences whose transition sequences are shorter than this length are padded, and sentences whose transition sequences are longer than this length are cropped.
 
-\todo{[SB] Why cropping is sensible if done well.}\\
+Padding, if done properly, should not impact the output of the model significantly:\footnote{The tracking LSTM can cause a slight dependence of the final representation on unrolling length.
+Because of this, we always train and test the model with a fixed unrolling length.} a model with the fixed parameters should produce the same representation for a 10-word sentence whether it is unrolled for 19 transitions (the minimum to avoid cropping) or 100.
+
+Cropping necessarily discards information, but if done properly, it is nonetheless possible to learn good representations for cropped sentences.
+When a transition sequence is cropped, the number of removed \shift~transitions is tracked, and an equal number of word tokens is removed from the tail (left) end of the buffer.
+This makes it possible for a transition sequence to begin with a \reduce~transition, or to otherwise use a \reduce~transition on a stack that does not have two elements to reduce.
+In this instance, we simply feed the composition function one or more zero vectors, which represent nonexistent stack elements.
+If the composition learns to interpret these zero vectors properly, they can be taken to be unknown or incomplete nodes in an otherwise complete tree that is constructed according to the intended parse structure.
+\todo{[SB] Make a figure showing the results of cropping if there's room.}
 
@@
 
-\begin{itemize}
-\item \cite{chen2014}
-\item \cite{dyer-EtAl:2015:ACL-IJCNLP}, \cite{dyer2016rnn}
-\end{itemize}
+There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing \cite{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP}.
+In addition, there has been recent work \cite{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
+To our knowledge, our model is the first to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
 
@@
 \item The advantages of attention carry over to tree-structured models.
+\item Future work: Encode the stack, after \citealt{dyer-EtAl:2015:ACL-IJCNLP}
 \item Future work: Incorporate the parser

### 2016-03-02 rembed/models/fat_classifier.py
Initialize both c and h states in tracking unit
Both h and c states in the tracking unit in hypothesis model are initilialized from the premise model.

@@ -190,7 +190,7 @@ def build_sentence_pair_model(cls, vocab_size, seq_length, tokens, transitions,
         initialize_hyp_tracking_state=FLAGS.initialize_hyp_tracking_state)
 
     premise_stack_tops = premise_model.stack_tops if FLAGS.use_attention != "None" else None
-    premise_tracking_c_state_final = premise_model.tracking_c_state_final
+    premise_tracking_state_final = premise_model.tracking_state_final
     hypothesis_model = cls(
         FLAGS.model_dim, FLAGS.word_embedding_dim, vocab_size, seq_length,
         compose_network, embedding_projection_network, training_mode, ground_truth_transitions_visible, vs,
@@ -209,7 +209,7 @@ def build_sentence_pair_model(cls, vocab_size, seq_length, tokens, transitions,
         premise_stack_tops=premise_stack_tops,
         is_hypothesis=True,
         initialize_hyp_tracking_state=FLAGS.initialize_hyp_tracking_state,
-        premise_tracking_c_state_final=premise_tracking_c_state_final)
+        premise_tracking_state_final=premise_tracking_state_final)
 
     # Extract top element of final stack timestep.
     if FLAGS.use_attention == "None" or FLAGS.use_difference_feature or FLAGS.use_product_feature:
@@ -744,8 +744,8 @@ if __name__ == '__main__':
     gflags.DEFINE_boolean("connect_tracking_comp", True,
         "Connect tracking unit and composition unit. Can only be true if using LSTM in both units.")
     gflags.DEFINE_boolean("initialize_hyp_tracking_state", False,
-        "Initialize the c state of the tracking unit of hypothesis model with the final"
-        "tracking unit c state of the premise model.")
+        "Initialize the h and c state of the tracking unit of hypothesis model with the final"
+        "tracking unit states of the premise model.")
 
     # Optimization settings.
     gflags.DEFINE_integer("training_steps", 500000, "Stop training after this point.")

### 2016-03-02 rembed/fat_stack.py
Initialize both c and h states in tracking unit
Both h and c states in the tracking unit in hypothesis model are initilialized from the premise model.

@@ -87,7 +87,7 @@ class HardStack(object):
                  attention_unit=None,
                  is_hypothesis=False,
                  initialize_hyp_tracking_state=False,
-                 premise_tracking_c_state_final=None):
+                 premise_tracking_state_final=None):
         """
         Construct a HardStack.
 
@@ -146,7 +146,7 @@ class HardStack(object):
             is_hypothesis: Whether we're processing the premise or the hypothesis (for SNLI)
             initialize_hyp_tracking_state: Initialize the c state of the tracking unit of hypothesis
                 model with the final tracking unit c state of the premise model.
-            premise_tracking_c_state_final: The final c state of the tracking unit in premise model.
+            premise_tracking_state_final: The final h and c state of the tracking unit in premise model.
         """
 
         self.model_dim = model_dim
@@ -205,9 +205,9 @@ class HardStack(object):
         else:
             self._attention_unit = None
         self.initialize_hyp_tracking_state = initialize_hyp_tracking_state
-        self.premise_tracking_c_state_final = premise_tracking_c_state_final
+        self.premise_tracking_state_final = premise_tracking_state_final
         if initialize_hyp_tracking_state:
-            assert not is_hypothesis or premise_tracking_c_state_final is not None, \
+            assert not is_hypothesis or premise_tracking_state_final is not None, \
                 "Must supply initial c states in hypothesis model" 
 
         # Check whether we're processing the hypothesis or the premise
@@ -396,9 +396,8 @@ class HardStack(object):
         # Initialize the hidden state for the tracking LSTM, if needed.
         if self.use_tracking_lstm:
             if self.initialize_hyp_tracking_state and self.is_hypothesis:
-                # Initialize the c state of tracking unit from the c states of premise model.
-                h_state_init = T.zeros((batch_size, self.tracking_lstm_hidden_dim))
-                hidden_init = T.concatenate([h_state_init, self.premise_tracking_c_state_final], axis=1)
+                # Initialize the h and c states of tracking unit from the premise model.
+                hidden_init = self.premise_tracking_state_final
             else:
                 hidden_init = T.zeros((batch_size, self.tracking_lstm_hidden_dim * 2))
         else:
@@ -478,10 +477,10 @@ class HardStack(object):
                 self.final_weighed_representation = scan_ret[0][stack_ind][-1][:,0,2*h_dim:3*h_dim]
 
[…]

### 2016-03-02 rembed/models/fat_classifier.py

@@ -189,7 +189,7 @@ def build_sentence_pair_model(cls, vocab_size, seq_length, tokens, transitions,
         initialize_hyp_tracking_state=FLAGS.initialize_hyp_tracking_state)
 
     premise_stack_tops = premise_model.stack_tops if FLAGS.use_attention != "None" else None
-    premise_tracking_c_state_final = premise_model.tracking_c_state_final
+    premise_tracking_state_final = premise_model.tracking_state_final
     hypothesis_model = cls(
         FLAGS.model_dim, FLAGS.word_embedding_dim, vocab_size, seq_length,
         compose_network, embedding_projection_network, training_mode, ground_truth_transitions_visible, vs,
@@ -208,7 +208,7 @@ def build_sentence_pair_model(cls, vocab_size, seq_length, tokens, transitions,
         premise_stack_tops=premise_stack_tops,
         is_hypothesis=True,
         initialize_hyp_tracking_state=FLAGS.initialize_hyp_tracking_state,
-        premise_tracking_c_state_final=premise_tracking_c_state_final)
+        premise_tracking_state_final=premise_tracking_state_final)
 
     # Extract top element of final stack timestep.
     if FLAGS.use_attention == "None" or FLAGS.use_difference_feature or FLAGS.use_product_feature:
@@ -727,8 +727,8 @@ if __name__ == '__main__':
     gflags.DEFINE_boolean("connect_tracking_comp", True,
         "Connect tracking unit and composition unit. Can only be true if using LSTM in both units.")
     gflags.DEFINE_boolean("initialize_hyp_tracking_state", False,
-        "Initialize the c state of the tracking unit of hypothesis model with the final"
-        "tracking unit c state of the premise model.")
+        "Initialize the h and c state of the tracking unit of hypothesis model with the final"
+        "tracking unit states of the premise model.")
 
     # Optimization settings.
     gflags.DEFINE_integer("training_steps", 500000, "Stop training after this point.")

### 2016-03-02 rembed/stack.py
Tentative gradient fix for thin_stack.

@@ -475,8 +475,7 @@ class ThinStack(object):
         input_ndim += [len(extra_output_shape) + 1
                        for extra_output_shape in self.recurrence.extra_outputs]
 
-        wrt = [self._vs.vars[key] for key in self._vars
-               if self._vs.vars[key] != self.embeddings]
+        wrt = self._vs.trainable_vars
 
         # TODO handle gradient of action prediction
         # TODO would it pay off to force these to have the same concrete

### 2016-03-02 rembed/stack.py
Fix, take 2.

@@ -475,7 +475,8 @@ class ThinStack(object):
         input_ndim += [len(extra_output_shape) + 1
                        for extra_output_shape in self.recurrence.extra_outputs]
 
-        wrt = self._vs.trainable_vars
+        wrt = wrt = [self._vs.vars[key] for key in self._vars
+               if key in self._vs.trainable_vars]
 
         # TODO handle gradient of action prediction
         # TODO would it pay off to force these to have the same concrete

### 2016-03-02 writing/hard_stack_paper/paper.tex
Paper update.

@@
 \vspace{14em}
-\todo{[SB] Figure showing why trees can't be batched.}
+\todo{[SB] Figure showing why trees can't be batched (modeled after slides).}
 
@@
 \vspace{14em}
-\todo{[SB] Figure showing why trees can't be batched.}
+\todo{[SB] Figure showing a batch of sequences (modeled after slides).}
 
@@
 \todo{[SB, JG] Brief into to transition-based parsing.}
+
+\vspace{10em}
 
@@
 
+\vspace{8em}
+
 \paragraph{Optimization and hyperparameters}
 
-We used random search to tune \todo{[SB] List tuned hyperparameters.}
+We use the RMSProp optimizer (\todo{Cite.}) with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
+We apply both dropout \cite{srivastava2014dropout} and batch normalization \cite{2015SIoffeCSzegedy} to the word embeddings in the buffer (after the linear projection is applied) and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
+In addition, we apply an L2 regularization penalty to all parameters.
 
-We trained each model for 250k steps, using a minibatch size of 64 for each step.
+We used random search to tune many of the hyperparameters, including the learning rate, the dropout rate, the L2 regularization weight, the hidden dimension of the tracking LSTM and the number of layers in the MLP \todo{[SB] Report the ranges searched for each parameter and the number of runs for each results.}
+
+We trained each model for 250k steps in each run, using a minibatch size of 64 for each step.
 We tracked each model's performance on the development set during training and saving parameters when this performance reached a new peak.
@@
                         & 3.5m               &   83.4   &   82.1       \\
-\todo{[SB] Import baselines from page}  
-                        & ?                  &   ?      &   81?       \\
 \midrule

### 2016-03-03 writing/hard_stack_paper/paper.tex
Filling out more of the paper.

@@
 \makeatother
-\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black, draft]{hyperref}
+\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black]{hyperref}
 
@@
 \item We present the transition-based TreeRNN implementation -- makes efficient training possible, show that it helps relative to seq models.
-\item Further, attention has proven powerful for extending sequence models. We introduce constit-by-constit attention, and find it to be effective here, too.
+\item Further, attention has proven powerful for extending sequence models. We introduce constituent-by-constituent attention, and find it to be effective here, too.
 \end{itemize}
@@
 \end{itemize}
+
+\item Outline of the paper
+\begin{itemize}
+\item Propose a model
+\item Show how the model can be used for NLI
+\item Show how attention can be used on the NLI model
+\item Experiments and discussion
+\end{itemize}
+
 \end{itemize}
@@
 
+\todo{[SB] Finish this subsection.}
+
 From \citealt{tai2015improved}.
+
+When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\ii composition function}, which is a neural network function that produces
 
@@
 
-\paragraph{The classifier}
+\paragraph{Creating a sentence pair classifier}
 
@@
 
-\paragraph{The Stanford Natural Language Inference task} 
-We evaluate our model on the Stanford Natural Language Inference task \cite{snli:emnlp2015}.
+\subsection{Natural language inference and SNLI}
 
-550k/10k/10k split
+We evaluate our model on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as here:
 
-We skip unlabeled examples.
+\begin{quote}
+Premise: {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}
 
-\todo{[SB] Pull out an example from SNLI.}
+Hypothesis: {\it A girl outside plays in the snow.}
+
+Correct label: {\it entailment}
+\end{quote}
+
+Even though NLI is framed as a relatively simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
+In order for a model to reliably perform well on NLI across a range of sentence types and text genres, it must be able to represent and reason with all of the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
+
+In particular, we use the Stanford NLI corpus (SNLI, \citealt{snli:emnlp2015}).
+SNLI is a corpus of 570k human-labeled pairs of scene descriptions like the one above.
+We use the standard train--test split and ignore unlabeled examples, which leaves about 549k examples for training, 9,842 for development, and 9,824 for testing.
+SNLI labels are roughly balanced, with the most frequent label, {\it entailment}, making up 34.2\% of the test set.
+
+\subsection{Models evaluated}
 
@@
 \multicolumn{4}{c}{New results}\\
-\midrule150d LSTM RNN           & ?                  &   ?      &   81?       \\
-150d TB TreeLSTM    
-                        & ?                  &   ?      &   82?       \\
-150d TB TreeLSTM w/ tracking    
-                        & ?                  &   ?      &   82?       \\
-150d TB TreeLSTM w/ tracking, sequence-based attn.        
-                        & ?                  &   ?      &   82?       \\
-150d TB TreeLSTM w/ tracking, tree-based attn.            
-                        & ?                  &   ?      &   \textbf{82?}\\
+\midrule
+\result{150d LSTM RNN      }     & ?                  &   ?      &   ?       \\
+\result{150d TB TreeLSTM}   
+                        & ?                  &   ?      &   ?       \\
+\result{150d TB TreeLSTM w/ tracking    }
+                        & ?                  &   ?      &   ?       \\
+\result{150d TB TreeLSTM w/ tracking, sequence-based attn.   }     
+                        & ?                  &   ?      &   ?       \\
+\result{150d TB TreeLSTM w/ tracking, tree-based attn. }           
+                        & ?                  &   ?      &   \textbf{?}\\
     \bottomrule
@@
 All reported figures are percent accuracy.
-\todo{Add seconds-per-step *if* we get a reasonably well-optimized model in time.}}
+\result{[JG]Add seconds-per-step *if* we get a reasonably well-optimized model in time.}}
 \end{table*}
@@
 Comparing model step time to the plain RNN of \citealt{li2015tree}. We use the small \citealt{pang2004sentimental} sentiment corpus that they use, and train with identical hyperparameters: ...
-
-Evaluation metrics: Time per minibatch on a jag machine, with and without GPU access.
 
@@
 \item The advantages of attention carry over to tree-structured models.
-\item Future work: Encode the stack, after \citealt{dyer-EtAl:2015:ACL-IJCNLP}
-\item Future work: Incorporate the parser
-\item Future work: Latent parses
 \end{itemize}
+
+\paragraph{Future work} The tracking LSTM that we incorporated into our model uses only simple, quick to compute features drawn from the head of the buffer and the head of the stack.
+It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors in the sentence string and its likely position in the final tree, and that that context information could support better sentence encoding.
+One promising way to pursue this goal would be to encode the full contents of the stack and buffer at each time step following the method used by \citealt{dyer-EtAl:2015:ACL-IJCNLP} for parsing.
+The model presented here is only capable of operating over sentences which have already been assigned a binary constituency tree by a parser.
+This is not necessarily a practical limitation, since modern parsers can assign reasonably high quality parses in so little time as to not dramatically impact the performance characteristics of our model \todo{[SB,JG,CM] How fast are fast constit parsers?}.
+However, depending upon an external parser can be a missed opportunity.
+
+In future work, we aim to incorporate a parser into our model, allowing it to simultaneously learn to parse sentences (using a conventional parsing objective function) and to use those parses to learn semantic representations for those sentences (using a classification objective function, as in this work).
+Implementing such a model on top of our proposed stack-based design would learn a model that could productively share representations between the two tasks.
+Implementing such a model on top of a differentiable version of our stack-based design (possibly following \citealt{grefenstette2015learning} \todo{[SB] Cite FAIR stack paper.}) would further make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing  learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 
@@
 
-(some of) the Tesla K40(s) used for this research was/were donated by the NVIDIA Corporation.
+Some of the Tesla K40(s) used for this research was/were donated by the NVIDIA Corporation.
+\todo{[CM,CP] Acknowledge other grants.}
 
@@
 \bibliography{../MLSemantics}
+\todo{[SB] Make bibliography style uniform.}
 

### 2016-03-04 writing/hard_stack_paper/paper.tex
More paper-fleshing.

@@
 \todo{Anonymize.}
+\todo{Name the damn model!}
 
@@
 \end{subfigure}
+\caption{An illustration of a conventionally-implemented TreeRNN model (a) and our equivalent model (b).}
 \end{figure}
@@
 \paragraph{The composition function}
+When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\ii composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
+This new node is then pushed on to the stack.
 
-\todo{[SB] Finish this subsection.}
+Our composition function is based closely on the TreeLSTM of \citealt{tai2015improved}.
+The TreeLSTM generalizes the LSTM neural network layer \cite{hochreiter1997long} to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a two-part vector representation containing an $\vec{h}$ vector and a $\vec{c}$ vector, where the latter is meant to serve as a slower-changing long-term memory.
 
-From \citealt{tai2015improved}.
-
-When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\ii composition function}, which is a neural network function that produces
-
-The inputs to the composition function are the representations of the two words or phrases closest to the head of the, $(\vec{h}^1_s, \vec{c}^1_s)$ and $(\vec{h}^2_s, \vec{c}^2_s)$, which are treated as the two children of the tree node whose representation the composition function is constructing.
-Each vector is of the same dimension $D$.
+The precise formulation that we use is as follows.
 
@@
 
-$\vec{e}$ is an optional input argument which is either the empty vector $[]$ or a vector from an external source like the tracking LSTM (see below).
+The results of this function are the pair $(\vec{h}, \vec{c})$, which are placed back on the stack.
+The two input tree nodes popped off the stack are represented as the pairs $(\vec{h}^1_s, \vec{c}^1_s)$ and $(\vec{h}^2_s, \vec{c}^2_s)$.
+In addition, $\vec{e}$ is an optional input argument which is either the empty vector $[]$ or a vector from an external source like the tracking LSTM (see below).
 $*$ denotes the elementwise product.
-Each vector-valued variable listed here is of dimension $D$.
-The values $(h, c)$ represent the local subtree, and both will be placed on the stack during a \reduce~operation.
+Each vector-valued variable listed here is of dimension $D$, except $\vec{e}$, which is of dimensions $D_t$, which can be chosen independently.
 
@@
 \paragraph{Software infrastructure} We will make Theano code available.
+
+\subsection{TreeRNN-equivalence and efficiency}
+
+In its bare form, without the addition of the tracking LSTM, our model computes the precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
+While it adds no expressive power on its own, our model design improves upon the standard TreeRNN implementation in three ways:
+
+\begin{itemize}
+\item It provides a straightforward strategy for implementing tree-structured composition which allows for batched computation, and provides near-optimal run times on typical computer systems.
+While our design includes some wasted computation induced by the full application of the composition function at every step, this is made up for by avoiding most of the (potentially expensive) scattered memory access that would be needed for batched computation strategies that more closely mimic the standard tree-based architecture.
+
+\item It provides a simple mechanism for integrating the use of left-context into tree-structured models through the tracking LSTM, providing a fast mechanism to help these models handle lexical ambiguity.
+
+\item By designing the model around the standard transition set used in neural network syntactic parsers, it opens the door to tight integration between parsing and sentence interpretation, including the possibility of joint learning, or even of learning a parser trained entirely on a sentence classification or interpretation objective.
+\end{itemize}
 

### 2016-03-04 writing/hard_stack_paper/paper.tex
Paper update.

@@
 
-Our model is based on a formalism adapted from transition-based parsing (\todo{What's a good classic citation for this?}).
+Our model is based on a formalism adapted from transition-based parsing (\question{What's a good classic citation for this?}).
 
@@
 \input{../model0_fig.tex}
-
-\paragraph{Equivalence with existing models}
-If we add no additional features, this model computes the same function as a plain TreeRNN.
-However, we expect it to be substantially faster than conventional TreeRNN implementations.
-Unlike a TreeRNN, the Model 0 computation graph is essentially static across examples, so examples of varying structures and lengths can be batched together and run on the same graph in a single step.
-This simply requires ensuring that the graph is run for enough timesteps to finish all of the sentences.
-This involves some wasted computation, since the composition function will be run $2N-1$ times (with the output of composition at non-\reduce~steps discarded), rather than $N-1$ times in a TreeRNN.
-However, this loss can be dramatically offset by the gains of batching, which stem from the ability to exploit highly optimized CPU and GPU libraries for batched matrix multiiplication.
-\todo{Update.}
 
@@
 
-This feature vector is then passed to a series of two (\todo{Update}) ReLU neural network layers, then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+This feature vector is then passed to a series of two (\todo{Update w/ final number}) ReLU neural network layers, then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
 
@@
 
-The tracking LSTM captures left context. \todo{Why?}
+Lexical ambiguity is omnipresent in natural language.
+Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
+Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word (first incorporates it into a partial sentence representation), the model has access to a state vector that summarizes all of the words to the left of the current word, and that state vector can serve as a crude but effective form of context to help with disambiguation.
+In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is strictly less than the sequence model has access to, and is often just a single additional word.
+
+In addition to our basic model, we additionally propose an extension that adds to the model (and exploits its sequential structure) by building in a parallel sequence-based model to track left context.
+This auxiliary model, called the tracking LSTM, maintains a low dimensional representation that is updated after every transition, thereby maintaining a representation of the left context at each time point.
+That representation is fed into the composition function as an additional input to allow for more efficient disambiguation.
+As long as the representations produced by the tracking LSTM are substantially smaller than those produced by the main composition function, the model is heavily biased towards computing sentence meaning compositionally according to the assigned tree structure, and is simply given the ability to do so with a better awareness of context.
 
@@
 We evaluate our model on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as here:
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus (SNLI, \citealt{snli:emnlp2015}), which we use for training and evaluation:
 
@@
 
-In particular, we use the Stanford NLI corpus (SNLI, \citealt{snli:emnlp2015}).
 SNLI is a corpus of 570k human-labeled pairs of scene descriptions like the one above.
@@
 \vspace{14em}
-\caption{\label{fig:induced-alignment}The alignment induced between two sentences by \todo{some model.}. Compare with a similar alignment presented by \protect\citealt{DBLP:journals/corr/WangJ15b}.}
+\caption{\label{fig:induced-alignment}The alignment induced between two sentences by \result{one of our models.}.
+Compare with a similar alignment presented by \protect\citealt{DBLP:journals/corr/WangJ15b}.}
 \end{figure}

### 2016-03-05 writing/hard_stack_paper/batching_fig.tex
Incremental paper update.

%!TEX root = paper.tex

\begin{figure}[t]
\begin{subfigure}[t]{0.9\columnwidth}
\begin{center}
 \begin{tikzpicture}
    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]    
    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]

    \begin{scope}[shift={(0.9in,-0.8in)}, frontier/.style={distance from root=60pt}]

    \Tree [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ]

    \end{scope}

    \begin{scope}[shift={(0in,0in)}]

    \Tree [.\node[nonleaf](2thekittenmeowed){the old cat meowed}; [.\node[nonleaf](2thekitten){the old cat}; \node[word](2the){the}; [.\node[nonleaf](2bigkitten){old cat}; \node[word](2kitten){old}; \node[word](2kitten){cat}; ] ] \node[word](2meowed){meowed}; ]

    \end{scope}

\end{tikzpicture}
\end{center}

\caption{\label{fig:batching:bad}Trees are hard to batch.}
\end{subfigure}

\vspace{2em}

\begin{subfigure}[t]{0.9\columnwidth}
 \begin{tikzpicture}
    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]    
    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]    
    \def\dx{21pt}
    \def\dy{11pt}
    \def\sy{7*\dy}
    \def\oxb{5.5*\dx}
    \def\by{1*\dy}
    \def\ox{0*\oxb}

    \node[word]  (w1) at (\ox+0*\dx,\by-1.5*\dy) {the};
    \node[word]  (w2) at (\ox+0*\dx,\by+0*\dy) {cat};
    \node[word]  (w3) at (\ox+0*\dx,\by+1*\dy) {sat};

    \begin{scope}[shift={(0.9in,-0.8in)}, frontier/.style={distance from root=60pt}]

    \Tree [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ]

    \end{scope}

    \begin{scope}[shift={(0in,0in)}]

    \Tree [.\node[nonleaf](2thekittenmeowed){the old cat meowed}; [.\node[nonleaf](2thekitten){the old cat}; \node[word](2the){the}; [.\node[nonleaf](2bigkitten){old cat}; \node[word](2kitten){old}; \node[word](2kitten){cat}; ] ] \node[word](2meowed){meowed}; ]

    \end{scope}

\end{tikzpicture}

\caption{\label{fig:batching:good}Sequences are easy to batch.}
\end{subfigure}
\caption{An illustration of a conventionally-implemented TreeRNN model (a) and our equivalent model (b). \todo{[SB] Finish.}}
\end{figure}

### 2016-03-05 writing/hard_stack_paper/paper.tex
Incremental paper update.

@@
 
-\begin{figure}[t]
-\begin{subfigure}[t]{0.9\columnwidth}
-\vspace{14em}
-\todo{[SB] Figure showing why trees can't be batched (modeled after slides).}
-
-\caption{\label{fig:batching:bad}Trees are hard to batch.}
-\end{subfigure}
-
-\vspace{2em}
-
-\begin{subfigure}[t]{0.9\columnwidth}
-\vspace{14em}
-\todo{[SB] Figure showing a batch of sequences (modeled after slides).}
-
-\caption{\label{fig:batching:good}Sequences are easy to batch.}
-\end{subfigure}
-\caption{An illustration of a conventionally-implemented TreeRNN model (a) and our equivalent model (b).}
-\end{figure}
+\input{batching_fig.tex}
 
@@
 
-\paragraph{Accumulating results sequentially} (After Rocktäschel, Wang and Jiang)
+\todo{[RG,AR] Accumulating results sequentially  (After Rocktäschel, Wang and Jiang)}
 
-\paragraph{Accumulating results using a second tree layer}
+\todo{[RG,SB] Accumulating results using a second tree layer}
 
@@
 \midrule
-\result{150d LSTM RNN      }     & ?                  &   ?      &   ?       \\
+\result{150d LSTM RNN      }     & ?                  &   ?      &   81?       \\
 \result{150d TB TreeLSTM}   
-                        & ?                  &   ?      &   ?       \\
+                        & ?                  &   ?      &   81?       \\
 \result{150d TB TreeLSTM w/ tracking    }
-                        & ?                  &   ?      &   ?       \\
+                        & ?                  &   ?      &   83?       \\
 \result{150d TB TreeLSTM w/ tracking, sequence-based attn.   }     
-                        & ?                  &   ?      &   ?       \\
+                        & ?                  &   ?      &   85+?       \\
 \result{150d TB TreeLSTM w/ tracking, tree-based attn. }           
@@
 
-\paragraph{Patterns in the examples that the tree model got right that the LSTM didn't}
+\result{Patterns in the examples that the tree model got right that the LSTM didn't}
 
-\paragraph{Induced alignments}
+\result{Induced alignments}
 
@@
 Implementing such a model on top of our proposed stack-based design would learn a model that could productively share representations between the two tasks.
-Implementing such a model on top of a differentiable version of our stack-based design (possibly following \citealt{grefenstette2015learning} \todo{[SB] Cite FAIR stack paper.}) would further make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing  learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+Implementing such a model on top of a differentiable version of our stack-based design (possibly following \citealt{grefenstette2015learning,joulin2015inferring}) would further make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing  learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 

### 2016-03-07 writing/hard_stack_paper/batching_fig.tex
Paper updates.

@@
     \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]
+    \tikzstyle{alt}=[fill=orange!40]    
+    \pgfsetarrowsend{latex}
+    \tikzset{edge from parent/.append style={<-, line width=1pt}}
+
+    \begin{scope}[shift={(0in,0in)}]
+
+    \Tree [.\node[nonleaf,alt](2thekittenmeowed){the old cat meowed}; [.\node[nonleaf,alt](2thekitten){the old cat}; \node[word,alt](2the){the}; [.\node[nonleaf,alt](2bigkitten){old cat}; \node[word,alt](2kitten){old}; \node[word,alt](2kitten){cat}; ] ] \node[word,alt](2meowed){meowed}; ]
+
+    \end{scope}
 
@@
     \Tree [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ]
-
-    \end{scope}
-
-    \begin{scope}[shift={(0in,0in)}]
-
-    \Tree [.\node[nonleaf](2thekittenmeowed){the old cat meowed}; [.\node[nonleaf](2thekitten){the old cat}; \node[word](2the){the}; [.\node[nonleaf](2bigkitten){old cat}; \node[word](2kitten){old}; \node[word](2kitten){cat}; ] ] \node[word](2meowed){meowed}; ]
 
@@
     \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black] 
-    \def\dx{21pt}
+    \tikzstyle{alt}=[fill=orange!40]
+    \pgfsetarrowsend{latex}
+    \tikzstyle{fwd} = [draw=black, line width=1pt]
+
+    \def\dx{26pt}
     \def\dy{11pt}
@@
 
-    \node[word]  (w1) at (\ox+0*\dx,\by-1.5*\dy) {the};
-    \node[word]  (w2) at (\ox+0*\dx,\by+0*\dy) {cat};
-    \node[word]  (w3) at (\ox+0*\dx,\by+1*\dy) {sat};
+    \begin{scope}[shift={(0in,0in)}, frontier/.style={distance from root=60pt}]
 
-    \begin{scope}[shift={(0.9in,-0.8in)}, frontier/.style={distance from root=60pt}]
+    \node[word,alt]  (w1) at (\ox+-3*\dx,\by+0*\dy) {the};
+    \node[word,alt]  (w2) at (\ox+-1*\dx,\by+0*\dy) {old};
+    \node[word,alt]  (w3) at (\ox+1*\dx,\by+0*\dy) {cat};
+    \node[word,alt]  (w4) at (\ox+3*\dx,\by+0*\dy) {meowed};
 
-    \Tree [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ]
+    \node[word,alt]  (n1) at (\ox+-3*\dx,\by+3*\dy) {~~~~~};
+    \node[word,alt]  (n2) at (\ox+-1*\dx,\by+3*\dy) {~~~~~};
+    \node[word,alt]  (n3) at (\ox+1*\dx,\by+3*\dy) {~~~~~};
+    \node[word,alt]  (n4) at (\ox+3*\dx,\by+3*\dy) {~~~~~};
+
+   \draw [fwd] (w1) -- (n1);
+   \draw [fwd] (w2) -- (n2);
+   \draw [fwd] (w3) -- (n3);
+   \draw [fwd] (w4) -- (n4);
+
+   \draw [fwd] (n1) -- (n2);
+   \draw [fwd] (n2) -- (n3);
+   \draw [fwd] (n3) -- (n4);
 
@@
 
-    \begin{scope}[shift={(0in,0in)}]
+    \begin{scope}[shift={(0.35in,-0.25in)}, frontier/.style={distance from root=60pt}]
 
-    \Tree [.\node[nonleaf](2thekittenmeowed){the old cat meowed}; [.\node[nonleaf](2thekitten){the old cat}; \node[word](2the){the}; [.\node[nonleaf](2bigkitten){old cat}; \node[word](2kitten){old}; \node[word](2kitten){cat}; ] ] \node[word](2meowed){meowed}; ]
+    \node[word]  (w1) at (\ox+-3*\dx,\by+0*\dy) {the};
+    \node[word]  (w2) at (\ox+-1*\dx,\by+0*\dy) {cat};
+    \node[word]  (w3) at (\ox+1*\dx,\by+0*\dy) {sat};
+    \node[word]  (w4) at (\ox+3*\dx,\by+0*\dy) {down};
+
+    \node[word]  (n1) at (\ox+-3*\dx,\by+3*\dy) {~~~~~};
+    \node[word]  (n2) at (\ox+-1*\dx,\by+3*\dy) {~~~~~};
+    \node[word]  (n3) at (\ox+1*\dx,\by+3*\dy) {~~~~~};
+    \node[word]  (n4) at (\ox+3*\dx,\by+3*\dy) {~~~~~};
+
+   \draw [fwd] (w1) -- (n1);
+   \draw [fwd] (w2) -- (n2);
+   \draw [fwd] (w3) -- (n3);
+   \draw [fwd] (w4) -- (n4);
+
+   \draw [fwd] (n1) -- (n2);
+   \draw [fwd] (n2) -- (n3);
+   \draw [fwd] (n3) -- (n4);
 
@@
 \end{subfigure}
-\caption{An illustration of a conventionally-implemented TreeRNN model (a) and our equivalent model (b). \todo{[SB] Finish.}}
+\caption{An illustration of a conventionally-implemented TreeRNN model (a) and a sequence-based RNN model (b).
 \end{figure}

### 2016-03-08 writing/hard_stack_paper/paper.tex
Paper updates.
Added citation.

@@
 \item Our proposed model builds on ideas from transition-based parsing to build a model which implements the same computations as a tree-structured model, but which can be efficiently batched.
-See Figure~\ref{fig:batching:good}.
 
@@
 
-There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing \cite{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP}.
+There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing \cite{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
 In addition, there has been recent work \cite{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.

### 2016-03-09 metal_stack/thin-stack.cc
Sketch kernel headers

@@ -61,6 +61,8 @@ ThinStack::forward() {
 
 ThinStack::step(int t) {
 
+  // TODO sync after kernel calls.
+
   // Extract top buffer values.
   // TODO subtensor with idxs buffer_cur_t + 0 * 1 + buffer_shift
   // == buffer_cur_t * 1 + (batch_range * seq_length)

### 2016-03-10 writing/hard_stack_paper/paper.tex
Initial abstract.

@@
 
-\title{Tree-structured neural networks revisited}
+\title{Combining parsing with sentence understanding}
 
@@
 \begin{abstract}
-\begin{itemize}
-\item TreeRNNs are great, don't work for big data.
-\item We present the transition-based TreeRNN implementation -- makes efficient training possible, show that it helps relative to seq models.
-\item Further, attention has proven powerful for extending sequence models. We introduce constituent-by-constituent attention, and find it to be effective here, too.
-\end{itemize}
+
+Tree-structured neural networks are linguistically principled models for sentence encoding that have proven effective on several natural language understanding tasks.
+However, these models can only operate on parsed sentences, and do not support batched computation, making them prohibitively slow for large-scale tasks.
+We present the parser--interpreter neural network, or PINN, which implements the sentence encoding computations of a tree-structured neural network within the structure of a transition-based neural constituency parser.
+This model addresses both of the weaknesses of existing tree-structured models: the transition-based architecture of the parser makes it possible to perform tree-structured sentence interpretation in an efficient, batched, setting, and the deep integration with the parser allows the model to handle interpret unparsed data at test time with nearly no loss in accuracy.
+Finally, we show that this architecture supports a novel and effective method for handling local context in sentence interpretation, and integrates will with soft attention methods for sentence pair tasks.
+
 \end{abstract}
@@
 \todo{Anonymize.}
-\todo{Name the damn model!}
+\todo{Fill in name where needed.}
+\todo{SST!}
+\todo{Model 1!}
 
@@
 
-\paragraph{Representing the stack in memory} Or, thin stack. 
+\paragraph{Representing the stack in memory} Or, thin stack. And the cursor strategy for the buffer. 
 \todo{[JG] Explain.}
@@
 
-We evaluate five models on the SNLI task. Each uses 150d hidden states: 
+We evaluate five models on the SNLI task. Each uses 300d hidden states: 
 \begin{itemize}
@@
 \midrule
-\result{150d LSTM RNN      }     & ?                  &   ?      &   81?       \\
-\result{150d TB TreeLSTM}   
+\result{300d LSTM RNN      }     & ?                  &   ?      &   81?       \\
+\result{300d TB TreeLSTM}   
                         & ?                  &   ?      &   81?       \\
-\result{150d TB TreeLSTM w/ tracking    }
+\result{300d TB TreeLSTM w/ tracking    }
                         & ?                  &   ?      &   83?       \\
-\result{150d TB TreeLSTM w/ tracking, sequence-based attn.   }     
+\result{300d TB TreeLSTM w/ tracking, sequence-based attn.   }     
                         & ?                  &   ?      &   85+?       \\
-\result{150d TB TreeLSTM w/ tracking, tree-based attn. }           
+\result{300d TB TreeLSTM w/ tracking, tree-based attn. }           
                         & ?                  &   ?      &   \textbf{?}\\

### 2016-03-10 writing/hard_stack_paper/paper.tex
Abstract update.

--- 2016-03-08 writing/hard_stack_paper/paper.tex
@@
 \begin{abstract}
-\begin{itemize}
-\item TreeRNNs are great, don't work for big data.
-\item We present the transition-based TreeRNN implementation -- makes efficient training possible, show that it helps relative to seq models.
-\item Further, attention has proven powerful for extending sequence models. We introduce constituent-by-constituent attention, and find it to be effective here, too.
-\end{itemize}
+
+Tree-structured neural networks are highly as effective sentence models, but they can only operate on parsed sentences, and do not support batched computation, making them prohibitively slow for large-scale tasks.
+This paper presents the parser-interpreter neural network, which reimplements a tree-structured network within a transition-based neural constituency parser.
+This allows us to address both weaknesses: the transition-based architecture of the parser makes it possible to perform tree-structured interpretation in an efficient batched setting, and the integrated parser allows the model to interpret unparsed data with nearly no loss in accuracy.
+In addition, our model supports a novel and effective method for handling local context in sentence interpretation and integrates will with soft attention methods for sentence pair tasks.
+
 \end{abstract}
@@
 \todo{Anonymize.}
-\todo{Name the damn model!}
+
+\todo{Fill in name where needed.}
+
+\todo{SST!}
+
+\todo{Model 1!}
 

### 2016-03-10 writing/hard_stack_paper/paper.tex
Abstract update.

@@
 
-Tree-structured neural networks are highly as effective sentence models, but they can only operate on parsed sentences, and do not support batched computation, making them prohibitively slow for large-scale tasks.
-This paper presents the parser-interpreter neural network, which reimplements a tree-structured network within a transition-based neural constituency parser.
-This allows us to address both weaknesses: the transition-based architecture of the parser makes it possible to perform tree-structured interpretation in an efficient batched setting, and the integrated parser allows the model to interpret unparsed data with nearly no loss in accuracy.
-In addition, our model supports a novel and effective method for handling local context in sentence interpretation and integrates will with soft attention methods for sentence pair tasks.
+Tree-structured neural networks integrate valuable syntactic parse information into the process of interpreting sentence meaning, but suffer two key technical problems.
+They can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks.
+We address these issues by introducing the Parser-Interpreter Neural Network (PINN), which simulates tree-structured sentence interpretation within the sequential structure of a transition-based neural parser.
+This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
+We evaluate it on the Stanford NLI entailment task, show that it outperforms other sentence encoding models, and show that it can be integrated effectively with soft attention for sentence pair tasks like entailment for further gains in performance.
 

### 2016-03-12 metal_stack/test.cc
add gtest for S S M feedforward. It works..

@@ -1,0 +1,102 @@
+#include <cuda_runtime.h>
+#include "cublas_v2.h"
+
+#include "gmock/gmock.h"
+#include "gtest/gtest.h"
+
+#include "thin-stack.h"
+#include "util.h"
+
+using namespace testing;
+
+static ThinStack make_stack(ModelSpec spec) {
+  // Make up random parameters.
+  float *compose_W_l, *compose_W_r;
+  cudaMalloc(&compose_W_l, spec.model_dim * spec.model_dim * sizeof(float));
+  cudaMalloc(&compose_W_r, spec.model_dim * spec.model_dim * sizeof(float));
+  fill_rand_matrix(compose_W_l, spec.model_dim, spec.model_dim);
+  fill_rand_matrix(compose_W_r, spec.model_dim, spec.model_dim);
+
+  ThinStackParameters params = {
+    NULL, NULL, // projection
+    NULL, NULL, // buffer batch-norm
+    NULL, NULL, NULL, // tracking
+    compose_W_l, compose_W_r, NULL, NULL // composition
+  };
+  cublasHandle_t handle = getCublasHandle();
+  ThinStack ts(spec, params, handle);
+  return ts;
+}
+
+static void free_stack(ThinStack s) {
+  cudaFree(s.params.compose_W_l);
+  cudaFree(s.params.compose_W_r);
+  cublasDestroy(s.handle);
+}
+
+static inline void assert_matrices_equal(const float *m1, const float *m2,
+    int M, int N) {
+  cudaDeviceSynchronize();
+  float *h_m1 = (float *) malloc(2 * M * N * sizeof(float));
+  float *h_m2 = &h_m1[M * N];
+
+  cudaMemcpy(h_m1, m1, M * N * sizeof(float), cudaMemcpyDeviceToHost);
+  cudaMemcpy(h_m2, m2, M * N * sizeof(float), cudaMemcpyDeviceToHost);
+  cudaDeviceSynchronize();
+
+  for (int i = 0; i < M; i++) {
+    for (int j = 0; j < N; j++) {
+      ASSERT_THAT(h_m1[j * M + i], FloatEq(h_m1[j * M + i]));
+    }
+  }
+
+  free(h_m1);
+}
+
+
+// Test simple shift-shift-merge feedforward with live random weights.
+TEST(ThinStackTest, ShiftShiftMerge) {
+
+  // TODO: Try with larger batch size, model dim
[…]

### 2016-03-12 metal_stack/thin-stack.h

@@ -33,6 +33,10 @@ class ThinStack {
 
     ~ThinStack();
 
+    ModelSpec spec;
+    ThinStackParameters params;
+    cublasHandle_t handle;
+
     void forward();
 
     // Embedding inputs, of dimension `model_dim * (batch_size * seq_length)` --
@@ -61,10 +65,6 @@ class ThinStack {
 
     void init_helpers();
     void free_helpers();
-
-    ModelSpec spec;
-    ThinStackParameters params;
-    cublasHandle_t handle;
 
     size_t stack_size;
 

### 2016-03-12 metal_stack/util.cc

@@ -25,6 +25,17 @@ float *load_weights_cuda(string filename, int N) {
 }
 
 
+cublasHandle_t getCublasHandle() {
+  cublasHandle_t handle;
+  cublasStatus_t stat = cublasCreate(&handle);
+  if (stat != CUBLAS_STATUS_SUCCESS) {
+    cout << "CUBLAS initialization failed (" << stat << ")" << endl;
+    return NULL;
+  }
+  return handle;
+}
+
+
 // Print a column-major matrix stored on device.
 void print_device_matrix(const float *m, int M, int N) {
   cudaDeviceSynchronize();

### 2016-03-12 metal_stack/util.h

@@ -7,6 +7,7 @@
 
 #include <cuda_runtime.h>
 #include <curand.h>
+#include "cublas_v2.h"
 
 using namespace std;
 
@@ -25,6 +26,8 @@ typedef struct ModelSpec {
 
 float *load_weights(string filename, int N);
 float *load_weights_cuda(string filename, int N);
+
+cublasHandle_t getCublasHandle();
 
 void print_device_matrix(const float *m, int M, int N);
 void fill_rand_matrix(float *m, int M, int N);

### 2016-03-12 writing/hard_stack_paper/batching_fig.tex
Fix bug
Abstract update.
Pre-coffee incremental paper commit.

@@
 
-    \Tree [.\node[nonleaf,alt](2thekittenmeowed){the old cat meowed}; [.\node[nonleaf,alt](2thekitten){the old cat}; \node[word,alt](2the){the}; [.\node[nonleaf,alt](2bigkitten){old cat}; \node[word,alt](2kitten){old}; \node[word,alt](2kitten){cat}; ] ] \node[word,alt](2meowed){meowed}; ]
+    \Tree [.\node[nonleaf,alt](2thekittenate){the old cat ate}; [.\node[nonleaf,alt](2thekitten){the old cat}; \node[word,alt](2the){the}; [.\node[nonleaf,alt](2bigkitten){old cat}; \node[word,alt](2kitten){old}; \node[word,alt](2kitten){cat}; ] ] \node[word,alt](2ate){ate}; ]
 
@@
 
-\caption{\label{fig:batching:bad}Trees are hard to batch.}
+\caption{\label{fig:batching:bad}A conventional TreeRNN, instantiated for two sentences.}
 \end{subfigure}
@@
     \node[word,alt]  (w3) at (\ox+1*\dx,\by+0*\dy) {cat};
-    \node[word,alt]  (w4) at (\ox+3*\dx,\by+0*\dy) {meowed};
+    \node[word,alt]  (w4) at (\ox+3*\dx,\by+0*\dy) {ate};
 
@@
 
-\caption{\label{fig:batching:good}Sequences are easy to batch.}
+\caption{\label{fig:batching:good}A conventional sequence-based RNN, instantiated for two sentences.}
 \end{subfigure}
-\caption{An illustration of a conventionally-implemented TreeRNN model (a) and a sequence-based RNN model (b).
+\caption{An illustration of two standard designs for sentence encoders.
+Note that the TreeRNN, unlike the sequence-based RNN, requires a substantially different connection structure for each sentence.}
 \end{figure}

### 2016-03-12 writing/hard_stack_paper/paper.tex
Pre-coffee incremental paper commit.

@@
 
-\title{Combining parsing with sentence understanding}
+\title{A unified model for parsing and sentence understanding}
 
-\author{
-Samuel R.\ Bowman$^{1,2,5,}$\thanks{The first two authors contributed equally.} \\
-\texttt{\small sbowman@stanford.edu} \\
-\And
-Jon Gauthier$^{2,3,5,*}$ \\
-\texttt{\small jgauthie@stanford.edu} \\
-\And
-Abhinav Rastogi$^{4,5}$ \\
-\texttt{\small arastogi@stanford.edu} \\
-\AND
-Raghav Gupta$^{6}$ \\
-\texttt{\small rgupta93@stanford.edu} \\
-\And
-Christopher D.\ Manning$^{1,2,5,6}$\\
-\texttt{\small manning@stanford.edu}\\
-\And
-Christopher Potts$^{1}$\\
-\texttt{\small cgpotts@stanford.edu}
-\AND\\[-3ex]
-{$^{1}$Stanford Linguistics\quad
-$^{2}$Stanford NLP Group\quad
-$^{3}$Stanford Symbolic Systems}\\
-{$^{4}$Stanford Electrical Engineering\quad
-$^{5}$Stanford AI Lab\quad
-$^{6}$Stanford Computer Science}
-}
+%    \author{
+%    Samuel R.\ Bowman$^{1,2,5,}$\thanks{The first two authors contributed equally.} \\
+%    \texttt{\small sbowman@stanford.edu} \\
+%    \And
+%    Jon Gauthier$^{2,3,5,*}$ \\
+%    \texttt{\small jgauthie@stanford.edu} \\
+%    \And
+%    Abhinav Rastogi$^{4,5}$ \\
+%    \texttt{\small arastogi@stanford.edu} \\
+%    \AND
+%    Raghav Gupta$^{6}$ \\
+%    \texttt{\small rgupta93@stanford.edu} \\
+%    \And
+%    Christopher D.\ Manning$^{1,2,5,6}$\\
+%    \texttt{\small manning@stanford.edu}\\
+%    \And
+%    Christopher Potts$^{1}$\\
+%    \texttt{\small cgpotts@stanford.edu}
+%    \AND\\[-3ex]
+%    {$^{1}$Stanford Linguistics\quad
+%    $^{2}$Stanford NLP Group\quad
+%    $^{3}$Stanford Symbolic Systems}\\
+%    {$^{4}$Stanford Electrical Engineering\quad
+%    $^{5}$Stanford AI Lab\quad
+%    $^{6}$Stanford Computer Science}
+%    }
 
@@
 We evaluate it on the Stanford NLI entailment task, show that it outperforms other sentence encoding models, and show that it can be integrated effectively with soft attention for sentence pair tasks like entailment for further gains in performance.
-
+\todo{[SB] Revise according to comments.}
 \end{abstract}
 
-\todo{Anonymize.}
-
-\todo{Fill in name where needed.}
-
-\todo{SST!}
-
-\todo{Model 1!}
+\question{Parser--Interpreter or Parser-Interpreter (hyphen or en-dash)?}
 
@@
 As the schematic in Figure~\ref{fig:batching:bad} indicates, different sentences are assigned different trees, and there is no straightforward way of aligning these trees to synchronize computation in the conventional style of implementation.
-\item Our proposed model builds on ideas from transition-based parsing to build a model which implements the same computations as a tree-structured model, but which can be efficiently batched.
+\item In addition, the requirement of an external parser complicated pipeline-building, and slows down inference.
+This is not necessarily a practical limitation, since modern parsers can assign reasonably high quality parses in so little time as to not dramatically impact the performance characteristics of our model \todo{[SB,JG,CM] How fast are fast constit parsers?}.
+
+\end{itemize}
+
+\item Our model
+\begin{itemize}
+\item In this paper, we present the Parser--Interpreter NN (PINN)
 
@@
 
-Our model is based on a formalism adapted from transition-based parsing (\question{What's a good classic citation for this?}).
+The PINN is based on a formalism adapted from transition-based parsing (\question{What's a good classic citation for this?}).
 
@@
 
-Our model, which is shown in Figure~\ref{fig:model:0}, takes the form of a transition-based parser, but it is designed to produce sentence representations given parse structures, rather than producing parse structures as its output, leading to the following differences:
+The PINN, which is shown in Figure~\ref{fig:model:0}, takes the form of a transition-based parser, but it is designed to produce sentence representations given parse structures, rather than producing parse structures as its output, leading to the following differences: \todo{[SB] Swap in Model 1 fig.}
 \begin{itemize}
@@
 \item A novel tracking LSTM component is added to improve the model's ability to efficiently track the context in which each step is being performed.
+\item A variant, PINN-GT uses ground truth parse information at training and test time, and omits the prediction layer.
 \end{itemize}
@@
 
-In our model, each row of the stack contains a pair of vectors $(h, c)$, which together represent some node in the parse tree of the sentence.
+In a PINN, each row of the stack contains a pair of vectors $(h, c)$, which together represent some node in the parse tree of the sentence.
 Since the \shift~operation simply copies entries from the buffer onto the stack, this means that the word representations in the buffer also have two parts, $h$ and $c$.
@@
 
+\subsection{Parsing: Predicting transitions from the tracking state}
+
+\todo{[SB] Fill in math, Explain.}
+
 \subsection{Implementing the transition-based sentence model}
@@
 
-In its bare form, without the addition of the tracking LSTM, our model computes the precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
-While it adds no expressive power on its own, our model design improves upon the standard TreeRNN implementation in three ways:
+In its bare form, before the addition of the tracking LSTM, the PINN computes the precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
+While it adds no expressive power on its own, this design improves upon the standard TreeRNN implementation in three ways:
 
@@
 In addition, there has been recent work \cite{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
-To our knowledge, our model is the first to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
+To our knowledge, the PINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
 
@@
 
-In our model, attention is performed at each step $t$ of the premise encoder. At step $t$, the query vector that drives attention will be $S^t_0$, the top of the stack.
+In the PINN, attention is performed at each step $t$ of the premise encoder. At step $t$, the query vector that drives attention will be $S^t_0$, the top of the stack.
 
@@
 
-We evaluate our model on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
+We evaluate the PINN on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
 NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus (SNLI, \citealt{snli:emnlp2015}), which we use for training and evaluation:
@@
 
-We evaluate five models on the SNLI task. Each uses 300d hidden states: 
+We evaluate five models on the SNLI task. Each uses 300d hidden states: \todo{[SB] Reframe around M1.}
 \begin{itemize}
@@
     \toprule
-Model                   & Trained params.    &   Train  &   Test \\
+Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
 \midrule
@@
 Lexicalized classifier \cite{snli:emnlp2015}
-                        & --                 &   99.7   &   78.2      \\
+                        & --                & --                    &   99.7   &   78.2      \\
 \midrule
@@
 100d LSTM encoders \cite{snli:emnlp2015}
-                        & 221k               &   84.8   &   77.6      \\
+                        & 221k               & --               &   84.8   &   77.6      \\
 1024d pretrained GRU encoders \cite{DBLP:journals/corr/VendrovKFU15}
-                        & 15m                &   98.8   &   81.4       \\
+                        & 15m                & --              &   98.8   &   81.4       \\
 300d Tree-based CNN encoders \cite{mou2015recognizing}
-                        & 3.5m               &   83.4   &   82.1       \\
+                        & 3.5m                & --             &   83.4   &   82.1       \\
 \midrule
@@
 100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
-                        & 252k               &   85.3   &   83.5       \\
+                        & 252k               & --              &   85.3   &   83.5       \\
 300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
-                        & 1.9m               &   92.0   &   86.1      \\
-300d LSTMN with deep attention fusion \cite{cheng2016long} (1.4m params) 92.3    89.1
-
-                        & 1.4m               &   92.3   &   89.1      \\
+                        & 1.9m               & --             &   92.0   &   86.1      \\
+300d LSTMN with deep attention fusion \cite{cheng2016long}
+                        & 1.4m               & --                &   92.3   &   89.1      \\
 \midrule
@@
 \midrule
-\result{300d LSTM RNN      }     & ?                  &   ?      &   81?       \\
-\result{300d TB TreeLSTM}   
-                        & ?                  &   ?      &   81?       \\
-\result{300d TB TreeLSTM w/ tracking    }
-                        & ?                  &   ?      &   83?       \\
-\result{300d TB TreeLSTM w/ tracking, sequence-based attn.   }     
-                        & ?                  &   ?      &   85+?       \\
-\result{300d TB TreeLSTM w/ tracking, tree-based attn. }           
-                        & ?                  &   ?      &   \textbf{?}\\
+\result{300d LSTM RNN      } & ?             & --                &   ?      &   81?       \\
+\result{300d PINN-GT w/o tracking}   
+                        & ?                  & --                &   ?      &   81?       \\
+\result{300d PINN-GT }
+                        & ?                  & --                &   ?      &   83?       \\
+\result{300d PINN }
+                        & ?                  & ??                  &   ?      &   83?       \\                        
+\result{300d PINN-GT, sequence-based attn.   }     
+                        & ?                  & --                &   ?      &   85+?       \\
+\result{300d PINN-GT, tree-based attn. }           
+                        & ?                  & --                &   ?      &   \textbf{?}\\
     \bottomrule
@@
   \protect\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification.
-All reported figures are percent accuracy.
+Params.
+is the approximate number of trained parameters (excluding word embeddings for models where they are trained).
+Trans.
+acc.
+is the model's accuracy in predicting parsing transitions.
+Train and test are SNLI classification accuracy.
 \result{[JG]Add seconds-per-step *if* we get a reasonably well-optimized model in time.}}
@@
 
-\paragraph{Future work} The tracking LSTM that we incorporated into our model uses only simple, quick to compute features drawn from the head of the buffer and the head of the stack.
+\paragraph{Future work} The tracking LSTM that we incorporated into the PINN uses only simple, quick to compute features drawn from the head of the buffer and the head of the stack.
 It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors in the sentence string and its likely position in the final tree, and that that context information could support better sentence encoding.
@@
 
+\todo{[SB] Remove Model 1 references.}
+
 In future work, we aim to incorporate a parser into our model, allowing it to simultaneously learn to parse sentences (using a conventional parsing objective function) and to use those parses to learn semantic representations for those sentences (using a classification objective function, as in this work).
@@
 
-\subsubsection*{Acknowledgments}
-
-Some of the Tesla K40(s) used for this research was/were donated by the NVIDIA Corporation.
-\todo{[CM,CP] Acknowledge other grants.}
+%    \subsubsection*{Acknowledgments}
+%    
+%    Some of the Tesla K40(s) used for this research was/were donated by the NVIDIA Corporation.
+%    \todo{[CM,CP] Acknowledge other grants.}
 

### 2016-03-12 metal_stack/test.cc
Refactor tests

@@ -53,6 +53,23 @@ static inline void assert_matrices_equal(const float *m1, const float *m2,
   free(h_m1);
 }
 
+static float *compose(float *dst, ThinStack ts, const float *l,
+    const float *r) {
+  // W_l l
+  float alpha = 1.0f, beta = 0.0f;
+  cublasSgemm(ts.handle, CUBLAS_OP_N, CUBLAS_OP_N, ts.spec.model_dim,
+      ts.spec.batch_size, ts.spec.model_dim, &alpha, ts.params.compose_W_l,
+      ts.spec.model_dim, l, ts.spec.model_dim, &beta, dst, ts.spec.model_dim);
+
+  // += W_r r
+  float beta2 = 1.0f;
+  cublasSgemm(ts.handle, CUBLAS_OP_N, CUBLAS_OP_N, ts.spec.model_dim,
+      ts.spec.batch_size, ts.spec.model_dim, &alpha, ts.params.compose_W_r,
+      ts.spec.model_dim, r, ts.spec.model_dim, &beta2, dst, ts.spec.model_dim);
+
+  return dst;
+}
+
 
 // Test simple shift-shift-merge feedforward with live random weights.
 TEST(ThinStackTest, ShiftShiftMerge) {
@@ -80,19 +97,9 @@ TEST(ThinStackTest, ShiftShiftMerge) {
   float *expected;
   cudaMalloc(&expected, spec.model_dim * spec.batch_size * sizeof(float));
 
-  // W_l l
-  float alpha = 1.0f, beta = 0.0f;
   float *left_child = &ts.X[0];
-  cublasSgemm(ts.handle, CUBLAS_OP_N, CUBLAS_OP_N, spec.model_dim, spec.batch_size,
-      spec.model_dim, &alpha, ts.params.compose_W_l, spec.model_dim,
-      left_child, spec.model_dim, &beta, expected, spec.model_dim);
-
-  // += W_r r
-  float beta2 = 1.0f;
   float *right_child = &ts.X[spec.model_dim * spec.batch_size];
-  cublasSgemm(ts.handle, CUBLAS_OP_N, CUBLAS_OP_N, spec.model_dim, spec.batch_size,
-      spec.model_dim, &alpha, ts.params.compose_W_r, spec.model_dim,
-      right_child, spec.model_dim, &beta2, expected, spec.model_dim);
+  compose(expected, ts, left_child, right_child);
 
   float *output = &ts.stack[0];
[…]

### 2016-03-12 writing/hard_stack_paper/batching_fig.tex
Post-coffee paper progress.

@@
 \begin{figure}[t]
-\begin{subfigure}[t]{0.9\columnwidth}
-\begin{center}
- \begin{tikzpicture}
-    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]    
-    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]
-    \tikzstyle{alt}=[fill=orange!40]    
-    \pgfsetarrowsend{latex}
-    \tikzset{edge from parent/.append style={<-, line width=1pt}}
 
-    \begin{scope}[shift={(0in,0in)}]
-
-    \Tree [.\node[nonleaf,alt](2thekittenate){the old cat ate}; [.\node[nonleaf,alt](2thekitten){the old cat}; \node[word,alt](2the){the}; [.\node[nonleaf,alt](2bigkitten){old cat}; \node[word,alt](2kitten){old}; \node[word,alt](2kitten){cat}; ] ] \node[word,alt](2ate){ate}; ]
-
-    \end{scope}
-
-    \begin{scope}[shift={(0.9in,-0.8in)}, frontier/.style={distance from root=60pt}]
-
-    \Tree [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ]
-
-    \end{scope}
-
-\end{tikzpicture}
-\end{center}
-
-\caption{\label{fig:batching:bad}A conventional TreeRNN, instantiated for two sentences.}
-\end{subfigure}
-
-\vspace{2em}
-
-\begin{subfigure}[t]{0.9\columnwidth}
+\begin{subfigure}[t]{\columnwidth}
  \begin{tikzpicture}
@@
 
-    \def\dx{26pt}
+    \def\dx{23pt}
     \def\dy{11pt}
@@
     \node[word,alt]  (n4) at (\ox+3*\dx,\by+3*\dy) {~~~~~};
+    \node[]  (n5) at (\ox+4.5*\dx,\by+3*\dy) {...};
 
@@
    \draw [fwd] (n3) -- (n4);
+   \draw [fwd] (n4) -- (n5);
 
@@
     \node[word]  (n4) at (\ox+3*\dx,\by+3*\dy) {~~~~~};
+    \node[]  (n5) at (\ox+4.5*\dx,\by+3*\dy) {...};
 
@@
 \end{subfigure}
+
+\vspace{2em}
+
+\begin{subfigure}[t]{\columnwidth}
+\begin{center}
+ \begin{tikzpicture}
+    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]    
+    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]
+    \tikzstyle{alt}=[fill=orange!40]    
+    \pgfsetarrowsend{latex}
+    \tikzset{edge from parent/.append style={<-, line width=1pt}}
+
+    \begin{scope}[shift={(0in,0in)}]
+
+    \Tree [.\node[](root){...}; [.\node[nonleaf,alt](2thekittenate){the old cat ate}; [.\node[nonleaf,alt](2thekitten){the old cat}; \node[word,alt](2the){the}; [.\node[nonleaf,alt](2bigkitten){old cat}; \node[word,alt](2kitten){old}; \node[word,alt](2kitten){cat}; ] ] \node[word,alt](2ate){ate}; ] ]
+
+    \end{scope}
+
+    \begin{scope}[shift={(0.9in,-0.8in)}]
+
+    \Tree [.\node[](root){...}; [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ] ]
+
+    \end{scope}
+
+\end{tikzpicture}
+\end{center}
+
+\caption{\label{fig:batching:bad}A conventional TreeRNN, instantiated for two sentences.}
+\end{subfigure}
+
 \caption{An illustration of two standard designs for sentence encoders.

### 2016-03-12 writing/hard_stack_paper/paper.tex
Post-coffee paper progress.

@@
 
+     \author{}
 %    \author{
@@
 
+A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \cite{socher2011semi,sutskever2014sequence} \todo{[SB] Cite Oriol's conversation paper.}.
+This component, the sentence encoder, is generally formulated as a learned parametric function from word vectors to sentence vectors, and this function can take a range of different forms.
+Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (\citealt{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \cite{kalchbrenner2014convolutional} \todo{Cite LeCun CharConvNet}, which pool information over short localsequences of words, and tree-structured recursive neural networks (TreeRNNs \citealt{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary tree structure over the words.
+Of these, there are tidy theoretical arguments suggesting that the TreeRNN is the principled choice: meaning in natural language sentences is widely believed to be constructed incrementally according to a tree-structure \cite{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time recovering the process by which meaning is constructed.
+
+In practice, TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons.
+Batched computation--performing synchronized computation across many examples at once--is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets \question{What, if anything, can we cite here?}.
+Batching yields order-of-magnitude improvements in standard CPU implementation, and substantial additional gains on GPUs.
+TreeRNNs build the parse structure of each sentence they process into the structure of the network itself, forcing the model to perform a different sequence of computations for each sentence, and thereby requiring that each sentence be processed one-by-one, making batched computation impossible.
+One aim of this paper is to make these large-scale evaluations possible.
+Beyond even this large handicap, TreeRNNs can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.
+
+In spite of the unwieldiness of current standard TreeRNNs, there is still value in the idea of using syntactic parse trees to guide the construction of meaning in neural network models.
+Tree-structured models have shown substantial performance gains on some smaller-scale NLP tasks \cite{tai2015improved,li2015tree}, and their theoretical foundations are strong.
+Tree-structured compositionality is a property of natural language, and simpler sequence-based RNNs have a difficult time exploiting this property---while sequence-based models should, theoretically, be able to learn to simulate tree-structured composition, this behavior does not emerge in practice on reasonably-sized tasks \cite{bowman2015trees}.
+
+In this paper, we introduce the Parser--Interpreter Neural Network, or PINN.
+The PINN is a model which reimplements tree-structured composition within an alternative design that incorporates a neural network parser.
+This design improves upon the previous TreeRNN approach to compositionality in three ways:
 \begin{itemize}
-
-\item Brief history of TreeRNNs
-\begin{itemize}
-\item Lx theory suggests should be better than comparable sequence models via compositionality \cite{Partee84,Janssen97}.
-\item In practice, results mixed but positive.
-\citealt{tai2015improved} shows dramatic gains on sentiment analysis using tree-structured models over sequence models.
-\citealt{li2015tree} shows mixed results.
-Both of these evaluations are limited to small datasets with only hundreds or thousands of training examples.
-\item Larger evaluations have been impossible due to issues of computational efficiency. One aim of this paper is to mitigate these issues to make larger-scale experiments possible.
-\item Note, though, that while tree structure can be seen as a bias, simply adding more data within reason may not be enough to get a sequence model to learn without that bias \citealt{bowman2015trees}.
+\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
+\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
+\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
 \end{itemize}
 
-\item Speed issues
-\begin{itemize}
-\item Efficient batched computation is a crucial enabling technology in allowing NNs to be used on large datasets.
-Since NNs require large datasets to be able to model complex problems like language understanding, this is crucial in NLP.
-\item Tree structured neural networks can't be batched.
-As the schematic in Figure~\ref{fig:batching:bad} indicates, different sentences are assigned different trees, and there is no straightforward way of aligning these trees to synchronize computation in the conventional style of implementation.
-\item In addition, the requirement of an external parser complicated pipeline-building, and slows down inference.
-This is not necessarily a practical limitation, since modern parsers can assign reasonably high quality parses in so little time as to not dramatically impact the performance characteristics of our model \todo{[SB,JG,CM] How fast are fast constit parsers?}.
+\todo{[SB] Introduce attention here if appropriate.}
 
-\end{itemize}
-
-\item Our model
-\begin{itemize}
-\item In this paper, we present the Parser--Interpreter NN (PINN)
-
-\end{itemize}
-
-\item Attention issues
+\textcolor{gray}{
 \begin{itemize}
@@
 \end{itemize}
+}
 
-\item Outline of the paper
-\begin{itemize}
-\item Propose a model
-\item Show how the model can be used for NLI
-\item Show how attention can be used on the NLI model
-\item Experiments and discussion
-\end{itemize}
+The paper is structured as follows: We first introduce the PINN, then demonstrate its application to the task of natural language inference.
+We then discuss applications of neural attention to the PINN.
+Finally, we present experimental results and discussion.
 
-\end{itemize}
+\input{../model1_fig.tex}
 
@@
 
-\subsection{The model}
+\subsection{Composition and representation}
 
-The PINN, which is shown in Figure~\ref{fig:model:0}, takes the form of a transition-based parser, but it is designed to produce sentence representations given parse structures, rather than producing parse structures as its output, leading to the following differences: \todo{[SB] Swap in Model 1 fig.}
+The PINN, which is shown in Figure~\ref{fig:model:1d}, takes the form of a transition-based parser, but it is designed to produce sentence representations given parse structures, rather than producing parse structures as its output, leading to the following differences:
 \begin{itemize}
-\item The model takes a sequence of transitions (\shift s or \reduce s) as input, rather than incorporating a decision function which chooses which transition to perform at each step.
 \item The intermediate representations in the parser, including the representations of the tree structures on the stack, are vectors of a fixed length.
 \item The \reduce~operation, which combines two trees or tree nodes on the stack into a single larger tree, is represented using a parametric neural network layer function (in particular, a TreeLSTM layer, after \citealt{tai2015improved}).
-\item The intermediate states of the stack and buffer during parsing are stored during training.
+\item The intermediate states of the stack and buffer during parsing are stored during the processing of each sentence at training time.
 This makes it possible to learn the parameters of the composition function and the input word representations that are fed into the buffer using backpropagation on an objective function that depends on the final sentence representation that emerges at the head of the stack.
 \item A novel tracking LSTM component is added to improve the model's ability to efficiently track the context in which each step is being performed.
-\item A variant, PINN-GT uses ground truth parse information at training and test time, and omits the prediction layer.
+\item The model can construct sentence representations on sentences which have already been parsed by an external parser.
+In this setting, which we call PINN-Ext, the model takes a sequence of transitions (\shift s or \reduce s) as an additional input.
+
 \end{itemize}
 
-\input{../model0_fig.tex}
+This subsection describes the core functions and representations of the PINN.
 
@@
 
-\paragraph{The representations in the stack}
+\paragraph{The stack and buffer}
 
@@
 
-Our composition function is based closely on the TreeLSTM of \citealt{tai2015improved}.
+Our composition function is based on the TreeLSTM of \citealt{tai2015improved}.
 The TreeLSTM generalizes the LSTM neural network layer \cite{hochreiter1997long} to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a two-part vector representation containing an $\vec{h}$ vector and a $\vec{c}$ vector, where the latter is meant to serve as a slower-changing long-term memory.
-
-The precise formulation that we use is as follows.
+It is formulated as:
 
@@
 \\
-\vec{c} = \vec{f}_l * \vec{c}_s^2 + \vec{f}_r * \vec{c}_s^1 + \vec{i} * \vec{g}  
+\vec{c} = \vec{f}_l * \vec{c}_s^{\,2} + \vec{f}_r * \vec{c}_s^{\,1} + \vec{i} * \vec{g}  
 \\
@@
 The results of this function are the pair $(\vec{h}, \vec{c})$, which are placed back on the stack.
-The two input tree nodes popped off the stack are represented as the pairs $(\vec{h}^1_s, \vec{c}^1_s)$ and $(\vec{h}^2_s, \vec{c}^2_s)$.
+The two input tree nodes popped off the stack are represented as the pairs $(\vec{h}^1_s, \vec{c}^{\,1}_s)$ and $(\vec{h}^2_s, \vec{c}^{\,2}_s)$.
 In addition, $\vec{e}$ is an optional input argument which is either the empty vector $[]$ or a vector from an external source like the tracking LSTM (see below).
@@
 
-\subsection{Tracking left context with an auxiliary LSTM}
+\subsection{Tracking left context with an auxiliary LSTM} \todo{[SB] Recheck.}
 
@@
 
-\todo{[SB] Fill in math, Explain.}
+The model described so far is the PINN-Ext model, which uses sequences of \shift~and \reduce~operations from an external parser.
+To build the fully independent PINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
+
+\begin{equation}
+
+$\vec{p}_{trans}$ = \text{softmax}(M_{trans}\vec{h}_{track})
+
+\end{equation}
+
+\todo{[SB] Objective function, soft v. hard max.}
 
@@
 \result{300d LSTM RNN      } & ?             & --                &   ?      &   81?       \\
-\result{300d PINN-GT w/o tracking}   
+\result{300d PINN-Ext w/o tracking}   
                         & ?                  & --                &   ?      &   81?       \\
-\result{300d PINN-GT }
+\result{300d PINN-Ext }
                         & ?                  & --                &   ?      &   83?       \\
@@
                         & ?                  & ??                  &   ?      &   83?       \\                        
-\result{300d PINN-GT, sequence-based attn.   }     
+\result{300d PINN-Ext, sequence-based attn.   }     
                         & ?                  & --                &   ?      &   85+?       \\
-\result{300d PINN-GT, tree-based attn. }           
+\result{300d PINN-Ext, tree-based attn. }           
                         & ?                  & --                &   ?      &   \textbf{?}\\
@@
 Comparing model step time to the plain RNN of \citealt{li2015tree}. We use the small \citealt{pang2004sentimental} sentiment corpus that they use, and train with identical hyperparameters: ...
+
+\todo{Comment on how long top-quality stand-alone greedy parsers take, for comparison.}
 

### 2016-03-12 metal_stack/test.cc
pass thin-stack by reference in composition helper. what a horrible bug

@@ -53,7 +53,7 @@ static inline void assert_matrices_equal(const float *m1, const float *m2,
   free(h_m1);
 }
 
-static float *compose(float *dst, ThinStack ts, const float *l,
+static float *compose(float *dst, ThinStack& ts, const float *l,
     const float *r) {
   // W_l l
   float alpha = 1.0f, beta = 0.0f;

### 2016-03-12 metal_stack/test.cc
Expand test seq length to 5; add safe dummy transitions for short model

@@ -79,7 +79,7 @@ class ThinStackTest : public ::testing::Test {
     ThinStack ts;
 
     ThinStackTest() :
-      spec({5, 5, 2, 10, 3, 5}),
+      spec({5, 5, 2, 10, 5, 5}),
       ts(make_stack(spec)) {
 
       fill_rand_matrix(ts.X, spec.model_dim, spec.seq_length * spec.batch_size);
@@ -99,7 +99,9 @@ TEST_F(ThinStackTest, ShiftShiftMerge) {
   float h_transitions[] = {
     0.0f, 0.0f,
     0.0f, 0.0f,
-    1.0f, 1.0f
+    1.0f, 1.0f,
+    0.0f, 0.0f, // DUMMY
+    0.0f, 0.0f, // DUMMY
   };
   cublasSetVector(spec.seq_length * spec.batch_size, sizeof(float),
       h_transitions, 1, ts.transitions, 1);

### 2016-03-12 metal_stack/test.cc
Remove debugprint.

@@ -118,8 +118,6 @@ TEST_F(ThinStackTest, ShiftShiftMerge) {
   float *right_child = &ts.X[spec.model_dim * spec.batch_size];
   compose(expected, ts, left_child, right_child);
 
-  print_device_matrix(ts.stack, spec.model_dim, spec.batch_size * 3);
-  print_device_matrix(expected, spec.model_dim, spec.batch_size);
   float *output = &ts.stack[2 * spec.model_dim * spec.batch_size];
   assert_matrices_equal(output, expected, spec.model_dim, spec.batch_size);
 

### 2016-03-13 writing/hard_stack_paper/paper.tex
Drop attention from paper.

@@
 This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
-We evaluate it on the Stanford NLI entailment task, show that it outperforms other sentence encoding models, and show that it can be integrated effectively with soft attention for sentence pair tasks like entailment for further gains in performance.
+We evaluate it on the Stanford NLI entailment task, and show that it outperforms other sentence encoding models.
 \todo{[SB] Revise according to comments.}
@@
 
-\todo{[SB] Introduce attention here if appropriate.}
-
-\textcolor{gray}{
-\begin{itemize}
-\item The technique of soft attention has proven to be exceptionally effective as a way of training neural network models for tasks like translation \cite{bahdanau2014neural,luong-pham-manning:2015:EMNLP} and natural language inference \cite{rocktaschel2015reasoning}.
-\item This paper proposes two complementary extensions of the idea of attention to tree-structured models:
-\begin{itemize}
-\item Constituent-by-constituent attention, in which each node in the tree of one sentence is given a soft alignment to some node or nodes in a second sentence.
-\item The mTreeLSTM, which uses a third tree structured model to accumulate information about the alignment between two sentences into a single feature vector.
-This builds on the notion of a matching LSTM or mLSTM from \citealt{DBLP:journals/corr/WangJ15b}.
-\end{itemize}
-\end{itemize}
-}
-
 The paper is structured as follows: We first introduce the PINN, then demonstrate its application to the task of natural language inference.
-We then discuss applications of neural attention to the PINN.
-Finally, we present experimental results and discussion.
+We then present experimental results and discussion.
+\todo{[SB] Update.}
 
@@
 \begin{equation}
-
-$\vec{p}_{trans}$ = \text{softmax}(M_{trans}\vec{h}_{track})
-
+\vec{p}_{trans} = \text{softmax}(M_{trans}\vec{h}_{track})
 \end{equation}
@@
 
-\section{Attention over trees}
-
-\input{../tree_attn_fig.tex}
-
-\paragraph{Constituent-by-constituent attention}
-We aim to build on the results of \citealt{rocktaschel2015reasoning} and \citealt{wang2015learning}, who find that neural soft attention models is an extremely effective technique for learning natural language inference.
-In particular, both papers use versions of word-by-word entailment, in which a latent alignment is learned from every word in the hypothesis to one or more words of the premise.
-We propose to borrow this basic idea, but to adapt it to a tree-structured setting, proposing \textit{constituent-by-constituent} attention.
-While these models do attention over a matrix $\mathbf{Y}$ of word-in-context representations from the premise encoder, we will perform attention instead over our own primary data structure, $\mathbf{Y}^{st}$, the matrix of vectors that have appeared at the top of the stack during premise encoding, which correspond one-to-one to the constituents in the tree structure representing the premise.
-Similarly, while the previous models perform one instance of soft attention conditioned on each word in the hypothesis, we perform one instance of soft attention conditioned on each stack top in the hypothesis encoder, representing the constituents of the hypothesis tree.
-
-In the PINN, attention is performed at each step $t$ of the premise encoder. At step $t$, the query vector that drives attention will be $S^t_0$, the top of the stack.
-
-\todo{[AR, RG] Write up the attention equations that you use.}
-
-\subsection{Accumulating the results of attention}
-
-\todo{[RG,AR] Accumulating results sequentially  (After Rocktäschel, Wang and Jiang)}
-
-\todo{[RG,SB] Accumulating results using a second tree layer}
-
 \section{Experiments}
@@
 \item The transition-based sentence model with the tracking LSTM.
-\item The transition-based sentence model with the tracking LSTM and sequence-based attention over premise states.
-\item The transition-based sentence model with the tracking LSTM and tree-based attention over premise states.
 \end{itemize}
@@
 \midrule
-\multicolumn{4}{c}{Previous NN results without attention}\\
+\multicolumn{4}{c}{Previous sentence encoder-based NN results}\\
 \midrule
@@
 \midrule
-\multicolumn{4}{c}{Previous NN results with attention}\\
+\multicolumn{4}{c}{Other previous NN results}\\
 \midrule
@@
 300d LSTMN with deep attention fusion \cite{cheng2016long}
-                        & 1.4m               & --                &   92.3   &   89.1      \\
+                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
 \midrule
@@
 \midrule
-\result{300d LSTM RNN      } & ?             & --                &   ?      &   81?       \\
+\result{300d LSTM RNN      } & ?             & --                &   83.9      &   80.6       \\
 \result{300d PINN-Ext w/o tracking}   
-                        & ?                  & --                &   ?      &   81?       \\
+                        & ?                  & --                &   ?      &   81.9dv       \\
 \result{300d PINN-Ext }
-                        & ?                  & --                &   ?      &   83?       \\
+                        & ?                  & --                &   ?      &   \underline{84.8dv}       \\
 \result{300d PINN }
-                        & ?                  & ??                  &   ?      &   83?       \\                        
-\result{300d PINN-Ext, sequence-based attn.   }     
-                        & ?                  & --                &   ?      &   85+?       \\
-\result{300d PINN-Ext, tree-based attn. }           
-                        & ?                  & --                &   ?      &   \textbf{?}\\
+                        & ?                  & ??                  &   ?    &   82.6+dv       \\          
+% \result{300d PINN-Ext, sequence-based attn.   }     
+%                         & ?                  & --                &   ?      &   86.3dv         \\
+% \result{300d PINN-Ext, tree-based attn. }           
+%                         & ?                  & --                &   ?      &   \textbf{?}\\
     \bottomrule
@@
 Train and test are SNLI classification accuracy.
-\result{[JG]Add seconds-per-step *if* we get a reasonably well-optimized model in time.}}
+\result{[JG]Add seconds-per-step *if* we get a reasonably well-optimized model in time.} \todo{[SB]: Estimate param counts.}}
 \end{table*}
@@
 \item Tree structure helps in large-scale language learning tasks. Not massively, but measurably.
-\item The advantages of attention carry over to tree-structured models.
+\item An external parser is not necessary.
 \end{itemize}

### 2016-03-13 writing/hard_stack_paper/paper.tex
Mid-afternoon paper update.

@@
 
-A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \cite{socher2011semi,sutskever2014sequence} \todo{[SB] Cite Oriol's conversation paper.}.
+A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \cite{socher2011semi,sutskever2014sequence,vinyals2015neural}.
 This component, the sentence encoder, is generally formulated as a learned parametric function from word vectors to sentence vectors, and this function can take a range of different forms.
@@
 
-The paper is structured as follows: We first introduce the PINN, then demonstrate its application to the task of natural language inference.
-We then present experimental results and discussion.
-\todo{[SB] Update.}
-
 \input{../model1_fig.tex}
@@
 
-\todo{[SB, JG] Brief into to transition-based parsing.}
+\todo{[JG, SB] Brief into to transition-based parsing.}
 
@@
 
-\subsection{Tracking left context with an auxiliary LSTM} \todo{[SB] Recheck.}
+\subsection{Tracking left context with an auxiliary LSTM}
 
-Lexical ambiguity is omnipresent in natural language.
+While it is possible to use the model that we have described thus far as a sentence encoder, at least when an external parser is available to supply \shift~and \reduce~instructions (and we do evaluate such a model), our full model includes an additional component: the tracking LSTM.
+The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which is what allows the model to stand alone as a parser, and it additionally supplies a secondary input to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
+
+Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the two two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
+Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function.
+
+\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) the PINN will use needs to take as its input some representation of the current state of the model.
+Because the tracking LSTM directly takes the buffer top as input, it can directly inform that function of what word is under consideration for a possible \shift.
+Because the tracking LSTM sees the top of the stack at every step, and information only enters the stack by being at the top at some step, the tracking LSTM is able to supply that function with a summary of the state of the stack without having to iterate over the full stack at each step.
+
+\paragraph{Tracking left context for interpretation} Lexical ambiguity is omnipresent in natural language.
 Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
@@
 In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is strictly less than the sequence model has access to, and is often just a single additional word.
-
-In addition to our basic model, we additionally propose an extension that adds to the model (and exploits its sequential structure) by building in a parallel sequence-based model to track left context.
-This auxiliary model, called the tracking LSTM, maintains a low dimensional representation that is updated after every transition, thereby maintaining a representation of the left context at each time point.
-That representation is fed into the composition function as an additional input to allow for more efficient disambiguation.
-As long as the representations produced by the tracking LSTM are substantially smaller than those produced by the main composition function, the model is heavily biased towards computing sentence meaning compositionally according to the assigned tree structure, and is simply given the ability to do so with a better awareness of context.
-
-The tracking LSTM is a standard LSTM RNN.
-At each step it takes three vectors as inputs: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the two two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
-Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step.
+Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
 
@@
 
-\todo{[SB] Objective function, soft v. hard max.}
+The use of a probabilistic classifier is simply a convenience for training.
+Transition decisions are hard, and only a single transition is chosen at each step (i.e., it is not well defined in our model to perform only three-quarters of a \shift operation).
+
+We train the model to mimic the decisions of an external parser.
+In particular, we use the transition sequences corresponding to the Stanford Parser parses included with the SNLI corpus.
+At training time, the classifier is trained using a cross-entropy objective using a supervision signal from those parses, and we use those external parses to guide the behavior of the model.
+At test time, a transition is chosen at each step by taking a hard max over the output of this function.
+We did not find the technique of scheduled sampling \cite{bengio2015scheduled}, or allowing the model to use its own transition decisions in some instances at training time, to be helpful.
 
@@
 
+\todo{[SB] Describe metrics, esp parsing.}
+
 \begin{table*}[t]
@@
 \result{300d PINN }
-                        & ?                  & ??                  &   ?    &   82.6+dv       \\          
+                        & ?                  & 92.2+dv                  &   ?    &   83.3+dv       \\          
 % \result{300d PINN-Ext, sequence-based attn.   }     
@@
 
+\todo{Contrast sentence encoders w/ other models}
+
 \result{Patterns in the examples that the tree model got right that the LSTM didn't}
-
-\result{Induced alignments}
-
-\begin{figure}
-\vspace{14em}
-\caption{\label{fig:induced-alignment}The alignment induced between two sentences by \result{one of our models.}.
-Compare with a similar alignment presented by \protect\citealt{DBLP:journals/corr/WangJ15b}.}
-\end{figure}
 
@@
 
-\paragraph{Future work} The tracking LSTM that we incorporated into the PINN uses only simple, quick to compute features drawn from the head of the buffer and the head of the stack.
-It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors in the sentence string and its likely position in the final tree, and that that context information could support better sentence encoding.
+\paragraph{Future work} Because our aim for this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness of the SNLI task.
+However, the choice of whether to use soft attention and the choice of what representations to perform attention over are orthogonal, and we expect that it should be possible to productively combine our model with soft attention to yield state-of-the-art performance.
+
+The tracking LSTM that we incorporated into the PINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
+It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors in the sentence string and its likely position in the final tree, and that that context information could support both better parsing and better sentence encoding.
 One promising way to pursue this goal would be to encode the full contents of the stack and buffer at each time step following the method used by \citealt{dyer-EtAl:2015:ACL-IJCNLP} for parsing.
-The model presented here is only capable of operating over sentences which have already been assigned a binary constituency tree by a parser.
-This is not necessarily a practical limitation, since modern parsers can assign reasonably high quality parses in so little time as to not dramatically impact the performance characteristics of our model \todo{[SB,JG,CM] How fast are fast constit parsers?}.
-However, depending upon an external parser can be a missed opportunity.
 
-\todo{[SB] Remove Model 1 references.}
-
-In future work, we aim to incorporate a parser into our model, allowing it to simultaneously learn to parse sentences (using a conventional parsing objective function) and to use those parses to learn semantic representations for those sentences (using a classification objective function, as in this work).
 Implementing such a model on top of our proposed stack-based design would learn a model that could productively share representations between the two tasks.

### 2016-03-13 metal_stack/testrun.cc
Exact comparison with drsv

@@ -13,10 +13,11 @@ ThinStackParameters load_params(ModelSpec spec) {
       spec.model_dim * spec.model_dim);
   float *compose_W_r = load_weights_cuda("params/compose_W_r.txt",
       spec.model_dim * spec.model_dim);
+  float *compose_b = load_weights_cuda("params/compose_b.txt", spec.model_dim);
 
   ThinStackParameters ret = {
     NULL, NULL, NULL, // tracking
-    compose_W_l, compose_W_r, NULL, NULL // composition
+    compose_W_l, compose_W_r, NULL, compose_b, // composition
   };
 
   return ret;
@@ -25,10 +26,11 @@ ThinStackParameters load_params(ModelSpec spec) {
 void destroy_params(ThinStackParameters params) {
   cudaFree(params.compose_W_l);
   cudaFree(params.compose_W_r);
+  cudaFree(params.compose_b);
 }
 
 int main() {
-  ModelSpec spec = {5, 5, 2, 10, 3, 5};//{300, 300, 256, 10, 25, 300};//{5, 5, 2, 10, 3, 5};
+  ModelSpec spec = {50, 50, 1, 10, 71, 50};//{300, 300, 256, 10, 25, 300};//{5, 5, 2, 10, 3, 5};
   ThinStackParameters params = load_params(spec);
 
   cublasHandle_t handle;
@@ -42,29 +44,16 @@ int main() {
 
   // Set model inputs.
   cout << "X:" << endl;
-  fill_rand_matrix(ts.X, spec.model_dim, spec.batch_size * spec.seq_length);
-#if DEBUG
-  print_device_matrix(ts.X, spec.model_dim, spec.batch_size * spec.seq_length);
-#endif
+  int num_tokens = (spec.seq_length + 1) / 2;
+  load_weights_cuda("/scr/jgauthie/projects/rembed/metal_stack/sst_one_buffer.txt", spec.model_dim * spec.batch_size * num_tokens, ts.X);
+/* #if DEBUG */
+/*   print_device_matrix(ts.X, spec.model_dim, spec.batch_size * num_tokens); */
+/* #endif */
 
   cout << "transitions:" << endl;
-  float *h_transitions = (float *) malloc(spec.seq_length * spec.batch_size * sizeof(float));
-  for (int i = 0; i < spec.seq_length * spec.batch_size; i++) {
-    float val;
-    if (i < spec.batch_size * 2) {
-      val = 0.0f;
[…]

### 2016-03-13 metal_stack/thin-stack.cc

@@ -105,6 +105,7 @@ void ThinStack::forward() {
   // TODO embedding projection
   buffer = X;
   reset();
+  cudaDeviceSynchronize();
 
   for (int t = 0; t < spec.seq_length; t++) {
     step(t);
@@ -206,6 +207,14 @@ void ThinStack::step(int t) {
 
 void ThinStack::recurrence(const float *stack_1_t, const float *stack_2_t,
     const float *buffer_top_t) {
+
+#if DEBUG
+  cout << "left child:" << endl;
+  print_device_matrix(stack_2_t, spec.model_dim, spec.batch_size);
+
+  cout << "right child:" << endl;
+  print_device_matrix(stack_1_t, spec.model_dim, spec.batch_size);
+#endif
 
   // merge_out = W_l l
   float alpha = 1.0f;

### 2016-03-13 metal_stack/util.cc

@@ -14,14 +14,18 @@ float *load_weights(string filename, int N) {
   return ret;
 }
 
-float *load_weights_cuda(string filename, int N) {
+float *load_weights_cuda(string filename, int N, float *target) {
   float *h_weights = load_weights(filename, N);
-  float *d_weights;
-  cudaMalloc(&d_weights, N * sizeof(float));
-  cudaMemcpy(d_weights, h_weights, N * sizeof(float),
-      cudaMemcpyHostToDevice);
+
+  if (!target) {
+    float *d_weights;
+    cudaMalloc(&d_weights, N * sizeof(float));
+    target = d_weights;
+  }
+  cudaMemcpy(target, h_weights, N * sizeof(float),
+        cudaMemcpyHostToDevice);
   free(h_weights);
-  return d_weights;
+  return target;
 }
 
 
@@ -50,7 +54,7 @@ void print_device_matrix(const float *m, int M, int N) {
 
     for (int j = 0; j < N; j++) {
       float val = h_m[j * M + i];
-      printf(" %+.03f, ", val);
+      printf(" %+.05f, ", val);
     }
 
     cout << " ],";

### 2016-03-13 metal_stack/util.h

@@ -12,7 +12,7 @@
 using namespace std;
 
 
-#define DEBUG 0
+#define DEBUG 1
 
 typedef struct ModelSpec {
   size_t model_dim;
@@ -25,7 +25,7 @@ typedef struct ModelSpec {
 
 
 float *load_weights(string filename, int N);
-float *load_weights_cuda(string filename, int N);
+float *load_weights_cuda(string filename, int N, float *target=NULL);
 
 cublasHandle_t getCublasHandle();
 

### 2016-03-13 metal_stack/thin-stack.cc
Remove unnecessary sync

@@ -109,7 +109,6 @@ void ThinStack::forward() {
 
   for (int t = 0; t < spec.seq_length; t++) {
     step(t);
-    cudaDeviceSynchronize();
 #if DEBUG
     cout << endl << "======================" << endl << endl;
 #endif
@@ -130,7 +129,6 @@ void ThinStack::forward() {
 void ThinStack::step(int t) {
 
   // TODO sync after kernel calls.
-  cout << t * spec.batch_size << endl;
   float *transitions_t = &transitions[t * spec.batch_size];
 #if DEBUG
   cout << "transitions " << t << endl;

### 2016-03-13 writing/hard_stack_paper/paper.tex
More paper content.

@@
 
-This feature vector is then passed to a series of two (\todo{Update w/ final number}) ReLU neural network layers, then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+This feature vector is then passed to a series of two (\todo{Update w/ final number}) ReLU neural network layers (following \citealt{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
 
@@
 Transition decisions are hard, and only a single transition is chosen at each step (i.e., it is not well defined in our model to perform only three-quarters of a \shift operation).
+Because of this, the transition prediction function is trained solely using a parsing-related objective, and cannot be directly trained to optimize for the downstream semantic objective, even though it uses the same word and phrase representations that are used for that downstream objective.
+We leave the project of true joint learning---learning syntactic parsing from a semantic objective---to future work.
 
@@
 
-We evaluate five models on the SNLI task. Each uses 300d hidden states: \todo{[SB] Reframe around M1.}
+We evaluate four models on the SNLI task. Each uses 300d hidden states:
 \begin{itemize}
 \item A baseline LSTM model (similar to that of \citealt{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM RNN sequence model.
-\item The unaugmented transition-based sentence model.
-\item The transition-based sentence model with the tracking LSTM.
+\item The minimal PINN-Ext model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition. This model is strictly equivalent to a TreeLSTM.
+\item The full PINN-Ext model, which adds in the use of a tracking LSTM to guide composition, making it a tree-sequence hybrid.
+\item The full PINN model, which makes its own parsing decisions, and does not depend on an external parser at test time, making it slightly weaker but a practical choice in applied settings.
 \end{itemize}
 
-\todo{[SB] Describe metrics, esp parsing.}
+We compare our models against a range of baselines, including the strongest published non-neural network-based result from \cite{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
+We should note that these attention-based models are by far the strongest published on SNLI to date, but we do not wish to compare our model with those models directly.
+SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
+We follow that approach, and focus on sentence encoding, rather than on attentional models, which are much more narrowly adapted to the sentence-pair structure of SNLI examples.
 
@@
 \midrule
-\multicolumn{4}{c}{Previous non-NN results}\\
+\multicolumn{5}{c}{Previous non-NN results}\\
 \midrule
@@
 \midrule
-\multicolumn{4}{c}{Previous sentence encoder-based NN results}\\
+\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
 \midrule
@@
 \midrule
-\multicolumn{4}{c}{Other previous NN results}\\
+\multicolumn{5}{c}{Our results}\\
 \midrule
-100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
-                        & 252k               & --              &   85.3   &   83.5       \\
-300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
-                        & 1.9m               & --             &   92.0   &   86.1      \\
-300d LSTMN with deep attention fusion \cite{cheng2016long}
-                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
-\midrule
-\multicolumn{4}{c}{New results}\\
-\midrule
-\result{300d LSTM RNN } & ?                  & --                &   83.9      &   80.6       \\
+300d LSTM RNN encoders          & 1.7m                  & --                &   83.9      &   80.6       \\
 \result{300d PINN-Ext w/o tracking}   
@@
 %                         & ?                  & --                &   ?      &   \textbf{?}\\
+\midrule
+\multicolumn{5}{c}{Other previous NN results}\\
+\midrule
+100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
+                        & 252k               & --              &   85.3   &   83.5       \\
+300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
+                        & 1.9m               & --             &   92.0   &   86.1      \\
+300d LSTMN with deep attention fusion \cite{cheng2016long}
+                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
     \bottomrule
@@
 Train and test are SNLI classification accuracy.
-\result{[JG]Add seconds-per-step *if* we get a reasonably well-optimized model in time.} \todo{[SB]: Estimate param counts.}}
+\todo{[SB]: Estimate param counts as numbers come in.}}
 \end{table*}
 
-\paragraph{Results}
+\paragraph{Results} Table~\ref{tab:results} shows our results.
+Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
+For the full PINN model, we also report a measure of the agreement of its generated parses with the Stanford Parser's parses of the same data, expressed in terms of the model's accuracy in making two-way transition decisions.
+In addition, we report the number of trained parameters in each model as well as training set accuracy.
 
-Table~\ref{tab:results} shows our results.
+We find that the bare PINN-Ext model performs relatively poorly, but that the full PINN-Ext with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models, indicating that our tree-sequence hybrid approach succeeds in combining the strengths of both approaches to encoding.
+The fully independent PINN model with its relatively weak internal parser performs more poorly, but nonetheless exceeds the performance of the LSTM baseline.
+None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task.
 
@@
 
-Implementing such a model on top of our proposed stack-based design would learn a model that could productively share representations between the two tasks.
-Implementing such a model on top of a differentiable version of our stack-based design (possibly following \citealt{grefenstette2015learning,joulin2015inferring}) would further make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing  learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+For a more ambitious goal, we expect that it should be possible to implement a variant of the PINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citealt{grefenstette2015learning,joulin2015inferring}) would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 

### 2016-03-13 writing/hard_stack_paper/paper.tex
More paper content.

@@
 300d LSTM RNN encoders          & 1.7m                  & --                &   83.9      &   80.6       \\
-\result{300d PINN-Ext w/o tracking}   
+\result{300d PINN-Ext-NoTracking}   
                         & ?                  & --                &   ?      &   81.9dv       \\
@@
 
-\todo{[JG,SB] Does anyone know of a better baseline TreeRNN implementation that we can use on Jagupard?
-We can use the CoreNLP SST model, but using a Java model as a baseline seems worrying unless we're guaranteed that it's competitively fast.}
+\todo{[JG] Write me!}
 
-Comparing model step time to the plain RNN of \citealt{li2015tree}. We use the small \citealt{pang2004sentimental} sentiment corpus that they use, and train with identical hyperparameters: ...
-
-\todo{Comment on how long top-quality stand-alone greedy parsers take, for comparison.}
+\vspace{10em}
 
@@
 
-\todo{Contrast sentence encoders w/ other models}
+\todo{[SB] Contrast sentence encoders w/ other models}
 
-\result{Patterns in the examples that the tree model got right that the LSTM didn't}
+\result{[SB] Patterns in the examples that the tree model got right that the LSTM didn't}
 
@@
 
+We find that:
 \begin{itemize}
-\item It is possible to train a tree-structured model, without approximation, on large-scale language learning tasks.
-\item Tree structure helps in large-scale language learning tasks. Not massively, but measurably.
-\item An external parser is not necessary.
+\item Our PINN-Ext-NoTracking model precisely implements tree-structured compositionality (as in a TreeLSTM), but is capable of exploiting batched computation for dramatic \todo{[JG] Numbers} improvements in inference speed.
+\item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our tree-sequence hybrid model (PINN-Ext) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.
+\item It is possible to remove the dependence on an external parser from our model ...
 \end{itemize}

### 2016-03-13 writing/hard_stack_paper/paper.tex
shift-reduce -> push-merge

@@
 
-\newcommand{\shift}{\textsc{shift}}
-\newcommand{\reduce}{\textsc{reduce}}
+\newcommand{\shift}{\textsc{push}}
+\newcommand{\reduce}{\textsc{merge}}
 
@@
 \paragraph{The size of the stack}
-The size of the stack should be $N$ for sentences of $N$ words, in case the first reduce merges the final two words. The size of the buffer should be $N$.
+The size of the stack should be $N$ for sentences of $N$ words, in case the first \reduce~operation merges the final two words. The size of the buffer should be $N$.
 
@@
 When a transition sequence is cropped, the number of removed \shift~transitions is tracked, and an equal number of word tokens is removed from the tail (left) end of the buffer.
-This makes it possible for a transition sequence to begin with a \reduce~transition, or to otherwise use a \reduce~transition on a stack that does not have two elements to reduce.
+This makes it possible for a transition sequence to begin with a \reduce~transition, or to otherwise use a \reduce~transition on a stack that does not have two elements to merge.
 In this instance, we simply feed the composition function one or more zero vectors, which represent nonexistent stack elements.

### 2016-03-14 writing/hard_stack_paper/paper.tex
Pausing further paper work. Switching to eval stuff.

@@
 \midrule
-300d LSTM RNN encoders          & 1.7m                  & --                &   83.9      &   80.6       \\
+300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
 \result{300d PINN-Ext-NoTracking}   
@@
 
+\todo{Note that the full PINN (i.e. Model 1) is in the same complexity class as PINN-Ext (what we're evaluating here) with very similar constants.}
+
 \vspace{10em}
@@
 
-\todo{[SB] Contrast sentence encoders w/ other models}
+\result{[SB] Patterns in the examples that the tree model got right that the LSTM didn't}
 
-\result{[SB] Patterns in the examples that the tree model got right that the LSTM didn't}
+\vspace{10em}
 
@@
 \item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our tree-sequence hybrid model (PINN-Ext) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.
-\item It is possible to remove the dependence on an external parser from our model ...
+\item It is possible to remove the dependence on an external parser from our model and replace it with a fast model-internal parser (as in the full PINN) with only a small decrease in model performance.
 \end{itemize}

### 2016-03-14 writing/hard_stack_paper/paper.tex
Minor math improvements

@@
     {\vec{o}}
-= \sigma(
+= \sigma\left(
 W_{\text{iffo}}
@@
     {\vec{e}}
-)
+\right)
 \\
 \vec{g}
-= \tanh(
+= \tanh\left(
 W_{g}
@@
     {\vec{e}}
-)
+\right)
 \\
-\vec{c} = \vec{f}_l * \vec{c}_s^{\,2} + \vec{f}_r * \vec{c}_s^{\,1} + \vec{i} * \vec{g}  
+\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
 \\
-\vec{h} = \vec{o} * \vec{c}
+\vec{h} = \vec{o} \odot \vec{c}
 \end{gather}
 
-The results of this function are the pair $(\vec{h}, \vec{c})$, which are placed back on the stack.
-The two input tree nodes popped off the stack are represented as the pairs $(\vec{h}^1_s, \vec{c}^{\,1}_s)$ and $(\vec{h}^2_s, \vec{c}^{\,2}_s)$.
+The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack.
+The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$.
 In addition, $\vec{e}$ is an optional input argument which is either the empty vector $[]$ or a vector from an external source like the tracking LSTM (see below).
-$*$ denotes the elementwise product.
+$\odot$ denotes the elementwise product.
 Each vector-valued variable listed here is of dimension $D$, except $\vec{e}$, which is of dimensions $D_t$, which can be chosen independently.
@@
     {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
-    {\vec{h}_{\text{premise}} * \vec{h}_{\text{hypothesis}}}
+    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
 \end{equation}
@@
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
+\todo{Include figure of post-order traversal sequence made by the tracking LSTM?}
 

### 2016-03-14 writing/hard_stack_paper/paper.tex
Clarify transition prediction section

@@
 To build the fully independent PINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
-
 \begin{equation}
-\vec{p}_{trans} = \text{softmax}(M_{trans}\vec{h}_{track})
+\vec{p}_{trans} = \text{softmax}(W_{trans}\vec{h}_{track})
 \end{equation}
 
-The use of a probabilistic classifier is simply a convenience for training.
-Transition decisions are hard, and only a single transition is chosen at each step (i.e., it is not well defined in our model to perform only three-quarters of a \shift operation).
-Because of this, the transition prediction function is trained solely using a parsing-related objective, and cannot be directly trained to optimize for the downstream semantic objective, even though it uses the same word and phrase representations that are used for that downstream objective.
-We leave the project of true joint learning---learning syntactic parsing from a semantic objective---to future work.
+The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift or \reduce at the current timestep).
+In the full PINN model, we follow whichever transition has a higher predicted probability in this distribution.
+This prediction module is trained to mimic the decisions of an external parser.
+In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus.
+The classifier is trained using a cross-entropy objective:
+\begin{equation}
+  L_{trans} = - \log \vec{p}_{trans}[t^*]
+\end{equation}
+where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.
 
-We train the model to mimic the decisions of an external parser.
-In particular, we use the transition sequences corresponding to the Stanford Parser parses included with the SNLI corpus.
-At training time, the classifier is trained using a cross-entropy objective using a supervision signal from those parses, and we use those external parses to guide the behavior of the model.
+At training time, the model follows the transitions corresponding to the gold parse.
 At test time, a transition is chosen at each step by taking a hard max over the output of this function.

### 2016-03-14 writing/hard_stack_paper/paper.tex
First stab at thin-stack explanation

@@
 \usepackage{framed}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 
@@
 
-\paragraph{Representing the stack in memory} Or, thin stack. And the cursor strategy for the buffer. 
-\todo{[JG] Explain.}
-
-\vspace{8em}
-
 \paragraph{Optimization and hyperparameters}
@@
 \paragraph{Software infrastructure} We will make Theano code available.
+
+\subsection{Efficient stack representation}
+
+A naive implementation of the PINN model would require simulating a stack of size $N$ for each input sentence.
+In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations.
+This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$.
+Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift and \reduce operations only affect the top of the stack).
+
+We designed an alternative minimal-space stack algorithm inspired by the zipper technique \cite{huet1997zipper}.
+For a single input sentence, we represent the stack over the entire feedforward with a $T \times D$ matrix $S$.
+Each row $S_t$ represents the top of the actual stack at timestep $t$.
+We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time.
+Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation.
+It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$.
+Table~\ref{tbl:thin-stack} shows an example run of this algorithm.
+
+\begin{algorithm}
+\caption{The thin-stack algorithm}
+\label{alg:thin-stack}
+\begin{algorithmic}[1]
+  \Function{Step}{bufferTop, op, $t$, $S$, $Q$}
+    \If{op = \shift}
+      \State $S$[$t$] := bufferTop
+    \ElsIf{op = \reduce}
+      \State right := $S$[$Q$.pop()]
+      \State left := $S$[$Q$.pop()]
+      \State $S$[$t$] := \Call{Compose}{left, right}
+    \EndIf
+    \State $Q$.push($t$)
+  \EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+\begin{table}
+\centering
+\begin{tabular}{c|cl}
+  \toprule
+  $t$ & $S$[$t$] & $Q_t$ \\
+  \midrule
+  0 & $a$ & 0 \\
+  1 & $b$ & 0 1 \\
+  2 & $c$ & 0 1 2 \\
+  3 & $M(c, b)$ & 0 3 \\
+  4 & $M(M(c, b), a)$ & 4 \\
+  \bottomrule
+\end{tabular}
+\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce sequence on the input sentence $(a, b, c)$.
+$S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$.
+The last two elements of $Q$ specify the rows $t$ which should be involved in a \reduce~operation at the next timestep.}
+\label{tbl:thin-stack}
+\end{table}
 

### 2016-03-14 writing/hard_stack_paper/paper.tex
Sell thin-stack a bit..

@@
 
+This stack representation requires substantially less space.
+It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation.
+Furthermore, all of the updates to $S$ and $Q$ can be performed in-place on a GPU.
+These features allow this stack model to run efficiently on a GPU.
+We describe speed results in Section~\ref{sec:speed}.
+
 \begin{table}
@@
 
-\subsection{Step time}
+\subsection{Runtime performance}
+\label{sec:speed}
 

### 2016-03-14 writing/hard_stack_paper/paper.tex
Initial tweaks and responses to JG paper edits.

@@
 
-\newcommand{\shift}{\textsc{push}}
-\newcommand{\reduce}{\textsc{merge}}
+\newcommand{\shift}{\textsc{shift}}
+\newcommand{\reduce}{\textsc{reduce}}
 
@@
 
-The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift or \reduce at the current timestep).
+The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep).
 In the full PINN model, we follow whichever transition has a higher predicted probability in this distribution.
@@
 In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations.
-This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$.
-Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift and \reduce operations only affect the top of the stack).
+This implies a total space requirement of $B \times N \times T \times D$, which is prohibitively large for significant batch sizes $B$ or sentence lengths $N$.
+Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).
 
 We designed an alternative minimal-space stack algorithm inspired by the zipper technique \cite{huet1997zipper}.
-For a single input sentence, we represent the stack over the entire feedforward with a $T \times D$ matrix $S$.
-Each row $S_t$ represents the top of the actual stack at timestep $t$.
+For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$.
+\todo{[JG] Either expand or remove references to batch size.} Each row $S_t$ represents the top of the actual stack at timestep $t$.
 We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time.
@@
   2 & $c$ & 0 1 2 \\
-  3 & $M(c, b)$ & 0 3 \\
-  4 & $M(M(c, b), a)$ & 4 \\
+  3 & $(c~b)$ & 0 3 \\
+  4 & $((c~b)~a)$ & 4 \\
   \bottomrule
 \end{tabular}
-\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce sequence on the input sentence $(a, b, c)$.
+\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce~sequence on the input sentence $(a, b, c)$.
 $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$.
@@
 \section{Experiments}
+
+\begin{table*}[t]
+  \centering\small
+  \begin{tabular}{lcccc} 
+    \toprule
+Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
+\midrule
+\multicolumn{5}{c}{Previous non-NN results}\\
+\midrule
+Lexicalized classifier \cite{snli:emnlp2015}
+                        & --                & --                    &   99.7   &   78.2      \\
+\midrule
+\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
+\midrule
+100d LSTM encoders \cite{snli:emnlp2015}
+                        & 221k               & --               &   84.8   &   77.6      \\
+1024d pretrained GRU encoders \cite{DBLP:journals/corr/VendrovKFU15}
+                        & 15m                & --              &   98.8   &   81.4       \\
+300d Tree-based CNN encoders \cite{mou2015recognizing}
+                        & 3.5m                & --             &   83.4   &   82.1       \\
+\midrule
+\multicolumn{5}{c}{Our results}\\
+\midrule
+300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
+\result{300d PINN-Ext-NoTracking}   
+                        & 3.4m                  & --                &   ?      &   81.9dv       \\
+\result{300d PINN-Ext }
+                        & 3.7m                  & --                &   ?      &   \underline{84.8dv}       \\
+\result{300d PINN }
+                        & ??                  & 92.2+dv            &   ?    &   83.3+dv       \\          
+% \result{300d PINN-Ext, sequence-based attn.   }     
+%                         & ?                  & --                &   ?      &   86.3dv         \\
+% \result{300d PINN-Ext, tree-based attn. }           
+%                         & ?                  & --                &   ?      &   \textbf{?}\\
+\midrule
+\multicolumn{5}{c}{Other previous NN results}\\
+\midrule
+100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
+                        & 252k               & --              &   85.3   &   83.5       \\
+300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
+                        & 1.9m               & --             &   92.0   &   86.1      \\
+300d LSTMN with deep attention fusion \cite{cheng2016long}
+                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
+    \bottomrule
+  \end{tabular}
+  \protect\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification.
+Params.
+is the approximate number of trained parameters (excluding word embeddings for models where they are trained).
+Trans.
+acc.
+is the model's accuracy in predicting parsing transitions.
+Train and test are SNLI classification accuracy.}
+\end{table*}
 
@@
 We follow that approach, and focus on sentence encoding, rather than on attentional models, which are much more narrowly adapted to the sentence-pair structure of SNLI examples.
-
-\begin{table*}[t]
-  \centering\small
-  \begin{tabular}{lcccc} 
-    \toprule
-Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
-\midrule
-\multicolumn{5}{c}{Previous non-NN results}\\
-\midrule
-Lexicalized classifier \cite{snli:emnlp2015}
-                        & --                & --                    &   99.7   &   78.2      \\
-\midrule
-\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
-\midrule
-100d LSTM encoders \cite{snli:emnlp2015}
-                        & 221k               & --               &   84.8   &   77.6      \\
-1024d pretrained GRU encoders \cite{DBLP:journals/corr/VendrovKFU15}
-                        & 15m                & --              &   98.8   &   81.4       \\
-300d Tree-based CNN encoders \cite{mou2015recognizing}
-                        & 3.5m                & --             &   83.4   &   82.1       \\
-\midrule
-\multicolumn{5}{c}{Our results}\\
-\midrule
-300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
-\result{300d PINN-Ext-NoTracking}   
-                        & ?                  & --                &   ?      &   81.9dv       \\
-\result{300d PINN-Ext }
-                        & ?                  & --                &   ?      &   \underline{84.8dv}       \\
-\result{300d PINN }
-                        & ?                  & 92.2+dv            &   ?    &   83.3+dv       \\          
-% \result{300d PINN-Ext, sequence-based attn.   }     
-%                         & ?                  & --                &   ?      &   86.3dv         \\
-% \result{300d PINN-Ext, tree-based attn. }           
-%                         & ?                  & --                &   ?      &   \textbf{?}\\
-\midrule
-\multicolumn{5}{c}{Other previous NN results}\\
-\midrule
-100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
-                        & 252k               & --              &   85.3   &   83.5       \\
-300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
-                        & 1.9m               & --             &   92.0   &   86.1      \\
-300d LSTMN with deep attention fusion \cite{cheng2016long}
-                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
-    \bottomrule
-  \end{tabular}
-  \protect\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification.
-Params.
-is the approximate number of trained parameters (excluding word embeddings for models where they are trained).
-Trans.
-acc.
-is the model's accuracy in predicting parsing transitions.
-Train and test are SNLI classification accuracy.
-\todo{[SB]: Estimate param counts as numbers come in.}}
-\end{table*}
 

### 2016-03-14 writing/hard_stack_paper/paper.tex
Slightly revised abstract, still work in progress.

@@
 
-Tree-structured neural networks integrate valuable syntactic parse information into the process of interpreting sentence meaning, but suffer two key technical problems.
-They can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks.
-We address these issues by introducing the Parser-Interpreter Neural Network (PINN), which simulates tree-structured sentence interpretation within the sequential structure of a transition-based neural parser.
+Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences.
+However, they suffer two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks.
+We address these issues by introducing the Parser-Interpreter Neural Network (PINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser.
 This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
-We evaluate it on the Stanford NLI entailment task, and show that it outperforms other sentence encoding models.
-\todo{[SB] Revise according to comments.}
+We evaluate it on the Stanford NLI entailment task, show that it outperforms other sentence encoding models, and show that it can be integrated effectively with soft attention for sentence pair tasks like entailment for further gains in performance.
+
 \end{abstract}
@@
 
-\todo{[JG] Write me!}
+\todo{[JG] Write me! In short: We're an order of magnitude faster than plain C++ CPU TreeRNNs w/ reasonable model sizes.}
 

### 2016-03-14 writing/hard_stack_paper/paper.tex
Add some more detail to The Big Sentence in the introduction

@@
 In this paper, we introduce the Parser--Interpreter Neural Network, or PINN.
-The PINN is a model which reimplements tree-structured composition within an alternative design that incorporates a neural network parser.
+The PINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
 This design improves upon the previous TreeRNN approach to compositionality in three ways:

### 2016-03-14 writing/hard_stack_paper/paper.tex
Try to clarify "Composition and representation" intro

@@
 
-The PINN, which is shown in Figure~\ref{fig:model:1d}, takes the form of a transition-based parser, but it is designed to produce sentence representations given parse structures, rather than producing parse structures as its output, leading to the following differences:
+The full PINN model is shown in Figure~\ref{fig:model:1d}.
+The PINN takes the form of a transition-based parser, but is designed to produce sentence representations given parse structures, rather than producing parse structures as its output.
+It extends the standard transition-based model in the following ways:
 \begin{itemize}
-\item The intermediate representations in the parser, including the representations of the tree structures on the stack, are vectors of a fixed length.
-\item The \reduce~operation, which combines two trees or tree nodes on the stack into a single larger tree, is represented using a parametric neural network layer function (in particular, a TreeLSTM layer, after \citealt{tai2015improved}).
-\item The intermediate states of the stack and buffer during parsing are stored during the processing of each sentence at training time.
-This makes it possible to learn the parameters of the composition function and the input word representations that are fed into the buffer using backpropagation on an objective function that depends on the final sentence representation that emerges at the head of the stack.
-\item A novel tracking LSTM component is added to improve the model's ability to efficiently track the context in which each step is being performed.
-\item The model can construct sentence representations on sentences which have already been parsed by an external parser.
-In this setting, which we call PINN-Ext, the model takes a sequence of transitions (\shift s or \reduce s) as an additional input.
-
+\item The intermediate representations in the parser, including the representations of partial tree structures on the stack, are fixed-length vectors.
+\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
+\item After processing a sentence, the model has a single vector representation of the sentence on its stack.
+This representation is a differentiable function of all prior representations.
+We can thus learn the weights of the model via backpropagation of some downstream error.
 \end{itemize}
@@
 This subsection describes the core functions and representations of the PINN.
+We begin with the most basic model, PINN-Ext, which processes input sentences with gold parses provided by some external parser.
 

### 2016-03-14 writing/hard_stack_paper/paper.tex
Add shift-reduce parsing intro

@@
 
-\section{The transition-based sentence model}
+\section{The shift-reduce sentence model}
 
-\subsection{Intuition: Transition-based parsing}
+\subsection{Shift-reduce parsing}
 
-The PINN is based on a formalism adapted from transition-based parsing (\question{What's a good classic citation for this?}).
+The PINN is inspired by the shift-reduce parsing formalism \cite{aho1986compilers}, which builds a tree structure over a sequence (e.g.
+a natural language sentence) by a single left-to-right scan over the tokens of the sequence.
+Shift-reduce parsing has been successfully applied in algorithms for natural language parsing \cite{nivre2003efficient}.
 
-\todo{[JG, SB] Brief into to transition-based parsing.}
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$.
+It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
+(For a binary-branching tree structure, this requires $T = 2N - 1$ operations.)
 
-\vspace{10em}
+The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed.
+We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$.
+Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state.
+We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations.
+Below $\mid$ denotes an infix \textit{cons} function.
+\begin{description}
+  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
+  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$.
+This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
+\end{description}
 
@@
 
-The full PINN model is shown in Figure~\ref{fig:model:1d}.
-The PINN takes the form of a transition-based parser, but is designed to produce sentence representations given parse structures, rather than producing parse structures as its output.
-It extends the standard transition-based model in the following ways:
+The core PINN model implements a sequential shift-reduce parser which operates over sentences.
+It is designed to produce a vector representation of the sentence as output (as opposed to a tree).
+This is accomplished by the following modifications to the shift-reduce algorithm:
 \begin{itemize}
@@
 \item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
-\item After processing a sentence, the model has a single vector representation of the sentence on its stack.
-This representation is a differentiable function of all prior representations.
-We can thus learn the weights of the model via backpropagation of some downstream error.
 \end{itemize}

### 2016-03-14 writing/hard_stack_paper/paper.tex
Add messy runtime plot

@@
 \usepackage{tikz-qtree}
+\usepackage{pgfplots}
 \usepackage{ifthen}
@@
 
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[xlabel=Batch size, ylabel=Feedforward time (sec),
+			 legend style={anchor=north west}]
+  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
+  \addlegendentry{CPU}
+  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
+  \addlegendentry{Thin-stack GPU}
+\end{axis}
+\end{tikzpicture}
+\caption{Runtime performance results with the thin-stack GPU algorithm for tree-structured neural networks. \todo{[JG] Add some comments, avg points, add larger batch sizes for thin-stack}}
+\end{figure}
+
 \todo{[JG] Write me! In short: We're an order of magnitude faster than plain C++ CPU TreeRNNs w/ reasonable model sizes.}

### 2016-03-14 writing/hard_stack_paper/paper.tex
Improve runtime results report
axis labels

@@
 \begin{figure}
+\resizebox{\linewidth}{!}{
 \begin{tikzpicture}
-\begin{axis}[xlabel=Batch size, ylabel=Feedforward time (sec),
-			 legend style={anchor=north west}]
+\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
+			 xtick={0,256,512,1024,2048}, legend style={at={(1, 1)},anchor=north east}]
   \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
-  \addlegendentry{CPU}
+  \addlegendentry{CPU \cite{irsoy2014deep}}
   \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
@@
 \end{tikzpicture}
-\caption{Runtime performance results with the thin-stack GPU algorithm for tree-structured neural networks. \todo{[JG] Add some comments, avg points, add larger batch sizes for thin-stack}}
+}
+\caption{Runtime performance results with the thin-stack GPU algorithm for tree-structured neural networks. \todo{[JG] Add some comments}}
 \end{figure}

### 2016-03-14 writing/hard_stack_paper/paper.tex
TODO tweaks.

@@
 
+\subsection{Parsing performance}
+
+\todo{[AR] Fill in some details here.}
+\todo{[SB] Read AR's report and work out what to include.}
+
 \section{Discussion}
 
-\result{[SB] Patterns in the examples that the tree model got right that the LSTM didn't}
+\question{What kind of error analysis would be interesting for this model?}
+
+\result{[SB] Compare the types of errors in the different runs.}
 

### 2016-03-14 writing/hard_stack_paper/paper.tex
Model name => SPINN

@@
 
-\title{A unified model for parsing and sentence understanding}
+\title{A fast unified model for parsing and sentence understanding}
 
@@
 However, they suffer two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks.
-We address these issues by introducing the Parser-Interpreter Neural Network (PINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser.
+We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser.
 This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
@@
 \end{abstract}
-
-\question{Parser--Interpreter or Parser-Interpreter (hyphen or en-dash)?}
 
@@
 
-In this paper, we introduce the Parser--Interpreter Neural Network, or PINN.
-The PINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
+In this paper, we introduce the Stack-augmented Parser-Interpreter Neural Network, or SPINN.
+The SPINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
 This design improves upon the previous TreeRNN approach to compositionality in three ways:
@@
 
-The PINN is inspired by the shift-reduce parsing formalism \cite{aho1986compilers}, which builds a tree structure over a sequence (e.g.
+The SPINN is inspired by the shift-reduce parsing formalism \cite{aho1986compilers}, which builds a tree structure over a sequence (e.g.
 a natural language sentence) by a single left-to-right scan over the tokens of the sequence.
@@
 
-The core PINN model implements a sequential shift-reduce parser which operates over sentences.
+The core SPINN model implements a sequential shift-reduce parser which operates over sentences.
 It is designed to produce a vector representation of the sentence as output (as opposed to a tree).
@@
 
-This subsection describes the core functions and representations of the PINN.
-We begin with the most basic model, PINN-Ext, which processes input sentences with gold parses provided by some external parser.
+This subsection describes the core functions and representations of the SPINN.
+We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences with gold parses provided by some external parser.
 
@@
 
-In a PINN, each row of the stack contains a pair of vectors $(h, c)$, which together represent some node in the parse tree of the sentence.
+In a SPINN, each row of the stack contains a pair of vectors $(h, c)$, which together represent some node in the parse tree of the sentence.
 Since the \shift~operation simply copies entries from the buffer onto the stack, this means that the word representations in the buffer also have two parts, $h$ and $c$.
@@
 
-\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) the PINN will use needs to take as its input some representation of the current state of the model.
+\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) the SPINN will use needs to take as its input some representation of the current state of the model.
 Because the tracking LSTM directly takes the buffer top as input, it can directly inform that function of what word is under consideration for a possible \shift.
@@
 
-The model described so far is the PINN-Ext model, which uses sequences of \shift~and \reduce~operations from an external parser.
-To build the fully independent PINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
+The model described so far is the SPINN-NP model, which uses sequences of \shift~and \reduce~operations from an external parser.
+To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
 \begin{equation}
@@
 The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep).
-In the full PINN model, we follow whichever transition has a higher predicted probability in this distribution.
+In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution.
 This prediction module is trained to mimic the decisions of an external parser.
@@
 
-A naive implementation of the PINN model would require simulating a stack of size $N$ for each input sentence.
+A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence.
 In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations.
@@
 
-In its bare form, before the addition of the tracking LSTM, the PINN computes the precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
+In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) computes the precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
 While it adds no expressive power on its own, this design improves upon the standard TreeRNN implementation in three ways:
@@
 In addition, there has been recent work \cite{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
-To our knowledge, the PINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
+To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
 
@@
 300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
-\result{300d PINN-Ext-NoTracking}   
+\result{300d SPINN-NP-NT}   
                         & 3.4m                  & --                &   ?      &   81.9dv       \\
-\result{300d PINN-Ext }
+\result{300d SPINN-NP}
                         & 3.7m                  & --                &   ?      &   \underline{84.8dv}       \\
-\result{300d PINN }
+\result{300d SPINN }
                         & ??                  & 92.2+dv            &   ?    &   83.3+dv       \\          
-% \result{300d PINN-Ext, sequence-based attn.   }     
+% \result{300d SPINN-NP, sequence-based attn.   }     
 %                         & ?                  & --                &   ?      &   86.3dv         \\
-% \result{300d PINN-Ext, tree-based attn. }           
+% \result{300d SPINN-NP, tree-based attn. }           
 %                         & ?                  & --                &   ?      &   \textbf{?}\\
@@
 
-We evaluate the PINN on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
+We evaluate the SPINN on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
 NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus (SNLI, \citealt{snli:emnlp2015}), which we use for training and evaluation:
@@
 \item A baseline LSTM model (similar to that of \citealt{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM RNN sequence model.
-\item The minimal PINN-Ext model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition. This model is strictly equivalent to a TreeLSTM.
-\item The full PINN-Ext model, which adds in the use of a tracking LSTM to guide composition, making it a tree-sequence hybrid.
-\item The full PINN model, which makes its own parsing decisions, and does not depend on an external parser at test time, making it slightly weaker but a practical choice in applied settings.
+\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition.
+This model is strictly equivalent to a TreeLSTM.
+\item The full SPINN-NP model, which adds in the use of a tracking LSTM to guide composition, making it a tree-sequence hybrid.
+\item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time, making it slightly weaker but a practical choice in applied settings.
 \end{itemize}
@@
 Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
-For the full PINN model, we also report a measure of the agreement of its generated parses with the Stanford Parser's parses of the same data, expressed in terms of the model's accuracy in making two-way transition decisions.
+For the full SPINN model, we also report a measure of the agreement of its generated parses with the Stanford Parser's parses of the same data, expressed in terms of the model's accuracy in making two-way transition decisions.
 In addition, we report the number of trained parameters in each model as well as training set accuracy.
 
-We find that the bare PINN-Ext model performs relatively poorly, but that the full PINN-Ext with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models, indicating that our tree-sequence hybrid approach succeeds in combining the strengths of both approaches to encoding.
-The fully independent PINN model with its relatively weak internal parser performs more poorly, but nonetheless exceeds the performance of the LSTM baseline.
+We find that the bare SPINN-NP-NT model performs relatively poorly, but that the SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models, indicating that our tree-sequence hybrid approach succeeds in combining the strengths of both approaches to encoding.
+The fully independent SPINN model with its relatively weak internal parser performs more poorly, but nonetheless exceeds the performance of the LSTM baseline.
 None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task.
@@
 
-\todo{Note that the full PINN (i.e. Model 1) is in the same complexity class as PINN-Ext (what we're evaluating here) with very similar constants.}
+\todo{Note that the full SPINN (i.e. Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
 
@@
 \begin{itemize}
-\item Our PINN-Ext-NoTracking model precisely implements tree-structured compositionality (as in a TreeLSTM), but is capable of exploiting batched computation for dramatic \todo{[JG] Numbers} improvements in inference speed.
-\item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our tree-sequence hybrid model (PINN-Ext) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.
-\item It is possible to remove the dependence on an external parser from our model and replace it with a fast model-internal parser (as in the full PINN) with only a small decrease in model performance.
+\item Our SPINN-NP-NT model precisely implements tree-structured compositionality (as in a TreeLSTM), but is capable of exploiting batched computation for dramatic \todo{[JG] Numbers} improvements in inference speed.
+\item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our base tree-sequence hybrid model (SPINN-NP) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.
+\item It is possible to remove the dependence on an external parser from our model and replace it with a fast model-internal parser (as in the full SPINN) with only a small decrease in model performance.
 \end{itemize}
@@
 
-The tracking LSTM that we incorporated into the PINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
+The tracking LSTM that we incorporated into the SPINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
 It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors in the sentence string and its likely position in the final tree, and that that context information could support both better parsing and better sentence encoding.
@@
 
-For a more ambitious goal, we expect that it should be possible to implement a variant of the PINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citealt{grefenstette2015learning,joulin2015inferring}) would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+For a more ambitious goal, we expect that it should be possible to implement a variant of the SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citealt{grefenstette2015learning,joulin2015inferring}) would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 

### 2016-03-15 writing/hard_stack_paper/paper.tex
Typo / diction fixes; be stronger in the anti-attention-model paragraph

@@
 Even though NLI is framed as a relatively simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
-In order for a model to reliably perform well on NLI across a range of sentence types and text genres, it must be able to represent and reason with all of the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
+In order for a model to perform reliably well on NLI across a range of sentence types and text genres, it must be able to represent and reason with all of the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
 
@@
 \begin{itemize}
-\item A baseline LSTM model (similar to that of \citealt{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM RNN sequence model.
+\item A baseline LSTM model (similar to that of \citealt{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
 \item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition.
-This model is strictly equivalent to a TreeLSTM.
-\item The full SPINN-NP model, which adds in the use of a tracking LSTM to guide composition, making it a tree-sequence hybrid.
+This model is equivalent to a TreeLSTM \cite{tai2015improved}.
+\item The SPINN-NP model, which includes a tracking LSTM which guides composition, making it a hybrid tree-sequence model.
 \item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time, making it slightly weaker but a practical choice in applied settings.
@@
 SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
-We follow that approach, and focus on sentence encoding, rather than on attentional models, which are much more narrowly adapted to the sentence-pair structure of SNLI examples.
+The attentional models are narrowly adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders.
+We focus on models which do support general sentence encoding in our analysis.
 

### 2016-03-15 writing/hard_stack_paper/paper.tex
Incremental update.

--- 2016-03-14 writing/hard_stack_paper/paper.tex
@@
 
-     \author{}
-%    \author{
-%    Samuel R.\ Bowman$^{1,2,5,}$\thanks{The first two authors contributed equally.} \\
-%    \texttt{\small sbowman@stanford.edu} \\
-%    \And
-%    Jon Gauthier$^{2,3,5,*}$ \\
-%    \texttt{\small jgauthie@stanford.edu} \\
-%    \And
-%    Abhinav Rastogi$^{4,5}$ \\
-%    \texttt{\small arastogi@stanford.edu} \\
-%    \AND
-%    Raghav Gupta$^{6}$ \\
-%    \texttt{\small rgupta93@stanford.edu} \\
-%    \And
-%    Christopher D.\ Manning$^{1,2,5,6}$\\
-%    \texttt{\small manning@stanford.edu}\\
-%    \And
-%    Christopher Potts$^{1}$\\
-%    \texttt{\small cgpotts@stanford.edu}
-%    \AND\\[-3ex]
-%    {$^{1}$Stanford Linguistics\quad
-%    $^{2}$Stanford NLP Group\quad
-%    $^{3}$Stanford Symbolic Systems}\\
-%    {$^{4}$Stanford Electrical Engineering\quad
-%    $^{5}$Stanford AI Lab\quad
-%    $^{6}$Stanford Computer Science}
-%    }
+%     \author{}
+    \author{
+    Samuel R.\ Bowman$^{1,2,5,}$\thanks{The first two authors contributed equally.} \\
+    \texttt{\small sbowman@stanford.edu} \\
+    \And
+    Jon Gauthier$^{2,3,5,*}$ \\
+    \texttt{\small jgauthie@stanford.edu} \\
+    \And
+    Abhinav Rastogi$^{4,5}$ \\
+    \texttt{\small arastogi@stanford.edu} \\
+    \AND
+    Raghav Gupta$^{6}$ \\
+    \texttt{\small rgupta93@stanford.edu} \\
+    \And
+    Christopher D.\ Manning$^{1,2,5,6}$\\
+    \texttt{\small manning@stanford.edu}\\
+    \And
+    Christopher Potts$^{1}$\\
+    \texttt{\small cgpotts@stanford.edu}
+    \AND\\[-3ex]
+    {$^{1}$Stanford Linguistics\quad
+    $^{2}$Stanford NLP Group\quad
+    $^{3}$Stanford Symbolic Systems}\\
+    {$^{4}$Stanford Electrical Engineering\quad
+    $^{5}$Stanford AI Lab\quad
+    $^{6}$Stanford Computer Science}
+    }
 
@@
 This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
-We evaluate it on the Stanford NLI entailment task, show that it outperforms other sentence encoding models, and show that it can be integrated effectively with soft attention for sentence pair tasks like entailment for further gains in performance.
+We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
 
@@
 This component, the sentence encoder, is generally formulated as a learned parametric function from word vectors to sentence vectors, and this function can take a range of different forms.
-Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (\citealt{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \cite{kalchbrenner2014convolutional} \todo{Cite LeCun CharConvNet}, which pool information over short localsequences of words, and tree-structured recursive neural networks (TreeRNNs \citealt{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary tree structure over the words.
-Of these, there are tidy theoretical arguments suggesting that the TreeRNN is the principled choice: meaning in natural language sentences is widely believed to be constructed incrementally according to a tree-structure \cite{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time recovering the process by which meaning is constructed.
+Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citealt{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citealt{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citealt{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words.
+Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \cite{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.
 
-In practice, TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons.
-Batched computation--performing synchronized computation across many examples at once--is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets \question{What, if anything, can we cite here?}.
-Batching yields order-of-magnitude improvements in standard CPU implementation, and substantial additional gains on GPUs.
-TreeRNNs build the parse structure of each sentence they process into the structure of the network itself, forcing the model to perform a different sequence of computations for each sentence, and thereby requiring that each sentence be processed one-by-one, making batched computation impossible.
-One aim of this paper is to make these large-scale evaluations possible.
-Beyond even this large handicap, TreeRNNs can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.
+TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons.
+Batched computation---performing synchronized computation across many examples at once---is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets (\question{What, if anything, can we cite here?}).
+Batching yields order-of-magnitude improvements in standard CPU implementations of neural network models, and substantial additional gains on GPUs.
+TreeRNNs build the parse structure of each sentence they process into the structure of the network itself, forcing the model to perform a different sequence of computations for each sentence, and thereby requiring that each sentence be processed one-by-one, making batched computation impossible in standard implementations.
+One aim of this paper is to restructure the computations involved in a TreeRNN in order to enable batching and make these large-scale evaluations possible.
+In addition to their slow speed, TreeRNNs have an additional feature that makes them impractical in many settings: they can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.
+This paper additionally aims to integrate parsing into TreeRNN processing to circumvent this issue.
 
@@
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
-\todo{Include figure of post-order traversal sequence made by the tracking LSTM?}
+
+\todo{[JG] Include figure of post-order traversal sequence made by the tracking LSTM?}
+\todo{[SB] Say more about joint learning of phrase representations for the two objectives.}
+
+\paragraph{The full objective function} \todo{[SB] Fill out. In short, there are two weighted terms in the objective: a semantic term and a syntactic term.}
 
@@
 
-We use the RMSProp optimizer (\todo{Cite.}) with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
+We use the RMSProp optimizer (\todo{[SB] Cite.}) with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
 We apply both dropout \cite{srivastava2014dropout} and batch normalization \cite{2015SIoffeCSzegedy} to the word embeddings in the buffer (after the linear projection is applied) and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
@@
 
-We used random search to tune many of the hyperparameters, including the learning rate, the dropout rate, the L2 regularization weight, the hidden dimension of the tracking LSTM and the number of layers in the MLP \todo{[SB] Report the ranges searched for each parameter and the number of runs for each results.}
+We used random search to tune many of the hyperparameters, including the learning rate, the dropout rate, the L2 regularization weight, the hidden dimension of the tracking LSTM and the number of layers in the MLP \todo{[SB] Report the ranges searched for each parameter and the number of runs for each results.
+Some of this can go in an appendix.}
 
-We trained each model for 250k steps in each run, using a minibatch size of 64 for each step.
+We trained each model for 250k steps in each run, using a minibatch size of 32 for each step.
 We tracked each model's performance on the development set during training and saving parameters when this performance reached a new peak.
@@
 
+\todo{[SB] Discuss relationship between our model and the TBCNN.}
+
 \subsection{Runtime performance}

### 2016-03-15 writing/hard_stack_paper/paper.tex
Edit results section

--- 2016-03-15 writing/hard_stack_paper/paper.tex
@@
 Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
-For the full SPINN model, we also report a measure of the agreement of its generated parses with the Stanford Parser's parses of the same data, expressed in terms of the model's accuracy in making two-way transition decisions.
-In addition, we report the number of trained parameters in each model as well as training set accuracy.
+The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations).
+We also report the number of trained parameters in each model and training set accuracy.
 
-We find that the bare SPINN-NP-NT model performs relatively poorly, but that the SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models, indicating that our tree-sequence hybrid approach succeeds in combining the strengths of both approaches to encoding.
-The fully independent SPINN model with its relatively weak internal parser performs more poorly, but nonetheless exceeds the performance of the LSTM baseline.
+We find that the bare SPINN-NP-NT model performs relatively poorly, but that the SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
+The success of the SPINN-NP model, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
+
+The full SPINN model with its relatively weak internal parser performs more poorly, but nonetheless exceeds the performance of the LSTM baseline.
 None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task.
@@
 \begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
-			 xtick={0,256,512,1024,2048}, legend style={at={(1, 1)},anchor=north east}]
+			 xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
   \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};

### 2016-03-15 writing/hard_stack_paper/paper.tex
Fix bibtex errrors.
Draft mode.

--- 2016-03-15 writing/hard_stack_paper/paper.tex
@@
     \author{
-    Samuel R.\ Bowman$^{1,2,5,}$\thanks{The first two authors contributed equally.} \\
+    Samuel R.\ Bowman$^{1,2,5,}$\thanks{~\,The first two authors contributed equally.} \\
     \texttt{\small sbowman@stanford.edu} \\
@@
 
+\input{../model1_fig.tex}
+
 TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons.
@@
 \end{itemize}
-
-\input{../model1_fig.tex}
 

### 2016-03-15 writing/hard_stack_paper/paper.tex
Hasty pre-NLP Group edits.

@@
 \input{batching_fig.tex}
+\input{../model1_fig.tex}
 
@@
 Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \cite{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.
-
-\input{../model1_fig.tex}
 
@@
 The SPINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
-This design improves upon the previous TreeRNN approach to compositionality in three ways:
+This design improves upon the TreeRNN architecture in three ways:
 \begin{itemize}
@@
 \end{itemize}
+
+\todo{[SB] Move the introduction of the SPINN earlier in the introduction. Mention that we'll be using SNLI as our sole evaluation task.}
 
@@
 It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
-(For a binary-branching tree structure, this requires $T = 2N - 1$ operations.)
+For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.
 
@@
 Below $\mid$ denotes an infix \textit{cons} function.
+\todo{[JG] Clarify prev.
+sentence.}
 \begin{description}
@@
 The core SPINN model implements a sequential shift-reduce parser which operates over sentences.
-It is designed to produce a vector representation of the sentence as output (as opposed to a tree).
-This is accomplished by the following modifications to the shift-reduce algorithm:
+It is designed to produce a vector representation of the sentence as output,as opposed to a tree as in standard shift-reduce \textit{parsing}.
+This relies on two key modifications to the shift-reduce algorithm:
 \begin{itemize}
-\item The intermediate representations in the parser, including the representations of partial tree structures on the stack, are fixed-length vectors.
+\item The parser's intermediate representations, including the representations of partial tree structures on the stack, are fixed-length vectors.
 \item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
@@
 This subsection describes the core functions and representations of the SPINN.
-We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences with gold parses provided by some external parser.
+We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences using parses provided by some external parser.
 
-\paragraph{The buffer and word representations}
+\paragraph{Word representations}
 
-We draw our word representations from the standard 300-dimensional vector package provided with GloVe \cite{pennington2014glove}.
-We do not update these vectors during training.
-Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors that can be used as inputs to a TreeLSTM composition function:
+The buffer is populated with the words in each sentence before the first step of computation.
+We use word representations from the standard 300-dimensional vector package provided with GloVe \cite{pennington2014glove}.
+We do not update these representations during training.
+Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors that are stored side-by-side in the buffer and used as inputs to a TreeLSTM composition function:
 
@@
 
-In a SPINN, each row of the stack contains a pair of vectors $(h, c)$, which together represent some node in the parse tree of the sentence.
-Since the \shift~operation simply copies entries from the buffer onto the stack, this means that the word representations in the buffer also have two parts, $h$ and $c$.
+In the SPINN, each row of the stack and each row of the buffer contains a pair of vectors $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence.
+\todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}
 
@@
 Our composition function is based on the TreeLSTM of \citealt{tai2015improved}.
-The TreeLSTM generalizes the LSTM neural network layer \cite{hochreiter1997long} to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a two-part vector representation containing an $\vec{h}$ vector and a $\vec{c}$ vector, where the latter is meant to serve as a slower-changing long-term memory.
+The TreeLSTM generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a vector pair $\langle \vec{h}, \vec{c}\rangle$, of which the latter vector is meant to serve as a slower-changing long-term memory.
 It is formulated as:
@@
 The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$.
-In addition, $\vec{e}$ is an optional input argument which is either the empty vector $[]$ or a vector from an external source like the tracking LSTM (see below).
+In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see below).
 $\odot$ denotes the elementwise product.
-Each vector-valued variable listed here is of dimension $D$, except $\vec{e}$, which is of dimensions $D_t$, which can be chosen independently.
+Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.
 
@@
 
-This paper presents the transition-based sentence model within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task.
+This paper evaluates the SPINN within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task.
 To classify a sentence pair, a feature vector is first constructed.
@@
 
-This feature vector is then passed to a series of two (\todo{Update w/ final number}) ReLU neural network layers (following \citealt{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+This feature vector is then passed to a series of two (\todo{[SB] Update number}) ReLU neural network layers (following \citealt{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+The model is trained in part using a cross-entropy objective over this distribution.
 
-\subsection{Tracking left context with an auxiliary LSTM}
+\subsection{The tracking LSTM}
 
 While it is possible to use the model that we have described thus far as a sentence encoder, at least when an external parser is available to supply \shift~and \reduce~instructions (and we do evaluate such a model), our full model includes an additional component: the tracking LSTM.
-The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
+(\todo{[SB] Update framing of prev sentence.}) The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
 It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which is what allows the model to stand alone as a parser, and it additionally supplies a secondary input to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
 
-Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the two two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
+Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
 Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function.
@@
 \todo{[JG] Include figure of post-order traversal sequence made by the tracking LSTM?}
-\todo{[SB] Say more about joint learning of phrase representations for the two objectives.}
-
-\paragraph{The full objective function} \todo{[SB] Fill out. In short, there are two weighted terms in the objective: a semantic term and a syntactic term.}
 
@@
 
-\subsection{Implementing the transition-based sentence model}
+\subsection{Implementing the SPINN}
+
+\paragraph{The full objective function} \todo{[SB] Fill out. In short, there are two weighted terms in the objective: a semantic term and a syntactic term.}
+
+\todo{[SB] Say more about joint learning of phrase representations for the two objectives.}
 
@@
 
-We trained each model for 250k steps in each run, using a minibatch size of 32 for each step.
+We trained each model for 250k steps in each run, using a batch size of 32 for each step.
 We tracked each model's performance on the development set during training and saving parameters when this performance reached a new peak.
@@
 In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations.
-This implies a total space requirement of $B \times N \times T \times D$, which is prohibitively large for significant batch sizes $B$ or sentence lengths $N$.
+This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$.
 Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).

### 2016-03-15 writing/hard_stack_paper/paper.tex
Add speedy stuff

--- 2016-03-15 writing/hard_stack_paper/paper.tex
@@
 }
-\caption{Runtime performance results with the thin-stack GPU algorithm for tree-structured neural networks. \todo{[JG] Add some comments}}
+\caption{Runtime performance results with the thin-stack GPU algorithm.}
+\label{fig:speed}
 \end{figure}
 
-\todo{[JG] Write me! In short: We're an order of magnitude faster than plain C++ CPU TreeRNNs w/ reasonable model sizes.}
+We implemented the feedforward of the SPINN-NP-NT model (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution \cite{irsoy2014deep}.
+Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}.
+Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 61 transitions).
+We fix the model dimensionality and word embedding dimensionality at 300 for all tests.
+
+The CPU reference model is implemented in C++ and uses a single-core Eigen linear algebra backend.
+The implementation is a simplified version of the TreeRNN from \citealt{irsoy2014deep}\footnote{The original code for this model is available at \url{https://github.com/oir/deep-recursive}.
+We plan to release our modified code.}, modified to exactly match the computation performed by our GPU model.
+We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled.
+Our GPU model is implemented in C++ and CUDA, and will be released with the rest of our code.
+We test its performance on an NVIDIA Titan X GPU.
+
+At batch sizes of interest for test-time applications (e.g., 512 or 1024 examples), we observe a 10--20x speedup relative to the standard CPU implementation.
+Figure~\ref{speed} shows a substantial difference in runtime scaling as batch sizes increase which is typical of GPU computation.
+We hope that this improved batchable TreeRNN algorithm will enable further rapid research in tree-structured neural networks.
 
 \todo{Note that the full SPINN (i.e. Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
-
-\vspace{10em}
 
@@
 \begin{itemize}
-\item Our SPINN-NP-NT model precisely implements tree-structured compositionality (as in a TreeLSTM), but is capable of exploiting batched computation for dramatic \todo{[JG] Numbers} improvements in inference speed.
+\item Our SPINN-NP-NT model is exactly equivalent to a tree-structured network in what it computes, but is 10--20 times faster than a standard tree network at inference time.
 \item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our base tree-sequence hybrid model (SPINN-NP) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.

### 2016-03-15 writing/hard_stack_paper/batching_fig.tex
Slightly revised abstract, still work in progress.

@@
 
-    \begin{scope}[shift={(0.35in,-0.25in)}, frontier/.style={distance from root=60pt}]
+    \begin{scope}[shift={(0.25in,-0.2in)}, frontier/.style={distance from root=60pt}]
 
@@
 
-\caption{\label{fig:batching:good}A conventional sequence-based RNN, instantiated for two sentences.}
+\caption{\label{fig:batching:good}A conventional sequence-based RNN for two sentences.}
 \end{subfigure}
@@
 
-\caption{\label{fig:batching:bad}A conventional TreeRNN, instantiated for two sentences.}
+\caption{\label{fig:batching:bad}A conventional TreeRNN for two sentences.}
 \end{subfigure}
@@
 \caption{An illustration of two standard designs for sentence encoders.
-Note that the TreeRNN, unlike the sequence-based RNN, requires a substantially different connection structure for each sentence.}
+Note that the TreeRNN, unlike the sequence-based RNN, requires a substantially different connection structure for each sentence, making batched computation impractical.}
 \end{figure}

### 2016-03-15 writing/hard_stack_paper/paper.tex
Switch to real ACL template and paper size.

@@
-\documentclass[11pt,letterpaper]{article}
-\usepackage{../acl2015}
+\documentclass[11pt]{article}
+\usepackage{../acl2016}
+
+%\aclfinalcopy % Uncomment this line for the final submission
+%\def\aclpaperid{***} %  Enter the acl Paper ID here
+%\setlength\titlebox{5cm}
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
+
+\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black, draft]{hyperref}
+\usepackage{natbib}
 \usepackage{times}
@@
 \usepackage{qtree}
-\usepackage{graphicx} 
 \usepackage{booktabs}
-\usepackage{multirow}
-\usepackage{colortbl}
 \usepackage{caption}
@@
 \usepackage{color}
-\usepackage{xcolor}
 \usepackage{tikz}
@@
 \usepackage{pgfplots}
-\usepackage{ifthen}
-\usepackage{framed}
 \usepackage{algorithm}
 \usepackage[noend]{algpseudocode}
+\usepackage{stmaryrd}
+\usepackage{gb4e}
 
@@
 
-\newcommand{\nateq}{\equiv}
-\newcommand{\natind}{\mathbin{\#}}
-\newcommand{\natneg}{\mathbin{^{\wedge}}}
-\newcommand{\natfor}{\sqsubset}
-\newcommand{\natrev}{\sqsupset}
-\newcommand{\natalt}{\mathbin{|}}
-\newcommand{\natcov}{\mathbin{\smallsmile}}
-
-\newcommand{\plneg}{\mathop{\textit{not}}}
-\newcommand{\pland}{\mathbin{\textit{and}}}
-\newcommand{\plor}{\mathbin{\textit{or}}}
-
 \newcommand{\shift}{\textsc{shift}}
 \newcommand{\reduce}{\textsc{reduce}}
-
-% Strikeout
-\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
-\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}
-
-\newcommand{\True}{\texttt{T}}
-\newcommand{\False}{\texttt{F}}
-\usepackage{stmaryrd}
-\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}
-
-\newcommand{\mynote}[1]{{\color{blue}#1}}
-\newcommand{\tbchecked}[1]{{\color{red}#1}}
-
-\usepackage{gb4e}
-\noautomath
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%% Code to simulate natbib's citealt, which prints citations with
-%%%%% no parentheses:
-
-\makeatletter
-\def\citealt{\def\citename##1{{\frenchspacing##1} }\@internalcitec}
-\def\@citexc[#1]#2{\if@filesw\immediate\write\@auxout{\string\citation{#2}}\fi
-  \def\@citea{}\@citealt{\@for\@citeb:=#2\do
-    {\@citea\def\@citea{;\penalty\@m\ }\@ifundefined
-       {b@\@citeb}{{\bf ?}\@warning
-       {Citation `\@citeb' on page \thepage \space undefined}}%
-{\csname b@\@citeb\endcsname}}}{#1}}
-\def\@internalcitec{\@ifnextchar [{\@tempswatrue\@citexc}{\@tempswafalse\@citexc[]}}
-\def\@citealt#1#2{{#1\if@tempswa, #2\fi}}
-\makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\noautomath
 
@@
 
-\makeatletter
-\newcommand{\@BIBLABEL}{\@emptybiblabel}
-\newcommand{\@emptybiblabel}[1]{}
-\definecolor{black}{rgb}{0,0,0}
-\makeatother
-\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black, draft]{hyperref}
-
-\def\t#1{#1}
-\def\b#1{\t{\textbf{#1}}}
-\def\colspaceS{2.25mm}
-\def\colspaceM{4.0mm}
-\def\colspaceL{4.25mm}
-
 \begin{document}
@@
 
-A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \cite{socher2011semi,sutskever2014sequence,vinyals2015neural}.
+A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence,vinyals2015neural}.
 This component, the sentence encoder, is generally formulated as a learned parametric function from word vectors to sentence vectors, and this function can take a range of different forms.
-Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citealt{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citealt{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citealt{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words.
-Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \cite{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.
+Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citet{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citet{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citet{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words.
+Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \citep{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.
 
@@
 In spite of the unwieldiness of current standard TreeRNNs, there is still value in the idea of using syntactic parse trees to guide the construction of meaning in neural network models.
-Tree-structured models have shown substantial performance gains on some smaller-scale NLP tasks \cite{tai2015improved,li2015tree}, and their theoretical foundations are strong.
-Tree-structured compositionality is a property of natural language, and simpler sequence-based RNNs have a difficult time exploiting this property---while sequence-based models should, theoretically, be able to learn to simulate tree-structured composition, this behavior does not emerge in practice on reasonably-sized tasks \cite{bowman2015trees}.
+Tree-structured models have shown substantial performance gains on some smaller-scale NLP tasks \citep{tai2015improved,li2015tree}, and their theoretical foundations are strong.
+Tree-structured compositionality is a property of natural language, and simpler sequence-based RNNs have a difficult time exploiting this property---while sequence-based models should, theoretically, be able to learn to simulate tree-structured composition, this behavior does not emerge in practice on reasonably-sized tasks \citep{bowman2015trees}.
 
@@
 
-The SPINN is inspired by the shift-reduce parsing formalism \cite{aho1986compilers}, which builds a tree structure over a sequence (e.g.
+The SPINN is inspired by the shift-reduce parsing formalism \citep{aho1986compilers}, which builds a tree structure over a sequence (e.g.
 a natural language sentence) by a single left-to-right scan over the tokens of the sequence.
-Shift-reduce parsing has been successfully applied in algorithms for natural language parsing \cite{nivre2003efficient}.
+Shift-reduce parsing has been successfully applied in algorithms for natural language parsing \citep{nivre2003efficient}.
 
@@
 The buffer is populated with the words in each sentence before the first step of computation.
-We use word representations from the standard 300-dimensional vector package provided with GloVe \cite{pennington2014glove}.
+We use word representations from the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}.
 We do not update these representations during training.
@@
 
-Our composition function is based on the TreeLSTM of \citealt{tai2015improved}.
+Our composition function is based on the TreeLSTM of \citet{tai2015improved}.
 The TreeLSTM generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a vector pair $\langle \vec{h}, \vec{c}\rangle$, of which the latter vector is meant to serve as a slower-changing long-term memory.
@@
 
-This feature vector is then passed to a series of two (\todo{[SB] Update number}) ReLU neural network layers (following \citealt{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+This feature vector is then passed to a series of two (\todo{[SB] Update number}) ReLU neural network layers (following \citet{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
 The model is trained in part using a cross-entropy objective over this distribution.
@@
 At test time, a transition is chosen at each step by taking a hard max over the output of this function.
-We did not find the technique of scheduled sampling \cite{bengio2015scheduled}, or allowing the model to use its own transition decisions in some instances at training time, to be helpful.
+We did not find the technique of scheduled sampling \citep{bengio2015scheduled}, or allowing the model to use its own transition decisions in some instances at training time, to be helpful.
 
@@
 We use the RMSProp optimizer (\todo{[SB] Cite.}) with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
-We apply both dropout \cite{srivastava2014dropout} and batch normalization \cite{2015SIoffeCSzegedy} to the word embeddings in the buffer (after the linear projection is applied) and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
+We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the word embeddings in the buffer (after the linear projection is applied) and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
 In addition, we apply an L2 regularization penalty to all parameters.
@@
 
-We designed an alternative minimal-space stack algorithm inspired by the zipper technique \cite{huet1997zipper}.
+We designed an alternative minimal-space stack algorithm inspired by the zipper technique \citep{huet1997zipper}.
 For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$.
@@
 
-There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing \cite{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
-In addition, there has been recent work \cite{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
+There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing \citep{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
+In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
 To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
@@
 \midrule
-Lexicalized classifier \cite{snli:emnlp2015}
+Lexicalized classifier \citep{snli:emnlp2015}
                         & --                & --                    &   99.7   &   78.2      \\
@@
 \midrule
-100d LSTM encoders \cite{snli:emnlp2015}
+100d LSTM encoders \citep{snli:emnlp2015}
                         & 221k               & --               &   84.8   &   77.6      \\
-1024d pretrained GRU encoders \cite{DBLP:journals/corr/VendrovKFU15}
+1024d pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                         & 15m                & --              &   98.8   &   81.4       \\
-300d Tree-based CNN encoders \cite{mou2015recognizing}
+300d Tree-based CNN encoders \citep{mou2015recognizing}
                         & 3.5m                & --             &   83.4   &   82.1       \\
@@
 \result{300d SPINN-NP-NT}   
-                        & 3.4m                  & --                &   ?      &   81.9dv       \\
+                        & 3.4m                  & --                &   ?      &   80.8       \\
 \result{300d SPINN-NP}
-                        & 3.7m                  & --                &   ?      &   \underline{84.8dv}       \\
+                        & 3.7m                  & --                &   ?      &   \underline{83.?}       \\
 \result{300d SPINN }
-                        & ??                  & 92.2+dv            &   ?    &   83.3+dv       \\          
+                        & ??                  & 92.2+dv            &   ?    &   83.?      \\          
 % \result{300d SPINN-NP, sequence-based attn.   }     
@@
 \midrule
-100d LSTM w/ word-by-word attention \cite{rocktaschel2015reasoning}
+100d LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
                         & 252k               & --              &   85.3   &   83.5       \\
-300d mLSTM word-by-word attention model \cite{DBLP:journals/corr/WangJ15b}
+300d mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
                         & 1.9m               & --             &   92.0   &   86.1      \\
-300d LSTMN with deep attention fusion \cite{cheng2016long}
+300d LSTMN with deep attention fusion \citep{cheng2016long}
                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
@@
   \end{tabular}
-  \protect\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification.
+\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification.
 Params.
@@
 We evaluate the SPINN on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus (SNLI, \citealt{snli:emnlp2015}), which we use for training and evaluation:
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus (SNLI, \citet{snli:emnlp2015}), which we use for training and evaluation:
 
@@
 \begin{itemize}
-\item A baseline LSTM model (similar to that of \citealt{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
+\item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
 \item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition.
-This model is equivalent to a TreeLSTM \cite{tai2015improved}.
+This model is equivalent to a TreeLSTM \citep{tai2015improved}.
 \item The SPINN-NP model, which includes a tracking LSTM which guides composition, making it a hybrid tree-sequence model.
@@
 
-We compare our models against a range of baselines, including the strongest published non-neural network-based result from \cite{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
+We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
 We should note that these attention-based models are by far the strongest published on SNLI to date, but we do not wish to compare our model with those models directly.
@@
   \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
-  \addlegendentry{CPU \cite{irsoy2014deep}}
+  \addlegendentry{CPU \citep{irsoy2014deep}}
   \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
@@
 It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors in the sentence string and its likely position in the final tree, and that that context information could support both better parsing and better sentence encoding.
-One promising way to pursue this goal would be to encode the full contents of the stack and buffer at each time step following the method used by \citealt{dyer-EtAl:2015:ACL-IJCNLP} for parsing.
+One promising way to pursue this goal would be to encode the full contents of the stack and buffer at each time step following the method used by \citet{dyer-EtAl:2015:ACL-IJCNLP} for parsing.
 
-For a more ambitious goal, we expect that it should be possible to implement a variant of the SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citealt{grefenstette2015learning,joulin2015inferring}) would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+For a more ambitious goal, we expect that it should be possible to implement a variant of the SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citet{grefenstette2015learning,joulin2015inferring}) would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 
@@
 
-\bibliographystyle{../acl}
+\bibliographystyle{../acl_natbib}
 \bibliography{../MLSemantics}

### 2016-03-15 writing/hard_stack_paper/paper.tex
Typing up Abi's comments.

@@
 Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences.
-However, they suffer two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks.
+However, they suffer from two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks.
 We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser.
@@
 A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence,vinyals2015neural}.
-This component, the sentence encoder, is generally formulated as a learned parametric function from word vectors to sentence vectors, and this function can take a range of different forms.
+This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms.
 Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citet{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citet{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citet{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words.
@@
 For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.
+\todo{[JG] Try to introduce transitions before referring to the transition var $\mathbf t$.
+Maybe say that each $t\in\{\shift, \reduce\}$ }
 
@@
 The core SPINN model implements a sequential shift-reduce parser which operates over sentences.
-It is designed to produce a vector representation of the sentence as output,as opposed to a tree as in standard shift-reduce \textit{parsing}.
+It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing.
 This relies on two key modifications to the shift-reduce algorithm:
@@
 We do not update these representations during training.
-Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors that are stored side-by-side in the buffer and used as inputs to a TreeLSTM composition function:
+Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to a TreeLSTM composition function:
 
@@
 
-In the SPINN, each row of the stack and each row of the buffer contains a pair of vectors $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence.
+In the SPINN, each row of the stack and each row of the buffer contains a pair $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence.
 \todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}
@@
 This paper evaluates the SPINN within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task.
-To classify a sentence pair, a feature vector is first constructed.
+To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed.
 This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition.
 In particular, the $\vec{h}$ portions of these representations are used.
-The final feature vector consists of the concatenation of these two vectors, $\vec{x}_{\text{premise}}$ and $\vec{x}_{\text{hypothesis}}$, their difference, and their elementwise product:
+The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:
 
 \begin{equation}
-\vec{x}_{classifier} = 
+\vec{x}_{\text{classifier}} = 
 \colvec{4}
@@
 
+\todo{[SB] Good point to reference fig 2a?}
+
 \paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) the SPINN will use needs to take as its input some representation of the current state of the model.
@@
 \begin{equation}
-\vec{p}_{trans} = \text{softmax}(W_{trans}\vec{h}_{track})
+\vec{p}_{\text{trans}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{track}})
 \end{equation}
 
-The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep).
+The resulting vector $\vec{p}_{\text{trans}}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep).
 In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution.
@@
 \begin{equation}
-  L_{trans} = - \log \vec{p}_{trans}[t^*]
+  L_{\text{trans}} = - \log \vec{p}_{\text{trans}}[t^*]
 \end{equation}
@@
 \paragraph{The size of the stack}
-The size of the stack should be $N$ for sentences of $N$ words, in case the first \reduce~operation merges the final two words. The size of the buffer should be $N$.
+The size of the stack should be $N$ for sentences of $N$ words, in case the first \reduce~operation merges the final two words.
+The size of the buffer should be $N$.
+\todo{[SB] Height, not storage size.}
 
 \paragraph{Converting parses to transition sequences}
+\todo{[SB] Remove all Model N references.}
 
@@
 Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).
+\todo{[JG] Remind reader what T and D are here.}
 
@@
 
-In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) computes the precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
+\todo{[SB] Make equivalence claim more explicit. Clean up bullets.}
+
+In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) computes precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
 While it adds no expressive power on its own, this design improves upon the standard TreeRNN implementation in three ways:
@@
 \subsection{Related work: Transition-based parsing and neural networks.}
+\todo{[SB] Move!}
 
@@
 \subsection{Natural language inference and SNLI}
+\todo{[SB] Move up.}
 
@@
 
-We evaluate four models on the SNLI task. Each uses 300d hidden states:
+We evaluate four models on the SNLI task. Each uses 300d hidden states: \todo{[SB] d or D?}
 \begin{itemize}
@@
 We should note that these attention-based models are by far the strongest published on SNLI to date, but we do not wish to compare our model with those models directly.
-SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
+\todo{[SB] But we do.
+Rephrase.} SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
 The attentional models are narrowly adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders.
@@
 Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
-The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations).
+The full SPINN model also generates parsing predictions, and we report a measure of agreement \todo{[SB] Clarify.} between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations).
 We also report the number of trained parameters in each model and training set accuracy.
@@
 At batch sizes of interest for test-time applications (e.g., 512 or 1024 examples), we observe a 10--20x speedup relative to the standard CPU implementation.
-Figure~\ref{speed} shows a substantial difference in runtime scaling as batch sizes increase which is typical of GPU computation.
+Figure~\ref{fig:speed} shows a substantial difference in runtime scaling as batch sizes increase which is typical of GPU computation.
 We hope that this improved batchable TreeRNN algorithm will enable further rapid research in tree-structured neural networks.
@@
 
-\paragraph{Future work} Because our aim for this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness of the SNLI task.
-However, the choice of whether to use soft attention and the choice of what representations to perform attention over are orthogonal, and we expect that it should be possible to productively combine our model with soft attention to yield state-of-the-art performance.
+\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness of the SNLI task.
+However, the choice of whether to use soft attention and the choice of what representations to perform attention over are orthogonal \todo{[SB] Rephrase prev clause.}, and we expect that it should be possible to productively combine our model with soft attention to yield state-of-the-art performance.
 
@@
 For a more ambitious goal, we expect that it should be possible to implement a variant of the SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citet{grefenstette2015learning,joulin2015inferring}) would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+
+\todo{[SB] Fix run-ons here.}
 

### 2016-03-15 writing/hard_stack_paper/paper.tex
Incorporate Hieu's comments.

@@
 We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
+\todo{[SB] Quantify success.}
 
@@
 TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons.
-Batched computation---performing synchronized computation across many examples at once---is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets (\question{What, if anything, can we cite here?}).
+Batched computation---performing synchronized computation across many examples at once---is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets (\question{What, if anything, can we cite here?} \todo{[SB] Bengio 2003?}).
 Batching yields order-of-magnitude improvements in standard CPU implementations of neural network models, and substantial additional gains on GPUs.
@@
 \todo{[JG] Clarify prev.
-sentence.}
+sentence.
+Indicate which side of the stack/buffer is the head below.}
 \begin{description}
@@
 
-In the SPINN, each row of the stack and each row of the buffer contains a pair $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence.
+In the SPINN, each element of the stack and each element of the buffer contains a pair $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence.
 \todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}
@@
 The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$.
-In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see below).
+In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see below \todo{[SB] Ref.}).
 $\odot$ denotes the elementwise product.
@@
 (\todo{[SB] Update framing of prev sentence.}) The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
-It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which is what allows the model to stand alone as a parser, and it additionally supplies a secondary input to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (\todo{[SB] Refer to eq no.
+here and below.}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
 
@@
 Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function.
+\todo{[SB] Trim.}
 
@@
 
-\paragraph{Tracking left context for interpretation} Lexical ambiguity is omnipresent in natural language.
+\paragraph{Tracking left context for interpretation} \todo{[SB] Trim.} Lexical ambiguity is omnipresent in natural language.
 Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.

### 2016-03-15 writing/hard_stack_paper/paper.tex
Comments from me.
One more Q.

@@
 
-\subsection{Models evaluated}
+\subsection{Models evaluated} \todo{[SB] Condense.}
 
@@
 Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}.
-Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 61 transitions).
+Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 61 transitions \todo{[JG] Not 59?}).
 We fix the model dimensionality and word embedding dimensionality at 300 for all tests.
+\todo{[JG] Explain why we implemented *just* feedforward in this style.
+(Something like: Theano is convenient for training but very non-optimal for our model.)}
 

### 2016-03-15 writing/hard_stack_paper/paper.tex
Injecting more results, comment from Jiwei.
Writing up CM's comments.

@@
 Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences.
-However, they suffer from two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks.
-We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser.
+However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they can only operate on parsed sentences and do not directly support batched computation.
+We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser.
 This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
 We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
-\todo{[SB] Quantify success.}
+\todo{[SB] Quantify success.
+Emphasize novelty.}
 
@@
 
-A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence,vinyals2015neural}.
+A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence}.
 This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms.
@@
 Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \citep{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.
+\todo{[SB] Trim this para.}
 
@@
 This paper additionally aims to integrate parsing into TreeRNN processing to circumvent this issue.
+\todo{[SB] Trim this para.
+Reframe parsing claim around Socher 2011.}
 
@@
 
-The SPINN is inspired by the shift-reduce parsing formalism \citep{aho1986compilers}, which builds a tree structure over a sequence (e.g.
+The SPINN is inspired by the shift-reduce parsing formalism \citep{aho1986compilers} \todo{[SB] Switch to 1976 paper.}, which builds a tree structure over a sequence (e.g.
 a natural language sentence) by a single left-to-right scan over the tokens of the sequence.
-Shift-reduce parsing has been successfully applied in algorithms for natural language parsing \citep{nivre2003efficient}.
+Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.][]{nivre2003efficient}.
 
@@
 sentence.
-Indicate which side of the stack/buffer is the head below.}
+Indicate which side of the stack/buffer is the head below.
+Indicate that we'll eventually be predicting $t$.}
 \begin{description}
@@
 Our composition function is based on the TreeLSTM of \citet{tai2015improved}.
-The TreeLSTM generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with that older design the idea of representing intermediate states in a computation using a vector pair $\langle \vec{h}, \vec{c}\rangle$, of which the latter vector is meant to serve as a slower-changing long-term memory.
+The TreeLSTM generalizes the LSTM neural network layer to tree- rather than sequence-based inputs; it shares the idea of representing intermediate states in a computation using a vector pair $\langle \vec{h}, \vec{c}\rangle$, of which the latter vector is meant to serve as a slower-changing memory component.
 It is formulated as:
@@
 
-This paper evaluates the SPINN within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task.
+This paper evaluates the SPINN on a sentence pair classification task.
 To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed.
@@
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
+\todo{[SB] Mention that humans can do incremental parsing with left context only.}
 
@@
 
+\todo{[SB] Mention that our features are weaker than those in competitive parsers.}
+
 \subsection{Implementing the SPINN}
@@
 
-For Models 0--3, all training data must be parsed in advance into an unlabeled binary constituency tree.
-In addition, Model 0 requires that  parses be available at test time as well.
-For both SST and SNLI we use the parses included with the corpus distributions whenever parses are needed.
-
 For model 0, training data can be prepared by linearizing the provided parse, then deleting left brackets and replacing right brackets with \reduce~instructions.
@@
 
-The input for models 1--4 is simply the word sequence from the parse, with the first two words moved into the stack.
-The syntactic supervision labels for models 1--3 are simply a binarized version of the Model 0 inputs, with the first two tokens (which are necessarily \shift~\shift) omitted:
+The input for models 1--4 is simply the word sequence from the parse. The syntactic supervision labels for models 1--3 are simply a binarized version of the Model 0 inputs:
 
@@
 
+\todo{[SB] Trim this section. Just say that it can be done.}
 To efficiently implement batched computation, we must stipulate a fixed number of transitions (50) that the model can perform before producing an output.
@@
 While our design includes some wasted computation induced by the full application of the composition function at every step, this is made up for by avoiding most of the (potentially expensive) scattered memory access that would be needed for batched computation strategies that more closely mimic the standard tree-based architecture.
+\todo{[SB] Clarify.}
 
@@
 
-\item By designing the model around the standard transition set used in neural network syntactic parsers, it opens the door to tight integration between parsing and sentence interpretation, including the possibility of joint learning, or even of learning a parser trained entirely on a sentence classification or interpretation objective.
+\item It opens the door to tight integration between parsing and sentence interpretation, including the possibility of joint learning, or even of learning a parser trained entirely on a sentence classification or interpretation objective.
 \end{itemize}
@@
 
-There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing \citep{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
+There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy} \todo{[SB] Order citations by year}.
 In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
@@
 300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
-\result{300d SPINN-NP-NT}   
-                        & 3.4m                  & --                &   ?      &   80.8       \\
-\result{300d SPINN-NP}
-                        & 3.7m                  & --                &   ?      &   \underline{83.?}       \\
-\result{300d SPINN }
+300d SPINN-NP-NT encoders
+                        & 3.4m                  & --                &   84.4      &   80.9       \\
+300d SPINN-NP encoders
+                        & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
+\result{300d SPINN encoders }
                         & ??                  & 92.2+dv            &   ?    &   83.?      \\          
@@
 
-We evaluate the SPINN on the task of natural language inference (NLI), also known as recognizing textual entailment (RTE).
+We evaluate the SPINN on the task of natural language inference (NLI) \todo{[SB] Cite.}, also known as recognizing textual entailment (RTE).
 NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus (SNLI, \citet{snli:emnlp2015}), which we use for training and evaluation:
@@
 \todo{[SB] But we do.
-Rephrase.} SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
+Rephrase.
+Move before naming of models.} SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
 The attentional models are narrowly adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders.
@@
 Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
-The full SPINN model also generates parsing predictions, and we report a measure of agreement \todo{[SB] Clarify.} between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations).
+The full SPINN model also generates parsing predictions, and we report a measure of agreement \todo{[SB] Clarify.} between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations) \todo{[SB] Which Stanford parser.}.
 We also report the number of trained parameters in each model and training set accuracy.
 
-We find that the bare SPINN-NP-NT model performs relatively poorly, but that the SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
+We find that the bare SPINN-NP-NT model performs little better than the RNN baseline, but that the SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
 The success of the SPINN-NP model, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
 
-The full SPINN model with its relatively weak internal parser performs more poorly, but nonetheless exceeds the performance of the LSTM baseline.
+The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline.
 None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task.
 
-\todo{[SB] Discuss relationship between our model and the TBCNN.}
+\todo{[SB] Discuss relationship between our model and the TBCNN. Add significance numbers.}
 
@@
 At batch sizes of interest for test-time applications (e.g., 512 or 1024 examples), we observe a 10--20x speedup relative to the standard CPU implementation.
-Figure~\ref{fig:speed} shows a substantial difference in runtime scaling as batch sizes increase which is typical of GPU computation.
+Figure~\ref{fig:speed} shows a substantial difference in runtime scaling as batch sizes increase, which is typical of GPU computation.
 We hope that this improved batchable TreeRNN algorithm will enable further rapid research in tree-structured neural networks.
@@
 
-\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness of the SNLI task.
+\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness on the SNLI task.
 However, the choice of whether to use soft attention and the choice of what representations to perform attention over are orthogonal \todo{[SB] Rephrase prev clause.}, and we expect that it should be possible to productively combine our model with soft attention to yield state-of-the-art performance.
@@
 The tracking LSTM that we incorporated into the SPINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
-It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors in the sentence string and its likely position in the final tree, and that that context information could support both better parsing and better sentence encoding.
+It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors \todo{[SB] Rewrite.} in the sentence string and its likely position in the final tree, and that that context information could support both better parsing and better sentence encoding.
 One promising way to pursue this goal would be to encode the full contents of the stack and buffer at each time step following the method used by \citet{dyer-EtAl:2015:ACL-IJCNLP} for parsing.
 
-For a more ambitious goal, we expect that it should be possible to implement a variant of the SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citet{grefenstette2015learning,joulin2015inferring}) would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+For a more ambitious goal, we expect that it should be possible to implement a variant of the SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citet{grefenstette2015learning,joulin2015inferring}).
+This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 
@@
 \bibliography{../MLSemantics}
-\todo{[SB] Make bibliography style uniform.}
+\todo{[SB] Make bibliography style uniform, esp BatchNorm.}
 

### 2016-03-15 writing/hard_stack_paper/paper.tex
Initial revision of abstract and Section 1.

@@
 However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they can only operate on parsed sentences and do not directly support batched computation.
-We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser.
-This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
+We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single tree-sequence hybrid model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser.
+Our model supports batched computation for a speedup of up to 20x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
 We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
-\todo{[SB] Quantify success.
-Emphasize novelty.}
-
 \end{abstract}
@@
 This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms.
-Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citet{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citet{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citet{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words.
-Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \citep{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.
-\todo{[SB] Trim this para.}
+Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory \citep[LSTM,][see Figure~\ref{fig:batching:good}]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011semi}, which propagate information up a binary parse tree structure over words.
 
-TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons.
-Batched computation---performing synchronized computation across many examples at once---is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets (\question{What, if anything, can we cite here?} \todo{[SB] Bengio 2003?}).
-Batching yields order-of-magnitude improvements in standard CPU implementations of neural network models, and substantial additional gains on GPUs.
-TreeRNNs build the parse structure of each sentence they process into the structure of the network itself, forcing the model to perform a different sequence of computations for each sentence, and thereby requiring that each sentence be processed one-by-one, making batched computation impossible in standard implementations.
-One aim of this paper is to restructure the computations involved in a TreeRNN in order to enable batching and make these large-scale evaluations possible.
-In addition to their slow speed, TreeRNNs have an additional feature that makes them impractical in many settings: they can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.
-This paper additionally aims to integrate parsing into TreeRNN processing to circumvent this issue.
-\todo{[SB] Trim this para.
-Reframe parsing claim around Socher 2011.}
+Of these, it is easy to argue that the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}.
+However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
+Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets.
+Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations.
+In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.
 
-In spite of the unwieldiness of current standard TreeRNNs, there is still value in the idea of using syntactic parse trees to guide the construction of meaning in neural network models.
-Tree-structured models have shown substantial performance gains on some smaller-scale NLP tasks \citep{tai2015improved,li2015tree}, and their theoretical foundations are strong.
-Tree-structured compositionality is a property of natural language, and simpler sequence-based RNNs have a difficult time exploiting this property---while sequence-based models should, theoretically, be able to learn to simulate tree-structured composition, this behavior does not emerge in practice on reasonably-sized tasks \citep{bowman2015trees}.
-
-In this paper, we introduce the Stack-augmented Parser-Interpreter Neural Network, or SPINN.
-The SPINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
+This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}.
+The SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
 This design improves upon the TreeRNN architecture in three ways:
@@
 
-\todo{[SB] Move the introduction of the SPINN earlier in the introduction. Mention that we'll be using SNLI as our sole evaluation task.}
+We evaluate the SPINN on the Stanford Natural Language Inference sentence understanding task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of upwards of 10x over comparable TreeRNN models.
 
-\section{The shift-reduce sentence model}
+\subsection{Related work}
+
+There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
+In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
+To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
+
+\citet{socher2011dynamic} presents a version of the TreeRNN model which is capable of operating over unparsed inputs.
+However, this method is unsupervised, limiting its performance, and it requires an expensive search process at test time.
+Our model presents an alternative approach to combining interpretation and parsing that supports supervised training for the parsing component, and that is dramatically faster than any previous tree-structured model.
+
+\section{Our model: The SPINN}
 
@@
 
-\subsection{Related work: Transition-based parsing and neural networks.}
-\todo{[SB] Move!}
-
-There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy} \todo{[SB] Order citations by year}.
-In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
-To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
-
 \section{Experiments}

### 2016-03-15 writing/hard_stack_paper/paper.tex
Update results with better numbers (rerun on jag10)

@@
 Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}.
-Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 61 transitions \todo{[JG] Not 59?}).
+Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 59 transitions.
 We fix the model dimensionality and word embedding dimensionality at 300 for all tests.

### 2016-03-15 writing/hard_stack_paper/paper.tex
Resolve todos in efficiency section

@@
 In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations.
-This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$.
-Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).
-\todo{[JG] Remind reader what T and D are here.}
+This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$.
+(Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).
 
@@
 For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$.
-\todo{[JG] Either expand or remove references to batch size.} Each row $S_t$ represents the top of the actual stack at timestep $t$.
+Each row $S_t$ represents the top of the actual stack at timestep $t$.
 We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time.

### 2016-03-15 writing/hard_stack_paper/paper.tex
More minor improvements re: comments

@@
 
-A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$.
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below).
 It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
 For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.
-\todo{[JG] Try to introduce transitions before referring to the transition var $\mathbf t$.
-Maybe say that each $t\in\{\shift, \reduce\}$ }
 
@@
 
-We implemented the feedforward of the SPINN-NP-NT model (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution \cite{irsoy2014deep}.
+We implemented the feedforward of the SPINN-NP-NT model (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution from \citet{irsoy2014deep}.
+Such a standard TreeRNN implementation operates on a single example at a time and is thus not feasible to run on a GPU.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as this is the more critical performance metric for real-world applications.}
+
 Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}.
-Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 59 transitions.
+Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 59 transitions).
 We fix the model dimensionality and word embedding dimensionality at 300 for all tests.
-\todo{[JG] Explain why we implemented *just* feedforward in this style.
-(Something like: Theano is convenient for training but very non-optimal for our model.)}
 
 The CPU reference model is implemented in C++ and uses a single-core Eigen linear algebra backend.
-The implementation is a simplified version of the TreeRNN from \citealt{irsoy2014deep}\footnote{The original code for this model is available at \url{https://github.com/oir/deep-recursive}.
+The implementation is a simplified version of the TreeRNN from \citet{irsoy2014deep}\footnote{The original code for this model is available at \url{https://github.com/oir/deep-recursive}.
 We plan to release our modified code.}, modified to exactly match the computation performed by our GPU model.

### 2016-03-15 writing/hard_stack_paper/paper.tex
Shift-reduce section: Minor notes on notation, transition decisions

@@
 A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below).
+In general a parser may also generate these predictions on-the-fly as it reads the tokens.
 It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
@@
 We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations.
-Below $\mid$ denotes an infix \textit{cons} function.
-\todo{[JG] Clarify prev.
-sentence.
-Indicate which side of the stack/buffer is the head below.
-Indicate that we'll eventually be predicting $t$.}
+Below the $\mid$ symbol denotes the \textit{cons} (concatenation) operator.
+We arbitrarily choose to always \textit{cons} on the left in the notation below.
 \begin{description}

### 2016-03-16 writing/hard_stack_paper/paper.tex
Initial revision of Section 2.

@@
 
-%     \author{}
-    \author{
-    Samuel R.\ Bowman$^{1,2,5,}$\thanks{~\,The first two authors contributed equally.} \\
-    \texttt{\small sbowman@stanford.edu} \\
-    \And
-    Jon Gauthier$^{2,3,5,*}$ \\
-    \texttt{\small jgauthie@stanford.edu} \\
-    \And
-    Abhinav Rastogi$^{4,5}$ \\
-    \texttt{\small arastogi@stanford.edu} \\
-    \AND
-    Raghav Gupta$^{6}$ \\
-    \texttt{\small rgupta93@stanford.edu} \\
-    \And
-    Christopher D.\ Manning$^{1,2,5,6}$\\
-    \texttt{\small manning@stanford.edu}\\
-    \And
-    Christopher Potts$^{1}$\\
-    \texttt{\small cgpotts@stanford.edu}
-    \AND\\[-3ex]
-    {$^{1}$Stanford Linguistics\quad
-    $^{2}$Stanford NLP Group\quad
-    $^{3}$Stanford Symbolic Systems}\\
-    {$^{4}$Stanford Electrical Engineering\quad
-    $^{5}$Stanford AI Lab\quad
-    $^{6}$Stanford Computer Science}
-    }
+     \author{}
+%    \author{
+%    Samuel R.\ Bowman$^{1,2,5,}$\thanks{~\,The first two authors contributed equally.} \\
+%    \texttt{\small sbowman@stanford.edu} \\
+%    \And
+%    Jon Gauthier$^{2,3,5,*}$ \\
+%    \texttt{\small jgauthie@stanford.edu} \\
+%    \And
+%    Abhinav Rastogi$^{4,5}$ \\
+%    \texttt{\small arastogi@stanford.edu} \\
+%    \AND
+%    Raghav Gupta$^{6}$ \\
+%    \texttt{\small rgupta93@stanford.edu} \\
+%    \And
+%    Christopher D.\ Manning$^{1,2,5,6}$\\
+%    \texttt{\small manning@stanford.edu}\\
+%    \And
+%    Christopher Potts$^{1}$\\
+%    \texttt{\small cgpotts@stanford.edu}
+%    \AND\\[-3ex]
+%    {$^{1}$Stanford Linguistics\quad
+%    $^{2}$Stanford NLP Group\quad
+%    $^{3}$Stanford Symbolic Systems}\\
+%    {$^{4}$Stanford Electrical Engineering\quad
+%    $^{5}$Stanford AI Lab\quad
+%    $^{6}$Stanford Computer Science}
+%    }
 
@@
 
-The SPINN is inspired by the shift-reduce parsing formalism \citep{aho1986compilers} \todo{[SB] Switch to 1976 paper.}, which builds a tree structure over a sequence (e.g.
+The SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g.
 a natural language sentence) by a single left-to-right scan over the tokens of the sequence.
@@
 \item The parser's intermediate representations, including the representations of partial tree structures on the stack, are fixed-length vectors.
-\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
+\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.
 \end{itemize}
@@
 We do not update these representations during training.
-Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to a TreeLSTM composition function:
+Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
 
@@
 
-\paragraph{The stack and buffer}
-
-In the SPINN, each element of the stack and each element of the buffer contains a pair $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence.
-\todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}
-
 \paragraph{The composition function}
@@
 
-Our composition function is based on the TreeLSTM of \citet{tai2015improved}.
-The TreeLSTM generalizes the LSTM neural network layer to tree- rather than sequence-based inputs; it shares the idea of representing intermediate states in a computation using a vector pair $\langle \vec{h}, \vec{c}\rangle$, of which the latter vector is meant to serve as a slower-changing memory component.
+Our composition is a version of the TreeLSTM of \citet{tai2015improved}.
+The TreeLSTM layer, which generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$.
 It is formulated as:
@@
     {\vec{e}}
-\right)
+\right) \label{eqn:lstm1}
 \\
@@
     {\vec{e}}
-\right)
+\right) \label{eqn:lstm2}
 \\
@@
 The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$.
-In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see below \todo{[SB] Ref.}).
+In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
 $\odot$ denotes the elementwise product.
 Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.
+
+\paragraph{The contents of the stack and buffer}
+
+The stack and the buffer can be represented as lists of $N$ elements each (for sentences of up to $N$ words), with two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ constituting each element.
 
@@
 
-This feature vector is then passed to a series of two (\todo{[SB] Update number}) ReLU neural network layers (following \citet{snli:emnlp2015}), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+Following \citep{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (the number of layers is a tuned hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
 The model is trained in part using a cross-entropy objective over this distribution.
 
-\subsection{The tracking LSTM}
+\subsection{The tracking LSTM}\label{sec:tracking}
 
-While it is possible to use the model that we have described thus far as a sentence encoder, at least when an external parser is available to supply \shift~and \reduce~instructions (and we do evaluate such a model), our full model includes an additional component: the tracking LSTM.
-(\todo{[SB] Update framing of prev sentence.}) The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
-It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (\todo{[SB] Refer to eq no.
-here and below.}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
+In addition to the stack, the buffer, and the composition function, our full model includes an additional component: the tracking LSTM.
+The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equations~\ref{eqn:lstm1}--\ref{eqn:lstm2}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
 
-Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
-Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function.
-\todo{[SB] Trim.}
-
-\todo{[SB] Good point to reference fig 2a?}
+The tracking LSTM takes three vectors as inputs at each step (highlighted in yellow in Figure~\ref{fig:m1-views}): the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
 
 \paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) the SPINN will use needs to take as its input some representation of the current state of the model.
-Because the tracking LSTM directly takes the buffer top as input, it can directly inform that function of what word is under consideration for a possible \shift.
-Because the tracking LSTM sees the top of the stack at every step, and information only enters the stack by being at the top at some step, the tracking LSTM is able to supply that function with a summary of the state of the stack without having to iterate over the full stack at each step.
+We use the hidden state of the tracking LSTM as that input.
+The tracking LSTM's limited access to the buffer makes this fast to compute, but somewhat more limited than the rich feature sets used in state-of-the-art parsers.
 
-\paragraph{Tracking left context for interpretation} \todo{[SB] Trim.} Lexical ambiguity is omnipresent in natural language.
+\paragraph{Tracking left context for interpretation} Lexical ambiguity is ubiquitous in natural language.
 Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
-Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word (first incorporates it into a partial sentence representation), the model has access to a state vector that summarizes all of the words to the left of the current word, and that state vector can serve as a crude but effective form of context to help with disambiguation.
-In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is strictly less than the sequence model has access to, and is often just a single additional word.
+Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which provides some cues for disambiguation.
+In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word.
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
-\todo{[SB] Mention that humans can do incremental parsing with left context only.}
+
+It would be straightforward to augment the SPINN to support the use of some amount of right context as well, but this would add complexity to the model that we think is largely unnecessary: humans are very effective at understanding the beginnings of sentences before having seen or heard the ends, suggesting that it is possible to get by without the unavailable right context.
 
@@
 
-\todo{[SB] Mention that our features are weaker than those in competitive parsers.}
-
 \subsection{Implementing the SPINN}
 
-\paragraph{The full objective function} \todo{[SB] Fill out. In short, there are two weighted terms in the objective: a semantic term and a syntactic term.}
+\paragraph{The full objective function} Our objective function combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentence and the hypothesis sentence at each step $t$, and an L2 regularization term on the trained parameters.
+The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
 
-\todo{[SB] Say more about joint learning of phrase representations for the two objectives.}
+\begin{equation}
+\begin{split}
+\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
+\end{split}
+\end{equation}
 
-\paragraph{The size of the stack}
-The size of the stack should be $N$ for sentences of $N$ words, in case the first \reduce~operation merges the final two words.
-The size of the buffer should be $N$.
-\todo{[SB] Height, not storage size.}
+\paragraph{Preparing the data} At training time, the SPINN requires both a transition sequences $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence.
+The token sequence is simply the words in the sentence in order.
+$\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:
 
-\paragraph{Converting parses to transition sequences}
-\todo{[SB] Remove all Model N references.}
+\vspace{0.5em}
+{\noindent\small
+{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) ) $\Rightarrow$\\
+{$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
+{$\mathbf x$}: the, cat, sat, down
+}
 
-For model 0, training data can be prepared by linearizing the provided parse, then deleting left brackets and replacing right brackets with \reduce~instructions.
-That is demonstrated here with the example sentence \ii{the cat sat down}:
+\paragraph{Handling variable sentence lengths} For any sentence model to be trained with batched computation, it is necessary to pad or crop sentences to a fixed length.
+We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI.
+Transition sequences $\mathbf t$ are cropped at the left or padded at the left with \shift s.
+Token sequences $\mathbf x$ are then cropped or padded with empty tokens at the left to match the number of \shift s added or removed from $\mathbf t$, and can then be padded with empty tokens at the right to meet the desired length $N$.
 
-\begin{quote}\small
-( ( the cat ) ( sat down ) )$\Rightarrow$\\
-the cat \reduce~sat down \reduce~\reduce
-\end{quote}
+\paragraph{Optimization, initialization, and hyperparameters}
 
-The input for models 1--4 is simply the word sequence from the parse. The syntactic supervision labels for models 1--3 are simply a binarized version of the Model 0 inputs:
+We use the RMSProp optimizer \citep{tieleman2012lecture} with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
+We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the word embeddings in the buffer (after the linear projection is applied) and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
 
-\begin{quote}\small
-( ( the cat ) ( sat down ) )$\Rightarrow$ \\
-stack: $\langle$the, cat$\rangle$\\
-buffer: $\langle$sat, down$\rangle$\\
-ops: \reduce~\shift~\shift~\reduce~\reduce
-\end{quote}
+We initialize the model parameters using the nonparametric strategy of \citet{DBLP:journals/corr/HeZR015}, with the exception of the softmax classifier layers, which we initialize uniformly from $[-0.005, 0.005]$.
 
-\paragraph{Handling variable sentence lengths}
-
-\todo{[SB] Trim this section. Just say that it can be done.}
-To efficiently implement batched computation, we must stipulate a fixed number of transitions (50) that the model can perform before producing an output.
-Sentences whose transition sequences are shorter than this length are padded, and sentences whose transition sequences are longer than this length are cropped.
-
-Padding, if done properly, should not impact the output of the model significantly:\footnote{The tracking LSTM can cause a slight dependence of the final representation on unrolling length.
-Because of this, we always train and test the model with a fixed unrolling length.} a model with the fixed parameters should produce the same representation for a 10-word sentence whether it is unrolled for 19 transitions (the minimum to avoid cropping) or 100.
-
-Cropping necessarily discards information, but if done properly, it is nonetheless possible to learn good representations for cropped sentences.
-When a transition sequence is cropped, the number of removed \shift~transitions is tracked, and an equal number of word tokens is removed from the tail (left) end of the buffer.
-This makes it possible for a transition sequence to begin with a \reduce~transition, or to otherwise use a \reduce~transition on a stack that does not have two elements to merge.
-In this instance, we simply feed the composition function one or more zero vectors, which represent nonexistent stack elements.
-If the composition learns to interpret these zero vectors properly, they can be taken to be unknown or incomplete nodes in an otherwise complete tree that is constructed according to the intended parse structure.
-\todo{[SB] Make a figure showing the results of cropping if there's room.}
-
-\paragraph{Optimization and hyperparameters}
-
-We use the RMSProp optimizer (\todo{[SB] Cite.}) with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
-We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the word embeddings in the buffer (after the linear projection is applied) and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
-In addition, we apply an L2 regularization penalty to all parameters.
-
-We used random search to tune many of the hyperparameters, including the learning rate, the dropout rate, the L2 regularization weight, the hidden dimension of the tracking LSTM and the number of layers in the MLP \todo{[SB] Report the ranges searched for each parameter and the number of runs for each results.
-Some of this can go in an appendix.}
+We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and than launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
+An appendix shows the precise values used on the strongest models.
 
@@
 
-\paragraph{Software infrastructure} We will make Theano code available.
+\paragraph{Software infrastructure} We will make source code available.
 
@@
 
-\subsection{TreeRNN-equivalence and efficiency}
+\subsection{TreeRNN-equivalence}
 
-\todo{[SB] Make equivalence claim more explicit. Clean up bullets.}
-
-In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) computes precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
-While it adds no expressive power on its own, this design improves upon the standard TreeRNN implementation in three ways:
-
-\begin{itemize}
-\item It provides a straightforward strategy for implementing tree-structured composition which allows for batched computation, and provides near-optimal run times on typical computer systems.
-While our design includes some wasted computation induced by the full application of the composition function at every step, this is made up for by avoiding most of the (potentially expensive) scattered memory access that would be needed for batched computation strategies that more closely mimic the standard tree-based architecture.
-\todo{[SB] Clarify.}
-
-\item It provides a simple mechanism for integrating the use of left-context into tree-structured models through the tracking LSTM, providing a fast mechanism to help these models handle lexical ambiguity.
-
-\item It opens the door to tight integration between parsing and sentence interpretation, including the possibility of joint learning, or even of learning a parser trained entirely on a sentence classification or interpretation objective.
-\end{itemize}
+In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
+In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
 
@@
 \subsection{Natural language inference and SNLI}
-\todo{[SB] Move up.}
 
-We evaluate the SPINN on the task of natural language inference (NLI) \todo{[SB] Cite.}, also known as recognizing textual entailment (RTE).
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus (SNLI, \citet{snli:emnlp2015}), which we use for training and evaluation:
+We evaluate the SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE][]{dagan2006pascal,MacCartney09}.
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus \citet[SNLI][]{snli:emnlp2015}), which we use for training and evaluation:
 
@@
 
+\begin{table*}[ht]\small
+\begin{center}
+\begin{tabular}{lrlrrrr}
+\toprule
+Param.     & Range & Strategy        & RNN       & SPINN-NP-NT   & SPINN-NP  & SPINN \\
+\midrule
+Initial LR & 0.0002--0.02 & \textsc{log} & 0.005  & 0.0003 & 0.007  & SPINN\\
+L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & SPINN\\
+Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & N/A & N/A & N/A  & SPINN\\
+Embedding transformation dropout keep rate & 80\%--95\% & \textsc{lin} & N/A & 83\% & 92\%  & SPINN\\
+Classifier MLP dropout keep rate & 80\%--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & SPINN\\
+Tracking LSTM size $D_t$ & 24--128 & \textsc{log} & N/A & N/A & 61  & SPINN\\
+Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & SPINN\\
+\bottomrule
+\end{tabular}
+\end{center}
+
+\caption{
+\label{tab:hyperparams}
+Hyperparameter ranges and values.
+\ii{Range} shows the hyperparameter ranges explored during random search.
+\ii{Strategy} indicates whether sampling from the range was uniform, or log--uniform.
+\result{[SB] Full model hyperparams.}
+}
+\end{table*}
+
+\appendix
+\section{Hyperparameters}
+
+Table \ref{tab:hyperparams} shows the hyperparameters used in the best run of each model.
+
 \end{document}

### 2016-03-16 writing/hard_stack_paper/paper.tex
Initial revision of Sections 3, 4, and Bib.

@@
 \input{../model1_fig.tex}
+\todo{[SB,JG,AR,RG] Trim mercilessly whenever you see anything redundant.}
 
@@
 
-\citet{socher2011dynamic} presents a version of the TreeRNN model which is capable of operating over unparsed inputs.
+\citet{socher2011semi} presents a version of the TreeRNN model which is capable of operating over unparsed inputs.
 However, this method is unsupervised, limiting its performance, and it requires an expensive search process at test time.
@@
 \midrule
-100d LSTM encoders \citep{snli:emnlp2015}
+100D LSTM encoders \citep{snli:emnlp2015}
                         & 221k               & --               &   84.8   &   77.6      \\
-1024d pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
+1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                         & 15m                & --              &   98.8   &   81.4       \\
-300d Tree-based CNN encoders \citep{mou2015recognizing}
+300D Tree-based CNN encoders \citep{mou2015recognizing}
                         & 3.5m                & --             &   83.4   &   82.1       \\
@@
 \midrule
-300d LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
-300d SPINN-NP-NT encoders
+300D LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
+300D SPINN-NP-NT encoders
                         & 3.4m                  & --                &   84.4      &   80.9       \\
-300d SPINN-NP encoders
+300D SPINN-NP encoders
                         & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
-\result{300d SPINN encoders }
+\result{300D SPINN encoders }
                         & ??                  & 92.2+dv            &   ?    &   83.?      \\          
-% \result{300d SPINN-NP, sequence-based attn.   }     
+% \result{300D SPINN-NP, sequence-based attn.   }     
 %                         & ?                  & --                &   ?      &   86.3dv         \\
-% \result{300d SPINN-NP, tree-based attn. }           
+% \result{300D SPINN-NP, tree-based attn. }           
 %                         & ?                  & --                &   ?      &   \textbf{?}\\
@@
 \midrule
-100d LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
+100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
                         & 252k               & --              &   85.3   &   83.5       \\
-300d mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
+300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
                         & 1.9m               & --             &   92.0   &   86.1      \\
-300d LSTMN with deep attention fusion \citep{cheng2016long}
+300D LSTMN with deep attention fusion \citep{cheng2016long}
                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
@@
 We evaluate the SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE][]{dagan2006pascal,MacCartney09}.
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus \citet[SNLI][]{snli:emnlp2015}), which we use for training and evaluation:
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus \citep[SNLI][]{snli:emnlp2015}, which we use for training and evaluation:
 
-\begin{quote}
+\begin{quote}\small 
 Premise: {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}
@@
 
-\subsection{Models evaluated} \todo{[SB] Condense.}
+\subsection{Models evaluated}
 
-We evaluate four models on the SNLI task. Each uses 300d hidden states: \todo{[SB] d or D?}
+We evaluate four models on the SNLI task. Each uses 300D hidden states:
 \begin{itemize}
 \item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
-\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition.
+\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser and has no tracking LSTM to inform semantic composition.
 This model is equivalent to a TreeLSTM \citep{tai2015improved}.
-\item The SPINN-NP model, which includes a tracking LSTM which guides composition, making it a hybrid tree-sequence model.
-\item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time, making it slightly weaker but a practical choice in applied settings.
+\item The SPINN-NP model, which adds the tracking LSTM to guide composition, making it a hybrid tree-sequence model.
+\item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time.
 \end{itemize}
@@
 We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
-We should note that these attention-based models are by far the strongest published on SNLI to date, but we do not wish to compare our model with those models directly.
-\todo{[SB] But we do.
-Rephrase.
-Move before naming of models.} SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
-The attentional models are narrowly adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders.
-We focus on models which do support general sentence encoding in our analysis.
+We should note that these attention-based models are by far the strongest published on SNLI to date, but these are not necessarily comparable to ours: SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
+The attentional models are specifically adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders.
+We focus on models which do support general sentence encoding in our evaluation.
 
@@
 Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
-The full SPINN model also generates parsing predictions, and we report a measure of agreement \todo{[SB] Clarify.} between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations) \todo{[SB] Which Stanford parser.}.
+The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the predictions associated with the Stanford PCFG Parser parses included with SNLI (calculated as classification accuracy over \shift~and \reduce~operations).
 We also report the number of trained parameters in each model and training set accuracy.
@@
 The success of the SPINN-NP model, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
-
 The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline.
@@
 
-\todo{[SB] Discuss relationship between our model and the TBCNN. Add significance numbers.}
+Both the SPINN-NP and the full SPINN outperform all previous sentence encoding models by a significant margin.
+Most notably our modes outperform the TBCNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
+Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (TBCNN), which in turn outperforms one which does not use it at all (LSTM RNN).
 
@@
 }
-\caption{Runtime performance results with the thin-stack GPU algorithm.}
+\caption{Runtime performance results with the thin-stack GPU algorithm. \todo{[JG] Shrink a bit.}}
 \label{fig:speed}
@@
 
-\todo{Note that the full SPINN (i.e. Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
+\todo{[JG] Note that the full SPINN (i.e. Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
 
-\subsection{Parsing performance}
+\paragraph{Parsing performance}
 
@@
 \paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness on the SNLI task.
-However, the choice of whether to use soft attention and the choice of what representations to perform attention over are orthogonal \todo{[SB] Rephrase prev clause.}, and we expect that it should be possible to productively combine our model with soft attention to yield state-of-the-art performance.
+However, we expect that it should be possible to productively combine our model with soft attention, following \citet{cheng2016long}, to yield state-of-the-art performance.
 
 The tracking LSTM that we incorporated into the SPINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
-It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context of each tree node, including information about that nodes neighbors \todo{[SB] Rewrite.} in the sentence string and its likely position in the final tree, and that that context information could support both better parsing and better sentence encoding.
-One promising way to pursue this goal would be to encode the full contents of the stack and buffer at each time step following the method used by \citet{dyer-EtAl:2015:ACL-IJCNLP} for parsing.
+It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context at each tree node, yielding both better parsing and better sentence encoding.
+One promising way to pursue this goal would be to encode the full contents of the stack and buffer at each time step following the method used by \citet{dyer-EtAl:2015:ACL-IJCNLP}.
 
@@
 This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
-
-\todo{[SB] Fix run-ons here.}
 
@@
 \bibliography{../MLSemantics}
-\todo{[SB] Make bibliography style uniform, esp BatchNorm.}
 
-\begin{table*}[ht]\small
+\begin{table*}\small
 \begin{center}

### 2016-03-16 writing/hard_stack_paper/paper.tex
One last tweak.

@@
 
-We find that:
+This paper shows that:
 \begin{itemize}
-\item Our SPINN-NP-NT model is exactly equivalent to a tree-structured network in what it computes, but is 10--20 times faster than a standard tree network at inference time.
+\item The SPINN-NP-NT model is exactly equivalent to a tree-structured network in what it computes, but is 10--20 times faster than a standard tree network at inference time.
 \item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our base tree-sequence hybrid model (SPINN-NP) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.

### 2016-03-16 writing/hard_stack_paper/batching_fig.tex
Initial revision of abstract and Section 1.
Initial revision of Sections 3, 4, and Bib.
Writing up CP's comments.

@@
 \begin{subfigure}[t]{\columnwidth}
+\begin{center}
+\scalebox{0.75}{
  \begin{tikzpicture}
@@
 
-\end{tikzpicture}
+\end{tikzpicture}}
+\end{center}
 
@@
 
-\vspace{2em}
-
 \begin{subfigure}[t]{\columnwidth}
 \begin{center}
+\scalebox{0.75}{
  \begin{tikzpicture}
@@
 
-    \begin{scope}[shift={(0.9in,-0.8in)}]
+    \begin{scope}[shift={(1.5in,0in)}]
 
@@
 
-\end{tikzpicture}
+\end{tikzpicture}}
 \end{center}
@@
 \caption{An illustration of two standard designs for sentence encoders.
-Note that the TreeRNN, unlike the sequence-based RNN, requires a substantially different connection structure for each sentence, making batched computation impractical.}
+The TreeRNN, unlike the sequence-based RNN, requires a substantially different connection structure for each sentence, making batched computation impractical.}
 \end{figure}

### 2016-03-16 writing/hard_stack_paper/paper.tex
Results tweaks.
Writing up CP's comments.

@@
 Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations.
-In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.
+In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.
 
 This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}.
-The SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
+SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
 This design improves upon the TreeRNN architecture in three ways:
@@
 
-We evaluate the SPINN on the Stanford Natural Language Inference sentence understanding task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of upwards of 10x over comparable TreeRNN models.
+We evaluate SPINN on the Stanford Natural Language Inference sentence understanding task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of upwards of 10x over comparable TreeRNN models.
 
-\subsection{Related work}
+\section{Related work}
 
@@
 In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
-To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
+To our knowledge, SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
 
@@
 
-\section{Our model: The SPINN}
+\section{Our model: SPINN}
 
@@
 
-The SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g.
-a natural language sentence) by a single left-to-right scan over the tokens of the sequence.
-Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.][]{nivre2003efficient}.
+SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over the tokens.
+Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{nivre2003efficient}.
 
@@
 \begin{itemize}
-\item The parser's intermediate representations, including the representations of partial tree structures on the stack, are fixed-length vectors.
+\item The parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
 \item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.
@@
 
-This subsection describes the core functions and representations of the SPINN.
+This subsection describes the core functions and representations of SPINN.
 We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences using parses provided by some external parser.
@@
 Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
-
 \begin{equation}
@@
 \end{equation}
-
 \paragraph{The composition function}
-When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\ii composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
+When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
 This new node is then pushed on to the stack.
@@
 It is formulated as:
-
 \begin{gather}
@@
 \end{gather}
-
 The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack.
@@
 
-This paper evaluates the SPINN on a sentence pair classification task.
+This paper evaluates SPINN on a sentence pair classification task.
 To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed.
@@
 The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:
-
 \begin{equation}
@@
 \end{equation}
-
-Following \citep{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (the number of layers is a tuned hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (the number of layers is a tuned hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
 The model is trained in part using a cross-entropy objective over this distribution.
@@
 
-The tracking LSTM takes three vectors as inputs at each step (highlighted in yellow in Figure~\ref{fig:m1-views}): the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation.
+The tracking LSTM takes three vectors as inputs at each step (highlighted in yellow in Figure~\ref{fig:m1-views}): the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.
 
-\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) the SPINN will use needs to take as its input some representation of the current state of the model.
+\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) SPINN will use needs to take as its input some representation of the current state of the model.
 We use the hidden state of the tracking LSTM as that input.
@@
 
-It would be straightforward to augment the SPINN to support the use of some amount of right context as well, but this would add complexity to the model that we think is largely unnecessary: humans are very effective at understanding the beginnings of sentences before having seen or heard the ends, suggesting that it is possible to get by without the unavailable right context.
+It would be straightforward to augment SPINN to support the use of some amount of right context as well, but this would add complexity to the model that we think is largely unnecessary: humans are very effective at understanding the beginnings of sentences before having seen or heard the ends, suggesting that it is possible to get by without the unavailable right context.
 
@@
 
-The model described so far is the SPINN-NP model, which uses sequences of \shift~and \reduce~operations from an external parser.
+The model described so far is SPINN-NP (\ii{no parsing}), which uses sequences of \shift~and \reduce~operations from an external parser.
 To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
@@
 \end{equation}
-
-The resulting vector $\vec{p}_{\text{trans}}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep).
+The resulting vector $\vec{p}_{\text{trans}}$ is a probability distribution over parsing decisions (i.e., predicting  \shift~or \reduce~at the current timestep).
 In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution.
@@
 
-\subsection{Implementing the SPINN}
+\subsection{Implementing SPINN}
 
@@
 The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
-
 \begin{equation}
@@
 \end{equation}
-
-\paragraph{Preparing the data} At training time, the SPINN requires both a transition sequences $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence.
+\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence.
 The token sequence is simply the words in the sentence in order.
@@
 
-A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence.
+A na\"ive implementation of SPINN would require simulating a stack of size $N$ for each input sentence.
 In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations.
@@
 (Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).
+\todo{[JG] Shorten this a tiny bit.}
 
@@
 
-In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
+In its bare form, before the addition of the tracking LSTM, SPINN (in particular the SPINN-NP-NT variant, for \textit{no parsing, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
 In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
+The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context information.
 
@@
 \midrule
-300D LSTM RNN encoders          & 3.0m                  & --                &   83.9      &   80.6       \\
-300D SPINN-NP-NT encoders
+\textbf{300D LSTM RNN encoders}          & 3.0m                  & --                &   83.9      &   80.6       \\
+\textbf{300D SPINN-NP-NT encoders}
                         & 3.4m                  & --                &   84.4      &   80.9       \\
-300D SPINN-NP encoders
+\textbf{300D SPINN-NP encoders}
                         & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
-\result{300D SPINN encoders }
-                        & ??                  & 92.2+dv            &   ?    &   83.?      \\          
+\result{\textbf{300D SPINN encoders} }
+                        & ??                  & 92?           &   ?    &   82?      \\          
 % \result{300D SPINN-NP, sequence-based attn.   }     
@@
 
-We evaluate the SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE][]{dagan2006pascal,MacCartney09}.
+We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE][]{dagan2006pascal,MacCartney09}.
 NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus \citep[SNLI][]{snli:emnlp2015}, which we use for training and evaluation:
@@
 Even though NLI is framed as a relatively simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
-In order for a model to perform reliably well on NLI across a range of sentence types and text genres, it must be able to represent and reason with all of the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
+In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
 
@@
 \item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
-\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser and has no tracking LSTM to inform semantic composition.
+\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}), which uses transitions from an external parser and has no tracking LSTM to inform semantic composition.
 This model is equivalent to a TreeLSTM \citep{tai2015improved}.
-\item The SPINN-NP model, which adds the tracking LSTM to guide composition, making it a hybrid tree-sequence model.
-\item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time.
+\item SPINN-NP, which adds the tracking LSTM to guide composition, making it a hybrid tree-sequence model.
+\item The full SPINN, which makes its own parsing decisions, and does not depend on an external parser at test time.
 \end{itemize}
@@
 We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
-We should note that these attention-based models are by far the strongest published on SNLI to date, but these are not necessarily comparable to ours: SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
+These attention-based models are by far the strongest published on SNLI to date, but these are not necessarily comparable to ours: SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
 The attentional models are specifically adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders.
@@
 
-We find that the bare SPINN-NP-NT model performs little better than the RNN baseline, but that the SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
-The success of the SPINN-NP model, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
+We find that the bare SPINN-NP-NT model performs little better than the RNN baseline, but that SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
+The success of SPINN-NP, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
 The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline.
 None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task.
+\todo{[SB] Add context here.}
 
-Both the SPINN-NP and the full SPINN outperform all previous sentence encoding models by a significant margin.
+Both SPINN-NP and the full SPINN outperform all previous sentence encoding models by a significant margin.
 Most notably our modes outperform the TBCNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
@@
 
-We implemented the feedforward of the SPINN-NP-NT model (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution from \citet{irsoy2014deep}.
+We implemented the feedforward of SPINN-NP-NT (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution from \citet{irsoy2014deep}.
 Such a standard TreeRNN implementation operates on a single example at a time and is thus not feasible to run on a GPU.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as this is the more critical performance metric for real-world applications.}
@@
 Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}.
-Each model is restricted to run on sentences of 30 tokens or fewer (for the SPINN model, this means we iterate over a sequence of 59 transitions).
+Each model is restricted to run on sentences of 30 tokens or fewer (for SPINN, this means we iterate over a sequence of 59 transitions).
 We fix the model dimensionality and word embedding dimensionality at 300 for all tests.
@@
 
-\todo{[JG] Note that the full SPINN (i.e. Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
+\todo{[JG] Note that the full SPINN (i.e., Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
 
@@
 \begin{itemize}
-\item The SPINN-NP-NT model is exactly equivalent to a tree-structured network in what it computes, but is 10--20 times faster than a standard tree network at inference time.
+\item SPINN-NP-NT is exactly equivalent to a tree-structured network in what it computes, but is 10--20 times faster than a standard tree network at inference time.
 \item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our base tree-sequence hybrid model (SPINN-NP) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.
@@
 
-The tracking LSTM that we incorporated into the SPINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
+The tracking LSTM that we incorporated into SPINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
 It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context at each tree node, yielding both better parsing and better sentence encoding.
@@
 
-For a more ambitious goal, we expect that it should be possible to implement a variant of the SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citet{grefenstette2015learning,joulin2015inferring}).
+For a more ambitious goal, we expect that it should be possible to implement a variant of SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citet{grefenstette2015learning,joulin2015inferring}).
 This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
@@
 
-\begin{table*}\small
+\begin{table*}[t]\small
 \begin{center}
@@
 \midrule 
-Initial LR & 0.0002--0.02 & \textsc{log} & 0.005  & 0.0003 & 0.007  & SPINN\\
+Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & SPINN\\
 L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & SPINN\\

### 2016-03-16 writing/hard_stack_paper/paper.tex
Low hanging fruit for length reduction.

@@
 
-Of these, it is easy to argue that the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}.
+Of these, the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}.
 However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
@@
 SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
-This design improves upon the TreeRNN architecture in three ways:
-\begin{itemize}
-\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
-\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
-\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
-\end{itemize}
+This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
+In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
+Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
 
-We evaluate SPINN on the Stanford Natural Language Inference sentence understanding task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of upwards of 10x over comparable TreeRNN models.
+We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of over 10x over a standard TreeRNN implementation.
 
@@
 It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing.
-This relies on two key modifications to the shift-reduce algorithm:
-\begin{itemize}
-\item The parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
-\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.
-\end{itemize}
+This relies on two key modifications to the shift-reduce algorithm: First, the parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
+In addition, the \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.
 
@@
 
-Our composition is a version of the TreeLSTM of \citet{tai2015improved}.
-The TreeLSTM layer, which generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$.
-It is formulated as:
+The TreeLSTM composition function generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$.
+Our version is formulated as:
 \begin{gather}
-\colvec{4}
+\colvec{5}
     {\vec{i}}
@@
     {\vec{o}}
-= \sigma\left(
+    {\vec{g}}
+= \colvec{5}
+    {\sigma}
+    {\sigma}
+    {\sigma}
+    {\sigma}
+    {\text{tanh}}
+\left(
 W_{\text{iffo}}
@@
 \\
-\vec{g}
-= \tanh\left(
-W_{g}
-\colvec{3}
-    {\vec{h}_s^1}
-    {\vec{h}_s^2}
-    {\vec{e}}
-\right) \label{eqn:lstm2}
-\\
 \vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
@@
 \end{gather}
+$\sigma$ is the sigmoid activation function.
 The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack.
@@
 
-\paragraph{Creating a sentence pair classifier}
+\paragraph{Creating a sentence pair classifier} \label{sec:classifier}
 
@@
 The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
-It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equations~\ref{eqn:lstm1}--\ref{eqn:lstm2}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
 
@@
 
-\todo{[JG] Include figure of post-order traversal sequence made by the tracking LSTM?}
+\todo{[SB] Shorten the previous paras.}
 
@@
 \begin{equation}
-\vec{p}_{\text{trans}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{track}})
+\vec{p}_{\text{t}} = \text{softmax}(W_{\text{transition}}\vec{h}_{\text{tracking}})
 \end{equation}
-The resulting vector $\vec{p}_{\text{trans}}$ is a probability distribution over parsing decisions (i.e., predicting  \shift~or \reduce~at the current timestep).
+The resulting vector $\vec{p}_{\text{t}}$ is a probability distribution over parsing decisions (i.e., predicting  \shift~or \reduce~at the current timestep).
 In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution.
@@
 In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus.
-The classifier is trained using a cross-entropy objective:
-\begin{equation}
-  L_{\text{trans}} = - \log \vec{p}_{\text{trans}}[t^*]
-\end{equation}
-where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.
 
@@
 
-\paragraph{The full objective function} Our objective function combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentence and the hypothesis sentence at each step $t$, and an L2 regularization term on the trained parameters.
+\paragraph{The objective function} Our objective function combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentence and the hypothesis sentence at each step $t$, and an L2 regularization term on the trained parameters.
 The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
@@
 We used early stopping, evaluating on the test set using the parameters that performed best on the development set.
-
-\paragraph{Software infrastructure} We will make source code available.
 
@@
 
-Even though NLI is framed as a relatively simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
+Although NLI is framed as a simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
 In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
@@
 
-We evaluate four models on the SNLI task. Each uses 300D hidden states:
-\begin{itemize}
-\item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
-\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}), which uses transitions from an external parser and has no tracking LSTM to inform semantic composition.
-This model is equivalent to a TreeLSTM \citep{tai2015improved}.
-\item SPINN-NP, which adds the tracking LSTM to guide composition, making it a hybrid tree-sequence model.
-\item The full SPINN, which makes its own parsing decisions, and does not depend on an external parser at test time.
-\end{itemize}
+We evaluate four models on the SNLI task.
+Each uses the sentence pair classifier architecture described in Section~\ref{sec:classifier}, and they differ only in the function computing the sentence encodings.
+First, a single-layer LSTM RNN \citep[similar to that of][]{snli:emnlp2015} serves as a baseline.
+Next, the minimal SPINN-NP-NT model (exactly equivalent to a TreeLSTM) introduces the SPINN model design.
+SPINN-NP adds the tracking LSTM to that design.
+Finally, the full SPINN adds the integrated parser.
 
@@
 \begin{figure}
-\resizebox{\linewidth}{!}{
+\resizebox{0.85\linewidth}{!}{
 \begin{tikzpicture}
@@
 }
-\caption{Runtime performance results with the thin-stack GPU algorithm. \todo{[JG] Shrink a bit.}}
+\caption{Runtime performance results with the thin-stack GPU algorithm.}}
 \label{fig:speed}
@@
 
-Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \todo{cite}.
+Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \citep{socher2013recursive}.
 Each model is restricted to run on sentences of 30 tokens or fewer (for SPINN, this means we iterate over a sequence of 59 transitions).
-We fix the model dimensionality and word embedding dimensionality at 300 for all tests.
+We fix the model dimension and the word embedding dimension at 300.
 
@@
 The implementation is a simplified version of the TreeRNN from \citet{irsoy2014deep}\footnote{The original code for this model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code.}, modified to exactly match the computation performed by our GPU model.
+We plan to release our modified code alongside the full SPINN source code.}, modified to exactly match the computation performed by our GPU model.
 We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled.
@@
 
-This paper shows that:
-\begin{itemize}
-\item SPINN-NP-NT is exactly equivalent to a tree-structured network in what it computes, but is 10--20 times faster than a standard tree network at inference time.
-\item While the use of tree-structure alone does not offer a substantial improvement over an LSTM baseline, our base tree-sequence hybrid model (SPINN-NP) substantially outperforms that baseline, suggesting that our model is able to successfully exploit tree structure, while using sequential context information to better resolve ambiguity.
-\item It is possible to remove the dependence on an external parser from our model and replace it with a fast model-internal parser (as in the full SPINN) with only a small decrease in model performance.
-\end{itemize}
+In this paper, we first introduce a model architecture (SPINN-NP-NT) that is precisely equivalent to a TreeLSTM, but an order of magnitude faster at test time.
+Next, we expand that architecture into a tree-sequence hybrid model (SPINN-NP), and show that this yields significant gains on the SNLI entailment task.
+Finally, we show that it is possible exploit the strengths of this model without the need for an external parser by integrating a fast parser into the model itself, as in the full SPINN.
 
@@
 
-For a more ambitious goal, we expect that it should be possible to implement a variant of SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations (possibly following \citet{grefenstette2015learning,joulin2015inferring}).
+For a more ambitious goal, we expect that it should be possible to implement a variant of SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations \citep[as in][]{grefenstette2015learning,joulin2015inferring}.
 This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.

### 2016-03-16 writing/hard_stack_paper/paper.tex
Second wave of trimming.

@@
 
-\subsection{Shift-reduce parsing}
+\subsection{Background: Shift-reduce parsing}
 
@@
 
-A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below).
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below).
 In general a parser may also generate these predictions on-the-fly as it reads the tokens.
@@
 
-This subsection describes the core functions and representations of SPINN.
-We begin with the most basic model, SPINN-NP (for \textit{no parser}), which processes input sentences using parses provided by some external parser.
-
-\paragraph{Word representations}
-
-The buffer is populated with the words in each sentence before the first step of computation.
-We use word representations from the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}.
-We do not update these representations during training.
-Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
-\begin{equation}
-\colvec{2}
-    {\vec{h}}
-    {\vec{c}}
-= W_{\text{wd}} \vec{x}
-\end{equation}
 \paragraph{The composition function}
@@
 
-\paragraph{The contents of the stack and buffer}
+\paragraph{The stack and buffer}
 
 The stack and the buffer can be represented as lists of $N$ elements each (for sentences of up to $N$ words), with two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ constituting each element.
+
+\paragraph{Word representations}
+
+We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}.
+We do not update these representations during training.
+Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
+\begin{equation}
+\colvec{2}
+    {\vec{h}}
+    {\vec{c}}
+= W_{\text{wd}} \vec{x}_{\text{GloVe}}
+\end{equation}
 
@@
 This paper evaluates SPINN on a sentence pair classification task.
-To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed.
-This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition.
+To classify a sentence pair, we first construct a feature vector $\vec{x}_{\text{classifier}}$ based on the SPINN's representation for each of the two sentences---the value at the head of the stacks after the final transition.
 In particular, the $\vec{h}$ portions of these representations are used.
@@
 \end{equation}
-Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (the number of layers is a tuned hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
-The model is trained in part using a cross-entropy objective over this distribution.
+Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (or MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
 
@@
 In addition to the stack, the buffer, and the composition function, our full model includes an additional component: the tracking LSTM.
-The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
-It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
+This is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
 
-The tracking LSTM takes three vectors as inputs at each step (highlighted in yellow in Figure~\ref{fig:m1-views}): the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.
+The tracking LSTM's inputs (highlighted in yellow in Figure~\ref{fig:m1-views}) are: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.
 
-\paragraph{Tracking left context for parsing} The decision function which determines which transitions (and thereby, which parses) SPINN will use needs to take as its input some representation of the current state of the model.
-We use the hidden state of the tracking LSTM as that input.
-The tracking LSTM's limited access to the buffer makes this fast to compute, but somewhat more limited than the rich feature sets used in state-of-the-art parsers.
-
-\paragraph{Tracking left context for interpretation} Lexical ambiguity is ubiquitous in natural language.
-Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
+\paragraph{Why a tree-sequence hybrid?} Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
 Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which provides some cues for disambiguation.
@@
 
-\todo{[SB] Shorten the previous paras.}
+\subsection{Parsing: Predicting transitions}
 
-\subsection{Parsing: Predicting transitions from the tracking state}
-
-The model described so far is SPINN-NP (\ii{no parsing}), which uses sequences of \shift~and \reduce~operations from an external parser.
-To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
+For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply this as part of the input.
+To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
 \begin{equation}
@@
 \end{equation}
-The resulting vector $\vec{p}_{\text{t}}$ is a probability distribution over parsing decisions (i.e., predicting  \shift~or \reduce~at the current timestep).
-In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution.
-This prediction module is trained to mimic the decisions of an external parser.
-In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus.
+In the full SPINN model, we follow whichever transition (i.e., \shift~or \reduce) has a higher predicted probability in this distribution.
+This prediction function is trained to mimic the decisions of an external parser.
+In particular, we use the transition sequences corresponding to Stanford PCFG Parser constituency parses included with SNLI.
 
@@
 {\noindent\small
-{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) ) $\Rightarrow$\\
+{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
 {$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
@@
 
-\begin{quote}\small 
-Premise: {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}
-
-Hypothesis: {\it A girl outside plays in the snow.}
-
-Correct label: {\it entailment}
-\end{quote}
+\begin{description}\small 
+\item[\bf Premise:] {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}
+\item[\bf Hypothesis:] {\it A girl outside plays in the snow.}
+\item[\bf Correct label:] entailment
+\end{description}
 
@@
 Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
-The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the predictions associated with the Stanford PCFG Parser parses included with SNLI (calculated as classification accuracy over \shift~and \reduce~operations).
+The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the parses included with SNLI (calculated as classification accuracy over transitions averaged across timesteps).
 We also report the number of trained parameters in each model and training set accuracy.
@@
 }
-\caption{Runtime performance results with the thin-stack GPU algorithm.}}
+\caption{Runtime performance results with the thin-stack GPU algorithm.}
 \label{fig:speed}

### 2016-03-16 writing/hard_stack_paper/paper.tex
Paper cuts continued.

@@
 
-\paragraph{Optimization, initialization, and hyperparameters}
+\paragraph{Initialization, optimization, and tuning}
+
+We initialize the model parameters using the nonparametric strategy of \citet{DBLP:journals/corr/HeZR015}, with the exception of the softmax classifier parameters, which we initialize using random uniform samples from $[-0.005, 0.005]$.
 
 We use the RMSProp optimizer \citep{tieleman2012lecture} with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
-We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the word embeddings in the buffer (after the linear projection is applied) and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
-
-We initialize the model parameters using the nonparametric strategy of \citet{DBLP:journals/corr/HeZR015}, with the exception of the softmax classifier layers, which we initialize uniformly from $[-0.005, 0.005]$.
+We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the output of the word embedding projection layer and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
 
 We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and than launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
-An appendix shows the precise values used on the strongest models.
+Appendix~\ref{sec:hyperparams} shows the precise values used on the strongest models.
 
-We trained each model for 250k steps in each run, using a batch size of 32 for each step.
-We tracked each model's performance on the development set during training and saving parameters when this performance reached a new peak.
-We used early stopping, evaluating on the test set using the parameters that performed best on the development set.
+We train each model for 250k steps in each run, using a batch size of 32.
+We track each model's performance on the development set during training and save parameters when this performance reached a new peak.
+We use early stopping, evaluating on the test set using the parameters that perform best on the development set.
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \caption{The thin-stack algorithm}
@@
 
-\begin{table}
+\subsection{TreeRNN-equivalence}
+
+In its bare form, before the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
+In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
+The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context information.
+
+\begin{table}[t]
 \centering
@@
 \end{table}
-
-\subsection{TreeRNN-equivalence}
-
-In its bare form, before the addition of the tracking LSTM, SPINN (in particular the SPINN-NP-NT variant, for \textit{no parsing, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
-In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
-The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context information.
 
@@
 \textbf{300D LSTM RNN encoders}          & 3.0m                  & --                &   83.9      &   80.6       \\
-\textbf{300D SPINN-NP-NT encoders}
+\textbf{300D SPINN-PI-NT encoders}
                         & 3.4m                  & --                &   84.4      &   80.9       \\
-\textbf{300D SPINN-NP encoders}
+\textbf{300D SPINN-PI encoders}
                         & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
@@
                         & ??                  & 92?           &   ?    &   82?      \\          
-% \result{300D SPINN-NP, sequence-based attn.   }     
+% \result{300D SPINN-PI, sequence-based attn.   }     
 %                         & ?                  & --                &   ?      &   86.3dv         \\
-% \result{300D SPINN-NP, tree-based attn. }           
+% \result{300D SPINN-PI, tree-based attn. }           
 %                         & ?                  & --                &   ?      &   \textbf{?}\\
@@
 
-We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE][]{dagan2006pascal,MacCartney09}.
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the Stanford NLI corpus \citep[SNLI][]{snli:emnlp2015}, which we use for training and evaluation:
+We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE;][]{dagan2006pascal,MacCartney09}.
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use for training and evaluation:
 
@@
 
-We evaluate four models on the SNLI task.
-Each uses the sentence pair classifier architecture described in Section~\ref{sec:classifier}, and they differ only in the function computing the sentence encodings.
-First, a single-layer LSTM RNN \citep[similar to that of][]{snli:emnlp2015} serves as a baseline.
-Next, the minimal SPINN-NP-NT model (exactly equivalent to a TreeLSTM) introduces the SPINN model design.
-SPINN-NP adds the tracking LSTM to that design.
+We evaluate four models.
+The four all use the sentence pair classifier architecture described in Section~\ref{sec:classifier}, and differ only in the function computing the sentence encodings.
+First, a single-layer LSTM RNN \citep[similar to that of][]{snli:emnlp2015} serves as a baseline encoder.
+Next, the minimal SPINN-PI-NT model (equivalent to a TreeLSTM) introduces the SPINN model design.
+SPINN-PI adds the tracking LSTM to that design.
 Finally, the full SPINN adds the integrated parser.
 
-We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
+We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citet{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
 These attention-based models are by far the strongest published on SNLI to date, but these are not necessarily comparable to ours: SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
@@
 
-\paragraph{Results} Table~\ref{tab:results} shows our results.
-Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set.
-The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the parses included with SNLI (calculated as classification accuracy over transitions averaged across timesteps).
-We also report the number of trained parameters in each model and training set accuracy.
+\paragraph{Results} Table~\ref{tab:results} shows our results on SNLI inference classification.
+For the full SPINN, we also report a measure of agreement between this model's parses and the parses included with SNLI, calculated as classification accuracy over transitions averaged across timesteps.
 
-We find that the bare SPINN-NP-NT model performs little better than the RNN baseline, but that SPINN-NP with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
-The success of SPINN-NP, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
+We find that the bare SPINN-PI-NT model performs little better than the RNN baseline, but that SPINN-PI with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
+The success of SPINN-PI, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
 The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline.
@@
 
-Both SPINN-NP and the full SPINN outperform all previous sentence encoding models by a significant margin.
-Most notably our modes outperform the TBCNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
-Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (TBCNN), which in turn outperforms one which does not use it at all (LSTM RNN).
+Both SPINN-PI and the full SPINN outperform all previous sentence encoding models by a significant margin.
+Most notably our modes outperform the Tree-based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
+Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-based CNN), which in turn outperforms one which does not use it at all (LSTM RNN).
 
@@
 }
-\caption{Runtime performance results with the thin-stack GPU algorithm.}
+\caption{Feedforward speed comparison.}
 \label{fig:speed}
@@
 
-We implemented the feedforward of SPINN-NP-NT (which is equivalent to a standard TreeRNN) in C++/CUDA in order to compare to the performance of a standard CPU-based TreeRNN solution from \citet{irsoy2014deep}.
-Such a standard TreeRNN implementation operates on a single example at a time and is thus not feasible to run on a GPU.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as this is the more critical performance metric for real-world applications.}
+In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion.
+While the full models evaluated above are implemented and trained using Theano \todo{[SB] Cite Theano.}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
+To do this, we reimplemented the feedforward of SPINN-PI-NT in C++/CUDA, and we compare with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, modified to perform exactly the same computations as our model.\footnote{The original code for this model is available at \url{https://github.com/oir/deep-recursive}.
+We plan to release our modified code alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as this is the more critical performance metric for real-world applications.}
 
-Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \citep{socher2013recursive}.
 Each model is restricted to run on sentences of 30 tokens or fewer (for SPINN, this means we iterate over a sequence of 59 transitions).
 We fix the model dimension and the word embedding dimension at 300.
-
-The CPU reference model is implemented in C++ and uses a single-core Eigen linear algebra backend.
-The implementation is a simplified version of the TreeRNN from \citet{irsoy2014deep}\footnote{The original code for this model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code alongside the full SPINN source code.}, modified to exactly match the computation performed by our GPU model.
 We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled.
@@
 
+Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \citep{socher2013recursive}.
 At batch sizes of interest for test-time applications (e.g., 512 or 1024 examples), we observe a 10--20x speedup relative to the standard CPU implementation.
 Figure~\ref{fig:speed} shows a substantial difference in runtime scaling as batch sizes increase, which is typical of GPU computation.
-We hope that this improved batchable TreeRNN algorithm will enable further rapid research in tree-structured neural networks.
+We hope that this dramatically accelerated approach to tree-structured models will accelerate research into this model class.
 
-\todo{[JG] Note that the full SPINN (i.e., Model 1) is in the same complexity class as SPINN-NP (what we're evaluating here) with very similar constants.}
+\todo{[JG] Note that the full SPINN (i.e., Model 1) is in the same complexity class as SPINN-PI (what we're evaluating here) with very similar constants.}
 
@@
 
-In this paper, we first introduce a model architecture (SPINN-NP-NT) that is precisely equivalent to a TreeLSTM, but an order of magnitude faster at test time.
-Next, we expand that architecture into a tree-sequence hybrid model (SPINN-NP), and show that this yields significant gains on the SNLI entailment task.
+In this paper, we first introduce a model architecture (SPINN-PI-NT) that is precisely equivalent to a TreeLSTM, but an order of magnitude faster at test time.
+Next, we expand that architecture into a tree-sequence hybrid model (SPINN-PI), and show that this yields significant gains on the SNLI entailment task.
 Finally, we show that it is possible exploit the strengths of this model without the need for an external parser by integrating a fast parser into the model itself, as in the full SPINN.
@@
 \toprule
-Param.     & Range & Strategy        & RNN       & SPINN-NP-NT   & SPINN-NP  & SPINN \\
+Param.     & Range & Strategy        & RNN       & SPINN-PI-NT   & SPINN-PI  & SPINN \\
 \midrule 
@@
 \appendix
-\section{Hyperparameters}
+\section{Hyperparameters}\label{sec:hyperparams}
 

### 2016-03-16 writing/hard_stack_paper/paper.tex
Inputting CM/CP comments from today's meeting.

@@
 
-Of these, the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}.
+Of these, the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}.
 However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
@@
 Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
+\todo{[SB] Weaken parsing claim.}
 
@@
 
-\citet{socher2011semi} presents a version of the TreeRNN model which is capable of operating over unparsed inputs.
-However, this method is unsupervised, limiting its performance, and it requires an expensive search process at test time.
+\citet{socher2011semi} and \citet{socher2011parsing} present versions of the TreeRNN model which are capable of operating over unparsed inputs.
+However, these methods require an expensive search process at test time.
 Our model presents an alternative approach to combining interpretation and parsing that supports supervised training for the parsing component, and that is dramatically faster than any previous tree-structured model.
@@
 \left(
-W_{\text{iffo}}
+W_{\text{comp}}
 \colvec{3}
@@
     {\vec{e}}
++ \vec{b}_{\text{comp}}
 \right) \label{eqn:lstm1}
@@
     {\vec{c}}
-= W_{\text{wd}} \vec{x}_{\text{GloVe}}
+= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
 \end{equation}
@@
 \begin{equation}
-\vec{p}_{\text{t}} = \text{softmax}(W_{\text{transition}}\vec{h}_{\text{tracking}})
+\vec{p}_{\text{t}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{tracking}} + \vec{b}_{\text{trans}})
 \end{equation}
@@
 \textbf{300D LSTM RNN encoders}          & 3.0m                  & --                &   83.9      &   80.6       \\
-\textbf{300D SPINN-PI-NT encoders}
+\textbf{300D SPINN-PI-NT (parsed input, no ...) encoders}
                         & 3.4m                  & --                &   84.4      &   80.9       \\
@@
 is the model's accuracy in predicting parsing transitions.
-Train and test are SNLI classification accuracy.}
+Train and test are SNLI classification accuracy.
+\todo{Move attention results to footnote.}}
 \end{table*}
@@
 
-\question{What kind of error analysis would be interesting for this model?}
-
-\result{[SB] Compare the types of errors in the different runs.}
+\result{[SB] Compare the types of errors in the different runs and cherry pick interesting examples/tenancies; measure frequency of broken parses.}
 

### 2016-03-16 writing/hard_stack_paper/paper.tex
Cut back attention comparisons.

@@
 SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
-This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
+This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser.
 In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
 Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
-\todo{[SB] Weaken parsing claim.}
 
@@
 However, these methods require an expensive search process at test time.
-Our model presents an alternative approach to combining interpretation and parsing that supports supervised training for the parsing component, and that is dramatically faster than any previous tree-structured model.
+Our model presents a fast alternative approach.
 
@@
   \centering\small
-  \begin{tabular}{lcccc} 
+  \begin{tabular}{lrrrr} 
     \toprule
-Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
+Model                   & Params.    & Trans. acc. (\%)  &   Train acc. (\%)  &   Test acc. (\%) \\
 \midrule
-\multicolumn{5}{c}{Previous non-NN results}\\
-\midrule
+\multicolumn{5}{c}{\textbf{Previous non-NN results}}\\
 Lexicalized classifier \citep{snli:emnlp2015}
@@
 \midrule
-\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
-\midrule
+\multicolumn{5}{c}{\textbf{Previous sentence encoder-based NN results}}\\
 100D LSTM encoders \citep{snli:emnlp2015}
@@
 \midrule
-\multicolumn{5}{c}{Our results}\\
-\midrule
-\textbf{300D LSTM RNN encoders}          & 3.0m                  & --                &   83.9      &   80.6       \\
-\textbf{300D SPINN-PI-NT (parsed input, no ...) encoders}
+\multicolumn{5}{c}{\textbf{Our results}}\\
+300D LSTM RNN encoders         & 3.0m                  & --                &   83.9      &   80.6       \\
+300D SPINN-PI-NT (\ii{parsed input, no tracking}) encoders
                         & 3.4m                  & --                &   84.4      &   80.9       \\
-\textbf{300D SPINN-PI encoders}
-                        & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
-\result{\textbf{300D SPINN encoders} }
+300D SPINN-PI (\ii{parsed input}) encoders
+                        & 3.7m                  & --                &   89.2      &   \textbf{83.2}       \\
+\result{300D SPINN (unparsed input) encoders }
                         & ??                  & 92?           &   ?    &   82?      \\          
@@
 %                         & ?                  & --                &   ?      &   \textbf{?}\\
-\midrule
-\multicolumn{5}{c}{Other previous NN results}\\
-\midrule
-100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
-                        & 252k               & --              &   85.3   &   83.5       \\
-300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
-                        & 1.9m               & --             &   92.0   &   86.1      \\
-300D LSTMN with deep attention fusion \citep{cheng2016long}
-                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
+% \midrule
+% \multicolumn{5}{c}{Other previous NN results}\\
+% \midrule
+% 100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
+%                         & 252k               & --              &   85.3   &   83.5       \\
+% 300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
+%                         & 1.9m               & --             &   92.0   &   86.1      \\
+% 300D LSTMN with deep attention fusion \citep{cheng2016long}
+%                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
     \bottomrule
@@
 is the model's accuracy in predicting parsing transitions.
-Train and test are SNLI classification accuracy.
-\todo{Move attention results to footnote.}}
+Train and test are SNLI classification accuracy.}
 \end{table*}
@@
 
-We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citet{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures.
-These attention-based models are by far the strongest published on SNLI to date, but these are not necessarily comparable to ours: SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks.
-The attentional models are specifically adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders.
-We focus on models which do support general sentence encoding in our evaluation.
+We compare our models against several baselines, including the strongest published non-neural network-based result from \citet{snli:emnlp2015} and previous neural network models built around several types of sentence encoder.
 
@@
 The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline.
-None of our models reached the performance of the strongest attention-based models, which were explicitly engineered for the sentence pair-based structure of the SNLI task.
-\todo{[SB] Add context here.}
 
@@
 
-\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness on the SNLI task.
-However, we expect that it should be possible to productively combine our model with soft attention, following \citet{cheng2016long}, to yield state-of-the-art performance.
+\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citep{cheng2016long} have shown accuracies as high as 89.1\% on SNLI, but are more narrowly engineered to suit the task, and do not produce sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention, following \citet{cheng2016long}, to yield state-of-the-art performance.
 

### 2016-03-17 writing/hard_stack_paper/paper.tex
Finishing this round of cuts.

@@
 
-\subsection{Runtime performance}
+\subsection{Inference speed}
 \label{sec:speed}
@@
 While the full models evaluated above are implemented and trained using Theano \todo{[SB] Cite Theano.}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
-To do this, we reimplemented the feedforward of SPINN-PI-NT in C++/CUDA, and we compare with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, modified to perform exactly the same computations as our model.\footnote{The original code for this model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as this is the more critical performance metric for real-world applications.}
+To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
+We plan to release our modified code at publication time, alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as feedforward/inference speed is the relevant performance metric for most practical applications.}
 
@@
 Our GPU model is implemented in C++ and CUDA, and will be released with the rest of our code.
-We test its performance on an NVIDIA Titan X GPU.
+We test it on a NVIDIA Titan X GPU.
+\todo{[SB, JG] Is Titan X sufficiently unambiguous?}
 
 Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \citep{socher2013recursive}.
-At batch sizes of interest for test-time applications (e.g., 512 or 1024 examples), we observe a 10--20x speedup relative to the standard CPU implementation.
-Figure~\ref{fig:speed} shows a substantial difference in runtime scaling as batch sizes increase, which is typical of GPU computation.
+Figure~\ref{fig:speed} shows a substantial difference in runtime that increases with batch size, as is typical of GPU computation.
+At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--20x speedup relative to the standard CPU implementation.
 We hope that this dramatically accelerated approach to tree-structured models will accelerate research into this model class.
 
-\todo{[JG] Note that the full SPINN (i.e., Model 1) is in the same complexity class as SPINN-PI (what we're evaluating here) with very similar constants.}
+Though this experiment only covers SPINN-PI-NT, the results should be similar for the full SPINN model: most of the computation involved in running SPINN is involved in populating the buffer, applying the composition function, and manipulating the buffer and the stack, with the low-dimensional tracking and parsing components adding only a small additional load.
 
@@
 
-In this paper, we first introduce a model architecture (SPINN-PI-NT) that is precisely equivalent to a TreeLSTM, but an order of magnitude faster at test time.
+In this paper, we first introduce a model architecture (SPINN-PI-NT) that is equivalent to a TreeLSTM, but an order of magnitude faster at test time.
 Next, we expand that architecture into a tree-sequence hybrid model (SPINN-PI), and show that this yields significant gains on the SNLI entailment task.
-Finally, we show that it is possible exploit the strengths of this model without the need for an external parser by integrating a fast parser into the model itself, as in the full SPINN.
+Finally, we show that it is possible exploit the strengths of this model without the need for an external parser by integrating a fast parser into the model (as in the full SPINN), and that the lack of external parse information yields little loss in accuracy.
 
-\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citep{cheng2016long} have shown accuracies as high as 89.1\% on SNLI, but are more narrowly engineered to suit the task, and do not produce sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention, following \citet{cheng2016long}, to yield state-of-the-art performance.
+\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.1\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention, following \citet{cheng2016long}, to yield state-of-the-art performance.
 
@@
 \bibliography{../MLSemantics}
+
+\todo{[SB] Track down mystery spaces.}
 

### 2016-03-17 writing/hard_stack_paper/paper.tex
Last few technical fixes.

@@
 \input{../model1_fig.tex}
-\todo{[SB,JG,AR,RG] Trim mercilessly whenever you see anything redundant.}
 
@@
 In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion.
-While the full models evaluated above are implemented and trained using Theano \todo{[SB] Cite Theano.}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
+While the full models evaluated above are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
 To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
@@
 We test it on a NVIDIA Titan X GPU.
-\todo{[SB, JG] Is Titan X sufficiently unambiguous?}
 
@@
 
-\todo{[SB] Track down mystery spaces.}
-
 \begin{table*}[t]\small
@@
 \midrule 
-Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & SPINN\\
-L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & SPINN\\
-Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & N/A & N/A & N/A  & SPINN\\
-Embedding transformation dropout keep rate & 80\%--95\% & \textsc{lin} & N/A & 83\% & 92\%  & SPINN\\
-Classifier MLP dropout keep rate & 80\%--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & SPINN\\
-Tracking LSTM size $D_t$ & 24--128 & \textsc{log} & N/A & N/A & 61  & SPINN\\
-Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & SPINN\\
+Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & 2e-3\\
+L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & 3e-5\\
+Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & --- & --- & ---  & 3.9    \\
+Embedding transformation dropout keep rate & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
+Classifier MLP dropout keep rate & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
+Tracking LSTM size $D_t$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
+Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & 1\\
 \bottomrule
@@
 \label{tab:hyperparams}
-Hyperparameter ranges and values.
-\ii{Range} shows the hyperparameter ranges explored during random search.
-\ii{Strategy} indicates whether sampling from the range was uniform, or log--uniform.
-\result{[SB] Full model hyperparams.}
+Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log--uniform.
 }

### 2016-03-17 writing/hard_stack_paper/paper.tex
Draft discussion section, other tidying.

@@
 
-A na\"ive implementation of SPINN would require simulating a stack of size $N$ for each input sentence.
-In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations.
-This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$.
-(Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).
-\todo{[JG] Shorten this a tiny bit.}
+A na\"ive implementation of SPINN would require representing a stack of size $N$ for each timestep of each input sentence at training time to support backpropagation.
+This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$.
+(Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep, since \shift~and \reduce~operations only affect the top of the stack.
 
-We designed an alternative minimal-space stack algorithm inspired by the zipper technique \citep{huet1997zipper}.
+We propose an alternative space-efficient stack representation technique inspired by the zipper technique \citep{huet1997zipper}.
 For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$.
@@
 \result{300D SPINN (unparsed input) encoders }
-                        & ??                  & 92?           &   ?    &   82?      \\          
+                        & ??                  & 9?           &   87.2    &   82.6      \\          
 % \result{300D SPINN-PI, sequence-based attn.   }     
@@
 
-\result{[SB] Compare the types of errors in the different runs and cherry pick interesting examples/tenancies; measure frequency of broken parses.}
+The use of tree structure improves the performance of sentence encoding models for SNLI.
+We suspect that this improvement is largely attributable to the more efficient learning of accurate generalizations overall, and not to any small set of phenomena.
+However, some patterns are identifiable in the results.
+In particular, it seems that tree-structured models are better able to focus on the key actors and events named in a sentence (possibly by learning an approximation of the linguistic notion of headedness), and that this allows them to effectively inferences based on these key actors.
+For example, this pair was classified successfully by all three SPINN variants, but not by the RNN baseline (emphasis added):
 
-\vspace{10em}
+\begin{description}\small 
+\item[\bf Premise:] {\it A \textbf{woman} in red blouse is \textbf{standing} with small blond \textbf{child} in front of a small folding chalkboard.}
+\item[\bf Hypothesis:] {\it a woman stands with her child}
+\item[\bf Correct label:] neutral
+\end{description}
+
+We suspect that tree-sequence hybrid models in particular are especially effective at comparing sentences with very different structures, where it is useful to use strict compositionality with the tree-structured component to get the meanings of the setence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures.
+For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT:
+
+\begin{description}\small 
+\item[\bf Premise:] {\it Nurses in a medical setting conversing over a plastic cup.}
+\item[\bf Hypothesis:] {\it The nurses are discussing a patient.}
+\item[\bf Correct label:] neutral
+\end{description}
+
+We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
+We suspect that it is acting somewhat like the Tree-based CNN, only with access to more larger trees: Using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Trim short lines.

@@
 
+\newcommand{\snli}[3]{{\vspace{0.25em}
+{\small \setlength{\parindent}{0.6em} \hangindent=1.2em  \textbf{Premise:} #1\par}\vspace{0.25em}
+{\small \setlength{\parindent}{0.6em} \hangindent=1.2em   \textbf{Hypothesis:} #2\par}\vspace{0.25em}
+{\small \setlength{\parindent}{0.6em}  \textbf{Label:} #3\par}
+}}
+
 \noautomath
@@
 This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}.
-SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
+SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
 This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser.
@@
 
-The stack and the buffer can be represented as lists of $N$ elements each (for sentences of up to $N$ words), with two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ constituting each element.
+The stack and the buffer can be represented as arrays of $N$ elements each (for sentences of up to $N$ words), with the two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ in each element.
 
@@
 We do not update these representations during training.
-Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
+Instead, we use a learned linear transformation to transform each input word vector $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
 \begin{equation}
@@
 \paragraph{Why a tree-sequence hybrid?} Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
-Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which provides some cues for disambiguation.
+Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
 In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word.
@@
 
-We propose an alternative space-efficient stack representation technique inspired by the zipper technique \citep{huet1997zipper}.
+We propose an alternative space-efficient stack representation technique inspired by the zipper \citep{huet1997zipper}.
 For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$.
@@
                         & 3.7m                  & --                &   89.2      &   \textbf{83.2}       \\
-\result{300D SPINN (unparsed input) encoders }
-                        & ??                  & 9?           &   87.2    &   82.6      \\          
+300D SPINN (unparsed input) encoders
+                        & 2.7m                  & \result{}           &   87.2    &   82.6      \\          
 % \result{300D SPINN-PI, sequence-based attn.   }     
@@
 We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE;][]{dagan2006pascal,MacCartney09}.
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use for training and evaluation:
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use:
 
-\begin{description}\small 
-\item[\bf Premise:] {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}
-\item[\bf Hypothesis:] {\it A girl outside plays in the snow.}
-\item[\bf Correct label:] entailment
-\end{description}
+\snli{Girl in a red coat, blue head wrap and jeans is making a snow angel.}
+{A girl outside plays in the snow.}
+{entailment}
 
@@
 To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code at publication time, alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as feedforward/inference speed is the relevant performance metric for most practical applications.}
+We plan to release our modified code at publication time, alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference in C++/CUDA, as inference speed is the relevant performance metric for most practical applications.}
 
@@
 
-\begin{description}\small 
-\item[\bf Premise:] {\it A \textbf{woman} in red blouse is \textbf{standing} with small blond \textbf{child} in front of a small folding chalkboard.}
-\item[\bf Hypothesis:] {\it a woman stands with her child}
-\item[\bf Correct label:] neutral
-\end{description}
+\snli
+{A \textit{woman} in red blouse is \textit{standing} with small blond \textit{child} in front of a small folding chalkboard.}
+{a woman stands with her child}
+{neutral}
 
@@
 
-\begin{description}\small 
-\item[\bf Premise:] {\it Nurses in a medical setting conversing over a plastic cup.}
-\item[\bf Hypothesis:] {\it The nurses are discussing a patient.}
-\item[\bf Correct label:] neutral
-\end{description}
+\snli{Nurses in a medical setting conversing over a plastic cup.}
+{The nurses are discussing a patient.}
+{neutral}
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Major cleanup/trimming/overhaul.

@@
 
-A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence}.
-This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms.
-Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory \citep[LSTM,][see Figure~\ref{fig:batching:good}]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011semi}, which propagate information up a binary parse tree structure over words.
+A wide range of current models in NLP are built around a neural network component that produces vector representations of sentence meaning \citep{tai2015improved,sutskever2014sequence}.
+This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vectors to a sentence vector, and this function can take a range of different forms.
+Common sentence encoders include sequence-based recurrent neural network models (RNNs, see Figure~\ref{fig:batching:good}) with Long Short-Term Memory \citep[LSTM,][]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011parsing}, which propagate information up a binary parse tree.
 
-Of these, the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}.
+Of these, the TreeRNN is the principled choice, since meaning in natural language sentences is constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}.
 However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
@@
 This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}.
-SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly.
+SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces can produce the required parse structure on-the-fly.
 This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser.
 In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
-Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
+Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over pure sequence- or tree-based models.
 
@@
 
-\citet{socher2011semi} and \citet{socher2011parsing} present versions of the TreeRNN model which are capable of operating over unparsed inputs.
+\citet{socher2011parsing,socher2011semi} present versions of the TreeRNN model which are capable of operating over unparsed inputs.
 However, these methods require an expensive search process at test time.
@@
 
-SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over the tokens.
+SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over its tokens.
 Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{nivre2003efficient}.
 
-A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below).
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below).
 In general a parser may also generate these predictions on-the-fly as it reads the tokens.
@@
 The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed.
-We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$.
+The parser is initialized with the stack empty and with the buffer containing the tokens $\mathbf x$ of the sentence in order.
 Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state.
-We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations.
+We next proceed through the transition sequence, where each transition $t_t$ selects one of the two following operations.
 Below the $\mid$ symbol denotes the \textit{cons} (concatenation) operator.
@@
 It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing.
-This relies on two key modifications to the shift-reduce algorithm: First, the parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
-In addition, the \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.
+This relies on two key modifications to the shift-reduce algorithm.
+First, the parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
+In addition, the \reduce~operation combines two such vectors from the stack into another vector using a neural network function.
 
@@
 
-The TreeLSTM composition function generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$.
+The TreeLSTM composition function \citep{tai2015improved} generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$.
 Our version is formulated as:
@@
 \end{gather}
-$\sigma$ is the sigmoid activation function.
-The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack.
+where $\sigma$ is the sigmoid activation function and $\odot$ is the elementwise product.
 The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$.
 In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
-$\odot$ denotes the elementwise product.
-Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.
+The result of this function is the pair $\langle\vec{h}, \vec{c}\rangle$, which is placed back on the stack.
+Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_{\text{tracking}}$.
 
@@
 
-The stack and the buffer can be represented as arrays of $N$ elements each (for sentences of up to $N$ words), with the two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ in each element.
+The stack and the buffer are arrays of $N$ elements each (for sentences of up to $N$ words), with the two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ in each element.
 
@@
 
-We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}.
+We use word representations based on the standard 300D vector package provided with GloVe \citep{pennington2014glove}.
 We do not update these representations during training.
@@
 
-\paragraph{Creating a sentence pair classifier} \label{sec:classifier}
-
-This paper evaluates SPINN on a sentence pair classification task.
-To classify a sentence pair, we first construct a feature vector $\vec{x}_{\text{classifier}}$ based on the SPINN's representation for each of the two sentences---the value at the head of the stacks after the final transition.
-In particular, the $\vec{h}$ portions of these representations are used.
-The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:
-\begin{equation}
-\vec{x}_{\text{classifier}} = 
-\colvec{4}
-    {\vec{h}_{\text{premise}}}
-    {\vec{h}_{\text{hypothesis}}}
-    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
-    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
-\end{equation}
-Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (or MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
-
 \subsection{The tracking LSTM}\label{sec:tracking}
@@
 This is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
-It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, which is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
 
-The tracking LSTM's inputs (highlighted in yellow in Figure~\ref{fig:m1-views}) are: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.
+The tracking LSTM's inputs (highlighted in yellow in Figure~\ref{fig:m1-views}) are the top element of the buffer $\vec{h}_b^1$ (which would be moved in a \shift~operation) and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ (which would be composed in a \reduce~operation).
 
-\paragraph{Why a tree-sequence hybrid?} Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
+\paragraph{Why a tree-sequence hybrid?}
+
+Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
 Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
@@
 
-For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply this as part of the input.
+For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply is as part of the input.
 To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
@@
 \end{equation}
-In the full SPINN model, we follow whichever transition (i.e., \shift~or \reduce) has a higher predicted probability in this distribution.
-This prediction function is trained to mimic the decisions of an external parser.
-In particular, we use the transition sequences corresponding to Stanford PCFG Parser constituency parses included with SNLI.
-
-At training time, the model follows the transitions corresponding to the gold parse.
-At test time, a transition is chosen at each step by taking a hard max over the output of this function.
+At test time, the model uses whichever transition (i.e., \shift~or \reduce) is assigned a higher probability.
+The prediction function is trained to mimic the decisions of an external parser, and the decisions of this external parser are used as the inputs to the model during training.
+For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus for this purpose.
 We did not find the technique of scheduled sampling \citep{bengio2015scheduled}, or allowing the model to use its own transition decisions in some instances at training time, to be helpful.
 
-\subsection{Implementing SPINN}
+\subsection{Implementation issues}
 
-\paragraph{The objective function} Our objective function combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentence and the hypothesis sentence at each step $t$, and an L2 regularization term on the trained parameters.
-The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
-\begin{equation}
-\begin{split}
-\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
-\end{split}
-\end{equation}
-\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence.
-The token sequence is simply the words in the sentence in order.
-$\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:
-
-\vspace{0.5em}
-{\noindent\small
-{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
-{$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
-{$\mathbf x$}: the, cat, sat, down
-}
-
-\paragraph{Handling variable sentence lengths} For any sentence model to be trained with batched computation, it is necessary to pad or crop sentences to a fixed length.
-We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI.
-Transition sequences $\mathbf t$ are cropped at the left or padded at the left with \shift s.
-Token sequences $\mathbf x$ are then cropped or padded with empty tokens at the left to match the number of \shift s added or removed from $\mathbf t$, and can then be padded with empty tokens at the right to meet the desired length $N$.
-
-\paragraph{Initialization, optimization, and tuning}
-
-We initialize the model parameters using the nonparametric strategy of \citet{DBLP:journals/corr/HeZR015}, with the exception of the softmax classifier parameters, which we initialize using random uniform samples from $[-0.005, 0.005]$.
-
-We use the RMSProp optimizer \citep{tieleman2012lecture} with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
-We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the output of the word embedding projection layer and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
-
-We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and than launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
-Appendix~\ref{sec:hyperparams} shows the precise values used on the strongest models.
-
-We train each model for 250k steps in each run, using a batch size of 32.
-We track each model's performance on the development set during training and save parameters when this performance reached a new peak.
-We use early stopping, evaluating on the test set using the parameters that perform best on the development set.
-
-\subsection{Efficient stack representation}
+\paragraph{Representing the stack efficiently}
 
@@
 This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$.
-(Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep, since \shift~and \reduce~operations only affect the top of the stack.
+Such a naive implementation would also require copying a largely unchanged stack at each timestep, since \shift~and \reduce~operations only affect the top of the stack.
 
@@
 
+\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence.
+The token sequence is simply the words in the sentence in order.
+$\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:
+
+\vspace{0.5em}
+{\noindent\small
+{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
+{$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
+{$\mathbf x$}: the, cat, sat, down
+}
+
+\paragraph{Handling variable sentence lengths} For any sentence model to be trained with batched computation, it is necessary to pad or crop sentences to a fixed length.
+We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI.
+Transition sequences $\mathbf t$ are cropped at the left or padded at the left with \shift s.
+Token sequences $\mathbf x$ are then cropped or padded with empty tokens at the left to match the number of \shift s added or removed from $\mathbf t$, and can then be padded with empty tokens at the right to meet the desired length $N$.
+
 \subsection{TreeRNN-equivalence}
 
-In its bare form, before the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
+Without the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
 In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
-The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context information.
+The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context through the tracking LSTM.
 
@@
   \midrule
-  0 & $a$ & 0 \\
-  1 & $b$ & 0 1 \\
-  2 & $c$ & 0 1 2 \\
-  3 & $(c~b)$ & 0 3 \\
-  4 & $((c~b)~a)$ & 4 \\
+  0 & $a$ & \underline{\hphantom{0} 0} \\
+  1 & $b$ & \hphantom{0} \underline{0 1} \\
+  2 & $c$ & \hphantom{0} 0 \underline{1 2} \\
+  3 & $(c~b)$ & \hphantom{0} \underline{0 3} \\
+  4 & $((c~b)~a)$ & \underline{\hphantom{0} 4} \\
   \bottomrule
@@
 $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$.
-The last two elements of $Q$ specify the rows $t$ which should be involved in a \reduce~operation at the next timestep.}
+The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce~operation at the next timestep.}
 \label{tbl:thin-stack}
@@
 
-\section{Experiments}
+\subsection{Inference speed}
+\label{sec:speed}
+
+\begin{figure}
+\resizebox{0.85\linewidth}{!}{
+\begin{tikzpicture} 
+\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
+             xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
+  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
+  \addlegendentry{CPU \citep{irsoy2014deep}}
+  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
+  \addlegendentry{Thin-stack GPU}
+\end{axis}
+\end{tikzpicture}
+}
+\caption{Feedforward speed comparison.}
+\label{fig:speed}
+\end{figure}
+
+In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion.
+While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
+To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
+We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}
+
+Each model is restricted to run on sentences of 30 tokens or fewer.
+We fix the model dimension and the word embedding dimension at 300.
+We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled.
+Our GPU model is implemented in C++ and CUDA.
+We test it on a NVIDIA Titan X GPU.
+
+Figure~\ref{fig:speed} compares the sentence encoding speed of the two models using data from the Stanford Sentiment Treebank \citep{socher2013recursive}.
+Figure~\ref{fig:speed} shows a substantial difference in runtime that increases with batch size, as is typical of GPU computation.
+At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--20x speedup relative to the standard CPU implementation.
+
+Though this experiment only covers SPINN-PI-NT, the results should be similar for the full SPINN model: most of the computation involved in running SPINN is involved in populating the buffer, applying the composition function, and manipulating the buffer and the stack, with the low-dimensional tracking and parsing components adding only a small additional load.
+
+\section{NLI Experiments}
+
+We evaluate SPINN on the task of natural language inference \citep[NLI, a.k.a.
+recognizing textual entailment, or RTE;][]{dagan2006pascal}.
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use:
+
+\snli{Girl in a red coat, blue head wrap and jeans is making a snow angel.}
+{A girl outside plays in the snow.}
+{entailment}
+
+Although NLI is framed as a simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
+In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
+
+SNLI is a corpus of 570k human-labeled pairs of scene descriptions like the one above.
+We use the standard train--test split and ignore unlabeled examples, which leaves about 549k examples for training, 9,842 for development, and 9,824 for testing.
+SNLI labels are roughly balanced, with the most frequent label, {\it entailment}, making up 34.2\% of the test set.
+
+\subsection{Applying SPINN to SNLI}
+
+\paragraph{Creating a sentence pair classifier} \label{sec:classifier}
+
+To classify an SNLI sentence pair, we run two copies of SPINN with shared parameters: one on the premise sentence and another the hypothesis sentence.
+We then use their outputs (the $\vec{h}$ states at the top of each stack at time $t=T$) to construct a feature vector $\vec{x}_{\text{classifier}}$ for the pair.
+This feature vector consists of the concatenation of these two sentence vectors, their difference, and their elementwise product:
+\begin{equation}
+\vec{x}_{\text{classifier}} = 
+\colvec{4}
+    {\vec{h}_{\text{premise}}}
+    {\vec{h}_{\text{hypothesis}}}
+    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
+    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
+\end{equation}
+Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
 
@@
 acc.
-is the model's accuracy in predicting parsing transitions.
+is the model's accuracy in predicting parsing transitions at test time.
 Train and test are SNLI classification accuracy.}
@@
 
-\subsection{Natural language inference and SNLI}
+\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters.
+The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
+\begin{equation}
+\begin{split}
+\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
+\end{split}
+\end{equation}
 
-We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE;][]{dagan2006pascal,MacCartney09}.
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use:
+\paragraph{Initialization, optimization, and tuning}
 
-\snli{Girl in a red coat, blue head wrap and jeans is making a snow angel.}
-{A girl outside plays in the snow.}
-{entailment}
+We initialize the model parameters using the nonparametric strategy of \citet{DBLP:journals/corr/HeZR015}, with the exception of the softmax classifier parameters, which we initialize using random uniform samples from $[-0.005, 0.005]$.
 
-Although NLI is framed as a simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
-In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
+We use the RMSProp optimizer \citep{tieleman2012lecture} with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
+We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the output of the word embedding projection layer and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
 
-SNLI is a corpus of 570k human-labeled pairs of scene descriptions like the one above.
-We use the standard train--test split and ignore unlabeled examples, which leaves about 549k examples for training, 9,842 for development, and 9,824 for testing.
-SNLI labels are roughly balanced, with the most frequent label, {\it entailment}, making up 34.2\% of the test set.
+We train each model for 250k steps in each run, using a batch size of 32.
+We track each model's performance on the development set during training and save parameters when this performance reached a new peak.
+We use early stopping, evaluating on the test set using the parameters that perform best on the development set.
+
+Appendix~\ref{sec:hyperparams} discusses hyperparameter tuning.
 
@@
 
-\paragraph{Results} Table~\ref{tab:results} shows our results on SNLI inference classification.
+\subsection{Results}
+
+Table~\ref{tab:results} shows our results on SNLI inference classification.
 For the full SPINN, we also report a measure of agreement between this model's parses and the parses included with SNLI, calculated as classification accuracy over transitions averaged across timesteps.
@@
 The success of SPINN-PI, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
-The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless exceeds the performance of the LSTM baseline.
+The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless robustly exceeds the performance of the LSTM baseline.
 
@@
 Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-based CNN), which in turn outperforms one which does not use it at all (LSTM RNN).
-
-\subsection{Inference speed}
-\label{sec:speed}
-
-\begin{figure}
-\resizebox{0.85\linewidth}{!}{
-\begin{tikzpicture} 
-\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
-			 xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
-  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
-  \addlegendentry{CPU \citep{irsoy2014deep}}
-  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
-  \addlegendentry{Thin-stack GPU}
-\end{axis}
-\end{tikzpicture}
-}
-\caption{Feedforward speed comparison.}
-\label{fig:speed}
-\end{figure}
-
-In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion.
-While the full models evaluated above are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
-To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code at publication time, alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference in C++/CUDA, as inference speed is the relevant performance metric for most practical applications.}
-
-Each model is restricted to run on sentences of 30 tokens or fewer (for SPINN, this means we iterate over a sequence of 59 transitions).
-We fix the model dimension and the word embedding dimension at 300.
-We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled.
-Our GPU model is implemented in C++ and CUDA, and will be released with the rest of our code.
-We test it on a NVIDIA Titan X GPU.
-
-Figure~\ref{fig:speed} compares the runtime performance of the two models computing representations on data from the Stanford Sentiment Treebank \citep{socher2013recursive}.
-Figure~\ref{fig:speed} shows a substantial difference in runtime that increases with batch size, as is typical of GPU computation.
-At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--20x speedup relative to the standard CPU implementation.
-We hope that this dramatically accelerated approach to tree-structured models will accelerate research into this model class.
-
-Though this experiment only covers SPINN-PI-NT, the results should be similar for the full SPINN model: most of the computation involved in running SPINN is involved in populating the buffer, applying the composition function, and manipulating the buffer and the stack, with the low-dimensional tracking and parsing components adding only a small additional load.
 
@@
 
-\section{Discussion}
+\subsection{Discussion}
 
 The use of tree structure improves the performance of sentence encoding models for SNLI.
-We suspect that this improvement is largely attributable to the more efficient learning of accurate generalizations overall, and not to any small set of phenomena.
+We suspect that this improvement is largely attributable to the more efficient learning of accurate generalizations overall, and not to any particular few phenomena.
 However, some patterns are identifiable in the results.
-In particular, it seems that tree-structured models are better able to focus on the key actors and events named in a sentence (possibly by learning an approximation of the linguistic notion of headedness), and that this allows them to effectively inferences based on these key actors.
+In particular, it seems that tree-structured models are better able to focus on the key actors and events named in a sentence (possibly by learning an approximation of the linguistic notion of headedness), and that this allows them to better handle inferences based on these key actors.
 For example, this pair was classified successfully by all three SPINN variants, but not by the RNN baseline (emphasis added):
@@
 
-We suspect that tree-sequence hybrid models in particular are especially effective at comparing sentences with very different structures, where it is useful to use strict compositionality with the tree-structured component to get the meanings of the setence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures.
-For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT:
+The tree-sequence hybrid models seem to be especially effective at comparing sentences with very different structures, where it is useful to use strict compositionality with the tree-structured component to get the meanings of the sentence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures.
+For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT: \todo{[SB] Lexical ambiguity example/comment.}
 
@@
 
-\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.1\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention, following \citet{cheng2016long}, to yield state-of-the-art performance.
+\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.1\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 
-The tracking LSTM that we incorporated into SPINN uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
+Our tracking LSTM uses only simple, quick-to-compute features drawn from the head of the buffer and the head of the stack.
 It is plausible that giving the tracking LSTM access to more information from the buffer and stack at each step would allow it to better represent the context at each tree node, yielding both better parsing and better sentence encoding.
@@
 Classifier MLP dropout keep rate & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
-Tracking LSTM size $D_t$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
+Tracking LSTM size $D_\text{tracking}$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
 Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & 1\\
@@
 
-Table \ref{tab:hyperparams} shows the hyperparameters used in the best run of each model.
+We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and than launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
+Table~\ref{tab:hyperparams} below shows the hyperparameters used in the best run of each model.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Tiny typos

@@
 
-We compare our models against several baselines, including the strongest published non-neural network-based result from \citet{snli:emnlp2015} and previous neural network models built around several types of sentence encoder.
+We compare our models against several baselines, including the strongest published non-neural network-based result from \citet{snli:emnlp2015} and previous neural network models built around several types of sentence encoders.
 
@@
 Both SPINN-PI and the full SPINN outperform all previous sentence encoding models by a significant margin.
-Most notably our modes outperform the Tree-based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
+Most notably our models outperform the Tree-based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
 Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-based CNN), which in turn outperforms one which does not use it at all (LSTM RNN).

### 2016-03-18 writing/hard_stack_paper/paper.tex
Center speed figure
Correct speed factor: 25x!

@@
 We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single tree-sequence hybrid model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser.
-Our model supports batched computation for a speedup of up to 20x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
+Our model supports batched computation for a speedup of up to 25x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
 We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
@@
 \begin{figure}
+\centering
 \resizebox{0.85\linewidth}{!}{
@@
 Figure~\ref{fig:speed} shows a substantial difference in runtime that increases with batch size, as is typical of GPU computation.
-At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--20x speedup relative to the standard CPU implementation.
+At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--25x speedup relative to the standard CPU implementation.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Simplify redundant sentence and save a line

@@
 This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vectors to a sentence vector, and this function can take a range of different forms.
-Common sentence encoders include sequence-based recurrent neural network models (RNNs, see Figure~\ref{fig:batching:good}) with Long Short-Term Memory \citep[LSTM,][]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011parsing}, which propagate information up a binary parse tree.
+Common sentence encoders include sequence-based recurrent neural network models (RNNs, see Figure~\ref{fig:batching:good}) with Long Short-Term Memory \citep[LSTM,][]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words or characters, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011parsing}, which propagate information up a binary parse tree.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Typo: more larger -> larger

@@
 We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
-We suspect that it is acting somewhat like the Tree-based CNN, only with access to more larger trees: Using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
+We suspect that it is acting somewhat like the Tree-based CNN, only with access to larger trees: Using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Tighten up intro prose

@@
 Of these, the TreeRNN is the principled choice, since meaning in natural language sentences is constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}.
-However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
+TreeRNNs have shown some promise \citep{tai2015improved,li2015tree,bowman2015trees}, but have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
 Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets.

### 2016-03-18 writing/hard_stack_paper/paper.tex
Capitalization, removing initials for first submission, adding title page.

@@
 
-\title{A fast unified model for parsing and sentence understanding}
+\title{A Fast Unified Model for Parsing and Sentence Understanding}
 
@@
 300D SPINN (unparsed input) encoders
-                        & 2.7m                  & \result{}           &   87.2    &   82.6      \\          
+                        & 2.7m                  & 92.4           &   87.2    &   82.6      \\          
 % \result{300D SPINN-PI, sequence-based attn.   }     
@@
 
-\todo{[AR] Fill in some details here.}
-\todo{[SB] Read AR's report and work out what to include.}
+\todo{[Fill in some details here.}
+\todo{Read report and work out what to include.}
 
@@
 The tree-sequence hybrid models seem to be especially effective at comparing sentences with very different structures, where it is useful to use strict compositionality with the tree-structured component to get the meanings of the sentence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures.
-For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT: \todo{[SB] Lexical ambiguity example/comment.}
+For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT: \todo{Lexical ambiguity example/comment.}
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
typo fix

@@
 This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}.
-SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces can produce the required parse structure on-the-fly.
+SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that can produce the required parse structure on-the-fly.
 This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser.

### 2016-03-18 writing/hard_stack_paper/paper.tex
Make the speed claim as strong as possible here, too

@@
 
-We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of over 10x over a standard TreeRNN implementation.
+We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of up to 25x over a standard TreeRNN implementation.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Minor improvements in shift-reduce section

@@
 
-A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below).
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies a parser transition (described below).
 In general a parser may also generate these predictions on-the-fly as it reads the tokens.
-It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
-For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.
+It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
+For any binary-branching tree structure over $N$ words, this requires $2N - 1$ transitions.
 
-The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed.
+The parser uses two auxiliary data structures: a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed.
 The parser is initialized with the stack empty and with the buffer containing the tokens $\mathbf x$ of the sentence in order.

### 2016-03-18 writing/hard_stack_paper/paper.tex
Trim section 3.2

@@
 It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing.
-This relies on two key modifications to the shift-reduce algorithm.
-First, the parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
-In addition, the \reduce~operation combines two such vectors from the stack into another vector using a neural network function.
+SPINN is an extension of the shift-reduce model whose intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
+Its \reduce~operation combines two vector representations from the stack into another vector using a neural network function.
 
 \paragraph{The composition function}
-When a \reduce~operation is performed, vector representations of two tree nodes are popped off of the head of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
+When a \reduce~operation is performed, the vector representations of two tree nodes are popped off of the head of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
 This new node is then pushed on to the stack.

### 2016-03-18 writing/hard_stack_paper/paper.tex
More section 3 edits

@@
 
-Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
+Lexical ambiguity is ubiquitous in natural language.
+Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
 Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
@@
 
-We propose an alternative space-efficient stack representation technique inspired by the zipper \citep{huet1997zipper}.
+We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}.
 For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$.
@@
 It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation.
-Furthermore, all of the updates to $S$ and $Q$ can be performed in-place on a GPU.
-These features allow this stack model to run efficiently on a GPU.
+Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains.
 We describe speed results in Section~\ref{sec:speed}.

### 2016-03-18 writing/hard_stack_paper/paper.tex
Proofreading the rest of the paper.

@@
 \end{equation}
-Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.
+Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution over the three labels.
 
@@
 We train each model for 250k steps in each run, using a batch size of 32.
-We track each model's performance on the development set during training and save parameters when this performance reached a new peak.
+We track each model's performance on the development set during training and save parameters when this performance reaches a new peak.
 We use early stopping, evaluating on the test set using the parameters that perform best on the development set.
@@
 Both SPINN-PI and the full SPINN outperform all previous sentence encoding models by a significant margin.
-Most notably our models outperform the Tree-based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
-Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-based CNN), which in turn outperforms one which does not use it at all (LSTM RNN).
+Most notably our models outperform the Tree-Based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
+Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-Based CNN), which in turn outperforms one which does not use it at all (LSTM RNN).
 
@@
 We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
-We suspect that it is acting somewhat like the Tree-based CNN, only with access to larger trees: Using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
+We suspect that it is acting somewhat like the Tree-Based CNN, only with access to larger trees: Using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
s/and than/and then/

@@
 
-We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and than launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
+We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and then launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
 Table~\ref{tab:hyperparams} below shows the hyperparameters used in the best run of each model.

### 2016-03-18 writing/hard_stack_paper/paper.tex
More typo fixes

@@
 
-For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply is as part of the input.
+For SPINN to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply it as part of the input.
 To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:

### 2016-03-18 writing/titlepage.tex

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Title page for:\\A Fast Unified Model for Parsing and Sentence Understanding}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle
A version of this paper will be posted as a non-archival manuscript to arXiv.org under the same title.
\end{document}

### 2016-03-18 writing/hard_stack_paper/paper.tex
Save a line
Space freeup in section 3

@@
 The parser uses two auxiliary data structures: a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed.
-The parser is initialized with the stack empty and with the buffer containing the tokens $\mathbf x$ of the sentence in order.
+The parser is initialized with the stack empty and the buffer containing the tokens $\mathbf x$ of the sentence in order.
 Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state.
@@
 \paragraph{The composition function}
-When a \reduce~operation is performed, the vector representations of two tree nodes are popped off of the head of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
-This new node is then pushed on to the stack.
+When a \reduce~operation is performed, the vector representations of two tree nodes are popped off of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
+This new node is pushed on to the stack.
 
@@
 where $\sigma$ is the sigmoid activation function and $\odot$ is the elementwise product.
-The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$.
-In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
-The result of this function is the pair $\langle\vec{h}, \vec{c}\rangle$, which is placed back on the stack.
-Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_{\text{tracking}}$.
+The pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack.
+In addition, $\vec{e}$ is an optional vector valued input argument which is either empty or from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
+The result of this function, the pair $\langle\vec{h}, \vec{c}\rangle$, is placed back on the stack.
+Each vector-valued variable listed is of the same dimension $D$ except $\vec{e}$, of the independent dimension $D_{\text{tracking}}$.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Summarize the parsing results.

@@
 
-\paragraph{Parsing performance}
-
-\todo{[Fill in some details here.}
-\todo{Read report and work out what to include.}
+The full SPINN performed moderately well at reproducing the Stanford Parser's parses of the SNLI data at a transition-by-transition level with 92.4\% accuracy at test time.
+However, its transition prediction errors were fairly evenly distributed across sentences, and most sentences were assigned partially invalid transition sequences that either left a few words out of the final representation or incorporated a few padding tokens into the final representation.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
A couple of edits-to-edits. More in 2h.

@@
 It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing.
-SPINN is an extension of the shift-reduce model whose intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
+SPINN is an extension of the shift-reduce model in which its intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
 Its \reduce~operation combines two vector representations from the stack into another vector using a neural network function.
@@
 The pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack.
-In addition, $\vec{e}$ is an optional vector valued input argument which is either empty or from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
+In addition, $\vec{e}$ is an optional vector valued input argument which is either empty or comes from from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
 The result of this function, the pair $\langle\vec{h}, \vec{c}\rangle$, is placed back on the stack.
-Each vector-valued variable listed is of the same dimension $D$ except $\vec{e}$, of the independent dimension $D_{\text{tracking}}$.
+Each vector-valued variable listed is of dimension $D$ except $\vec{e}$, of the independent dimension $D_{\text{tracking}}$.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Nixing TODO.

@@
 SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over its tokens.
-Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{nivre2003efficient}.
+Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{shieber:1983:ACL,nivre2003efficient}.
 
@@
 The tree-sequence hybrid models seem to be especially effective at comparing sentences with very different structures, where it is useful to use strict compositionality with the tree-structured component to get the meanings of the sentence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures.
-For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT: \todo{Lexical ambiguity example/comment.}
+For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT:
 
@@
 %    
-%    Some of the Tesla K40(s) used for this research was/were donated by the NVIDIA Corporation.
-%    \todo{[CM,CP] Acknowledge other grants.}
+%    We acknowledge financial support from a Google Faculty Research Award, the Stanford Data Science Initiative, and the National Science Foundation under grant no.~BCS 1456077.
+Some of the Tesla K40s used for this research were donated by the NVIDIA Corporation.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Corrected attn number.

@@
 
-\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.1\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
+\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Edits to pp. 1-4.

@@
 Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences.
-However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they can only operate on parsed sentences and do not directly support batched computation.
+However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they can only operate on parsed sentences and they do not directly support batched computation.
 We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single tree-sequence hybrid model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser.
-Our model supports batched computation for a speedup of up to 25x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy.
+Our model supports batched computation for a speedup of up to 25x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with little loss of accuracy.
 We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
@@
 
-A wide range of current models in NLP are built around a neural network component that produces vector representations of sentence meaning \citep{tai2015improved,sutskever2014sequence}.
+A wide range of current models in NLP are built around a neural network component that produces vector representations of sentence meaning \citep[e.g.,][]{sutskever2014sequence,tai2015improved}.
 This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vectors to a sentence vector, and this function can take a range of different forms.
@@
 
-Of these, the TreeRNN is the principled choice, since meaning in natural language sentences is constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}.
-TreeRNNs have shown some promise \citep{tai2015improved,li2015tree,bowman2015trees}, but have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
+Of these, the TreeRNN appears to be the principled choice, since meaning in natural language sentences is known to be constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}.
+TreeRNNs have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, but have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
 Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets.
-Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations.
+Because TreeRNNs use a different model structure for each sentence, as in Figure~\ref{fig:batching}, batching is impossible in standard implementations.
 In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.
@@
 This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}.
-SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that can produce the required parse structure on-the-fly.
+SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces the required parse structure on the fly.
 This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser.
@@
 Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over pure sequence- or tree-based models.
+% TRIMMABLE
 
@@
 
-There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
-In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well.
+There is a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a variant \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
+In addition, there has been recent work proposing models designed primarily for generative language modeling tasks that use these structures as well \citep{zhang2016top,dyer2016rnn}.
 To our knowledge, SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
+%TRIMMABLE
 
@@
 SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over its tokens.
-Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{shieber:1983:ACL,nivre2003efficient}.
+Shift-reduce parsing is widely used in natural language parsing \citep[e.g.,][]{shieber:1983:ACL,nivre2003efficient}.
 
-A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies a parser transition (described below).
-In general a parser may also generate these predictions on-the-fly as it reads the tokens.
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and consumes transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies one step of the parsing process.
+In general a parser may also generate these transitions on the fly as it reads the tokens.
 It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
@@
 Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state.
-We next proceed through the transition sequence, where each transition $t_t$ selects one of the two following operations.
+It next proceeds through the transition sequence, where each transition $t_t$ selects one of the two following operations.
 Below the $\mid$ symbol denotes the \textit{cons} (concatenation) operator.
 We arbitrarily choose to always \textit{cons} on the left in the notation below.
+%TRIMMABLE
 \begin{description}
@@
 
-The core SPINN model implements a sequential shift-reduce parser which operates over sentences.
-It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing.
-SPINN is an extension of the shift-reduce model in which its intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors.
-Its \reduce~operation combines two vector representations from the stack into another vector using a neural network function.
+SPINN is based on a shift-reduce parser, but it is designed to produce a vector representation of a sentence as its output, rather than a tree as in standard shift-reduce parsing.
+It modifies the shift-reduce formalism by using fixed length vectors to represent each entry in the stack and the buffer.
+Correspondingly, its \reduce~operation combines two vector representations from the stack into another vector using a neural network function.
 
@@
 \end{gather}
-where $\sigma$ is the sigmoid activation function and $\odot$ is the elementwise product.
-The pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack.
-In addition, $\vec{e}$ is an optional vector valued input argument which is either empty or comes from from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
+where $\sigma$ is the sigmoid activation function, $\odot$ is the elementwise product, the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack, and $\vec{e}$ is an optional vector valued input argument which is either empty or comes from from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
 The result of this function, the pair $\langle\vec{h}, \vec{c}\rangle$, is placed back on the stack.
@@
 We do not update these representations during training.
-Instead, we use a learned linear transformation to transform each input word vector $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
+Instead, we use a learned linear transformation to map each input word vector $\vec{x}_{\text{GloVe}}$ into a vector pair $\langle \vec{h}, \vec{c}\rangle$ that is stored in the buffer:
 \begin{equation}
@@
 
-The tracking LSTM's inputs (highlighted in yellow in Figure~\ref{fig:m1-views}) are the top element of the buffer $\vec{h}_b^1$ (which would be moved in a \shift~operation) and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ (which would be composed in a \reduce~operation).
+The tracking LSTM's inputs (yellow in Figure~\ref{fig:m1-views}) are the top element of the buffer $\vec{h}_b^1$ (which would be moved in a \shift~operation) and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ (which would be composed in a \reduce~operation).
 
@@
 For SPINN to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply it as part of the input.
-To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
+To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier whose input is the state of the tracking LSTM:
 \begin{equation}
@@
 For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus for this purpose.
-We did not find the technique of scheduled sampling \citep{bengio2015scheduled}, or allowing the model to use its own transition decisions in some instances at training time, to be helpful.
+We did not find the technique of scheduled sampling \citep{bengio2015scheduled}---allowing the model to use its own transition decisions in some instances at training time---to be helpful.
 
@@
 This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$.
-Such a naive implementation would also require copying a largely unchanged stack at each timestep, since \shift~and \reduce~operations only affect the top of the stack.
+Such a naive implementation would also require copying a largely unchanged stack at each timestep, since each \shift~or \reduce~operation writes only one new representation to the stack.
 
@@
 We describe speed results in Section~\ref{sec:speed}.
+
+A simpler variant of this technique can be used to represent the buffer, since information is not written to the buffer during model operation.
+It can be stored as a single fixed matrix with a scalar pointer variable indicating which element is the head at each step.
 
@@
 Lexicalized classifier \citep{snli:emnlp2015}
-                        & --                & --                    &   99.7   &   78.2      \\
+                        & ---                & ---                    &   99.7   &   78.2      \\
 \midrule
@@
 100D LSTM encoders \citep{snli:emnlp2015}
-                        & 221k               & --               &   84.8   &   77.6      \\
+                        & 221k               & ---               &   84.8   &   77.6      \\
 1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
-                        & 15m                & --              &   98.8   &   81.4       \\
+                        & 15m                & ---              &   98.8   &   81.4       \\
 300D Tree-based CNN encoders \citep{mou2015recognizing}
-                        & 3.5m                & --             &   83.4   &   82.1       \\
+                        & 3.5m                & ---             &   83.4   &   82.1       \\
 \midrule
 \multicolumn{5}{c}{\textbf{Our results}}\\
-300D LSTM RNN encoders         & 3.0m                  & --                &   83.9      &   80.6       \\
+300D LSTM RNN encoders         & 3.0m                  & ---                &   83.9      &   80.6       \\
 300D SPINN-PI-NT (\ii{parsed input, no tracking}) encoders
-                        & 3.4m                  & --                &   84.4      &   80.9       \\
+                        & 3.4m                  & ---                &   84.4      &   80.9       \\
 300D SPINN-PI (\ii{parsed input}) encoders
-                        & 3.7m                  & --                &   89.2      &   \textbf{83.2}       \\
+                        & 3.7m                  & ---                &   89.2      &   \textbf{83.2}       \\
 300D SPINN (unparsed input) encoders
@@
 
-\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
+\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0  \% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 

### 2016-03-18 metal_stack/thin-stack.cc
Add perf note

@@ -288,6 +288,10 @@ void ThinStack::update_buffer_cur(float *buffer_cur_t, float *transitions, int t
 
 
 void ThinStack::reset() {
+  // TODO: Technically these don't need to be explicitly zeroed out before
+  // every feedforward. They just get overwritten and their bad values are
+  // never used, provided that the feedforward uses a valid transition
+  // sequence.
   cudaMemset(stack, 0, stack_total_size * sizeof(float));
   cudaMemset(queue, 0, queue_total_size * sizeof(float));
 

### 2016-03-18 metal_stack/testrun.cc
Replace the equality check with perf code. People will want to run that more, I assume..

@@ -30,7 +30,7 @@ void destroy_params(ThinStackParameters params) {
 }
 
 int main() {
-  ModelSpec spec = {50, 50, 1, 10, 71, 50};//{300, 300, 256, 10, 25, 300};//{5, 5, 2, 10, 3, 5};
+  ModelSpec spec = {300, 300, (size_t) atoi(getenv("BATCH_SIZE")), 10, 59, 300};
   ThinStackParameters params = load_params(spec);
 
   cublasHandle_t handle;
@@ -45,19 +45,29 @@ int main() {
   // Set model inputs.
   cout << "X:" << endl;
   int num_tokens = (spec.seq_length + 1) / 2;
-  load_weights_cuda("/scr/jgauthie/projects/rembed/metal_stack/sst_one_buffer.txt", spec.model_dim * spec.batch_size * num_tokens, ts.X);
-/* #if DEBUG */
-/*   print_device_matrix(ts.X, spec.model_dim, spec.batch_size * num_tokens); */
-/* #endif */
+  fill_rand_matrix(ts.X, spec.model_dim, spec.batch_size * num_tokens);
 
   cout << "transitions:" << endl;
-  load_weights_cuda("/scr/jgauthie/tmp/deep-recursive/models/drsv_1_50_0_0.1_1_200_0.002_0.0001_0.9_0.transitions", spec.batch_size * spec.seq_length, ts.transitions);
+  float *h_transitions = (float *) malloc(spec.seq_length * spec.batch_size * sizeof(float));
+  for (int i = 0; i < spec.seq_length * spec.batch_size; i++) {
+    float val;
+    if (i < spec.batch_size * 2) {
+      val = 0.0f;
+    } else if (i >= spec.batch_size * 2 && i < spec.batch_size * 3) {
+      val = 1.0f;
+    } else {
+      val = rand() % 2 == 0 ? 1.0f : 0.0f;
+    }
+    h_transitions[i] = val;
+  }
+  cudaMemcpy(ts.transitions, h_transitions, spec.seq_length * spec.batch_size * sizeof(float), cudaMemcpyHostToDevice);
+  free(h_transitions);
 #if DEBUG
   print_device_matrix(ts.transitions, 1, spec.batch_size * spec.seq_length);
 #endif
 
   auto time_elapsed = chrono::microseconds::zero();
-  int n_batches = 1; // 50
+  int n_batches = 50;
   for (int t = 0; t < n_batches; t++) {
     auto start = chrono::high_resolution_clock::now();
     ts.forward();

### 2016-03-18 metal_stack/util.cc
Use a nicer random initialization to prevent exploding feedforwards

@@ -67,6 +67,14 @@ void print_device_matrix(const float *m, int M, int N) {
 }
 
 
+float *make_rand_matrix(int M, int N) {
+  float *m;
+  cudaMalloc(&m, M * N * sizeof(float));
+  fill_rand_matrix(m, M, N);
+  return m;
+}
+
+
 void fill_rand_matrix(float *m, int M, int N) {
   static curandGenerator_t prng;
   if (!prng) {
@@ -74,5 +82,5 @@ void fill_rand_matrix(float *m, int M, int N) {
     curandSetPseudoRandomGeneratorSeed(prng, (unsigned long long) clock());
   }
 
-  curandGenerateUniform(prng, m, M * N);
+  curandGenerateNormal(prng, m, M * N, 0.0f, 0.5f);
 }

### 2016-03-18 metal_stack/util.h

@@ -30,6 +30,7 @@ float *load_weights_cuda(string filename, int N, float *target=NULL);
 cublasHandle_t getCublasHandle();
 
 void print_device_matrix(const float *m, int M, int N);
+float *make_rand_matrix(int M, int N);
 void fill_rand_matrix(float *m, int M, int N);
 
 #endif

### 2016-03-18 metal_stack/bin/rnntest.cc
Integrate RNN / sequence model code

@@ -1,0 +1,59 @@
+#include <chrono>
+#include <iostream>
+
+#include <cuda_runtime.h>
+#include "cublas_v2.h"
+
+#include "rnn.h"
+#include "util.h"
+
+
+RNNParameters load_params(ModelSpec spec) {
+  float *U = make_rand_matrix(spec.model_dim, spec.model_dim);
+  float *W = make_rand_matrix(spec.model_dim, spec.word_embedding_dim);
+  float *b = make_rand_matrix(spec.model_dim, 1);
+
+  RNNParameters ret = {U, W, b};
+  return ret;
+}
+
+void destroy_params(RNNParameters params) {
+  cudaFree(params.U);
+  cudaFree(params.W);
+  cudaFree(params.b);
+}
+
+int main() {
+  ModelSpec spec = {10, 10, 1, 10, 30, 10};//{300, 300, (size_t) atoi(getenv("BATCH_SIZE")), 10, 30, 300};
+  RNNParameters params = load_params(spec);
+
+  cublasHandle_t handle;
+  cublasStatus_t stat = cublasCreate(&handle);
+  if (stat != CUBLAS_STATUS_SUCCESS) {
+    cout << "CUBLAS initialization failed" << endl;
+    return 1;
+  }
+
+  RNN rnn(spec, params, handle);
+
+  // Set model inputs.
+  cout << "X:" << endl;
+  fill_rand_matrix(rnn.X, spec.model_dim, spec.batch_size * spec.seq_length);
+
+  auto time_elapsed = chrono::microseconds::zero();
+  int n_batches = 50;
+  for (int t = 0; t < n_batches; t++) {
+    auto start = chrono::high_resolution_clock::now();
+    rnn.forward();
+    auto end = chrono::high_resolution_clock::now();
+    time_elapsed += chrono::duration_cast<chrono::microseconds>(end - start);
+  }
+
+  // Print the final representation.
+  cout << "Output:" << endl;
+  print_device_matrix(rnn.output, spec.model_dim, spec.batch_size);
+
+  cout << "Total time elapsed: " << time_elapsed.count() << endl;
+
+  destroy_params(params);
+}

### 2016-03-18 metal_stack/rnn.cc

@@ -1,0 +1,105 @@
+#include "rnn.h"
+using namespace std;
+namespace k = kernels;
+
+
+RNN::RNN(ModelSpec spec, RNNParameters params, cublasHandle_t handle)
+  : SequenceModel(spec), params(params), handle(handle) {
+
+  // Pre-allocate inputs.
+  cudaMalloc(&X_indices, spec.batch_size * spec.seq_length * sizeof(float));
+  cudaMalloc(&X, spec.batch_size * spec.seq_length * spec.model_dim * sizeof(float));
+
+  // Pre-allocate temporary containers.
+  cudaMalloc(&odd_output, spec.batch_size * spec.model_dim * sizeof(float));
+  cudaMalloc(&even_output, spec.batch_size * spec.model_dim * sizeof(float));
+
+  output = spec.seq_length % 2 == 0 ? even_output : odd_output;
+
+}
+
+
+RNN::~RNN() {
+
+  cout << "!!!!!!!!!!" << endl;
+  cout << "RNN dying!" << endl;
+  cout << "!!!!!!!!!!" << endl;
+
+  cudaFree(X_indices);
+  cudaFree(X);
+
+}
+
+
+void RNN::forward() {
+
+  // First timestep will read from odd output slot. Make sure it sees an empty
+  // (zero) state.
+  cudaMemset(odd_output, 0, spec.model_dim * spec.batch_size * sizeof(float));
+
+  for (int t = 0; t < spec.seq_length; t++) {
+    step(t);
+#if DEBUG
+    cout << endl << "======================" << endl << endl;
+#endif
+  }
+
+  // TODO: Don't need to sync here. Could have the client establish a lock on
+  // results and simultaneously begin the next batch + copy out results
+  cudaDeviceSynchronize();
+
+}
+
+
+void RNN::step(int t) {
+
+  const float *X_t = &X[t * spec.word_embedding_dim * spec.batch_size];
+#if DEBUG
+  cout << "X_t " << t << endl;
+  print_device_matrix(X_t, spec.model_dim, spec.batch_size);
+#endif
+
+  const float *state;
+  float *output;
+  if (t % 2 == 0) {
+    // t is even -- read from output at previous odd timestep, and write into
+    // even slot
+    state = odd_output;
+    output = even_output;
+  } else {
+    // t is odd -- read from output at previous even timestep, and write into
+    // odd slot
[…]

### 2016-03-18 metal_stack/rnn.h

@@ -1,0 +1,44 @@
+#ifndef _thin_stack_
+#define _thin_stack_
+
+#include <cuda_runtime.h>
+#include "cublas_v2.h"
+
+#include "sequence-model.h"
+#include "util.h"
+
+#include "kernels.cuh"
+
+
+typedef struct RNNParameters {
+  float *U; // hidden-to-hidden
+  float *W; // input-to-hidden
+  float *b;
+} RNNParameters;
+
+class RNN : public SequenceModel {
+  public:
+    RNN(ModelSpec spec, RNNParameters params, cublasHandle_t handle);
+    ~RNN();
+
+    RNNParameters params;
+    cublasHandle_t handle;
+
+    void forward();
+
+    float *output;
+
+  private:
+
+    void step(int t);
+
+    void recurrence(const float *state, const float *input, float *output);
+
+    // Containers for temporary (per-step) data
+    // RNN feedforward need only maintain two state cells. Just read from one
+    // and write to the other!
+    float *odd_output, *even_output;
+
+};
+
+#endif

### 2016-03-18 metal_stack/sequence-model.h

@@ -1,0 +1,36 @@
+#ifndef _sequence_model_
+#define _sequence_model_
+
+#include <cuda_runtime.h>
+
+#include "util.h"
+
+#include "kernels.cuh"
+
+class SequenceModel {
+
+  public:
+
+    ModelSpec spec;
+    SequenceModel(ModelSpec spec) : spec(spec) {};
+
+    // Embedding index inputs, of dimension `batch_size * seq_length` -- i.e.,
+    // we have `seq_length`-many concatenated vectors of embedding integers
+    float *X_indices;
+    // Embedding inputs, of dimension `model_dim * (batch_size * seq_length)` --
+    // i.e., along 2nd axis we have `seq_length`-many `model_dim * batch_size`
+    // matrices.
+    float *X;
+
+    void lookup_embeddings(float *embedding_source) {
+      kernels::subtensor1(X, embedding_source, X_indices, spec.vocab_size,
+          spec.seq_length * spec.batch_size, spec.model_dim, 0.0f, 1.0f, 0.0f,
+          NULL);
+    }
+
+    virtual void forward() = 0;
+
+};
+
+
+#endif

### 2016-03-18 metal_stack/bin/rnntest.cc
Use full model size in rnntest

@@ -24,7 +24,7 @@ void destroy_params(RNNParameters params) {
 }
 
 int main() {
-  ModelSpec spec = {10, 10, 1, 10, 30, 10};//{300, 300, (size_t) atoi(getenv("BATCH_SIZE")), 10, 30, 300};
+  ModelSpec spec = {300, 300, (size_t) atoi(getenv("BATCH_SIZE")), 10, 30, 300};
   RNNParameters params = load_params(spec);
 
   cublasHandle_t handle;

### 2016-03-18 writing/hard_stack_paper/paper.tex
Add RNN speed results (costs us five lines..)

@@
              xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
+  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
+  \addlegendentry{Thin-stack GPU}
   \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
   \addlegendentry{CPU \citep{irsoy2014deep}}
-  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
-  \addlegendentry{Thin-stack GPU}
+  \addplot table [x={Batch size}, y=RNN, col sep=tab] {runtime.csv};
+  \addlegendentry{RNN}
 \end{axis}
@@
 
-In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion.
+In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and a standard recurrent neural network sequence model.
 While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
 To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}
+We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} We also include speed results from a simple C++/CUDA implementation of a recurrent neural network.
 
@@
 We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled.
-Our GPU model is implemented in C++ and CUDA.
-We test it on a NVIDIA Titan X GPU.
+We test our thin-stack implementation and the RNN model on an NVIDIA Titan X GPU.
 
-Figure~\ref{fig:speed} compares the sentence encoding speed of the two models using data from the Stanford Sentiment Treebank \citep{socher2013recursive}.
-Figure~\ref{fig:speed} shows a substantial difference in runtime that increases with batch size, as is typical of GPU computation.
+Figure~\ref{fig:speed} compares the sentence encoding speed of the three models on random input data.
+Figure~\ref{fig:speed} shows a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size.
 At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--25x speedup relative to the standard CPU implementation.
+The thin-stack implementation is consistently about 4x slower than the recurrent neural network we tested here.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
CP's edits.

@@
 Our model supports batched computation for a speedup of up to 25x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with little loss of accuracy.
-We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
+We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence-encoding models.
 \end{abstract}
@@
 
-Of these, the TreeRNN appears to be the principled choice, since meaning in natural language sentences is known to be constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}.
+Of these, the TreeRNN appears to be the principled choice, since meaning in natural language sentences is known to be constructed recursively according to a tree structure \citep[][i.a.]{Dowty07DC}.
 TreeRNNs have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, but have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
@@
 
-We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of up to 25x over a standard TreeRNN implementation.
+We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence-encoding-based models, and that it yields speed increases of up to 25x over a standard TreeRNN implementation.
 
@@
 
-There is a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a variant \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
+There is a fairly long history of work on building neural network-based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a variant \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
 In addition, there has been recent work proposing models designed primarily for generative language modeling tasks that use these structures as well \citep{zhang2016top,dyer2016rnn}.
@@
 It next proceeds through the transition sequence, where each transition $t_t$ selects one of the two following operations.
-Below the $\mid$ symbol denotes the \textit{cons} (concatenation) operator.
+Below, the $\mid$ symbol denotes the \textit{cons} (concatenation) operator.
 We arbitrarily choose to always \textit{cons} on the left in the notation below.
@@
 It modifies the shift-reduce formalism by using fixed length vectors to represent each entry in the stack and the buffer.
-Correspondingly, its \reduce~operation combines two vector representations from the stack into another vector using a neural network function.
+Correspondingly, its \reduce\ operation combines two vector representations from the stack into another vector using a neural network function.
 
 \paragraph{The composition function}
-When a \reduce~operation is performed, the vector representations of two tree nodes are popped off of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
+When a \reduce\ operation is performed, the vector representations of two tree nodes are popped off of the stack and fed into a {\it composition function}, which is a neural network function that produces a representation for a new tree node that is the parent of the two popped nodes.
 This new node is pushed on to the stack.
@@
 
-The tracking LSTM's inputs (yellow in Figure~\ref{fig:m1-views}) are the top element of the buffer $\vec{h}_b^1$ (which would be moved in a \shift~operation) and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ (which would be composed in a \reduce~operation).
+The tracking LSTM's inputs (yellow in Figure~\ref{fig:m1-views}) are the top element of the buffer $\vec{h}_b^1$ (which would be moved in a \shift\ operation) and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ (which would be composed in a \reduce\ operation).
 
@@
 Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
-Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
+Simpler sequence-based sentence-encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
 In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word.
@@
 \end{equation}
-At test time, the model uses whichever transition (i.e., \shift~or \reduce) is assigned a higher probability.
+At test time, the model uses whichever transition (i.e., \shift\ or \reduce) is assigned a higher probability.
 The prediction function is trained to mimic the decisions of an external parser, and the decisions of this external parser are used as the inputs to the model during training.
-For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus for this purpose.
+For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus.
 We did not find the technique of scheduled sampling \citep{bengio2015scheduled}---allowing the model to use its own transition decisions in some instances at training time---to be helpful.
@@
 This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$.
-Such a naive implementation would also require copying a largely unchanged stack at each timestep, since each \shift~or \reduce~operation writes only one new representation to the stack.
+Such a naive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the stack.
 
@@
 Each row $S_t$ represents the top of the actual stack at timestep $t$.
-We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time.
+We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce\ operation at any given time.
 Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation.
@@
 The token sequence is simply the words in the sentence in order.
-$\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:
+$\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift\ transition and each `)' as a \reduce\ transition, as here:
 
@@
 \end{tabular}
-\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce~sequence on the input sentence $(a, b, c)$.
+\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce\ sequence on the input sentence $(a, b, c)$.
 $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$.
-The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce~operation at the next timestep.}
+The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next timestep.}
 \label{tbl:thin-stack}
@@
 
-We evaluate SPINN on the task of natural language inference \citep[NLI, a.k.a.
-recognizing textual entailment, or RTE;][]{dagan2006pascal}.
+We evaluate SPINN on the task of natural language inference \citep[NLI, a.k.a.\ recognizing textual entailment, or RTE;][]{dagan2006pascal}.
 NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use:
@@
 
-\paragraph{Creating a sentence pair classifier} \label{sec:classifier}
+\paragraph{Creating a sentence-pair classifier} \label{sec:classifier}
 
@@
 We evaluate four models.
-The four all use the sentence pair classifier architecture described in Section~\ref{sec:classifier}, and differ only in the function computing the sentence encodings.
+The four all use the sentence-pair classifier architecture described in Section~\ref{sec:classifier}, and differ only in the function computing the sentence encodings.
 First, a single-layer LSTM RNN \citep[similar to that of][]{snli:emnlp2015} serves as a baseline encoder.
@@
 
-We find that the bare SPINN-PI-NT model performs little better than the RNN baseline, but that SPINN-PI with the added tracking LSTM reaches state-of-the-art accuracy among sentence encoding based models.
+We find that the bare SPINN-PI-NT model performs little better than the RNN baseline, but that SPINN-PI with the added tracking LSTM reaches state-of-the-art accuracy among sentence-encoding-based models.
 The success of SPINN-PI, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
@@
 
-Both SPINN-PI and the full SPINN outperform all previous sentence encoding models by a significant margin.
+Both SPINN-PI and the full SPINN outperform all previous sentence-encoding models by a significant margin.
 Most notably our models outperform the Tree-Based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
@@
 
-The use of tree structure improves the performance of sentence encoding models for SNLI.
+The use of tree structure improves the performance of sentence-encoding models for SNLI.
 We suspect that this improvement is largely attributable to the more efficient learning of accurate generalizations overall, and not to any particular few phenomena.
@@
 In particular, it seems that tree-structured models are better able to focus on the key actors and events named in a sentence (possibly by learning an approximation of the linguistic notion of headedness), and that this allows them to better handle inferences based on these key actors.
-For example, this pair was classified successfully by all three SPINN variants, but not by the RNN baseline (emphasis added):
+For example, this pair was classified successfully by all three SPINN variants, but not by the RNN baseline (key words emphasized):
 
@@
 
-In this paper, we first introduce a model architecture (SPINN-PI-NT) that is equivalent to a TreeLSTM, but an order of magnitude faster at test time.
-Next, we expand that architecture into a tree-sequence hybrid model (SPINN-PI), and show that this yields significant gains on the SNLI entailment task.
+We introduce a model architecture (SPINN-PI-NT) that is equivalent to a TreeLSTM, but an order of magnitude faster at test time.
+W expand that architecture into a tree-sequence hybrid model (SPINN-PI), and show that this yields significant gains on the SNLI entailment task.
 Finally, we show that it is possible exploit the strengths of this model without the need for an external parser by integrating a fast parser into the model (as in the full SPINN), and that the lack of external parse information yields little loss in accuracy.
 
-\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0  \% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
+Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 
@@
 For a more ambitious goal, we expect that it should be possible to implement a variant of SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations \citep[as in][]{grefenstette2015learning,joulin2015inferring}.
-This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing it to learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
JG meta-edits, plus figure placement.

@@
 
+\begin{table}[t]
+\centering
+\begin{tabular}{c|cl}
+  \toprule
+  $t$ & $S$[$t$] & $Q_t$ \\
+  \midrule
+  0 & $a$ & \underline{\hphantom{0} 0} \\
+  1 & $b$ & \hphantom{0} \underline{0 1} \\
+  2 & $c$ & \hphantom{0} 0 \underline{1 2} \\
+  3 & $(c~b)$ & \hphantom{0} \underline{0 3} \\
+  4 & $((c~b)~a)$ & \underline{\hphantom{0} 4} \\
+  \bottomrule
+\end{tabular}
+\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce\ sequence on the input sentence $(a, b, c)$.
+$S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$.
+The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next timestep.}
+\label{tbl:thin-stack}
+\end{table}
+
 \paragraph{Representing the stack efficiently}
@@
 We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce\ operation at any given time.
-Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation.
+The thin-stack algorithm (Algorithm~\ref{alg:thin-stack}) describes the full mechanics of a stack feedforward in this compressed representation.
 It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$.
-Table~\ref{tbl:thin-stack} shows an example run of this algorithm.
+Table~\ref{tbl:thin-stack} shows an example run.
+
+This stack representation requires substantially less space.
+It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation.
+Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains.
+We describe speed results in Section~\ref{sec:speed}.
+
+A simpler variant of this technique can be used to represent the buffer, since information is not written to the buffer during model operation.
+It can be stored as a single fixed matrix with a scalar pointer variable indicating which element is the head at each step.
 
@@
 \end{algorithm}
-
-This stack representation requires substantially less space.
-It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation.
-Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains.
-We describe speed results in Section~\ref{sec:speed}.
-
-A simpler variant of this technique can be used to represent the buffer, since information is not written to the buffer during model operation.
-It can be stored as a single fixed matrix with a scalar pointer variable indicating which element is the head at each step.
 
@@
 
-\begin{table}[t]
-\centering
-\begin{tabular}{c|cl}
-  \toprule
-  $t$ & $S$[$t$] & $Q_t$ \\
-  \midrule
-  0 & $a$ & \underline{\hphantom{0} 0} \\
-  1 & $b$ & \hphantom{0} \underline{0 1} \\
-  2 & $c$ & \hphantom{0} 0 \underline{1 2} \\
-  3 & $(c~b)$ & \hphantom{0} \underline{0 3} \\
-  4 & $((c~b)~a)$ & \underline{\hphantom{0} 4} \\
-  \bottomrule
-\end{tabular}
-\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce\ sequence on the input sentence $(a, b, c)$.
-$S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$.
-The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next timestep.}
-\label{tbl:thin-stack}
-\end{table}
-
 \subsection{Inference speed}
@@
 
-In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and a standard recurrent neural network sequence model.
+In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and a standard RNN sequence model.
 While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
-To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} We also include speed results from a simple C++/CUDA implementation of a recurrent neural network.
+To do this, we reimplement the feedforward of SPINN-PI-NT and our RNN baseline in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
+We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}
 
@@
 Figure~\ref{fig:speed} shows a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size.
-At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--25x speedup relative to the standard CPU implementation.
-The thin-stack implementation is consistently about 4x slower than the recurrent neural network we tested here.
+At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), our model 10--25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Add a true t=0 step to the thin_stack fig.

@@
   \toprule
-  $t$ & $S$[$t$] & $Q_t$ \\
+  $t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} \\
   \midrule
-  0 & $a$ & \underline{\hphantom{0} 0} \\
-  1 & $b$ & \hphantom{0} \underline{0 1} \\
-  2 & $c$ & \hphantom{0} 0 \underline{1 2} \\
-  3 & $(c~b)$ & \hphantom{0} \underline{0 3} \\
-  4 & $((c~b)~a)$ & \underline{\hphantom{0} 4} \\
+  0 & & \underline{\hphantom{0} \hphantom{0}} \\
+  1 & $a$ & \hphantom{0} \underline{\hphantom{0} 1} \\
+  2 & $b$ & \hphantom{0} \hphantom{0} \underline{1 2} \\
+  3 & $c$ & \hphantom{0} \hphantom{0} 1 \underline{2 3} \\
+  4 & $(c~b)$ & \hphantom{0} \hphantom{0} \underline{1 4} \\
+  5 & $((c~b)~a)$ & \hphantom{0} \underline{\hphantom{0} 5} \\
   \bottomrule
 \end{tabular}
-\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce\ sequence on the input sentence $(a, b, c)$.
-$S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$.
-The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next timestep.}
+\caption{The thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce\ sequence on the input sentence $(a, b, c)$.
+$S$ is shown in the second column and represents the top of the stack at each step $t$.
+The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next step.}
 \label{tbl:thin-stack}

### 2016-03-18 writing/hard_stack_paper/paper.tex
Last minute edits.

@@
 This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$.
-Such a naive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the stack.
+Such a na\"ive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the stack.
 
@@
 
-Without the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics.
-In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
+Without the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and therefore also has the same learning dynamics.
+In both, the representation of each sentence consists of the representations of the words combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
 The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context through the tracking LSTM.
@@
 \label{sec:speed}
+
+In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and with a standard RNN sequence model.
+While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of all three models.
+To do this, we reimplement the feedforward of SPINN-PI-NT and our RNN baseline in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
+We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}
 
@@
 
-In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and a standard RNN sequence model.
-While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here.
-To do this, we reimplement the feedforward of SPINN-PI-NT and our RNN baseline in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}
-
 Each model is restricted to run on sentences of 30 tokens or fewer.
 We fix the model dimension and the word embedding dimension at 300.
-We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled.
+We run the CPU performance test on a 2.20-GHz 16-core Intel Xeon E5-2660 processor with hyperthreading enabled.
 We test our thin-stack implementation and the RNN model on an NVIDIA Titan X GPU.
@@
 Figure~\ref{fig:speed} shows a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size.
-At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), our model 10--25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.
+At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), our model is 10--25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.
 
@@
 
-To classify an SNLI sentence pair, we run two copies of SPINN with shared parameters: one on the premise sentence and another the hypothesis sentence.
+To classify an SNLI sentence pair, we run two copies of SPINN with shared parameters: one on the premise sentence and another on the hypothesis sentence.
 We then use their outputs (the $\vec{h}$ states at the top of each stack at time $t=T$) to construct a feature vector $\vec{x}_{\text{classifier}}$ for the pair.
-This feature vector consists of the concatenation of these two sentence vectors, their difference, and their elementwise product:
+This feature vector consists of the concatenation of these two sentence vectors, their difference, and their elementwise product \citep[following][]{mou2015recognizing}:
 \begin{equation}
@@
 Params.
-is the approximate number of trained parameters (excluding word embeddings for models where they are trained).
+is the approximate number of trained parameters (excluding word embeddings for all models).
 Trans.
@@
 
-Appendix~\ref{sec:hyperparams} discusses hyperparameter tuning.
+A supplement discusses hyperparameter tuning.
+% TODO for arXiv: Make to appendix.
 
@@
 
-We find that the bare SPINN-PI-NT model performs little better than the RNN baseline, but that SPINN-PI with the added tracking LSTM reaches state-of-the-art accuracy among sentence-encoding-based models.
-The success of SPINN-PI, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are complementary.
+We find that the bare SPINN-PI-NT model performs little better than the RNN baseline, but that SPINN-PI with the added tracking LSTM performs well.
+The success of SPINN-PI, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are at least partially complementary.
 The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless robustly exceeds the performance of the LSTM baseline.
 
-Both SPINN-PI and the full SPINN outperform all previous sentence-encoding models by a significant margin.
-Most notably our models outperform the Tree-Based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
+Both SPINN-PI and the full SPINN significantly outperform all previous sentence-encoding models.
+Most notably, these models outperform the Tree-Based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
 Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-Based CNN), which in turn outperforms one which does not use it at all (LSTM RNN).
@@
 However, some patterns are identifiable in the results.
-In particular, it seems that tree-structured models are better able to focus on the key actors and events named in a sentence (possibly by learning an approximation of the linguistic notion of headedness), and that this allows them to better handle inferences based on these key actors.
+In particular, it seems that tree-structured models are better able to focus on the key actors and events named in a sentence (possibly by learning an approximation of the linguistic notion of headedness), and that this allows them to better focus on the central actors and events described in a sentence.
 For example, this pair was classified successfully by all three SPINN variants, but not by the RNN baseline (key words emphasized):
@@
 {A \textit{woman} in red blouse is \textit{standing} with small blond \textit{child} in front of a small folding chalkboard.}
-{a woman stands with her child}
+{a \textit{woman} \textit{stands} with her \textit{child}}
 {neutral}
 
-The tree-sequence hybrid models seem to be especially effective at comparing sentences with very different structures, where it is useful to use strict compositionality with the tree-structured component to get the meanings of the sentence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures.
+The tree-sequence hybrid models seem to be especially good at reasoning over pairs of sentences with very different structures, for which it is helpful to use strict compositionality with the tree-structured component to get the meanings of the sentence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures of each sentences.
 For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT:
@@
 We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
-We suspect that it is acting somewhat like the Tree-Based CNN, only with access to larger trees: Using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
+We suspect that it is acting somewhat like the Tree-Based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
 
@@
 We introduce a model architecture (SPINN-PI-NT) that is equivalent to a TreeLSTM, but an order of magnitude faster at test time.
-W expand that architecture into a tree-sequence hybrid model (SPINN-PI), and show that this yields significant gains on the SNLI entailment task.
-Finally, we show that it is possible exploit the strengths of this model without the need for an external parser by integrating a fast parser into the model (as in the full SPINN), and that the lack of external parse information yields little loss in accuracy.
+We expand that architecture into a tree-sequence hybrid model (SPINN-PI), and show that this yields significant gains on the SNLI entailment task.
+Finally, we show that it is possible to exploit the strengths of this model without the need for an external parser by integrating a fast parser into the model (as in the full SPINN), and that the lack of external parse information yields little loss in accuracy.
 
@@
 
-\begin{table*}[t]\small
-\begin{center}
-\begin{tabular}{lrlrrrr}
-\toprule
-Param.     & Range & Strategy        & RNN       & SPINN-PI-NT   & SPINN-PI  & SPINN \\
-\midrule 
-Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & 2e-3\\
-L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & 3e-5\\
-Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & --- & --- & ---  & 3.9    \\
-Embedding transformation dropout keep rate & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
-Classifier MLP dropout keep rate & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
-Tracking LSTM size $D_\text{tracking}$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
-Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & 1\\
-\bottomrule
-\end{tabular}
-\end{center}
-
-\caption{
-\label{tab:hyperparams}
-Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log--uniform.
-}
-\end{table*}
-
-\appendix
-\section{Hyperparameters}\label{sec:hyperparams}
-
-We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and then launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
-Table~\ref{tab:hyperparams} below shows the hyperparameters used in the best run of each model.
-
 \end{document}

### 2016-03-18 writing/hard_stack_paper/paper.tex
Fix list punctuation in introduction

@@
 This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vectors to a sentence vector, and this function can take a range of different forms.
-Common sentence encoders include sequence-based recurrent neural network models (RNNs, see Figure~\ref{fig:batching:good}) with Long Short-Term Memory \citep[LSTM,][]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words or characters, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011parsing}, which propagate information up a binary parse tree.
+Common sentence encoders include sequence-based recurrent neural network models (RNNs, see Figure~\ref{fig:batching:good}) with Long Short-Term Memory \citep[LSTM,][]{hochreiter1997long}, which accumulate information over the sentence sequentially; convolutional neural networks \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words or characters; and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011parsing}, which propagate information up a binary parse tree.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Mention minibatch SGD

@@
 
-We use the RMSProp optimizer \citep{tieleman2012lecture} with a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
+We use minibatch SGD with the RMSProp optimizer \citep{tieleman2012lecture} and a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
 We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the output of the word embedding projection layer and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.

### 2016-03-18 writing/hard_stack_paper/paper.tex
Incremental updates.

@@
 This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser.
-In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
+In addition, it supports batched computation for both parsed and unparsed sentences, which yields dramatic speedups over standard TreeRNNs.
 Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over pure sequence- or tree-based models.
@@
 There is a fairly long history of work on building neural network-based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a variant \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}.
-In addition, there has been recent work proposing models designed primarily for generative language modeling tasks that use these structures as well \citep{zhang2016top,dyer2016rnn}.
-To our knowledge, SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.
+In addition, there has been recent work proposing models designed primarily for generative language modeling tasks that use this architecture as well \citep{zhang2016top,dyer2016rnn}.
+To our knowledge, SPINN is the first model to use this architecture for the purpose of sentence interpretation, rather than parsing or generation.
 %TRIMMABLE
@@
 \end{gather}
-where $\sigma$ is the sigmoid activation function, $\odot$ is the elementwise product, the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack, and $\vec{e}$ is an optional vector valued input argument which is either empty or comes from from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
+where $\sigma$ is the sigmoid activation function, $\odot$ is the elementwise product, the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack, and $\vec{e}$ is an optional vector-valued input argument which is either empty or comes from from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
 The result of this function, the pair $\langle\vec{h}, \vec{c}\rangle$, is placed back on the stack.
@@
 Simpler sequence-based sentence-encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
-In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word.
+In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word, and don't get additional context information until much later.
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
@@
 For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus.
-We did not find the technique of scheduled sampling \citep{bengio2015scheduled}---allowing the model to use its own transition decisions in some instances at training time---to be helpful.
+We did not find scheduled sampling \citep{bengio2015scheduled}---allowing the model to use its own transition decisions in some instances at training time---to help.
 
@@
 While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of all three models.
-To do this, we reimplement the feedforward of SPINN-PI-NT and our RNN baseline in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}
+To do this, we reimplement the feedforward\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} of SPINN-PI-NT and our RNN baseline in C++/CUDA, and compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
+We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.
 
@@
 Each model is restricted to run on sentences of 30 tokens or fewer.
-We fix the model dimension and the word embedding dimension at 300.
+We fix the model dimension $D$ and the word embedding dimension at 300.
 We run the CPU performance test on a 2.20-GHz 16-core Intel Xeon E5-2660 processor with hyperthreading enabled.

### 2016-03-18 writing/hard_stack_paper/paper.tex
Remove redundant sentence start
Name thin stack.

@@
 Simpler sequence-based sentence-encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
-In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word, and don't get additional context information until much later.
+In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with.
+This is often just a single additional word, and the model may not get much additional context information until much later.
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
@@
 
-We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}.
-For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$.
+We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}.
+For each input sentence, we represent the stack with a single $T \times D$ matrix $S$.
 Each row $S_t$ represents the top of the actual stack at timestep $t$.
-We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce\ operation at any given time.
-The thin-stack algorithm (Algorithm~\ref{alg:thin-stack}) describes the full mechanics of a stack feedforward in this compressed representation.
+We maintain a queue of backpointers onto $S$ that indicates which elements would be involved in a \reduce\ operation at any given step.
+The thin stack algorithm (Algorithm~\ref{alg:thin-stack}) describes the full mechanics of a stack feedforward in this compressed representation.
 It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$.
-Table~\ref{tbl:thin-stack} shows an example run.
+Table~\ref{tbl:thin-stack} shows a sample run.
 
@@
 \begin{algorithm}[t]
-\caption{The thin-stack algorithm}
+\caption{The thin stack algorithm}
 \label{alg:thin-stack}
@@
 Figure~\ref{fig:speed} compares the sentence encoding speed of the three models on random input data.
-Figure~\ref{fig:speed} shows a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size.
+We observe a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size.
 At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), our model is 10--25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.

### 2016-03-18 writing/hard_stack_paper/paper.tex
Remove induced redundancy.

@@
 We maintain a queue of backpointers onto $S$ that indicates which elements would be involved in a \reduce\ operation at any given step.
-The thin stack algorithm (Algorithm~\ref{alg:thin-stack}) describes the full mechanics of a stack feedforward in this compressed representation.
+Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation.
 It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$.
-Table~\ref{tbl:thin-stack} shows a sample run.
+Table~\ref{tbl:thin-stack} shows an example run.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Unvague the speed claim.

@@
 We observe a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size.
-At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), our model is 10--25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.
+With a large but practical batch size of 512, the largest we on which we tested the TreeRNN, our model is about 25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Rewrite the loss.

@@
 
-\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters.
+\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_t^p, \mathcal{L}_t^h\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters.
 The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
@@
 \begin{split}
-\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
+\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{t}}^{p} + \mathcal{L}_{\text{t}}^{h}) + \lambda \|\theta\|^2_2
 \end{split}

### 2016-03-18 writing/hard_stack_paper/paper.tex
Fix premature reference to baseline.

@@
 While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of all three models.
-To do this, we reimplement the feedforward\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} of SPINN-PI-NT and our RNN baseline in C++/CUDA, and compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
+To do this, we reimplement the feedforward\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} of SPINN-PI-NT and an LSTM RNN baseline in C++/CUDA, and compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as SPINN-PI-NT.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
 We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.
@@
 The success of SPINN-PI, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are at least partially complementary.
-The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless robustly exceeds the performance of the LSTM baseline.
+The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless robustly exceeds the performance of the RNN baseline.
 
@@
 Most notably, these models outperform the Tree-Based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
-Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-Based CNN), which in turn outperforms one which does not use it at all (LSTM RNN).
+Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-Based CNN), which in turn outperforms one which does not use it at all (RNN).
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Extraneous capitals.

@@
 Both SPINN-PI and the full SPINN significantly outperform all previous sentence-encoding models.
-Most notably, these models outperform the Tree-Based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
-Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (Tree-Based CNN), which in turn outperforms one which does not use it at all (RNN).
+Most notably, these models outperform the tree-based CNN of \citet{mou2015recognizing}, which also uses tree-structured composition for local feature extraction, but uses simpler pooling techniques to build sentence features in the interest of efficiency.
+Our results show that a model that uses tree-structured composition fully (SPINN) outperforms one which uses it only partially (tree-based CNN), which in turn outperforms one which does not use it at all (RNN).
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Two small edits.

@@
 
-The full SPINN performed moderately well at reproducing the Stanford Parser's parses of the SNLI data at a transition-by-transition level with 92.4\% accuracy at test time.
+The full SPINN performed moderately well at reproducing the Stanford Parser's parses of the SNLI data at a transition-by-transition level, with 92.4\% accuracy at test time.
 However, its transition prediction errors were fairly evenly distributed across sentences, and most sentences were assigned partially invalid transition sequences that either left a few words out of the final representation or incorporated a few padding tokens into the final representation.
@@
 
-Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
+Because this paper aims to introduce a general purpose model for sentence encoding, we do not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 

### 2016-03-18 writing/hard_stack_paper/paper.tex
Bib cleanup, one last detail.

@@
 \end{equation}
-Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution over the three labels.
+Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of 1024D ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution over the three labels.
 

### 2016-03-19 writing/hard_stack_paper/paper.tex
Edits from AR and RG.

@@
 TreeRNNs have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, but have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.
-Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets.
+Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is crucial in enabling neural networks to be trained efficiently on large datasets.
 Because TreeRNNs use a different model structure for each sentence, as in Figure~\ref{fig:batching}, batching is impossible in standard implementations.
@@
 SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over its tokens.
-Shift-reduce parsing is widely used in natural language parsing \citep[e.g.,][]{shieber:1983:ACL,nivre2003efficient}.
+The formalism is widely used in natural language parsing \citep[e.g.,][]{shieber:1983:ACL,nivre2003efficient}.
 
-A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and consumes transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies one step of the parsing process.
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_0, \dots, x_{N-1})$ and consumes transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies one step of the parsing process.
 In general a parser may also generate these transitions on the fly as it reads the tokens.
@@
 Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
-Simpler sequence-based sentence-encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
-In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with.
-This is often just a single additional word, and the model may not get much additional context information until much later.
+Even though TreeRNNs are much more effective at composing meanings in principle, this ambiguity can give simpler sequence-based sentence-encoding models an advantage: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
+In contrast, when a standard tree-structured model first processes a word, it only has access to the constituent that the word is merging with, which is often just a single additional word.
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
 
-It would be straightforward to augment SPINN to support the use of some amount of right context as well, but this would add complexity to the model that we think is largely unnecessary: humans are very effective at understanding the beginnings of sentences before having seen or heard the ends, suggesting that it is possible to get by without the unavailable right context.
+It would be straightforward to augment SPINN to support the use of some amount of right-side context as well, but this would add complexity to the model that we think is largely unnecessary: humans are very effective at understanding the beginnings of sentences before having seen or heard the ends, suggesting that it is possible to get by without the unavailable right-side context.
 
@@
 At test time, the model uses whichever transition (i.e., \shift\ or \reduce) is assigned a higher probability.
-The prediction function is trained to mimic the decisions of an external parser, and the decisions of this external parser are used as the inputs to the model during training.
+The prediction function is trained to mimic the decisions of an external parser, and these decisions are used as the inputs to the model during training.
 For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus.
@@
 We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
-We suspect that it is acting somewhat like the Tree-Based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
+We suspect that it is acting somewhat like the tree-based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
 

### 2016-03-29 writing/hard_stack_paper/paper.tex
Changes to sections 0, 1, 2

@@
 
+\usepackage[leqno, fleqn]{amsmath}
 \usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black, draft]{hyperref}
 \usepackage{natbib}
-\usepackage{times}
+% \usepackage{times}
+\usepackage{newtxtext,newtxmath}
 \usepackage{latexsym}
@@
 \usepackage{url}
-\usepackage[leqno, fleqn]{amsmath}
-\usepackage{amssymb}
+% \usepackage{amssymb}
 \usepackage{qtree}
@@
 
+\hyphenation{Image-Net}
+
+\pgfplotsset{compat=1.12}
+
 \noautomath
@@
 Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences.
-However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they can only operate on parsed sentences and they do not directly support batched computation.
+However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they usually operate on parsed sentences and they do not directly support batched computation.
 We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single tree-sequence hybrid model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser.
-Our model supports batched computation for a speedup of up to 25x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with little loss of accuracy.
+Our model supports batched computation for a speedup of up to 25$\times$ over other tree-structured models, and its integrated parser can operate on unparsed data with little loss in accuracy.
 We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence-encoding models.
@@
 Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is crucial in enabling neural networks to be trained efficiently on large datasets.
-Because TreeRNNs use a different model structure for each sentence, as in Figure~\ref{fig:batching}, batching is impossible in standard implementations.
-In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.
+Because TreeRNNs use a different model structure for each sentence, as in Figure~\ref{fig:batching}, efficient batching is impossible in standard implementations.
+Partly to address efficiency problems, standard TreeRNN models commonly only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.
 
@@
 SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces the required parse structure on the fly.
-This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser.
-In addition, it supports batched computation for both parsed and unparsed sentences, which yields dramatic speedups over standard TreeRNNs.
-Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over pure sequence- or tree-based models.
+This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser and saving its additional computational cost.
+Secondly, it supports batched computation for both parsed and unparsed sentences, yielding dramatic speedups over standard TreeRNNs.
+Finally, it supports a novel tree-sequence hybrid architecture for handling local linear context in sentence interpretation.
+This model is a plausible model of human sentence processing and yields substantial accuracy gains over pure sequence- or tree-based models.
 % TRIMMABLE
 
-We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence-encoding-based models, and that it yields speed increases of up to 25x over a standard TreeRNN implementation.
+We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence-encoding-based models, and that it yields speed increases of up to 25$\times$ over a standard TreeRNN implementation.
 
@@
 However, these methods require an expensive search process at test time.
-Our model presents a fast alternative approach.
+Our model presents a much faster alternative approach.
 

### 2016-03-29 writing/hard_stack_paper/paper.tex
Rest of the changes.

@@
 }}
+
+\providecommand{\norm}[1]{\lVert#1\rVert}
 
@@
 Finally, it supports a novel tree-sequence hybrid architecture for handling local linear context in sentence interpretation.
-This model is a plausible model of human sentence processing and yields substantial accuracy gains over pure sequence- or tree-based models.
+This model is a basically plausible model of human sentence processing and yields substantial accuracy gains over pure sequence- or tree-based models.
 % TRIMMABLE
@@
 
-SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over its tokens.
+SPINN is inspired by shift-reduce parsing \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over its tokens.
 The formalism is widely used in natural language parsing \citep[e.g.,][]{shieber:1983:ACL,nivre2003efficient}.
@@
 We arbitrarily choose to always \textit{cons} on the left in the notation below.
+\vspace*{-1ex} % dirty spacing trick that somehow seemed needed right now
 %TRIMMABLE
 \begin{description}
-  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
+  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it on to the top of the stack.
+\vspace*{-1ex}
   \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$.
-This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
+This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back on to the stack.
 \end{description}
@@
 = \colvec{5}
-    {\sigma}
-    {\sigma}
-    {\sigma}
-    {\sigma}
-    {\text{tanh}}
+     {\sigma\vphantom{\vec{i}}}
+     {\sigma\vphantom{\vec{f}_l}}
+     {\sigma\vphantom{\vec{f}_r}}
+     {\sigma\vphantom{\vec{o}}}
+    {\text{tanh}\vphantom{\vec{g}}}
 \left(
@@
 \end{gather}
-where $\sigma$ is the sigmoid activation function, $\odot$ is the elementwise product, the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack, and $\vec{e}$ is an optional vector-valued input argument which is either empty or comes from from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
+where $\sigma$ is the sigmoid activation function, $\odot$ is the elementwise product, the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack, and $\vec{e}$ is an optional vector-valued input argument which is either empty or comes from an external source like the tracking LSTM (see Section~\ref{sec:tracking}).
 The result of this function, the pair $\langle\vec{h}, \vec{c}\rangle$, is placed back on the stack.
@@
 
-We use word representations based on the standard 300D vector package provided with GloVe \citep{pennington2014glove}.
+We use word representations based on the standard 300D vectors provided with GloVe \citep{pennington2014glove}.
 We do not update these representations during training.
@@
 
-In addition to the stack, the buffer, and the composition function, our full model includes an additional component: the tracking LSTM.
-This is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
-It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, which is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.
+In addition to the stack, buffer, and composition function, our full model includes an additional component: the tracking LSTM.
+This is a simple sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, which is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$, see (\ref{eqn:lstm1}), to the composition function, allowing context information to enter the construction of sentence meaning,  forming what is effectively a tree-sequence hybrid model.
 
@@
 Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence.
-Even though TreeRNNs are much more effective at composing meanings in principle, this ambiguity can give simpler sequence-based sentence-encoding models an advantage: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
+Even though TreeRNNs are more effective at composing meanings in principle, this ambiguity can give simpler sequence-based sentence-encoding models an advantage: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation.
 In contrast, when a standard tree-structured model first processes a word, it only has access to the constituent that the word is merging with, which is often just a single additional word.
 Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
+Using left linear context to disambiguate is also a plausible model of human interpretation.
 
@@
 
-For SPINN to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply it as part of the input.
+For SPINN to operate on unparsed inputs, it needs to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply it as part of the input.
 To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier whose input is the state of the tracking LSTM:
@@
 At test time, the model uses whichever transition (i.e., \shift\ or \reduce) is assigned a higher probability.
-The prediction function is trained to mimic the decisions of an external parser, and these decisions are used as the inputs to the model during training.
+The prediction function is trained to mimic the decisions of an external parser.
+These decisions are used as the inputs to the model during training.
 For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus.
-We did not find scheduled sampling \citep{bengio2015scheduled}---allowing the model to use its own transition decisions in some instances at training time---to help.
+We did not find scheduled sampling \citep{bengio2015scheduled}---having the model use its own transition decisions sometimes at training time---to help.
 
@@
 \centering
-\begin{tabular}{c|cl}
+\begin{tabular}{c @{\hspace*{2em}}c @{\hspace*{2em}} l}
   \toprule
@@
 Each row $S_t$ represents the top of the actual stack at timestep $t$.
-We maintain a queue of backpointers onto $S$ that indicates which elements would be involved in a \reduce\ operation at any given step.
+We maintain a queue of backpointers into $S$ that indicates which elements would be involved in a \reduce\ operation at any given step.
 Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation.
-It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$.
+It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$.
 Table~\ref{tbl:thin-stack} shows an example run.
@@
 \paragraph{Handling variable sentence lengths} For any sentence model to be trained with batched computation, it is necessary to pad or crop sentences to a fixed length.
-We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI.
+We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI\@.
 Transition sequences $\mathbf t$ are cropped at the left or padded at the left with \shift s.
@@
 
-Without the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and therefore also has the same learning dynamics.
+Without the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and therefore it also has the same learning dynamics.
 In both, the representation of each sentence consists of the representations of the words combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function).
-The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context through the tracking LSTM.
+SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context through the tracking LSTM.
 
@@
 
-In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and with a standard RNN sequence model.
+In this section, we compare the test-time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and with a standard RNN sequence model.
 While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of all three models.
@@
 We observe a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size.
-With a large but practical batch size of 512, the largest we on which we tested the TreeRNN, our model is about 25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.
+With a large but practical batch size of 512, the largest on which we tested the TreeRNN, our model is about 25$\times$ faster than the standard CPU implementation, and about 4$\times$ slower than the RNN baseline.
 
@@
 \begin{split}
-\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{t}}^{p} + \mathcal{L}_{\text{t}}^{h}) + \lambda \|\theta\|^2_2
+\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{t}}^{p} + \mathcal{L}_{\text{t}}^{h}) + \lambda \norm{\theta}^2_2
 \end{split}
@@
 We find that the bare SPINN-PI-NT model performs little better than the RNN baseline, but that SPINN-PI with the added tracking LSTM performs well.
-The success of SPINN-PI, which is a hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are at least partially complementary.
+The success of SPINN-PI, which is the hybrid tree-sequence model, suggests that the tree- and sequence-based encoding methods are at least partially complementary, with the sequence model presumably providing useful local word disambiguation.
 The full SPINN model with its relatively weak internal parser performs slightly less well, but nonetheless robustly exceeds the performance of the RNN baseline.
@@
 
-The full SPINN performed moderately well at reproducing the Stanford Parser's parses of the SNLI data at a transition-by-transition level, with 92.4\% accuracy at test time.
-However, its transition prediction errors were fairly evenly distributed across sentences, and most sentences were assigned partially invalid transition sequences that either left a few words out of the final representation or incorporated a few padding tokens into the final representation.
+The full SPINN performed moderately well at reproducing the Stanford Parser's parses of the SNLI data at a transition-by-transition level, with 92.4\% accuracy at test time.\footnote
+  {Note that this is scoring the model against automatic parses, not a human-judged gold standard.}
+However, its transition prediction errors are fairly evenly distributed across sentences, and most sentences were assigned partially invalid transition sequences that either left a few words out of the final representation or incorporated a few padding tokens into the final representation.
 

### 2016-03-29 analyze_log.py
update script to plot from log

@@ -29,34 +29,46 @@ class EvalLine(object):
 
 class Log(object):
   def __init__(self, path):
+    self.evals = []
     with open(path) as f:
       lines = filter(lambda l : 'Step' in l, f.readlines())
       # count number of eval sets
       num_logs = 1
-      for i in xrange(1, len(lines)):
+      for i in xrange(len(lines)):
         if 'Acc' not in lines[i]:
           num_logs += 1
+          self.evals.append(lines[i].strip().split()[-1])
         else:
           break
       self.corpus = [[] for _ in xrange(num_logs)]
-      # read the costs and ccuracies
+      # read the costs and accuracies
       for i, line in enumerate(lines):
-        ind = i % num_logs
+        if 'Acc' in line:
+          ind = 0
+        else:
+          ind = self.evals.index(lines[i].strip().split()[-1]) + 1
+        # ind = i % num_logs
         if  ind == 0:
           self.corpus[0].append(TrainLine(line))
-        else:
+        elif 'Eval' in line:
           self.corpus[ind].append(EvalLine(line))
 
-def ShowPlots():
+def ShowPlots(subplot=False):
   for log_ind, path in enumerate(FLAGS.path.split(":")):
     log = Log(path)
+    if subplot:
+      plt.subplot(len(FLAGS.path.split(":")), 1, log_ind + 1)
     for index in FLAGS.index.split(","):
       index = int(index)
       for attr in ["pred_acc", "parse_acc", "total_cost", "xent_cost", "l2_cost", "action_cost"]:
         if getattr(FLAGS, attr):
           if "cost" in attr:
             assert index == 0, "costs only associated with training log"
-          steps, val = zip(*[(l.step, getattr(l, attr)) for l in log.corpus[index]])
+          steps, val = zip(*[(l.step, getattr(l, attr)) for l in log.corpus[index] if l.step < FLAGS.iters])
+          dct = {}
+          for k, v in zip(steps, val):
+            dct[k] = max(v, dct[k]) if k in dct else v
+          steps, val = zip(*sorted(dct.iteritems()))
           plt.plot(steps, val, label="Log%d:%s-%d" % (log_ind, attr, index))
     
   plt.xlabel("No. of training iteration")
@@ -77,9 +89,11 @@ if __name__ == '__main__':
   gflags.DEFINE_boolean("l2_cost", False, "L2 regularization cost, valid only if index == 0")
[…]

### 2016-04-01 writing/hard_stack_paper/paper.tex
Fig 3 updates. (Had to shift legend to a weird place to prevent too much figure-shrinking)

@@
 \begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
-             xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
+             xticklabels={0,32,64,128,256,512,1024,2048}, xmode=log, log basis x=2,
+             legend style={at={(0.55, 0.5)},anchor=north west}]
   \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};

### 2016-04-06 python/spinn/classifier.py
python -> python/spinn; rename modules; begin classifier file refactoring

@@ -1,23 +1,4 @@
-"""From the project root directory (containing data files), this can be run with:
-
-Boolean logic evaluation:
-python -m rembed.models.classifier --training_data_path bl-data/pbl_train.tsv \
-       --eval_data_path bl-data/pbl_dev.tsv
-
-SST sentiment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type sst --training_data_path sst-data/train.txt \
-       --eval_data_path sst-data/dev.txt --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
-       --model_dim 10 --word_embedding_dim 5
-
-SNLI entailment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
-       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
-       --model_dim 10 --word_embedding_dim 5
-
-Note: If you get an error starting with "TypeError: ('Wrong number of dimensions..." during development,
-    there may already be a saved checkpoint in ckpt_path that matches the name of the model you're developing.
-    Move or delete it as appropriate.
-"""
+"""Shared SPINN classifier code."""
 
 from functools import partial
 import os
@@ -25,31 +6,30 @@ import pprint
 import sys
 
 import gflags
-from theano import tensor as T
 import theano
-import numpy as np
+from theano import tensor as T
 
+import rembed
 from rembed import afs_safe_logger
-from rembed import recurrences, util
+from rembed import util
 from rembed.data.boolean import load_boolean_data
 from rembed.data.sst import load_sst_data
 from rembed.data.snli import load_snli_data
-from rembed.stack import ThinStack
-
-import rembed.plain_rnn
 
 
 FLAGS = gflags.FLAGS
 
 
-def build_sentence_model(cls, vocab_size, seq_length, tokens, transitions,
-                         num_classes, training_mode, ground_truth_transitions_visible, vs,
-                         initial_embeddings=None, project_embeddings=False, ss_mask_gen=None, ss_prob=0.0):
+def build_sentence_model(make_stack_fn, vocab_size, seq_length, tokens,
+                         transitions, num_classes, training_mode,
+                         ground_truth_transitions_visible, vs,
[…]

### 2016-04-06 python/spinn/models/classifier.py

@@ -1,0 +1,206 @@
+"""From the project root directory (containing data files), this can be run with:
+
+Boolean logic evaluation:
+python -m rembed.models.classifier --training_data_path bl-data/pbl_train.tsv \
+       --eval_data_path bl-data/pbl_dev.tsv
+
+SST sentiment (Demo only, model needs a full GloVe embeddings file to do well):
+python -m rembed.models.classifier --data_type sst --training_data_path sst-data/train.txt \
+       --eval_data_path sst-data/dev.txt --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+       --model_dim 10 --word_embedding_dim 5
+
+SNLI entailment (Demo only, model needs a full GloVe embeddings file to do well):
+python -m rembed.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
+       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+       --model_dim 10 --word_embedding_dim 5
+
+Note: If you get an error starting with "TypeError: ('Wrong number of dimensions..." during development,
+    there may already be a saved checkpoint in ckpt_path that matches the name of the model you're developing.
+    Move or delete it as appropriate.
+"""
+
+from copy import copy
+from functools import partial
+
+from rembed import recurrences, util
+from rembed.stack import ThinStack
+
+
+def make_sentence_stack(cls, embedding_projection_network, compose_network,
+                        vocab_size, seq_length, tokens, transitions,
+                        num_classes, training_mode,
+                        ground_truth_transitions_visible, vs, initial_embeddings,
+                        project_embeddings, ss_mask_gen, ss_prob):
+    model_visible_dim = FLAGS.model_dim / 2 if FLAGS.lstm_composition else FLAGS.model_dim
+    spec = util.ModelSpec(FLAGS.model_dim, FLAGS.word_embedding_dim,
+                          FLAGS.batch_size, vocab_size, seq_length,
+                          model_visible_dim=model_visible_dim)
+
+    # TODO: Check non-Model0 support.
+    recurrence = cls(spec, vs, compose_network,
+                     use_context_sensitive_shift=FLAGS.context_sensitive_shift,
+                     context_sensitive_use_relu=FLAGS.context_sensitive_use_relu,
+                     use_tracking_lstm=FLAGS.use_tracking_lstm,
[…]

### 2016-04-06 python/spinn/stack.py

@@ -473,7 +473,7 @@ class ThinStack(object):
         self.final_stack = self.scan_updates[self.stack]
         self.final_aux_stacks = [self.scan_updates[aux_stack]
                                  for aux_stack in self.aux_stacks]
-        self.sentence_embeddings = self.final_stack[-self.batch_size:]
+        self.final_representations = self.final_stack[-self.batch_size:]
 
         self.transitions_pred = None
         if self.recurrence.predicts_transitions:

### 2016-04-06 python/spinn/classifier.py
s/rembed/spinn/g

@@ -9,12 +9,12 @@ import gflags
 import theano
 from theano import tensor as T
 
-import rembed
-from rembed import afs_safe_logger
-from rembed import util
-from rembed.data.boolean import load_boolean_data
-from rembed.data.sst import load_sst_data
-from rembed.data.snli import load_snli_data
+import spinn
+from spinn import afs_safe_logger
+from spinn import util
+from spinn.data.boolean import load_boolean_data
+from spinn.data.sst import load_sst_data
+from spinn.data.snli import load_snli_data
 
 
 FLAGS = gflags.FLAGS
@@ -44,7 +44,7 @@ def build_sentence_model(make_stack_fn, vocab_size, seq_length, tokens,
     """
 
     # Prepare layer which performs stack element composition.
-    if cls is rembed.plain_rnn.RNN:
+    if cls is spinn.plain_rnn.RNN:
         compose_network = partial(util.LSTMLayer,
                                       initializer=util.HeKaimingInitializer())
         embedding_projection_network = None
@@ -100,7 +100,7 @@ def build_sentence_pair_model(make_stack_fn, vocab_size, seq_length, tokens, tra
     Construct a classifier which makes use of some hard-stack model.
 
     Args:
-      cls: Hard stack class to use (from e.g. `rembed.stack`)
+      cls: Hard stack class to use (from e.g. `spinn.stack`)
       vocab_size:
       seq_length: Length of each sequence provided to the stack model
       tokens: Theano batch (integer matrix), `batch_size * seq_length`
@@ -115,7 +115,7 @@ def build_sentence_pair_model(make_stack_fn, vocab_size, seq_length, tokens, tra
     """
 
     # Prepare layer which performs stack element composition.
-    if cls is rembed.plain_rnn.RNN:
+    if cls is spinn.plain_rnn.RNN:
         compose_network = partial(util.LSTMLayer,
                                       initializer=util.HeKaimingInitializer())
         embedding_projection_network = None
@@ -147,7 +147,7 @@ def build_sentence_pair_model(make_stack_fn, vocab_size, seq_length, tokens, tra
         premise_vector = premise_model.final_representations
         hypothesis_vector = hypothesis_model.final_representations
 
-        if FLAGS.lstm_composition or cls is rembed.plain_rnn.RNN:
+        if FLAGS.lstm_composition or cls is spinn.plain_rnn.RNN:
[…]

### 2016-04-06 python/spinn/data/boolean/load_boolean_data.py

@@ -8,7 +8,7 @@
 # sentence_label	( ( word word ) ( ( word word ) word ) )
 import numpy as np
 
-from rembed import util
+from spinn import util
 
 NUM_CLASSES = 2
 

### 2016-04-06 python/spinn/data/sst/load_sst_data.py

@@ -9,7 +9,7 @@
 import collections
 import numpy as np
 
-from rembed import util
+from spinn import util
 
 SENTENCE_PAIR_DATA = False
 

### 2016-04-06 python/spinn/fat_stack.py

@@ -6,7 +6,7 @@ import numpy as np
 import theano
 
 from theano import tensor as T
-from rembed import util
+from spinn import util
 
 
 def update_hard_stack(stack_t, stack_pushed, stack_merged, push_value,

### 2016-04-06 python/spinn/models/classifier.py

@@ -1,17 +1,17 @@
 """From the project root directory (containing data files), this can be run with:
 
 Boolean logic evaluation:
-python -m rembed.models.classifier --training_data_path bl-data/pbl_train.tsv \
+python -m spinn.models.classifier --training_data_path bl-data/pbl_train.tsv \
        --eval_data_path bl-data/pbl_dev.tsv
 
 SST sentiment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type sst --training_data_path sst-data/train.txt \
-       --eval_data_path sst-data/dev.txt --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type sst --training_data_path sst-data/train.txt \
+       --eval_data_path sst-data/dev.txt --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 SNLI entailment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
-       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
+       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 Note: If you get an error starting with "TypeError: ('Wrong number of dimensions..." during development,
@@ -22,8 +22,8 @@ Note: If you get an error starting with "TypeError: ('Wrong number of dimensions
 from copy import copy
 from functools import partial
 
-from rembed import recurrences, util
-from rembed.stack import ThinStack
+from spinn import recurrences, util
+from spinn.stack import ThinStack
 
 
 def make_sentence_stack(cls, embedding_projection_network, compose_network,

### 2016-04-06 python/spinn/models/fat_classifier.py

@@ -1,17 +1,17 @@
 """From the project root directory (containing data files), this can be run with:
 
 Boolean logic evaluation:
-python -m rembed.models.classifier --training_data_path bl-data/pbl_train.tsv \
+python -m spinn.models.classifier --training_data_path bl-data/pbl_train.tsv \
        --eval_data_path bl-data/pbl_dev.tsv
 
 SST sentiment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type sst --training_data_path sst-data/train.txt \
-       --eval_data_path sst-data/dev.txt --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type sst --training_data_path sst-data/train.txt \
+       --eval_data_path sst-data/dev.txt --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 SNLI entailment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
-       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
+       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 Note: If you get an error starting with "TypeError: ('Wrong number of dimensions..." during development,
@@ -29,14 +29,14 @@ from theano import tensor as T
 import theano
 import numpy as np
 
-from rembed import afs_safe_logger
-from rembed import util
-from rembed.data.boolean import load_boolean_data
-from rembed.data.sst import load_sst_data
-from rembed.data.snli import load_snli_data
+from spinn import afs_safe_logger
+from spinn import util
+from spinn.data.boolean import load_boolean_data
+from spinn.data.sst import load_sst_data
+from spinn.data.snli import load_snli_data
 
-import rembed.fat_stack
-import rembed.plain_rnn
+import spinn.fat_stack
+import spinn.plain_rnn
 
 
 FLAGS = gflags.FLAGS
[…]

### 2016-04-06 python/spinn/plain_rnn.py

@@ -4,7 +4,7 @@ import numpy as np
 import theano
 
 from theano import tensor as T
-from rembed import util
+from spinn import util
 
 
 class RNN(object):

### 2016-04-06 python/spinn/tests/test_plain_rnn.py

@@ -4,8 +4,8 @@ import numpy as np
 import theano
 from theano import tensor as T
 
-from rembed.plain_rnn import RNN
-from rembed.util import VariableStore, CropAndPad, IdentityLayer
+from spinn.plain_rnn import RNN
+from spinn.util import VariableStore, CropAndPad, IdentityLayer
 
 
 class RNNTestCase(unittest.TestCase):

### 2016-04-06 python/spinn/util.py

@@ -13,7 +13,7 @@ from theano.gof.fg import MissingInputError
 from theano.sandbox.cuda import HostFromGpu
 from theano.sandbox.rng_mrg import MRG_RandomStreams
 
-from rembed import cuda_util
+from spinn import cuda_util
 
 numpy_random = np.random.RandomState(1234)
 theano_random = MRG_RandomStreams(numpy_random.randint(999999))

### 2016-04-06 python/spinn/models/classifier.py
s/rembed/spinn/g

@@ -1,17 +1,17 @@
 """From the project root directory (containing data files), this can be run with:
 
 Boolean logic evaluation:
-python -m rembed.models.classifier --training_data_path bl-data/pbl_train.tsv \
+python -m spinn.models.classifier --training_data_path bl-data/pbl_train.tsv \
        --eval_data_path bl-data/pbl_dev.tsv
 
 SST sentiment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type sst --training_data_path sst-data/train.txt \
-       --eval_data_path sst-data/dev.txt --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type sst --training_data_path sst-data/train.txt \
+       --eval_data_path sst-data/dev.txt --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 SNLI entailment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
-       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
+       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 Note: If you get an error starting with "TypeError: ('Wrong number of dimensions..." during development,
@@ -29,14 +29,14 @@ from theano import tensor as T
 import theano
 import numpy as np
 
-from rembed import afs_safe_logger
-from rembed import util
-from rembed.data.boolean import load_boolean_data
-from rembed.data.sst import load_sst_data
-from rembed.data.snli import load_snli_data
+from spinn import afs_safe_logger
+from spinn import recurrences, util
+from spinn.data.boolean import load_boolean_data
+from spinn.data.sst import load_sst_data
+from spinn.data.snli import load_snli_data
+from spinn.stack import ThinStack
 
-import rembed.fat_stack
-import rembed.plain_rnn
+import spinn.plain_rnn
 
 
 FLAGS = gflags.FLAGS
[…]

### 2016-04-06 python/spinn/models/fat_classifier.py

@@ -1,17 +1,17 @@
 """From the project root directory (containing data files), this can be run with:
 
 Boolean logic evaluation:
-python -m rembed.models.classifier --training_data_path bl-data/pbl_train.tsv \
+python -m spinn.models.classifier --training_data_path bl-data/pbl_train.tsv \
        --eval_data_path bl-data/pbl_dev.tsv
 
 SST sentiment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type sst --training_data_path sst-data/train.txt \
-       --eval_data_path sst-data/dev.txt --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type sst --training_data_path sst-data/train.txt \
+       --eval_data_path sst-data/dev.txt --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 SNLI entailment (Demo only, model needs a full GloVe embeddings file to do well):
-python -m rembed.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
-       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path rembed/tests/test_embedding_matrix.5d.txt \
+python -m spinn.models.classifier --data_type snli --training_data_path snli_1.0/snli_1.0_dev.jsonl \
+       --eval_data_path snli_1.0/snli_1.0_dev.jsonl --embedding_data_path spinn/tests/test_embedding_matrix.5d.txt \
        --model_dim 10 --word_embedding_dim 5
 
 Note: If you get an error starting with "TypeError: ('Wrong number of dimensions..." during development,
@@ -29,14 +29,14 @@ from theano import tensor as T
 import theano
 import numpy as np
 
-from rembed import afs_safe_logger
-from rembed import recurrences, util
-from rembed.data.boolean import load_boolean_data
-from rembed.data.sst import load_sst_data
-from rembed.data.snli import load_snli_data
-from rembed.stack import ThinStack
+from spinn import afs_safe_logger
+from spinn import util
+from spinn.data.boolean import load_boolean_data
+from spinn.data.sst import load_sst_data
+from spinn.data.snli import load_snli_data
 
-import rembed.plain_rnn
+import spinn.fat_stack
+import spinn.plain_rnn
 
 
 FLAGS = gflags.FLAGS
[…]

### 2016-04-06 python/spinn/recurrences.py

@@ -2,14 +2,14 @@
 Implements the core recurrences for various stack models.
 
 The recurrences described here are unrolled into bona-fide stack models
-by `rembed.stack`.
+by `spinn.stack`.
 """
 
 from functools import partial
 
 from theano import tensor as T
 
-from rembed import util
+from spinn import util
 
 
 class Recurrence(object):

### 2016-04-06 python/spinn/stack.py

@@ -7,7 +7,7 @@ import theano
 from theano.ifelse import ifelse
 
 from theano import tensor as T
-from rembed import cuda_util, util
+from spinn import cuda_util, util
 
 
 def update_hard_stack(t, t_f, stack_t, push_value, merge_value, merge_queue_t,

### 2016-04-06 python/spinn/tests/test_cuda_util.py

@@ -3,7 +3,7 @@ import theano
 import theano.tensor as T
 
 # Import cuda_util in order to register the optimization
-from rembed import cuda_util
+from spinn import cuda_util
 
 
 def _test_gpu_rowwise_switch_inner(f, A, B, mask, expected):

### 2016-04-06 python/spinn/tests/test_stack.py

@@ -5,10 +5,10 @@ import numpy as np
 import theano
 from theano import tensor as T
 
-from rembed import cuda_util, util
-from rembed.stack import ThinStack
-from rembed.recurrences import Recurrence, Model0
-from rembed.util import VariableStore, CropAndPad, IdentityLayer, batch_subgraph_gradients
+from spinn import cuda_util, util
+from spinn.stack import ThinStack
+from spinn.recurrences import Recurrence, Model0
+from spinn.util import VariableStore, CropAndPad, IdentityLayer, batch_subgraph_gradients
 
 
 class ThinStackTestCase(unittest.TestCase):

### 2016-04-06 python/spinn/tests/test_util.py

@@ -2,10 +2,10 @@ import numpy as np
 
 from nose.tools import assert_equal
 
-from rembed import util
+from spinn import util
 
 
-TEST_EMBEDDING_MATRIX = "rembed/tests/test_embedding_matrix.5d.txt"
+TEST_EMBEDDING_MATRIX = "spinn/tests/test_embedding_matrix.5d.txt"
 
 
 def test_build_vocabulary_for_ascii_embedding_file():

### 2016-04-06 scripts/12_5_sst_2s.sh

@@ -3,12 +3,12 @@
 # SWEEP PARAMETERS: {'semantic_classifier_keep_rate': ('LIN', 0.4, 0.75), 'embedding_keep_rate': ('LIN', 0.4, 1.0), 'learning_rate': ('EXP', 0.0001, 0.0003), 'double_identity_init_range': ('EXP', 0.0005, 0.005), 'init_range': ('EXP', 0.001, 0.004), 'scheduled_sampling_exponent_base': ('SS_BASE', 2e-06, 0.0002), 'l2_lambda': ('EXP', 2e-07, 2e-05)}
 # FIXED_PARAMETERS: {'eval_seq_length': '150', 'clipping_max_value': '5.0', 'data_type': 'sst', 'training_data_path': 'sst-data/train_expanded.txt', 'batch_size': '32', 'embedding_data_path': '/scr/nlp/data/glove_vecs/glove.840B.300d.txt', 'model_dim': '50', 'seq_length': '100', 'ckpt_root': '/afs/cs.stanford.edu/u/sbowman/scr/', 'word_embedding_dim': '300', 'model_type': 'Model2S', 'lstm_composition': '', 'eval_data_path': 'sst-data/dev.txt:sst-data/train_sample.txt'}
 
-export REMBED_FLAGS=" --semantic_classifier_keep_rate 0.607345803836 --eval_seq_length 150 --clipping_max_value 5.0 --data_type sst --model_dim 50 --learning_rate 0.000229287344209 --training_data_path sst-data/train_expanded.txt --word_embedding_dim 300 --batch_size 32 --double_identity_init_range 0.00479811207292 --ckpt_root /afs/cs.stanford.edu/u/sbowman/scr/ --init_range 0.00173528232303 --embedding_keep_rate 0.626722888293 --seq_length 100 --embedding_data_path /scr/nlp/data/glove_vecs/glove.840B.300d.txt --model_type Model2S --scheduled_sampling_exponent_base 0.999840115881 --l2_lambda 3.27967310602e-06 --lstm_composition  --eval_data_path sst-data/dev.txt:sst-data/train_sample.txt --experiment_name 12_5_sweep_sst_Model2S_0-semantic_classifier_keep_rate0.61-learning_rate0.00023-double_identity_init_range0.0048-init_range0.0017-embedding_keep_rate0.63-scheduled_sampling_exponent_base1-l2_lambda3.3e-06"; export DEVICE=gpu5; qsub -v 
[…]

### 2016-04-06 scripts/make_snli_sweep.py

@@ -97,5 +97,5 @@ for run_id in range(sweep_runs):
                 val_disp = "%.2g" % value
             name += "-" + param + val_disp
     flags += " --experiment_name " + name
-    print "export REMBED_FLAGS=\"" + flags + "\"; export DEVICE=gpuX; qsub -v REMBED_FLAGS,DEVICE scripts/train_rembed_classifier.sh -q " + queue + " -l host=jagupardX"
+    print "export SPINN_FLAGS=\"" + flags + "\"; export DEVICE=gpuX; qsub -v SPINN_FLAGS,DEVICE scripts/train_spinn_classifier.sh -q " + queue + " -l host=jagupardX"
     print

### 2016-04-06 scripts/train_rembed_classifier.sh

@@ -1,22 +1,0 @@
-#!/bin/bash
-
-### Generic job script for all experiments.
-
-# Usage example:
-# export REMBED_FLAGS="--learning_rate 0.01 --batch_size 256"; export DEVICE=gpu2; export DEVICE=gpu0; qsub -v REMBED_FLAGS,DEVICE scripts/train_rembed_classifier.sh -l host=jagupard10
-
-# Change to the submission directory.
-cd $PBS_O_WORKDIR
-echo Lauching from working directory: $PBS_O_WORKDIR
-echo Flags: $REMBED_FLAGS
-echo Device: $DEVICE
-
-# Log what we're running and where.
-echo $PBS_JOBID - `hostname` - $DEVICE - at `git log --pretty=format:'%h' -n 1` - $REMBED_FLAGS >> ~/rembed_machine_assignments.txt
-
-# Use Jon's Theano install.
-source /u/nlp/packages/anaconda/bin/activate conda-common
-export LD_LIBRARY_PATH=/usr/local/cuda/lib64:$LD_LIBRARY_PATH
-export PYTHONPATH=/scr/jgauthie/tmp/theano-nshrdlu:$PYTHONPATH
-
-THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=$DEVICE,floatX=float32 python -m rembed.models.classifier $REMBED_FLAGS

### 2016-04-06 scripts/train_rembed_fat_classifier.sh

@@ -1,23 +1,0 @@
-#!/bin/bash
-
-### Generic job script for all experiments.
-
-# Usage example:
-# export REMBED_FLAGS="--learning_rate 0.01 --batch_size 256"; export DEVICE=gpu2; export DEVICE=gpu0; qsub -v REMBED_FLAGS,DEVICE scripts/train_rembed_classifier.sh -l host=jagupard10
-
-# Change to the submission directory.
-cd $PBS_O_WORKDIR
-echo Lauching from working directory: $PBS_O_WORKDIR
-echo Flags: $REMBED_FLAGS
-echo Device: $DEVICE
-
-# Log what we're running and where.
-echo $PBS_JOBID - `hostname` - $DEVICE - at `git log --pretty=format:'%h' -n 1` - $REMBED_FLAGS >> ~/rembed_machine_assignments.txt
-
-# Use Jon's Theano install.
-source /u/nlp/packages/anaconda/bin/activate conda-common
-export LD_LIBRARY_PATH=/usr/local/cuda/lib64:$LD_LIBRARY_PATH
-export PYTHONPATH=/scr/jgauthie/tmp/theano-nshrdlu:$PYTHONPATH
-export PATH=/usr/local/cuda/bin:$PATH
-
-THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=$DEVICE,floatX=float32 python -m rembed.models.fat_classifier $REMBED_FLAGS

### 2016-04-06 scripts/train_spinn_classifier.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+### Generic job script for all experiments.
+
+# Usage example:
+# export SPINN_FLAGS="--learning_rate 0.01 --batch_size 256"; export DEVICE=gpu2; export DEVICE=gpu0; qsub -v SPINN_FLAGS,DEVICE scripts/train_spinn_classifier.sh -l host=jagupard10
+
+# Change to the submission directory.
+cd $PBS_O_WORKDIR
+echo Lauching from working directory: $PBS_O_WORKDIR
+echo Flags: $SPINN_FLAGS
+echo Device: $DEVICE
+
+# Log what we're running and where.
+echo $PBS_JOBID - `hostname` - $DEVICE - at `git log --pretty=format:'%h' -n 1` - $SPINN_FLAGS >> ~/spinn_machine_assignments.txt
+
+# Use Jon's Theano install.
+source /u/nlp/packages/anaconda/bin/activate conda-common
+export LD_LIBRARY_PATH=/usr/local/cuda/lib64:$LD_LIBRARY_PATH
+export PYTHONPATH=/scr/jgauthie/tmp/theano-nshrdlu:$PYTHONPATH
+
+THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=$DEVICE,floatX=float32 python -m spinn.models.classifier $SPINN_FLAGS

### 2016-04-06 scripts/train_spinn_fat_classifier.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+### Generic job script for all experiments.
+
+# Usage example:
+# export SPINN_FLAGS="--learning_rate 0.01 --batch_size 256"; export DEVICE=gpu2; export DEVICE=gpu0; qsub -v SPINN_FLAGS,DEVICE scripts/train_spinn_classifier.sh -l host=jagupard10
+
+# Change to the submission directory.
+cd $PBS_O_WORKDIR
+echo Lauching from working directory: $PBS_O_WORKDIR
+echo Flags: $SPINN_FLAGS
+echo Device: $DEVICE
+
+# Log what we're running and where.
+echo $PBS_JOBID - `hostname` - $DEVICE - at `git log --pretty=format:'%h' -n 1` - $SPINN_FLAGS >> ~/spinn_machine_assignments.txt
+
+# Use Jon's Theano install.
+source /u/nlp/packages/anaconda/bin/activate conda-common
+export LD_LIBRARY_PATH=/usr/local/cuda/lib64:$LD_LIBRARY_PATH
+export PYTHONPATH=/scr/jgauthie/tmp/theano-nshrdlu:$PYTHONPATH
+export PATH=/usr/local/cuda/bin:$PATH
+
+THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=$DEVICE,floatX=float32 python -m spinn.models.fat_classifier $SPINN_FLAGS

### 2016-05-04 cpp/bin/stacktest.cc
clean up stacktest.cc a bit

@@ -1,3 +1,23 @@
+/**
+ * This binary runs a basic speed test of the C++ thin-stack implementation
+ * of SPINN in this package.
+ *
+ * The SPINN model and the thin-stack algorithm are described in our paper:
+ *
+ *     A Fast Unified Model for Sentence Parsing and Understanding.
+ *     Samuel R. Bowman, Jon Gauthier, Abhinav Rastogi, Raghav Gupta,
+ *     Christopher D. Manning, and Christopher Potts. arXiv March 2016.
+ *     http://arxiv.org/abs/1603.06021
+ *
+ * This script loads fixed parameters from the local `params` directory,
+ * generates random input data, and performs a SPINN feedforward as
+ * described in the paper. It outputs timing information and the final
+ * values at the top of the stack after the feedforward.
+ *
+ * For runtime instructions see the README in the root of this project
+ * directory.
+ */
+
 #include <chrono>
 #include <iostream>
 
@@ -30,7 +50,15 @@ void destroy_params(ThinStackParameters params) {
 }
 
 int main() {
-  ModelSpec spec = {300, 300, (size_t) atoi(getenv("BATCH_SIZE")), 10, 59, 300};
+  ModelSpec spec = {
+      300, // Dimension of stack values / constituent node values
+      300, // Word embedding dimension / tree leaf embedding dimension
+      (size_t) atoi(getenv("BATCH_SIZE")), // Batch size
+      10, // Vocabulary size
+      59, // Transition sequence length (== 2 * (sentence length) - 1)
+      300 // Ignore -- unused.
+  };
+
   ThinStackParameters params = load_params(spec);
 
   cublasHandle_t handle;
@@ -49,6 +77,10 @@ int main() {
 
   cout << "transitions:" << endl;
   float *h_transitions = (float *) malloc(spec.seq_length * spec.batch_size * sizeof(float));
+  // Build a batch of random transition sequences which somewhat resemble
+  // realistic transition sequences. (These are perhaps a bit more difficult
+  // than realistic transition sequence batches in that there will be less
+  // overlap w.r.t. merge locations within the sequence.)
   for (int i = 0; i < spec.seq_length * spec.batch_size; i++) {
     float val;
     if (i < spec.batch_size * 2) {
@@ -60,7 +92,9 @@ int main() {
     }
     h_transitions[i] = val;
   }
[…]

### 2016-05-04 cpp/rnn.cc
More documentation improvements

@@ -1,3 +1,11 @@
+/**
+ * Implements a basic RNN feedforward (ReLU activation) for speed comparison.
+ *
+ * You can run this code with the script in `bin/rnntest.cc`. See the README in
+ * the root directory of this project for full usage instructions.
+ */
+
+
 #include "rnn.h"
 using namespace std;
 namespace k = kernels;

### 2016-05-04 cpp/thin-stack.cc

@@ -1,3 +1,24 @@
+/**
+ * This file contains a C++/CUDA implementation of the thin-stack algorithm.
+ * Thin-stack computes the same function as a vanilla recursive neural network
+ * and as the SPINN-PI-NT model.
+ *
+ * The SPINN model and the thin-stack algorithm are described in our paper:
+ *
+ *     A Fast Unified Model for Sentence Parsing and Understanding.
+ *     Samuel R. Bowman, Jon Gauthier, Abhinav Rastogi, Raghav Gupta,
+ *     Christopher D. Manning, and Christopher Potts. arXiv March 2016.
+ *     http://arxiv.org/abs/1603.06021
+ *
+ * The exact model implemented here is a recursive neural network (equivalent
+ * to SPINN-PI-NT) with ReLU activations. It has been verified to compute the
+ * exact same function of its inputs as a recursive neural network.
+ *
+ * You can execute this code using the script `bin/stacktest`. See the README
+ * in the root directory of this project for more usage instructions.
+ */
+
+
 #include "thin-stack.h"
 using namespace std;
 namespace k = kernels;

### 2016-05-04 writing/hard_stack_paper/paper.tex
New paper TODO.
runtime.csv -> runtime.tsv

@@
 \input{../model1_fig.tex}
+
+\todo{[SB] Use $a$ for transitions.}
 
@@
              legend style={at={(0.55, 0.5)},anchor=north west}]
-  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
+  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.tsv};
   \addlegendentry{Thin-stack GPU}
-  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
+  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.tsv};
   \addlegendentry{CPU \citep{irsoy2014deep}}
-  \addplot table [x={Batch size}, y=RNN, col sep=tab] {runtime.csv};
+  \addplot table [x={Batch size}, y=RNN, col sep=tab] {runtime.tsv};
   \addlegendentry{RNN}

### 2016-05-27 python/spinn/util/__init__.py
Redo model 1 fix from 49967197f83e8 by rearranging util imports

@@ -7,8 +7,8 @@ if cuda_available:
     from spinn.util.cuda import *
 
 from spinn.util.theano_internal import *
-from spinn.util.blocks import *
 from spinn.util.data import *
+from spinn.util.blocks import *
 from spinn.util.variable_store import VariableStore
 
 

### 2016-05-27 python/spinn/util/blocks.py

@@ -8,11 +8,12 @@ import theano
 from theano import tensor as T
 from theano.sandbox.rng_mrg import MRG_RandomStreams
 
+from spinn.util import NUM_TRANSITION_TYPES
+
 
 numpy_random = np.random.RandomState(1234)
 theano_random = MRG_RandomStreams(numpy_random.randint(999999))
 
-NUM_TRANSITION_TYPES = 2
 
 def UniformInitializer(range):
     return lambda shape, **kwargs: np.random.uniform(-range, range, shape)

### 2016-05-27 writing/hard_stack_paper/paper.tex
Incremental paper update. Off to lunch.

@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
+\aclfinalcopy % Uncomment this line for the final submission
+\def\aclpaperid{265} %  Enter the acl Paper ID here
 %\setlength\titlebox{5cm}
@@
 
-     \author{}
-%    \author{
-%    Samuel R.\ Bowman$^{1,2,5,}$\thanks{~\,The first two authors contributed equally.} \\
-%    \texttt{\small sbowman@stanford.edu} \\
-%    \And
-%    Jon Gauthier$^{2,3,5,*}$ \\
-%    \texttt{\small jgauthie@stanford.edu} \\
-%    \And
-%    Abhinav Rastogi$^{4,5}$ \\
-%    \texttt{\small arastogi@stanford.edu} \\
-%    \AND
-%    Raghav Gupta$^{6}$ \\
-%    \texttt{\small rgupta93@stanford.edu} \\
-%    \And
-%    Christopher D.\ Manning$^{1,2,5,6}$\\
-%    \texttt{\small manning@stanford.edu}\\
-%    \And
-%    Christopher Potts$^{1}$\\
-%    \texttt{\small cgpotts@stanford.edu}
-%    \AND\\[-3ex]
-%    {$^{1}$Stanford Linguistics\quad
-%    $^{2}$Stanford NLP Group\quad
-%    $^{3}$Stanford Symbolic Systems}\\
-%    {$^{4}$Stanford Electrical Engineering\quad
-%    $^{5}$Stanford AI Lab\quad
-%    $^{6}$Stanford Computer Science}
-%    }
+\author{
+Samuel R.\ Bowman$^{1,2,5,}$\thanks{~\,The first two authors contributed equally.} \\
+\texttt{\small sbowman@stanford.edu} \\
+\And
+Jon Gauthier$^{2,3,5,*}$ \\
+\texttt{\small jgauthie@stanford.edu} \\
+\And
+Abhinav Rastogi$^{4,5}$ \\
+\texttt{\small arastogi@stanford.edu} \\
+\AND
+Raghav Gupta$^{6}$ \\
+\texttt{\small rgupta93@stanford.edu} \\
+\And
+Christopher D.\ Manning$^{1,2,5,6}$\\
+\texttt{\small manning@stanford.edu}\\
+\And
+Christopher Potts$^{1}$\\
+\texttt{\small cgpotts@stanford.edu}
+\AND\\[-3ex]
+{$^{1}$Stanford Linguistics\quad
+$^{2}$Stanford NLP Group\quad
+$^{3}$Stanford Symbolic Systems}\\
+{$^{4}$Stanford Electrical Engineering\quad
+$^{5}$Stanford AI Lab\quad
+$^{6}$Stanford Computer Science}
+}
 
@@
 
-We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence-encoding-based models, and that it yields speed increases of up to 25$\times$ over a standard TreeRNN implementation.
+We evaluate SPINN on the Stanford Natural Language Inference entailment task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence-encoding-based models, even with a relatively simple and underpowered implementation of the built-in parser.
+We also find that SPINN yields speed increases of up to 25$\times$ over a standard TreeRNN implementation.
 
@@
 
-For SPINN to operate on unparsed inputs, it needs to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply it as part of the input.
-To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier whose input is the state of the tracking LSTM:
+For SPINN to operate on unparsed inputs, it needs to produce its own transition sequence $\mathbf a$ rather than relying on an external parser to supply it as part of the input.
+To do this, the model predicts $a_t$ at each step using a simple two-way softmax classifier whose input is the state of the tracking LSTM:
 \begin{equation}
-\vec{p}_{\text{t}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{tracking}} + \vec{b}_{\text{trans}})
+\vec{p}_{\text{a}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{tracking}} + \vec{b}_{\text{trans}})
 \end{equation}
-At test time, the model uses whichever transition (i.e., \shift\ or \reduce) is assigned a higher probability.
+Note that this is the close to being simplest viable implementation of a transition decision function.
+In contrast, the decision functions in sate-of-the-art transition-based parsers tend to use significantly richer feature sets as inputs, including features containing information about several upcoming words on the buffer instead of just one.
+We use this simple parser here, and leave a more thorough investigation of this component of the model to future work.
+
+At test time, the model uses whichever transition (i.e., \shift\ or \reduce) is assigned a higher (unnormalized) probability.
 The prediction function is trained to mimic the decisions of an external parser.
@@
 {neutral}
+The three SPINN models appear to be 
+\todo{Replace above with example of negation, and stats}
+
+\todo{Replace bewol with comment/example about very long sentences exaggerating difference.}
 

### 2016-05-27 writing/hard_stack_paper/paper.tex
First awkward stab at lengthier thin stack explanation

@@
 Each row $S_t$ represents the top of the actual stack at timestep $t$.
-We maintain a queue of backpointers into $S$ that indicates which elements would be involved in a \reduce\ operation at any given step.
+At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element.
+To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$.
+We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack.
+To retrieve the top two elements of the stack, we look up the two rows indicated by the final two elements of the queue $Q$.
+These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack.
 Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation.

### 2016-05-27 writing/hard_stack_paper/paper.tex
Revised error analysis.

@@
 The use of tree structure improves the performance of sentence-encoding models for SNLI.
-We suspect that this improvement is largely attributable to the more efficient learning of accurate generalizations overall, and not to any particular few phenomena.
+We suspect that this improvement is largely due to the more efficient learning of accurate generalizations overall, and not to any particular few phenomena.
 However, some patterns are identifiable in the results.
-In particular, it seems that tree-structured models are better able to focus on the key actors and events named in a sentence (possibly by learning an approximation of the linguistic notion of headedness), and that this allows them to better focus on the central actors and events described in a sentence.
-For example, this pair was classified successfully by all three SPINN variants, but not by the RNN baseline (key words emphasized):
+
+While all four models under study have trouble with negation, the tree-structured SPINN models do quite substantially better on these pairs.
+This is likely due to the fact that parse trees make the scope of any instance of negation (the portion of the sentence's content that is negated) relatively easy to identify and separate from the rest of the sentence.
+For test set sentence pairs like the one below where negation (\word{not} or \word{n't}) does not appear in the premise but does appear in the hypothesis, the RNN shows $67\%$ accuracy, while all three tree-structured models exceed $73\%$.
+Only the RNN got the below example wrong:
 
 \snli
-{A \textit{woman} in red blouse is \textit{standing} with small blond \textit{child} in front of a small folding chalkboard.}
-{a \textit{woman} \textit{stands} with her \textit{child}}
-{neutral}
-The three SPINN models appear to be 
-\todo{Replace above with example of negation, and stats}
+{The rhythmic gymnast completes her floor exercise at the competition.}
+{The gymnast cannot finish her exercise.}
+{contradiction}
+\noindent Note that the presence of negation in the hypothesis is correlated with a label of \textit{contradiction} in SNLI, but not as strongly as one might intuit---only $45\%$ of these examples in the test set are labeled as contradictions.
 
-\todo{Replace bewol with comment/example about very long sentences exaggerating difference.}
+In addition, it seems that tree-structured models, and especially the tree-sequence hybrid models, are more effective than RNNs at extracting informative representations of long sentences.
+The RNN model falls off in test accuracy more quickly with increasing sentence length than SPINN-PI-NT, which in turn falls of substantially faster than the two hybrid models, repeating a pattern seen more dramatically on artificial data in \citet{bowman2015trees}.
+For example, all three SPINN models labeled this pair correctly, while the RNN did not:
 
-The tree-sequence hybrid models seem to be especially good at reasoning over pairs of sentences with very different structures, for which it is helpful to use strict compositionality with the tree-structured component to get the meanings of the sentence parts, but then to also accumulate a broad sentence summary that abstracts away from the precise structures of each sentences.
-For example, this pair is classified correctly by both hybrid models, but not by the RNN or the SPINN-PI-NT:
-
-\snli{Nurses in a medical setting conversing over a plastic cup.}
-{The nurses are discussing a patient.}
-{neutral}
+\snli
+{A man wearing glasses and a ragged costume is playing a Jaguar electric guitar and singing with the accompaniment of a drummer.}
+{A man with glasses and a disheveled outfit is playing a guitar and singing along with a drummer.}
+{entailment}
 
@@
 
-%    \subsubsection*{Acknowledgments}
-%
-%    We acknowledge financial support from a Google Faculty Research Award, the Stanford Data Science Initiative, and the National Science Foundation under grant no.~BCS 1456077.
+\subsubsection*{Acknowledgments}
+
+We acknowledge financial support from a Google Faculty Research Award, the Stanford Data Science Initiative, and the National Science Foundation under grant no.~BCS 1456077.
 Some of the Tesla K40s used for this research were donated by the NVIDIA Corporation.

### 2016-05-27 writing/hard_stack_paper/paper.tex
Summary of confusion matrix.

@@
 
+Finally, as is likely inevitable for models evaluated on SNLI, all four models under study did several percent worse on test examples whose ground truth label is \textit{neutral} than on examples of the other two classes.
+\textit{entailment}--\textit{neutral} and \textit{neutral}--\textit{contradiction} confusions appear to be much harder to avoid than \textit{entailment}--\textit{contradiction} confusions, where relatively superficial cues might be more readily useful.
+
 \section{Conclusions and future work}

### 2016-05-27 writing/hard_stack_paper/paper.tex
Fix transition notation; move tuning out of appendix.

@@
 
-\todo{[SB] Use $a$ for transitions.}
-
 A wide range of current models in NLP are built around a neural network component that produces vector representations of sentence meaning \citep[e.g.,][]{sutskever2014sequence,tai2015improved}.
@@
 
-A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_0, \dots, x_{N-1})$ and consumes transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies one step of the parsing process.
+A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_0, \dots, x_{N-1})$ and consumes transitions $\mathbf a = (a_0, \dots, a_{T-1})$, where each $a_t \in \{\shift, \reduce\}$ specifies one step of the parsing process.
 In general a parser may also generate these transitions on the fly as it reads the tokens.
 It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
-For any binary-branching tree structure over $N$ words, this requires $2N - 1$ transitions.
+For any binary-branching tree structure over $N$ words, this requires $T=2N-1$ transitions and $2N$ states.
 
@@
 Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state.
-It next proceeds through the transition sequence, where each transition $t_t$ selects one of the two following operations.
+It next proceeds through the transition sequence, where each transition $a_t$ selects one of the two following operations.
 Below, the $\mid$ symbol denotes the \textit{cons} (concatenation) operator.
@@
 We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}.
-For each input sentence, we represent the stack with a single $T \times D$ matrix $S$.
+For each input sentence, we represent the stack with a single $(T + 1) \times D$ matrix $S$.
 Each row $S_t$ represents the top of the actual stack at timestep $t$.
@@
 Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation.
-It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$.
+It operates on the single $(T + 1) \times D$ matrix $S$ and a backpointer queue $Q$.
 Table~\ref{tbl:thin-stack} shows an example run.
@@
 \begin{algorithmic}[1]
-  \Function{Step}{bufferTop, op, $t$, $S$, $Q$}
-    \If{op = \shift}
+  \Function{Step}{bufferTop, $a$, $t$, $S$, $Q$}
+    \If{$a$ = \shift}
       \State $S$[$t$] := bufferTop
-    \ElsIf{op = \reduce}
+    \ElsIf{$a$ = \reduce}
       \State right := $S$[$Q$.pop()]
@@
 
-\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence.
+\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf a$  and a token sequence $\mathbf x$ as its inputs for each sentence.
 The token sequence is simply the words in the sentence in order.
-$\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift\ transition and each `)' as a \reduce\ transition, as here:
+$\mathbf a$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift\ transition and each `)' as a \reduce\ transition, as here:
 
@@
 {\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
-{$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
+{$\mathbf a$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
 {$\mathbf x$}: the, cat, sat, down
@@
 We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI\@.
-Transition sequences $\mathbf t$ are cropped at the left or padded at the left with \shift s.
-Token sequences $\mathbf x$ are then cropped or padded with empty tokens at the left to match the number of \shift s added or removed from $\mathbf t$, and can then be padded with empty tokens at the right to meet the desired length $N$.
+Transition sequences $\mathbf a$ are cropped at the left or padded at the left with \shift s.
+Token sequences $\mathbf x$ are then cropped or padded with empty tokens at the left to match the number of \shift s added or removed from $\mathbf a$, and can then be padded with empty tokens at the right to meet the desired length $N$.
 
@@
 
+\begin{table*}[t]\small
+\begin{center}
+\begin{tabular}{lrlrrrr}
+\toprule
+Param.     & Range & Strategy        & RNN       & SPINN-PI-NT   & SPINN-PI  & SPINN \\
+\midrule 
+Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & 2e-3\\
+L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & 3e-5\\
+Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & --- & --- & ---  & 3.9    \\
+Embedding transformation dropout keep rate & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
+Classifier MLP dropout keep rate & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
+Tracking LSTM size $D_\text{tracking}$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
+Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & 1\\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:hyperparams}
+Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log--uniform.
+}
+\end{table*}
+
 \paragraph{Creating a sentence-pair classifier} \label{sec:classifier}
@@
 Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of 1024D ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution over the three labels.
+
+\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}^{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_t^{\text{p}}, \mathcal{L}_t^{\text{h}}\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters.
+The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
+\begin{equation}
+\begin{split}
+\mathcal{L}_{\text{m}} = &\mathcal{L}^{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_t^{\text{p}} + \mathcal{L}_t^{\text{h}}) + \lambda \norm{\theta}^2_2
+\end{split}
+\end{equation}
+
+\paragraph{Initialization, optimization, and tuning}
+
+We initialize the model parameters using the nonparametric strategy of \citet{DBLP:journals/corr/HeZR015}, with the exception of the softmax classifier parameters, which we initialize using random uniform samples from $[-0.005, 0.005]$.
+
+We use minibatch SGD with the RMSProp optimizer \citep{tieleman2012lecture} and a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
+We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the output of the word embedding projection layer and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
+
+We train each model for 250k steps in each run, using a batch size of 32.
+We track each model's performance on the development set during training and save parameters when this performance reaches a new peak.
+We use early stopping, evaluating on the test set using the parameters that perform best on the development set.
+
+We use random search to tune the hyperparameters of each model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and then launching eight copies of each experiment each with newly sampled hyperparameters from those ranges.
+Table~\ref{tab:hyperparams} shows the hyperparameters used in the best run of each model.
+
+\subsection{Models evaluated}
+
+We evaluate four models.
+The four all use the sentence-pair classifier architecture described in Section~\ref{sec:classifier}, and differ only in the function computing the sentence encodings.
+First, a single-layer LSTM RNN \citep[similar to that of][]{snli:emnlp2015} serves as a baseline encoder.
+Next, the minimal SPINN-PI-NT model (equivalent to a TreeLSTM) introduces the SPINN model design.
+SPINN-PI adds the tracking LSTM to that design.
+Finally, the full SPINN adds the integrated parser.
+
+We compare our models against several baselines, including the strongest published non-neural network-based result from \citet{snli:emnlp2015} and previous neural network models built around several types of sentence encoders.
 
@@
 \end{table*}
-
-\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_t^p, \mathcal{L}_t^h\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters.
-The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
-\begin{equation}
-\begin{split}
-\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{t}}^{p} + \mathcal{L}_{\text{t}}^{h}) + \lambda \norm{\theta}^2_2
-\end{split}
-\end{equation}
-
-\paragraph{Initialization, optimization, and tuning}
-
-We initialize the model parameters using the nonparametric strategy of \citet{DBLP:journals/corr/HeZR015}, with the exception of the softmax classifier parameters, which we initialize using random uniform samples from $[-0.005, 0.005]$.
-
-We use minibatch SGD with the RMSProp optimizer \citep{tieleman2012lecture} and a tuned starting learning rate that decays by a factor of 0.75 every 10k steps.
-We apply both dropout \citep{srivastava2014dropout} and batch normalization \citep{2015SIoffeCSzegedy} to the output of the word embedding projection layer and to the feature vectors that serve as the inputs and outputs to the MLP that precedes the final entailment classifier.
-
-We train each model for 250k steps in each run, using a batch size of 32.
-We track each model's performance on the development set during training and save parameters when this performance reaches a new peak.
-We use early stopping, evaluating on the test set using the parameters that perform best on the development set.
-
-A supplement discusses hyperparameter tuning.
-% TODO for arXiv: Make to appendix.
-
-\subsection{Models evaluated}
-
-We evaluate four models.
-The four all use the sentence-pair classifier architecture described in Section~\ref{sec:classifier}, and differ only in the function computing the sentence encodings.
-First, a single-layer LSTM RNN \citep[similar to that of][]{snli:emnlp2015} serves as a baseline encoder.
-Next, the minimal SPINN-PI-NT model (equivalent to a TreeLSTM) introduces the SPINN model design.
-SPINN-PI adds the tracking LSTM to that design.
-Finally, the full SPINN adds the integrated parser.
-
-We compare our models against several baselines, including the strongest published non-neural network-based result from \citet{snli:emnlp2015} and previous neural network models built around several types of sentence encoders.
 

### 2016-05-29 writing/hard_stack_paper/paper.tex
Affiliation update.
More thin stack description tweaks

@@
 \AND
-Raghav Gupta$^{6}$ \\
+Raghav Gupta$^{2,5,6}$ \\
 \texttt{\small rgupta93@stanford.edu} \\
@@
 To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$.
+In order to perform the \reduce\ operation, we need to retrieve the top two elements of the actual stack.
 We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack.
-To retrieve the top two elements of the stack, we look up the two rows indicated by the final two elements of the queue $Q$.
+The top two elements of the stack can be found by using the final two pointers in the queue $Q$.
 These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack.

### 2016-05-29 writing/hard_stack_paper/batching_fig.tex
Edits to pp. 1-4.
Incremental paper update. Off to lunch.
Re-style acknowledgments, batching figure.

@@
 \begin{center}
-\scalebox{0.75}{
+\scalebox{0.8}{
  \begin{tikzpicture}
-    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]    
+    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt]    
     \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black] 
@@
 \begin{center}
-\scalebox{0.75}{
+\scalebox{0.8}{
  \begin{tikzpicture}
-    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]    
-    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt,draw=black]
+    \tikzstyle{word}=[fill=yellow!40,text height=2mm,line width=1pt]    
+    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm,line width=1pt]
     \tikzstyle{alt}=[fill=orange!40]    
     \pgfsetarrowsend{latex}
-    \tikzset{edge from parent/.append style={<-, line width=1pt}}
+    \tikzset{edge from parent/.append style={<-, line width=1pt, >=latex}}
 
@@
 
-\caption{An illustration of two standard designs for sentence encoders.
+\caption{\label{fig:batching} An illustration of two standard designs for sentence encoders.
 The TreeRNN, unlike the sequence-based RNN, requires a substantially different connection structure for each sentence, making batched computation impractical.}

### 2016-05-29 writing/hard_stack_paper/paper.tex
Acknowledgments.
Re-style acknowledgments, batching figure.

@@
 \author{
-Samuel R.\ Bowman$^{1,2,5,}$\thanks{~\,The first two authors contributed equally.} \\
+Samuel R.\ Bowman$^{1,2,3,}$\thanks{~\,The first two authors contributed equally.} \\
 \texttt{\small sbowman@stanford.edu} \\
 \And
-Jon Gauthier$^{2,3,5,*}$ \\
+Jon Gauthier$^{2,3,4,*}$ \\
 \texttt{\small jgauthie@stanford.edu} \\
 \And
-Abhinav Rastogi$^{4,5}$ \\
+Abhinav Rastogi$^{3,5}$ \\
 \texttt{\small arastogi@stanford.edu} \\
 \AND
-Raghav Gupta$^{2,5,6}$ \\
+Raghav Gupta$^{2,3,6}$ \\
 \texttt{\small rgupta93@stanford.edu} \\
 \And
-Christopher D.\ Manning$^{1,2,5,6}$\\
+Christopher D.\ Manning$^{1,2,3,6}$\\
 \texttt{\small manning@stanford.edu}\\
@@
 $^{2}$Stanford NLP Group\quad
-$^{3}$Stanford Symbolic Systems}\\
-{$^{4}$Stanford Electrical Engineering\quad
-$^{5}$Stanford AI Lab\quad
+$^{3}$Stanford AI Lab}\\
+{$^{4}$Stanford Symbolic Systems\quad
+$^{5}$Stanford Electrical Engineering\quad
 $^{6}$Stanford Computer Science}
@@
 Some of the Tesla K40s used for this research were donated by the NVIDIA Corporation.
+We also thank Kelvin Gu, Noah Goodman, and many others in the Stanford NLP group for helpful comments.
 

### 2016-05-29 writing/hard_stack_paper/paper.tex
Misc paper revisions.

@@
 This is a simple sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step.
-It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, which is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$, see (\ref{eqn:lstm1}), to the composition function, allowing context information to enter the construction of sentence meaning,  forming what is effectively a tree-sequence hybrid model.
+It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, which is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ to the composition function---see (\ref{eqn:lstm1})---allowing context information to enter the construction of sentence meaning and forming what is effectively a tree-sequence hybrid model.
 
@@
 \end{equation}
-Note that this is the close to being simplest viable implementation of a transition decision function.
-In contrast, the decision functions in sate-of-the-art transition-based parsers tend to use significantly richer feature sets as inputs, including features containing information about several upcoming words on the buffer instead of just one.
-We use this simple parser here, and leave a more thorough investigation of this component of the model to future work.
+This is the close to being simplest viable implementation of a transition decision function.
+In contrast, the decision functions in sate-of-the-art transition-based parsers tend to use significantly richer feature sets as inputs, including features containing information about several upcoming words on the buffer instead of just one supplied to the tracking LSTM.
 
@@
 \centering
-\begin{tabular}{c @{\hspace*{2em}}c @{\hspace*{2em}} l}
+\begin{tabular}{cc @{\hspace*{2em}}c @{\hspace*{2em}} l}
   \toprule
-  $t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} \\
+  $t$ & $a_t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} \\
   \midrule
-  0 & & \underline{\hphantom{0} \hphantom{0}} \\
-  1 & $a$ & \hphantom{0} \underline{\hphantom{0} 1} \\
-  2 & $b$ & \hphantom{0} \hphantom{0} \underline{1 2} \\
-  3 & $c$ & \hphantom{0} \hphantom{0} 1 \underline{2 3} \\
-  4 & $(c~b)$ & \hphantom{0} \hphantom{0} \underline{1 4} \\
-  5 & $((c~b)~a)$ & \hphantom{0} \underline{\hphantom{0} 5} \\
+  0 & \shift &  & \underline{\hphantom{0} \hphantom{0}} \\
+  1 & \shift & \word{Spot} & \hphantom{0} \underline{\hphantom{0} 1} \\
+  2 & \shift & \word{sat} & \hphantom{0} \hphantom{0} \underline{1 2} \\
+  3 & \reduce & \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} \\
+  4 & \reduce & \word{sat down} & \hphantom{0} \hphantom{0} \underline{1 4} \\
+  5 & & \word{Spot} (\word{sat down}) & \hphantom{0} \underline{\hphantom{0} 5} \\
   \bottomrule
 \end{tabular}
-\caption{The thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce\ sequence on the input sentence $(a, b, c)$.
-$S$ is shown in the second column and represents the top of the stack at each step $t$.
+\caption{The thin-stack algorithm operating on the input sequence $\mathbf x = \text{(\word{Spot}, \word{sat}, \word{down})}$ and the transition sequence shown in the second column.
+$S$ is shown in the third column and represents the top of the stack at each step $t$.
 The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next step.}
@@
 We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}.
-For each input sentence, we represent the stack with a single $(T + 1) \times D$ matrix $S$.
-Each row $S_t$ represents the top of the actual stack at timestep $t$.
+For each input sentence, we represent the stack with a single $T \times D$ matrix $S$.
+Each row $S[t]$ (for $0<t\le T$) represents the top of the actual stack at timestep $t$.
 At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element.
@@
 Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation.
-It operates on the single $(T + 1) \times D$ matrix $S$ and a backpointer queue $Q$.
+It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$.
 Table~\ref{tbl:thin-stack} shows an example run.
@@
 {\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
-{$\mathbf a$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
-{$\mathbf x$}: the, cat, sat, down
+{$\mathbf x$}: \word{the}, \word{cat}, \word{sat}, \word{down}\\
+{$\mathbf a$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce
 }
@@
 In this section, we compare the test-time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and with a standard RNN sequence model.
-While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of all three models.
+While the full models evaluated below are implemented and trained using Theano \citep{theano}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of all three models.
 To do this, we reimplement the feedforward\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} of SPINN-PI-NT and an LSTM RNN baseline in C++/CUDA, and compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as SPINN-PI-NT.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.
+Our optimized C++/CUDA models and the Theano source code for the full SPINN are available at \url{https://github.com/hans/spinn}.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.
 
@@
 \centering
-\resizebox{0.85\linewidth}{!}{
+\resizebox{1\linewidth}{!}{
 \begin{tikzpicture}
@@
 \end{equation}
-Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of 1024D ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution over the three labels.
+This feature vector is then passed to a series of 1024D ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution over the three labels.
 
-\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}^{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_t^{\text{p}}, \mathcal{L}_t^{\text{h}}\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters.
+\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, cross-entropy objectives $\mathcal{L}^t_{\text{p}}$ and $\mathcal{L}^t_{\text{h}}$ for the parsing decision for each of the two sentences at each step $t$, and an L2 regularization term on the trained parameters.
 The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
@@
 \begin{split}
-\mathcal{L}_{\text{m}} = &\mathcal{L}^{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_t^{\text{p}} + \mathcal{L}_t^{\text{h}}) + \lambda \norm{\theta}^2_2
+\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}^t_{\text{p}} + \mathcal{L}^t_{\text{h}}) + \lambda \norm{\theta}^2_2
 \end{split}
@@
 Finally, as is likely inevitable for models evaluated on SNLI, all four models under study did several percent worse on test examples whose ground truth label is \textit{neutral} than on examples of the other two classes.
-\textit{entailment}--\textit{neutral} and \textit{neutral}--\textit{contradiction} confusions appear to be much harder to avoid than \textit{entailment}--\textit{contradiction} confusions, where relatively superficial cues might be more readily useful.
+\textit{Entailment}--\textit{neutral} and \textit{neutral}--\textit{contradiction} confusions appear to be much harder to avoid than \textit{entailment}--\textit{contradiction} confusions, where relatively superficial cues might be more readily useful.
 

### 2016-05-29 writing/hard_stack_paper/paper.tex
URL update

@@
 To do this, we reimplement the feedforward\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} of SPINN-PI-NT and an LSTM RNN baseline in C++/CUDA, and compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as SPINN-PI-NT.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}.
-Our optimized C++/CUDA models and the Theano source code for the full SPINN are available at \url{https://github.com/hans/spinn}.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.
+Our optimized C++/CUDA models and the Theano source code for the full SPINN are available at \url{https://github.com/stanfordnlp/spinn}.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.
 

### 2016-05-30 writing/hard_stack_paper/paper.tex
Incorporating Cheng/Lapata result update.

@@
 
-Because this paper aims to introduce a general purpose model for sentence encoding, we do not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 89.0\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
+Because this paper aims to introduce a general purpose model for sentence encoding, we do not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning,wang2015learning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 86.1\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 

### 2016-05-31 writing/hard_stack_paper/paper.tex
Typo smash!

@@
 This is the close to being simplest viable implementation of a transition decision function.
-In contrast, the decision functions in sate-of-the-art transition-based parsers tend to use significantly richer feature sets as inputs, including features containing information about several upcoming words on the buffer instead of just one supplied to the tracking LSTM.
+In contrast, the decision functions in state-of-the-art transition-based parsers tend to use significantly richer feature sets as inputs, including features containing information about several upcoming words on the buffer instead of just one supplied to the tracking LSTM.
 

### 2016-05-31 writing/hard_stack_paper/paper.tex
Reformulate the whole extra para on decision functions

--- 2016-05-30 writing/hard_stack_paper/paper.tex
@@
 \end{equation}
-This is the close to being simplest viable implementation of a transition decision function.
-In contrast, the decision functions in sate-of-the-art transition-based parsers tend to use significantly richer feature sets as inputs, including features containing information about several upcoming words on the buffer instead of just one supplied to the tracking LSTM.
+The above model is nearly the simplest viable implementation of a transition decision function.
+In contrast, the decision functions in state-of-the-art transition-based parsers tend to use significantly richer feature sets as inputs, including features containing information about several upcoming words on the buffer.
+The value $\vec{h}_{\text{tracking}}$ is a function of only the very top of the buffer and the top two stack elements at each timestep.
 

### 2016-05-31 writing/hard_stack_paper/paper.tex
Rewrite thin-stack intro

@@
 
-A na\"ive implementation of SPINN would require representing a stack of size $N$ for each timestep of each input sentence at training time to support backpropagation.
-This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$.
-Such a na\"ive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the stack.
+A na\"ive implementation of SPINN needs to handle a size $O(N)$ stack at each timestep in an example, any element of which may be involved in later computations.
+A na\"ive backpropagation implementation would then require storing each of the $O(N)$ stacks for a backward pass, leading to a per-example space requirement of $O(NTD)$ floats.
+This requirement is prohibitively large for significant batch sizes or sentence lengths $N$.
+Such a na\"ive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the top of the stack.
 
-We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}.
+We propose a space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}.
 For each input sentence, we represent the stack with a single $T \times D$ matrix $S$.

### 2016-05-31 writing/hard_stack_paper/paper.tex
Tweak shift-reduce description
Tweak introduction of 'states'.
More thin-stack tweaks

@@
 It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure.
-For any binary-branching tree structure over $N$ words, this requires $T=2N-1$ transitions and $2N$ states.
+For any binary-branching tree structure over $N$ words, this requires $T=2N-1$ transitions through a total of $T+1$ states.
 
@@
 
-We propose a space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}.
+We propose a space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper} that we call \textit{thin stack}.
 For each input sentence, we represent the stack with a single $T \times D$ matrix $S$.
@@
 The top two elements of the stack can be found by using the final two pointers in the queue $Q$.
-These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack.
+These retrieved elements are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack.
 Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation.
@@
 It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation.
-Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains.
+Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains over both a more na\"ive SPINN implementation and a standard TreeRNN implementation.
 We describe speed results in Section~\ref{sec:speed}.

### 2016-05-31 writing/hard_stack_paper/paper.tex
meta-meta-edit: Use a clearer sexpr notation in thin-stack example

@@
   3 & \reduce & \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} \\
-  4 & \reduce & \word{sat down} & \hphantom{0} \hphantom{0} \underline{1 4} \\
-  5 & & \word{Spot} (\word{sat down}) & \hphantom{0} \underline{\hphantom{0} 5} \\
+  4 & \reduce & (\word{sat down}) & \hphantom{0} \hphantom{0} \underline{1 4} \\
+  5 & & (\word{Spot} (\word{sat down})) & \hphantom{0} \underline{\hphantom{0} 5} \\
   \bottomrule

### 2016-05-31 writing/hard_stack_paper/paper.tex
thin-stack section: remove remark re: buffer representation
Reword NLI experiment section intro

@@
 
-A simpler variant of this technique can be used to represent the buffer, since information is not written to the buffer during model operation.
-It can be stored as a single fixed matrix with a scalar pointer variable indicating which element is the head at each step.
-
 \begin{algorithm}[t]
@@
 We evaluate SPINN on the task of natural language inference \citep[NLI, a.k.a.\ recognizing textual entailment, or RTE;][]{dagan2006pascal}.
-NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use:
+NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences.
+Below is an example sentence pair and judgement from the SNLI corpus which we use in our experiments:
 

### 2016-06-01 writing/hard_stack_paper/paper.tex
Add new NFS grant to acknowledgements

@@
 
-We acknowledge financial support from a Google Faculty Research Award, the Stanford Data Science Initiative, and the National Science Foundation under grant no.~BCS 1456077.
+We acknowledge financial support from a Google Faculty Research Award, the Stanford Data Science Initiative, and the National Science Foundation under grant nos.~BCS 1456077 and IIS 1514268.
 Some of the Tesla K40s used for this research were donated by the NVIDIA Corporation.

### 2016-06-06 writing/hard_stack_paper/paper.tex
CP's changes

@@
 \And
-Christopher Potts$^{1}$\\
+Christopher Potts$^{1,6}$\\
 \texttt{\small cgpotts@stanford.edu}
@@
 NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences.
-Below is an example sentence pair and judgement from the SNLI corpus which we use in our experiments:
+Below is an example sentence pair and judgment from the SNLI corpus which we use in our experiments:
 
@@
 \label{tab:hyperparams}
-Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log--uniform.
+Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log-uniform.
 }
@@
 \caption{\protect\label{tab:results}Results on SNLI 3-way inference classification.
-Params.
-is the approximate number of trained parameters (excluding word embeddings for all models).
-Trans.
-acc.
-is the model's accuracy in predicting parsing transitions at test time.
+Params.\ is the approximate number of trained parameters (excluding word embeddings for all models).
+Trans.\ acc.\ is the model's accuracy in predicting parsing transitions at test time.
 Train and test are SNLI classification accuracy.}
@@
 
-Because this paper aims to introduce a general purpose model for sentence encoding, we do not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention based models like \citet{rocktaschel2015reasoning,wang2015learning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 86.1\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
+Because this paper aims to introduce a general purpose model for sentence encoding, we do not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention-based models like \citet{rocktaschel2015reasoning,wang2015learning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 86.3\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 

### 2016-06-07 writing/hard_stack_paper/paper.tex
CM's changes

@@
 SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces the required parse structure on the fly.
-This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser and saving its additional computational cost.
+This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser at nearly no additional computational cost.
 Secondly, it supports batched computation for both parsed and unparsed sentences, yielding dramatic speedups over standard TreeRNNs.
@@
   \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$.
-This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back on to the stack.
+This operation pops the top two elements from the stack, merges them, and pushes the result back on to the stack.
 \end{description}
@@
 
-The TreeLSTM composition function \citep{tai2015improved} generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$.
+The TreeLSTM composition function \citep{tai2015improved} generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of an active state representation $\vec{h}$ and a memory representation $\vec{c}$.
 Our version is formulated as:
@@
 
-We use word representations based on the standard 300D vectors provided with GloVe \citep{pennington2014glove}.
+We use word representations based on the 300D vectors provided with GloVe \citep{pennington2014glove}.
 We do not update these representations during training.
@@
 The prediction function is trained to mimic the decisions of an external parser.
-These decisions are used as the inputs to the model during training.
+These decisions are used as inputs to the model during training.
 For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus.
@@
 \centering
-\begin{tabular}{cc @{\hspace*{2em}}c @{\hspace*{2em}} l}
+\begin{tabular}{c @{\hspace*{1.5em}}c @{\hspace*{1.5em}} lc}
   \toprule
-  $t$ & $a_t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} \\
+  $t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} & $a_t$ \\
   \midrule
-  0 & \shift &  & \underline{\hphantom{0} \hphantom{0}} \\
-  1 & \shift & \word{Spot} & \hphantom{0} \underline{\hphantom{0} 1} \\
-  2 & \shift & \word{sat} & \hphantom{0} \hphantom{0} \underline{1 2} \\
-  3 & \reduce & \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} \\
-  4 & \reduce & (\word{sat down}) & \hphantom{0} \hphantom{0} \underline{1 4} \\
-  5 & & (\word{Spot} (\word{sat down})) & \hphantom{0} \underline{\hphantom{0} 5} \\
+  0 &   & \underline{\hphantom{0} \hphantom{0}} & \shift \\
+  1 &  \word{Spot} & \hphantom{0} \underline{\hphantom{0} 1} & \shift \\
+  2 &  \word{sat} & \hphantom{0} \hphantom{0} \underline{1 2} & \shift \\
+  3  &   \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} & \reduce\\
+  4  &   (\word{sat down}) & \hphantom{0} \hphantom{0} \underline{1 4} & \reduce\\
+  5  & (\word{Spot} (\word{sat down})) & \hphantom{0} \underline{\hphantom{0} 5} &        \\
   \bottomrule
@@
 
-A na\"ive implementation of SPINN needs to handle a size $O(N)$ stack at each timestep in an example, any element of which may be involved in later computations.
+A na\"ive implementation of SPINN needs to handle a size $O(N)$ stack at each timestep, any element of which may be involved in later computations.
 A na\"ive backpropagation implementation would then require storing each of the $O(N)$ stacks for a backward pass, leading to a per-example space requirement of $O(NTD)$ floats.
@@
 \begin{tikzpicture}
-\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
+\begin{axis}[ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
              xticklabels={0,32,64,128,256,512,1024,2048}, xmode=log, log basis x=2,
@@
 We fix the model dimension $D$ and the word embedding dimension at 300.
-We run the CPU performance test on a 2.20-GHz 16-core Intel Xeon E5-2660 processor with hyperthreading enabled.
+We run the CPU performance test on a 2.20 GHz 16-core Intel Xeon E5-2660 processor with hyperthreading enabled.
 We test our thin-stack implementation and the RNN model on an NVIDIA Titan X GPU.
@@
 
-Although NLI is framed as a simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning.
-In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
-
-SNLI is a corpus of 570k human-labeled pairs of scene descriptions like the one above.
+SNLI is a corpus of 570k human-labeled pairs of scene descriptions like this one.
 We use the standard train--test split and ignore unlabeled examples, which leaves about 549k examples for training, 9,842 for development, and 9,824 for testing.
 SNLI labels are roughly balanced, with the most frequent label, {\it entailment}, making up 34.2\% of the test set.
+
+Although NLI is framed as a simple three-way classification task, it is nonetheless an effective way of evaluating the ability of a model to extract broadly informative representations of sentence meaning.
+In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.
 
@@
 \toprule
-Param.     & Range & Strategy        & RNN       & SPINN-PI-NT   & SPINN-PI  & SPINN \\
+Param.     & Range & Strategy        & RNN       & SP.-PI-NT   & SP.-PI  & SP. \\
 \midrule 
-Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & 2e-3\\
-L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & 3e-5\\
+Initial LR & $2\times 10^{-4}$--$2\times 10^{-2}$ & \textsc{log} & $5\times 10^{-3}$  & $3\times 10^{-4}$ & $7\times 10^{-3}$  & $2\times 10^{-3}$\\
+L2 regularization $\lambda$ & $8\times 10^{-7}$--$3\times 10^{-5}$   & \textsc{log} & $4\times 10^{-6}$  & $3\times 10^{-6}$   & $2\times 10^{-5}$  & $3\times 10^{-5}$\\
 Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & --- & --- & ---  & 3.9    \\
-Embedding transformation dropout keep rate & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
-Classifier MLP dropout keep rate & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
+Embedding transformation dropout & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
+Classifier MLP dropout & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
 Tracking LSTM size $D_\text{tracking}$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
@@
 \label{tab:hyperparams}
-Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log-uniform.
+Hyperparameter ranges and values.
+\ii{Range} shows the hyperparameter ranges explored during random search.
+\ii{Strategy} indicates whether sampling from the range was uniform, or log-uniform.
+Dropout parameters are expressed as keep rates rather than drop rates.
 }
@@
 
-Table~\ref{tab:results} shows our results on SNLI inference classification.
+Table~\ref{tab:results} shows our results on SNLI.
 For the full SPINN, we also report a measure of agreement between this model's parses and the parses included with SNLI, calculated as classification accuracy over transitions averaged across timesteps.
@@
 We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
-We suspect that it is acting somewhat like the tree-based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
+It may act somewhat like the tree-based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.
 
@@
 For a more ambitious goal, we expect that it should be possible to implement a variant of SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations \citep[as in][]{grefenstette2015learning,joulin2015inferring}.
-This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing it to learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
+This would make it possible for the model to learn to parse using guidance from the semantic representation objective, which currently is blocked from influencing the key parsing parameters by our use of hard \shift/\reduce\ decisions.
+This change would allow the model to learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.
 
@@
 Some of the Tesla K40s used for this research were donated by the NVIDIA Corporation.
-We also thank Kelvin Gu, Noah Goodman, and many others in the Stanford NLP group for helpful comments.
+We also thank Kelvin Guu, Noah Goodman, and many others in the Stanford NLP group for helpful comments.
 

### 2016-06-07 writing/hard_stack_paper/paper.tex
Slight additions to error analysis.

@@
 The RNN model falls off in test accuracy more quickly with increasing sentence length than SPINN-PI-NT, which in turn falls of substantially faster than the two hybrid models, repeating a pattern seen more dramatically on artificial data in \citet{bowman2015trees}.
-For example, all three SPINN models labeled this pair correctly, while the RNN did not:
+On pairs with premises of 20 or more words, the RNN's 76.7\% accuracy, while SPINN-PI reaches 80.2\%.
+All three SPINN models labeled the following example correctly, while the RNN did not:
 
@@
 
-We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
+We suspect that the hybrid nature of the full SPINN model is also responsible for its surprising ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses.
 It may act somewhat like the tree-based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.

### 2016-06-07 writing/hard_stack_paper/paper.tex
Tidying.
Figure placement tweak.
Footnote format fix

@@
 \usepackage[leqno, fleqn]{amsmath}
-\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black, draft]{hyperref}
+\usepackage[breaklinks, colorlinks, linkcolor=black, urlcolor=black, citecolor=black]{hyperref}
 \usepackage{natbib}
@@
 
-\begin{table}[t]
-\centering
-\begin{tabular}{c @{\hspace*{1.5em}}c @{\hspace*{1.5em}} lc}
-  \toprule
-  $t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} & $a_t$ \\
-  \midrule
-  0 &   & \underline{\hphantom{0} \hphantom{0}} & \shift \\
-  1 &  \word{Spot} & \hphantom{0} \underline{\hphantom{0} 1} & \shift \\
-  2 &  \word{sat} & \hphantom{0} \hphantom{0} \underline{1 2} & \shift \\
-  3  &   \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} & \reduce\\
-  4  &   (\word{sat down}) & \hphantom{0} \hphantom{0} \underline{1 4} & \reduce\\
-  5  & (\word{Spot} (\word{sat down})) & \hphantom{0} \underline{\hphantom{0} 5} &        \\
-  \bottomrule
-\end{tabular}
-\caption{The thin-stack algorithm operating on the input sequence $\mathbf x = \text{(\word{Spot}, \word{sat}, \word{down})}$ and the transition sequence shown in the second column.
-$S$ is shown in the third column and represents the top of the stack at each step $t$.
-The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next step.}
-\label{tbl:thin-stack}
-\end{table}
-
 \paragraph{Representing the stack efficiently}
@@
 
-This stack representation requires substantially less space.
-It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation.
-Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains over both a more na\"ive SPINN implementation and a standard TreeRNN implementation.
-We describe speed results in Section~\ref{sec:speed}.
-
 \begin{algorithm}[t]
@@
 
+This stack representation requires substantially less space.
+It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation.
+Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains over both a more na\"ive SPINN implementation and a standard TreeRNN implementation.
+We describe speed results in Section~\ref{sec:speed}.
+
 \paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf a$  and a token sequence $\mathbf x$ as its inputs for each sentence.
@@
 $\mathbf a$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift\ transition and each `)' as a \reduce\ transition, as here:
+
+\begin{table}[t]
+\centering
+\begin{tabular}{c @{\hspace*{1.5em}}c @{\hspace*{1.5em}} lc}
+  \toprule
+  $t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} & $a_t$ \\
+  \midrule
+  0 &   & \underline{\hphantom{0} \hphantom{0}} & \shift \\
+  1 &  \word{Spot} & \hphantom{0} \underline{\hphantom{0} 1} & \shift \\
+  2 &  \word{sat} & \hphantom{0} \hphantom{0} \underline{1 2} & \shift \\
+  3  &   \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} & \reduce\\
+  4  &   (\word{sat down}) & \hphantom{0} \hphantom{0} \underline{1 4} & \reduce\\
+  5  & (\word{Spot} (\word{sat down})) & \hphantom{0} \underline{\hphantom{0} 5} &        \\
+  \bottomrule
+\end{tabular}
+\caption{The thin-stack algorithm operating on the input sequence $\mathbf x = \text{(\word{Spot}, \word{sat}, \word{down})}$ and the transition sequence shown in the second column.
+$S$ is shown in the third column and represents the top of the stack at each step $t$.
+The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next step.}
+\label{tbl:thin-stack}
+\end{table}
 
@@
                         & 2.7m                  & 92.4           &   87.2    &   82.6      \\
-% \result{300D SPINN-PI, sequence-based attn.   }
-%                         & ?                  & --                &   ?      &   86.3dv         \\
-% \result{300D SPINN-PI, tree-based attn. }
-%                         & ?                  & --                &   ?      &   \textbf{?}\\
-% \midrule
-% \multicolumn{5}{c}{Other previous NN results}\\
-% \midrule
-% 100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
-%                         & 252k               & --              &   85.3   &   83.5       \\
-% 300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
-%                         & 1.9m               & --             &   92.0   &   86.1      \\
-% 300D LSTMN with deep attention fusion \citep{cheng2016long}
-%                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
     \bottomrule
@@
 
-Because this paper aims to introduce a general purpose model for sentence encoding, we do not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention-based models like \citet{rocktaschel2015reasoning,wang2015learning} and the unpublished \citet{cheng2016long} have shown accuracies as high as 86.3\% on SNLI, but are more narrowly engineered to suit the task, and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
+Because this paper aims to introduce a general purpose model for sentence encoding, we do not pursue the use of soft attention \citep{bahdanau2014neural,rocktaschel2015reasoning}, despite its demonstrated effectiveness on the SNLI task.\footnote{Attention-based models like \citet{rocktaschel2015reasoning}, \citet{wang2015learning}, and the unpublished \citet{cheng2016long} have shown accuracies as high as 86.3\% on SNLI, but are more narrowly engineered to suit the task and do not yield sentence encodings.} However, we expect that it should be possible to productively combine our model with soft attention to reach state-of-the-art performance.
 

### 2016-06-13 checkpoints/commands.sh
Minor commands.sh improvements

@@ -2,12 +2,15 @@
 
 # Script to do snli runs from pretrained models.
 
-# The directory where logs should be stored. Must exist.
+# The directory where logs should be stored.
 export LOG_DIR=logs
+mkdir -p $LOG_DIR
 # The path to pretrained embeddings.
 export EMBEDDING_PATH=../glove/glove.840B.300d.txt
 
 export PYTHONPATH=../python
+
+export THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=gpu,floatX=float32
 
 # The invocations below will load the pretrained models and continue training
 # by default. Use the flag --expanded_eval_only mode to do eval-only runs, and delete the flag --ckpt_path ... to train from scratch.
@@ -15,20 +18,20 @@ export PYTHONPATH=../python
 # RNN
 export REMBED_FLAGS="  --batch_size 32 --ckpt_path checkpoints/rnn.ckpt_best --data_type snli --embedding_data_path $EMBEDDING_PATH --embedding_keep_rate 0.852564448733 --eval_data_path snli_1.0/snli_1.0_dev.jsonl --eval_seq_length 25 --init_range 0.005 --l2_lambda 4.42556134893e-06 --learning_rate 0.00464868093302 --log_path $LOG_DIR  --model_dim 600 --model_type RNN --num_sentence_pair_combination_layers 2 --semantic_classifier_keep_rate 0.883392584372 --seq_length 25 --tracking_lstm_hidden_dim 33 --training_data_path snli_1.0/snli_1.0_train.jsonl --word_embedding_dim 300 --experiment_name rnn "
 
-THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=gpu,floatX=float32 python -m spinn.models.fat_classifier $REMBED_FLAGS
+python -m spinn.models.fat_classifier $REMBED_FLAGS
 
 # SPINN-PI-NT
[…]

### 2016-06-13 python/spinn/models/fat_classifier.py
fat_classifier: ignore on unused input to Theano function

@@ -381,12 +381,12 @@ def evaluate_expanded(eval_fn, eval_set, eval_path, logger, step, sentence_pair_
                 # write each predicted transition to file
                 for orig_transitions, pred_logit_hyp, pred_logit_prem, tokens, true_class, example_sem_logits \
                         in zip(eval_transitions_batch, logits_pred_hyp,
-                               logits_pred_prem, eval_X_batch, eval_y_batch, sem_logit_values):    
+                               logits_pred_prem, eval_X_batch, eval_y_batch, sem_logit_values):
                     if predict_transitions:
                         orig_hyp_transitions, orig_prem_transitions = orig_transitions.T
                         pred_hyp_transitions = pred_logit_hyp.argmax(axis=1)
                         pred_prem_transitions = pred_logit_prem.argmax(axis=1)
-                    else: 
+                    else:
                         orig_hyp_transitions = orig_prem_transitions = pred_hyp_transitions = pred_prem_transitions = None
 
                     hyp_tokens, prem_tokens = tokens.T
@@ -609,13 +609,13 @@ def run(only_forward=False):
             eval_fn = theano.function(
                 [X, transitions, y, num_transitions, training_mode, ground_truth_transitions_visible, ss_prob],
                 [acc, action_acc, logits, predicted_hypothesis_transitions, predicted_premise_transitions],
-                on_unused_input='warn',
+                on_unused_input='ignore',
                 allow_input_downcast=True)
         else:
             eval_fn = theano.function(
                 [X, transitions, y, num_transitions, training_mode, ground_truth_transitions_visible, ss_prob],
                 [acc, action_acc, logits, predicted_transitions],
-                on_unused_input='warn',
+                on_unused_input='ignore',
                 allow_input_downcast=True)
 
         # Generate the inverse vocabulary lookup table.
@@ -643,11 +643,11 @@ def run(only_forward=False):
             [X, transitions, y, num_transitions, lr, training_mode, ground_truth_transitions_visible, ss_prob],
             [total_cost, xent_cost, transition_cost, action_acc, l2_cost, acc],
             updates=new_values,
-            on_unused_input='warn',
+            on_unused_input='ignore',
             allow_input_downcast=True)
         logger.Log("Building eval function.")
         eval_fn = theano.function([X, transitions, y, num_transitions, training_mode, ground_truth_transitions_visible, ss_prob], [acc, action_acc],
-            on_unused_input='warn',
+            on_unused_input='ignore',
             allow_input_downcast=True)
         logger.Log("Training.")
 

### 2016-06-13 checkpoints/commands.sh
commands.sh: add friendly helpers to download SNLI, GloVe data

@@ -5,33 +5,57 @@
 # The directory where logs should be stored.
 export LOG_DIR=logs
 mkdir -p $LOG_DIR
+
 # The path to pretrained embeddings.
 export EMBEDDING_PATH=../glove/glove.840B.300d.txt
+if [ ! -e "$EMBEDDING_PATH" ]; then
+    echo "Could not find GloVe embeddings at $EMBEDDING_PATH."
+    read -p "Should we download them? (~2G download) (y/n) " yn
+    if echo "$yn" | grep -iq "[^yY]"; then
+        exit 1
+    fi
+
+    mkdir -p `dirname "$EMBEDDING_PATH"`
+    wget http://nlp.stanford.edu/data/glove.840B.300d.zip -O glove.zip \
+        || (echo "Failed to download GloVe embeddings." >&2 && exit 1)
+    unzip -d `dirname "$EMBEDDING_PATH"` glove.zip && rm glove.zip
+fi
+
+# Prepare SNLI data.
+export SNLI_DIR=../snli_1.0
+export SNLI_TRAIN_JSONL=$SNLI_DIR/snli_1.0_train.jsonl
+export SNLI_DEV_JSONL=$SNLI_DIR/snli_1.0_dev.jsonl
+export SNLI_TEST_JSONL=$SNLI_DIR/snli_1.0_test.jsonl
+if [ ! -d "$SNLI_DIR" ]; then
+    echo "Downloading SNLI data." >&2
+    wget http://nlp.stanford.edu/projects/snli/snli_1.0.zip -O snli_1.0.zip \
+        || (echo "Failed to download SNLI data." >&2 && exit 1)
+    unzip -d .. snli_1.0.zip && rm snli_1.0.zip
+fi
 
 export PYTHONPATH=../python
-
 export THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=gpu,floatX=float32
 
 # The invocations below will load the pretrained models and continue training
 # by default. Use the flag --expanded_eval_only mode to do eval-only runs, and delete the flag --ckpt_path ... to train from scratch.
 
 # RNN
[…]

### 2016-06-13 checkpoints/_init.sh
Split checkpoint runs into separate invocation files with common header

@@ -1,0 +1,37 @@
+#!/usr/bin/env bash
+
+# Header script to prepare to perform a SPINN SNLI run.
+
+# The directory where logs should be stored.
+export LOG_DIR=logs
+mkdir -p $LOG_DIR
+
+# The path to pretrained embeddings.
+export EMBEDDING_PATH=../glove/glove.840B.300d.txt
+if [ ! -e "$EMBEDDING_PATH" ]; then
+    echo "Could not find GloVe embeddings at $EMBEDDING_PATH."
+    read -p "Should we download them? (~2G download) (y/n) " yn
+    if echo "$yn" | grep -iq "[^yY]"; then
+        exit 1
+    fi
+
+    mkdir -p `dirname "$EMBEDDING_PATH"`
+    wget http://nlp.stanford.edu/data/glove.840B.300d.zip -O glove.zip \
+        || (echo "Failed to download GloVe embeddings." >&2 && exit 1)
+    unzip -d `dirname "$EMBEDDING_PATH"` glove.zip && rm glove.zip
+fi
+
+# Prepare SNLI data.
+export SNLI_DIR=../snli_1.0
+export SNLI_TRAIN_JSONL=$SNLI_DIR/snli_1.0_train.jsonl
+export SNLI_DEV_JSONL=$SNLI_DIR/snli_1.0_dev.jsonl
+export SNLI_TEST_JSONL=$SNLI_DIR/snli_1.0_test.jsonl
+if [ ! -d "$SNLI_DIR" ]; then
+    echo "Downloading SNLI data." >&2
+    wget http://nlp.stanford.edu/projects/snli/snli_1.0.zip -O snli_1.0.zip \
+        || (echo "Failed to download SNLI data." >&2 && exit 1)
+    unzip -d .. snli_1.0.zip && rm snli_1.0.zip
+fi
+
+export PYTHONPATH=../python
+export THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=gpu,floatX=float32

### 2016-06-13 checkpoints/commands.sh

@@ -1,61 +1,0 @@
-#!/usr/bin/env bash
-
-# Script to do snli runs from pretrained models.
-
-# The directory where logs should be stored.
-export LOG_DIR=logs
-mkdir -p $LOG_DIR
-
-# The path to pretrained embeddings.
-export EMBEDDING_PATH=../glove/glove.840B.300d.txt
-if [ ! -e "$EMBEDDING_PATH" ]; then
-    echo "Could not find GloVe embeddings at $EMBEDDING_PATH."
-    read -p "Should we download them? (~2G download) (y/n) " yn
-    if echo "$yn" | grep -iq "[^yY]"; then
-        exit 1
-    fi
-
-    mkdir -p `dirname "$EMBEDDING_PATH"`
-    wget http://nlp.stanford.edu/data/glove.840B.300d.zip -O glove.zip \
-        || (echo "Failed to download GloVe embeddings." >&2 && exit 1)
-    unzip -d `dirname "$EMBEDDING_PATH"` glove.zip && rm glove.zip
-fi
-
-# Prepare SNLI data.
-export SNLI_DIR=../snli_1.0
-export SNLI_TRAIN_JSONL=$SNLI_DIR/snli_1.0_train.jsonl
-export SNLI_DEV_JSONL=$SNLI_DIR/snli_1.0_dev.jsonl
-export SNLI_TEST_JSONL=$SNLI_DIR/snli_1.0_test.jsonl
-if [ ! -d "$SNLI_DIR" ]; then
-    echo "Downloading SNLI data." >&2
-    wget http://nlp.stanford.edu/projects/snli/snli_1.0.zip -O snli_1.0.zip \
-        || (echo "Failed to download SNLI data." >&2 && exit 1)
-    unzip -d .. snli_1.0.zip && rm snli_1.0.zip
-fi
-
-export PYTHONPATH=../python
-export THEANO_FLAGS=allow_gc=False,cuda.root=/usr/bin/cuda,warn_float64=warn,device=gpu,floatX=float32
-
-# The invocations below will load the pretrained models and continue training
-# by default. Use the flag --expanded_eval_only mode to do eval-only runs, and delete the flag --ckpt_path ... to train from scratch.
-
-# RNN
[…]

### 2016-06-13 checkpoints/rnn.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+
+source _init.sh
+
+# The invocation below will load the pretrained models and continue training
+# by default. Use the flag --expanded_eval_only mode to do eval-only runs, and delete the flag --ckpt_path ... to train from scratch.
+
+# RNN
+export REMBED_FLAGS="  --batch_size 32 --ckpt_path checkpoints/rnn.ckpt_best --data_type snli --embedding_data_path $EMBEDDING_PATH --embedding_keep_rate 0.852564448733 --eval_data_path $SNLI_DEV_JSONL --eval_seq_length 25 --init_range 0.005 --l2_lambda 4.42556134893e-06 --learning_rate 0.00464868093302 --log_path $LOG_DIR  --model_dim 600 --model_type RNN --num_sentence_pair_combination_layers 2 --semantic_classifier_keep_rate 0.883392584372 --seq_length 25 --tracking_lstm_hidden_dim 33 --training_data_path $SNLI_TRAIN_JSONL --word_embedding_dim 300 --experiment_name rnn "
+
+python -m spinn.models.fat_classifier $REMBED_FLAGS
+

### 2016-06-13 checkpoints/spinn.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+
+source _init.sh
+
+# The invocation below will load the pretrained models and continue training
+# by default. Use the flag --expanded_eval_only mode to do eval-only runs, and delete the flag --ckpt_path ... to train from scratch.
+
+# SPINN
+export REMBED_FLAGS="  --batch_size 32 --ckpt_path checkpoints/spinn.ckpt_best --connect_tracking_comp  --data_type snli --embedding_data_path $EMBEDDING_PATH --embedding_keep_rate 0.938514416034 --eval_data_path $SNLI_DEV_JSONL --eval_seq_length 50 --init_range 0.005 --l2_lambda 2.76018187539e-05 --learning_rate 0.00103428201391 --log_path $LOG_DIR  --model_dim 600 --model_type Model1 --num_sentence_pair_combination_layers 1 --semantic_classifier_keep_rate 0.949455648614 --seq_length 50 --tracking_lstm_hidden_dim 44 --training_data_path $SNLI_TRAIN_JSONL --transition_cost_scale 0.605159568546 --use_tracking_lstm  --word_embedding_dim 300 --experiment_name spinn "
+
+python -m spinn.models.fat_classifier $REMBED_FLAGS
+

### 2016-06-13 checkpoints/spinn_pi.sh

@@ -1,0 +1,13 @@
+#!/usr/bin/env bash
+
+source _init.sh
+
+# The invocation below will load the pretrained models and continue training
+# by default. Use the flag --expanded_eval_only mode to do eval-only runs, and delete the flag --ckpt_path ... to train from scratch.
+
+# SPINN-PI
+export REMBED_FLAGS="  --batch_size 32 --ckpt_path checkpoints/spinn_pi.ckpt_best --connect_tracking_comp  --data_type snli --embedding_data_path $EMBEDDING_PATH --embedding_keep_rate 0.917969380132 --eval_data_path $SNLI_DEV_JSONL --eval_seq_length 50 --init_range 0.005 --l2_lambda 2.00098223698e-05 --learning_rate 0.00701855401337 --log_path $LOG_DIR  --model_dim 600 --model_type Model0 --num_sentence_pair_combination_layers 2 --semantic_classifier_keep_rate 0.934741728838 --seq_length 50 --tracking_lstm_hidden_dim 61 --training_data_path $SNLI_TRAIN_JSONL --use_tracking_lstm  --word_embedding_dim 300 --experiment_name spinn_pi "
+
+python -m spinn.models.fat_classifier $REMBED_FLAGS
+
+

### 2016-06-13 checkpoints/spinn_pi_nt.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+
+source _init.sh
+
+# The invocation below will load the pretrained models and continue training
+# by default. Use the flag --expanded_eval_only mode to do eval-only runs, and delete the flag --ckpt_path ... to train from scratch.
+
+# SPINN-PI-NT
+export REMBED_FLAGS="  --batch_size 32 --ckpt_path checkpoints/spinn_pi_nt.ckpt_best --data_type snli --embedding_data_path $EMBEDDING_PATH --embedding_keep_rate 0.828528124124 --eval_data_path $SNLI_DEV_JSONL --eval_seq_length 50 --init_range 0.005 --l2_lambda 3.45058959758e-06 --learning_rate 0.000297682444894 --log_path $LOG_DIR  --model_dim 600 --model_type Model0 --noconnect_tracking_comp  --num_sentence_pair_combination_layers 2 --semantic_classifier_keep_rate 0.9437038157 --seq_length 50 --tracking_lstm_hidden_dim 57 --training_data_path $SNLI_TRAIN_JSONL --use_tracking_lstm  --word_embedding_dim 300 --experiment_name spinn_pi_nt "
+
+python -m spinn.models.fat_classifier $REMBED_FLAGS
+

### 2016-06-13 writing/hard_stack_paper/paper.tex
Bib for paper submission; fix caption error.

@@
 \end{tabular}
-\caption{The thin-stack algorithm operating on the input sequence $\mathbf x = \text{(\word{Spot}, \word{sat}, \word{down})}$ and the transition sequence shown in the second column.
-$S$ is shown in the third column and represents the top of the stack at each step $t$.
+\caption{The thin-stack algorithm operating on the input sequence $\mathbf x = \text{(\word{Spot}, \word{sat}, \word{down})}$ and the transition sequence shown in the rightmost column.
+$S$[$t$] shows the top of the stack at each step $t$.
 The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next step.}

### 2016-06-13 checkpoints/spinn_pi_nt.sh
Support running checkpoint scripts from root dir

@@ -1,5 +1,9 @@
 #!/usr/bin/env bash
 
+if [ -e README.md ]; then
+    cd checkpoints
+fi
+
 source _init.sh
 
 # The invocation below will load the pretrained models and continue training

### 2016-06-13 checkpoints/_init.sh
default log dir to root

@@ -3,7 +3,7 @@
 # Header script to prepare to perform a SPINN SNLI run.
 
 # The directory where logs should be stored.
-export LOG_DIR=logs
+export LOG_DIR=../logs
 mkdir -p $LOG_DIR
 
 # The path to pretrained embeddings.
